### 2018-10-15 notebooks/article_notebook.Rmd
added new figure on all resistance mechanisms

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This notebook is using the following packages:

- dplyr

- ggplot2

- purrr
```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

```{r Import data, include = FALSE}

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

two_reads <- read.table("../data/no_of_reads.txt",
                          sep = "\t",
                          header = F,
                          stringsAsFactors = F) %>%
  rename("ref" = V1,
         "reads" = V2) %>%
  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
  group_by(ref) %>%
  mutate(id = n()) %>%
  {. ->> four_reads} %>%
  filter(id == 2) %>%
  spread(type, reads) %>%
  rename("R1-1" = R1,
         "R2-1" = R2) %>%
  mutate("R1-2" = NA,
         "R2-2" = NA) %>%
  select(-id) %>%
  mutate_all(funs(as.character(.)))

four_reads <- four_reads %>%
  filter(id == 4) %>%
  group_by(ref) %>%
  mutate(id = 1:n()) %>%
  spread(id, reads) %>%
  summarise_all(funs(func_paste)) %>%
  rename("R1-1" = `1`,
         "R2-1" = `2`,
         "R1-2" = `3`,
         "R2-2" = `4`) %>%
  select(-type) %>%
  mutate_all(funs(as.character(.)))

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hf150bp <- read.table("../data/mlst/hiseq_flex_150bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hx125bp <- read.table("../data/mlst/hiseq_xt_125bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_hrx250bp <- read.table("../data/mlst/hiseqr_xt_250bp_mlst_summarized_results.tsv",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

mlst_data <- rbind(mlst_hf150bp, mlst_hx125bp, mlst_hrx250bp) %>%
  separate(ST.adk.fumC.gyrB.icd.mdh.purA.recA,
           into = c("x","ST","adk","fumC","gyrB","icd","mdh","purA","recA"),
           sep = " ") %>%
  select(-x) %>%
  rename("ref" = header) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)
```

# Sequence data
## Distribution of number of reads per sequencer
```{r}
sequence_data <- no_of_reads %>%
  left_join(isolate_data, by = "id") %>%
  mutate_at(vars(`R1-1`,`R2-1`,`R1-2`,`R2-2`),
            funs(as.numeric(as.character(.))))

ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = sequencer,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Sequencer")+
  theme(legend.position = c(0.7, 0.8))
```

## Distribution of number of reads per library prep type
```{r}
ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = libprep,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Library prep kit")+
  theme(legend.position = c(0.7, 0.8))
```

## Distribution of number of reads per animal species
```{r}
ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
  geom_density(aes(fill = species,
                   y = ..scaled..),
               alpha = 0.6)+
  labs(x = "Number of reads",
       y = "Density",
       fill = "Animal species")+
  theme(legend.position = c(0.7, 0.8))
```

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx")
mut_genes <- c("gyr","par","marR","soxR","acrR")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions

func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result)) %>%
    rename("gene" = gene_names)
  return(df)
}

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1),
           lwr = round(get_binCI(Present, Total)[1], 1),
           upr = round(get_binCI(Present, Total)[2], 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

acquired_filtered <- filter_acquired_table(acquired_table)

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref)

# ---------------------------------- Tables

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count()
```

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## List of genes of interest
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxs", cluster, ignore.case = T) == TRUE ~ "soxS",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxS" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump."),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

```

## Isolates per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

per_species
```

## Percent isolates with PMQR
```{r}
acquired_report %>%
  select(-id) %>%
  mutate_all(funs(as.numeric(.))) %>%
  mutate(total = rowSums(.),
         result = if_else(total != 0, 1, 0)) %>%
  group_by(result) %>%
  count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
  spread(result, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr))
```

## Percent isolates with specific qnr genes
```{r}
acquired_report %>%
  select(-id) %>%
  t() %>%
  as.data.frame(.) %>%
  mutate(Gene = rownames(.)) %>%
  mutate_at(vars(-Gene),
            funs(as.numeric(as.character(.)))) %>%
  mutate(Present = rowSums(.[,-286])) %>%
  select(c(Gene, Present)) %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent isolates with mutations in QRDR
```{r}
mut_report %>%
  select(-c(id, marR, nt_change)) %>%
  group_by_all() %>%
  count() %>%
  ungroup() %>%
  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
            .funs = funs(as.numeric)) %>%
  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
  select(n, type) %>%
  group_by(type) %>%
  summarise_all(funs(sum)) %>%
  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
  spread(type, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,2),
         lwr = round(get_binCI(Present, Total)[1], 2),
         upr = round(get_binCI(Present, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent of isolates with specific mutation combinations in QRDR
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count() %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,2),
         lwr = round(get_binCI(n, Total)[1], 2),
         upr = round(get_binCI(n, Total)[2], 2),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total))
```

## Percent isolates with specific nucleotide changes in QRDR
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  gather(gene, mut, -ref) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  select(-mut) %>%
  rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,1),
         lwr = round(get_binCI(n, Total)[1],1),
         upr = round(get_binCI(n, Total)[2],1),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total)) %>%
  rename("Mutation" = mut))
```

## Total combinations
```{r}
acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("marR","id")], by = "id") %>%
  left_join(isolate_data[c("id", "species")]) %>%
  select(-id) %>%
  group_by_all() %>%
  summarise(n = n()) %>%
  rowwise() %>%
  mutate(Total = 285,
         Percent = round(n/Total * 100,1),
         lwr = round(get_binCI(n, Total)[1], 1),
         upr = round(get_binCI(n, Total)[2], 1),
         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
  select(-c(lwr, upr, Total)) %>%
  arrange(desc(n))
```

# Figures
## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
```{r}
total_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 285,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1))

genes <- c(names(mut_report), names(acquired_report))
genes <- genes[!genes == "id"]

p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
  geom_col(color = "black")+
  labs(y = "Percent (%) isolates")+
  scale_fill_viridis(discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top",
        legend.spacing.x = unit(0.27, "cm"))

dot_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit()

p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
  geom_point(size = 2, pch = 21)+
  scale_fill_viridis(option = "B",
                     discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())

plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))

```

## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-c(1,2)]

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```

## Total resistance mechanisms per animal species
```{r}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

total_df <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("marR","id")], by = "id") %>%
  left_join(isolate_data[c("id", "species")])

col_proc <- c("gyrA","gyrB","parC","parE")

id <- total_df$id

cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  cbind(id) %>%
  gather(key, value, -id) %>%
  filter(grepl("(.*?)_0", key) == FALSE) %>%
  spread(key, value)

total_df2 <- total_df %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(cols_to_add, by = "id") %>%
  select(id, species, everything()) %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,3:9]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-id) %>%
  mutate_at(vars(-species, CIP),
            funs(as.numeric(as.character(.)))) %>%
  group_by(species, CIP) %>%
  summarise_all(funs(sum, n = n())) %>%
  select(-c(qnr_n,gyrA_D87G_n,gyrA_D87H_n,
            gyrA_D87N_n,gyrA_D87Y_n,gyrA_S83A_n,
            gyrA_S83L_n,parC_A56T_n,parC_E84V_n,
            parC_S57T_n,parC_S58I_n,parC_S80I_n,
            parC_S80R_n,parE_A512T_n,parE_D463N_n,
            parE_D475E_n,parE_H516Y_n,parE_L416F_n,
            parE_L488M_n,parE_S458A_n)) %>%
  gather(gene, value, -c(species, CIP, marR_n)) %>%
  rename("n" = marR_n) %>%
  mutate(gene2 = sub("(.*?)_(.*?)_sum", "\\1", gene),
         gene2 = sub("(.*?)_sum", "\\1", gene2),
         gene = sub("(.*?)_(.*?)_sum", "\\2", gene),
         gene = sub("(.*?)_sum", "\\1", gene),
         type = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "qnr" ~ 6),
         perc = value/n*100)

ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
  geom_tile(color = "grey70")+
  annotate("rect", xmin = 0, xmax = 6.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
  annotate("rect", xmin = 6.5, xmax = 12.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
  annotate("rect", xmin = 12.5, xmax = 19.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
  annotate("text", x = 3.25, y = -0.05, label = "gyrA", size = 2.5)+
  annotate("text", x = 9.5, y = -0.05, label = "parC", size = 2.5)+
  annotate("text", x = 16, y = -0.05, label = "parE", size = 2.5)+
  scale_fill_viridis(option = "A",
                     begin = 0.1,
                     guide = guide_colorbar(ticks = F,
                                            barheight = 0.5,
                                            barwidth = 10))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        legend.title = element_blank(),
        legend.position="bottom",
        legend.justification="center",
        legend.direction="horizontal",
        legend.text = element_text(size = 8),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        strip.text = element_text(size = 10))+
  facet_wrap(~species)+
  coord_fixed()
```

### 2018-10-15 notebooks/article_notebook.Rmd
updated figures

@@
 
-## Total combinations
-```{r}
-acquired_report %>%
+## Significant differences between resistance mechanisms for each animal species
+```{r, warning = F, message = F}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    unlist() %>% unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
+do_chisq_test <- function(df) {
+  df <- df %>%
+    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
+          filter(Var2 < Var1) %>% t) %>%
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    do(data_frame(gene = .$key,
+                  test = first(.$test),
+                  species1 = first(.$species),
+                  species2 = last(.$species),
+                  data = list(matrix(c(.$Absent,
+                                       .$Present),
+                                     ncol = 2)))) %>%
+    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
+    mutate(p.value = map_dbl(chi_test, "p.value"),
+           dupl = duplicated(test)) %>%
+    filter(dupl == FALSE) %>%
+    select(gene, species1, species2, p.value) %>%
+    filter(p.value < 0.05)
+  return(df)
+}
+
+total_df <- acquired_report %>%
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
   left_join(mut_report[c("marR","id")], by = "id") %>%
-  left_join(isolate_data[c("id", "species")]) %>%
+  left_join(isolate_data[c("id", "species")])
+
+col_proc <- c("gyrA","gyrB","parC","parE")
+
+id <- total_df$id
+
+cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
+                      X = seq_along(col_proc)) %>%
+  bind_cols() %>%
+  cbind(id) %>%
+  gather(key, value, -id) %>%
+  filter(grepl("(.*?)_0", key) == FALSE) %>%
+  spread(key, value)
+
+mechanism_per_species <- total_df %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  left_join(cols_to_add, by = "id") %>%
+  select(id, species, everything()) %>%
   select(-id) %>%
+  gather(key, value, -species) %>%
   group_by_all() %>%
-  summarise(n = n()) %>%
-  rowwise() %>%
-  mutate(Total = 285,
-         Percent = round(n/Total * 100,1),
-         lwr = round(get_binCI(n, Total)[1], 1),
-         upr = round(get_binCI(n, Total)[2], 1),
-         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
-  select(-c(lwr, upr, Total)) %>%
-  arrange(desc(n))
+  count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  select(key, species, Present, Absent) %>%
+  split(f = .$key)
+
+p_values_total <- suppressWarnings(lapply(mechanism_per_species, function(x) do_chisq_test(x))) %>%
+  bind_rows()
+
+p_values_total
 ```
@@
 ## Total resistance mechanisms per animal species
-```{r}
-AddCol <- function(df, col_name) {
-  # split rows by delimeters
-  string_to_proc <- df %>% select(!!col_name) %>%
-    unlist() %>% str_split(regex("\\, |\\,")) 
-  # find unique entries
-  unique_strings <- string_to_proc %>%
-    unlist() %>% unique()
-  # construct names of the new columns
-  cols_names <- paste(col_name, unique_strings, sep = "_")
-  # construct 0/1-content columns for each unique entry
-  cols_content <- sapply(function(i) {
-    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
-                             X = string_to_proc)))
-  }, X = seq_along(unique_strings))
-  res <- data.frame(cols_content)
-  names(res) <- cols_names
-  return(res)
-}
-
+```{r, warning=F, message=F}
 total_df <- acquired_report %>%
@@
         axis.text.y = element_text(size = 8),
-        axis.title = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size = 10),
         legend.title = element_blank(),
@@
         strip.text = element_text(size = 10))+
+  labs(y = "CIP MIC")+
   facet_wrap(~species)+

### 2018-10-18 notebooks/article_notebook.Rmd
updated last figure to include n

@@
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","oqx")
-mut_genes <- c("gyr","par","marR","soxR","acrR")
+ac_genes <- c("qnr","oqx","qep")
+mut_genes <- c("gyr","par","marR","soxR","acrR","mcb")
 
@@
 
+## Total combinations
+```{r}
+test <- acquired_report %>%
+  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
+  left_join(mut_report[c("marR","id")], by = "id") %>%
+  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  select(-id) %>%
+  group_by_at(vars(-c(CIP, NAL))) %>%
+  summarise(CIP = ifelse(min(as.numeric(CIP)) == max(as.numeric(CIP)),
+                         as.character(median(as.numeric(CIP))),
+                         paste(min(as.numeric(CIP)), max(as.numeric(CIP)), sep = " - ")),
+            NAL = ifelse(min(as.numeric(NAL)) == max(as.numeric(NAL)),
+                         as.character(median(as.numeric(NAL))),
+                         paste(min(as.numeric(NAL)), max(as.numeric(NAL)), sep = " - ")),
+            n = n()) %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1],1),
+         upr = round(get_binCI(n, Total)[2],1),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
 ## Significant differences between resistance mechanisms for each animal species
@@
 ## Total resistance mechanisms per animal species
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, fig.height=8, fig.width=12}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    unlist() %>% unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
 total_df <- acquired_report %>%
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,3:9]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,4:10]) == 0, 0, 1)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
@@
   summarise_all(funs(sum, n = n())) %>%
-  select(-c(qnr_n,gyrA_D87G_n,gyrA_D87H_n,
+  select(-c(qepA4_n,qnr_n,gyrA_D87G_n,gyrA_D87H_n,
             gyrA_D87N_n,gyrA_D87Y_n,gyrA_S83A_n,
@@
                           gene2 == "marR" ~ 5,
-                          gene2 == "qnr" ~ 6),
+                          gene2 == "qnr" ~ 6,
+                          gene2 == "qepA4" ~ 7),
          perc = value/n*100)
-
-ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
+# 
+# ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
+#   geom_tile(color = "grey70")+
+#   scale_fill_viridis(option = "A",
+#                      begin = 0.1,
+#                      guide = guide_colorbar(ticks = F,
+#                                             barheight = 0.5,
+#                                             barwidth = 10))+
+#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 7),
+#         axis.text.y = element_text(size = 8),
+#         axis.title.x = element_blank(),
+#         axis.title.y = element_text(size = 10),
+#         legend.title = element_blank(),
+#         legend.position="bottom",
+#         legend.justification="center",
+#         legend.direction="horizontal",
+#         legend.text = element_text(size = 8),
+#         strip.background = element_rect(colour = "black", fill = "grey70"),
+#         strip.text = element_text(size = 10))+
+#   labs(y = "CIP MIC")+
+#   facet_wrap(~species)+
+#   coord_fixed()
+
+p1 <- ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
   geom_tile(color = "grey70")+
-  annotate("rect", xmin = 0, xmax = 6.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
-  annotate("rect", xmin = 6.5, xmax = 12.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
-  annotate("rect", xmin = 12.5, xmax = 19.5, ymin = -0.5, ymax = 0.2, fill = "grey95", color = "black")+
-  annotate("text", x = 3.25, y = -0.05, label = "gyrA", size = 2.5)+
-  annotate("text", x = 9.5, y = -0.05, label = "parC", size = 2.5)+
-  annotate("text", x = 16, y = -0.05, label = "parE", size = 2.5)+
   scale_fill_viridis(option = "A",
-                     begin = 0.1,
-                     guide = guide_colorbar(ticks = F,
-                                            barheight = 0.5,
-                                            barwidth = 10))+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 8),
-        axis.text.y = element_text(size = 8),
+                     begin = 0.1)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
+        axis.text.y = element_text(size = 10),
         axis.title.x = element_blank(),
         axis.title.y = element_text(size = 10),
-        legend.title = element_blank(),
-        legend.position="bottom",
-        legend.justification="center",
-        legend.direction="horizontal",
-        legend.text = element_text(size = 8),
         strip.background = element_rect(colour = "black", fill = "grey70"),
-        strip.text = element_text(size = 10))+
+        strip.text = element_text(size = 10),
+        plot.margin = unit(c(0,0,0,0), "cm"))+
   labs(y = "CIP MIC")+
-  facet_wrap(~species)+
+  facet_wrap(~species, ncol = 1)+
   coord_fixed()
-```
+
+nplot_data <- total_df %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(species, CIP) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(species = paste0(species, " n"))
+
+p2 <- ggplot(nplot_data, aes(factor(CIP), n))+
+  geom_col()+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_text(size = 10),
+        axis.ticks.y = element_blank(),
+        strip.text = element_text(size = 10),
+        strip.background = element_rect(colour = "black", fill = "grey70"),
+        plot.margin = unit(c(0,0,-10,0), "cm"))+
+  facet_wrap(~species, ncol = 1)+
+  coord_flip()
+
+p1g <- ggplotGrob(p1)
+p2g <- ggplotGrob(p2)
+
+g <- cbind(p1g, p2g, size = "first")
+
+grid.arrange(g)
+
+```

### 2018-10-23 notebooks/article_notebook.Rmd
added interactive alignment plot for gyrB sequence alignment

@@
 ```
+
+## Correlation plot
+```{r}
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+total_df <- acquired_report %>%
+  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
+  left_join(mut_report[c("marR","id")], by = "id")
+
+col_proc <- c("gyrA","gyrB","parC","parE")
+
+id <- total_df$id
+
+cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
+                      X = seq_along(col_proc)) %>%
+  bind_cols() %>%
+  cbind(id) %>%
+  gather(key, value, -id) %>%
+  filter(grepl("(.*?)_0", key) == FALSE) %>%
+  spread(key, value)
+
+corr_df <- total_df %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  left_join(cols_to_add, by = "id") %>%
+  select(-id) %>%
+  mutate_all(funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,2:8]) == 0, 0, 1)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4))
+
+corr_matrix <- create_corr_matrix(corr_df) %>%
+    mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
+           gene2 = gsub("(.*?)_.+", "\\1", var2),
+           var1 = gsub(".+_(.*?)", "\\1", var1),
+           var2 = gsub(".+_(.*?)", "\\1", var2),
+           type1 = case_when(gene1 == "gyrA" ~ 1,
+                          gene1 == "gyrB" ~ 2,
+                          gene1 == "parC" ~ 3,
+                          gene1 == "parE" ~ 4,
+                          gene1 == "marR" ~ 5,
+                          gene1 == "qnr" ~ 6,
+                          gene1 == "qepA4" ~ 7),
+           type2 = case_when(gene2 == "gyrA" ~ 1,
+                          gene2 == "gyrB" ~ 2,
+                          gene2 == "parC" ~ 3,
+                          gene2 == "parE" ~ 4,
+                          gene2 == "marR" ~ 5,
+                          gene2 == "qnr" ~ 6,
+                          gene2 == "qepA4" ~ 7))
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(corr_matrix, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradientn(colors=cols,
+                       na.value="grey95",
+                       limits=c(-1, 1),
+                       guide=guide_colourbar(ticks=T,nbin=10,
+                                             barheight=.5,label=T, 
+                                             barwidth=10))+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 8),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank(),
+        legend.position="bottom",
+        legend.justification="center",
+        legend.direction="horizontal")+
+  coord_fixed()
+```
+
+# GyrB Protein sequence alignment
+```{r}
+alignment <- Biostrings::readAAMultipleAlignment("../alignment.fasta")
+msaR::msaR(alignment)
+```

### 2018-10-23 notebooks/article_notebook.Rmd
changed size of plot

@@
 ## Total resistance mechanisms per animal species
-```{r, warning=F, message=F, fig.height=8, fig.width=12}
+```{r, warning=F, message=F, fig.height=16, fig.width=20}
 AddCol <- function(df, col_name) {
@@
   guides(fill = FALSE)+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
-        axis.text.y = element_text(size = 10),
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 14),
+        axis.text.y = element_text(size = 14),
         axis.title.x = element_blank(),
-        axis.title.y = element_text(size = 10),
+        axis.title.y = element_text(size = 16),
         strip.background = element_rect(colour = "black", fill = "grey70"),
-        strip.text = element_text(size = 10),
+        strip.text = element_text(size = 14),
         plot.margin = unit(c(0,0,0,0), "cm"))+
@@
         axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 10),
+        axis.text.x = element_text(size = 14),
         axis.ticks.y = element_blank(),
-        strip.text = element_text(size = 10),
+        strip.text = element_text(size = 14),
         strip.background = element_rect(colour = "black", fill = "grey70"),
@@
 ```{r}
-alignment <- Biostrings::readAAMultipleAlignment("../alignment.fasta")
+alignment <- Biostrings::readAAMultipleAlignment("../data/alignment.fasta")
 msaR::msaR(alignment)

### 2018-10-23 alignment_script.R
script for creating gyrB alignments

@@ -1,0 +1,56 @@
+# Testing import of gyrB sequences from .gz files created by ARIBA
+library(Biostrings)
+library(msa)
+library(DECIPHER)
+
+# File locations
+path <- "data/ariba/megares_full/"
+
+## Functions
+# Identifies filenames in input folder
+file_names <- function(filepath) {
+  files <- list.files(path = filepath,
+                      pattern = "assembled_genes.fa.gz",
+                      recursive = TRUE)
+  return(files)
+}
+
+# Import data from assembled_genes.fa.gz files created by ariba
+get_seq_data <- function(filepath) {
+  files <- file_names(filepath)
+  
+  data_list <- lapply(files,
+                      FUN = function(file) {
+                        readDNAStringSet(paste0(filepath, "/", file))
+                      })
+  names(data_list) <- files
+  return(data_list)
+}
+
+# Filter out specific gene of interest
+filter_gyrB <- function(fasta_file, gene) {
+  gyrB_pos <- which(grepl(gene, names(fasta_file), ignore.case = TRUE) == TRUE)
+  gyrB_seq <- fasta_file[gyrB_pos]
+  return(gyrB_seq)
+}
+
+# Append sequences to fasta file
+append_fasta_file <- function(list) {
+  i <- 1
+  for (x in list) {
+    write.fasta(x, names(list[i]), "data/sequences.fasta", open = "a")
+    i <- i + 1
+  }
+}
+
+## Run functions
+sequences <- get_seq_data(path)
+gyrB_sequences <- lapply(sequences, function(x) filter_gyrB(x, "gyrb"))
+gyrB_ref <- readDNAStringSet("data/gyrB_ref.fasta")
+write.fasta(gyrB_ref, "gyrB_ref", "data/sequences.fasta", open = "w")
+append_fasta_file(gyrB_sequences)
+
+sequences_gyrB <- readDNAStringSet("data/sequences.fasta")
+trans_seqs <- Biostrings::translate(sequences_gyrB)
+alignment <- AlignSeqs(trans_seqs, iterations = 0, refinements = 0)
+writeXStringSet(alignment, "data/alignment.fasta")

### 2018-10-24 notebooks/article_notebook.Rmd
updated figure with legend

@@
 ```{r}
-test <- acquired_report %>%
+acquired_report %>%
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
@@
   scale_fill_viridis(option = "A",
-                     begin = 0.1)+
-  guides(fill = FALSE)+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 14),
-        axis.text.y = element_text(size = 14),
+                     begin = 0.1,
+                     guide = guide_colorbar(barheight = 15,
+                                            barwidth = 1,
+                                            label.position = "left",
+                                            label.hjust = 0.5,
+                                            label.vjust = 0.3,
+                                            ticks = TRUE))+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 18),
+        axis.text.y = element_text(size = 18),
         axis.title.x = element_blank(),
-        axis.title.y = element_text(size = 16),
+        axis.title.y = element_text(size = 20),
         strip.background = element_rect(colour = "black", fill = "grey70"),
-        strip.text = element_text(size = 14),
+        strip.text = element_text(size = 18),
+        legend.position = "left",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 14),
         plot.margin = unit(c(0,0,0,0), "cm"))+
@@
         axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 14),
+        axis.text.x = element_text(size = 18),
         axis.ticks.y = element_blank(),
-        strip.text = element_text(size = 14),
+        strip.text = element_text(size = 18),
         strip.background = element_rect(colour = "black", fill = "grey70"),

### 2018-10-25 notebooks/article_notebook.Rmd
added MLST plot and test stats

@@
 
+## Sequence types per resistance category
+```{r}
+ST_MIC <- mic_data %>%
+  left_join(mlst_data[c("id","ST")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  mutate(NAL_CIP_R = rowSums(.[,c(15,17)]),
+         antres = rowSums(.[,13:21])-NAL_CIP_R) %>%
+  select(-c(NAL_CIP_R,SMX,TMP,CIP,TET,NAL,CTX,CHL,AMP,GEN)) %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  gather(AB, res, -c(id, antres, ST, species)) %>%
+  group_by(AB, ST, res, species) %>%
+  count() %>%
+  spread(res, n, fill = 0) %>%
+  rename("res" = "1",
+         "nonres" = "0") %>%
+  group_by(AB, ST) %>%
+  mutate(p = res/(res+nonres)*100) %>%
+  ungroup() %>%
+  mutate(AB = gsub("_R", "", AB))
+
+ggplot(ST_MIC, aes(reorder(ST, p), reorder(AB, -p), fill = p))+
+  geom_tile()+
+  scale_fill_viridis(guide = guide_colorbar(barheight = 5,
+                                            barwidth = 0.5,
+                                            label.position = "left",
+                                            label.hjust = 0.5,
+                                            label.vjust = 0.3,
+                                            ticks = TRUE))+
+  labs(y = "Antimicrobials")+
+  theme(axis.text.y = element_text(size = 4),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 8),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        legend.position = "left",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 7))+
+  facet_wrap(~ species, scales = "free_x", ncol = 1)
+```
+
+# Statistics
+## Grouping of MIC values
+```{r}
+stat_MIC <- mic_data[c("id", "CIP")] %>%
+  left_join(total_df, by = "id") %>%
+  left_join(cols_to_add, by = "id") %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  mutate(category = ifelse(as.numeric(CIP) > 0.25, 1, 0)) %>%
+  select(-c(id, CIP))
+
+test <- glm(factor(category) ~ ., data = stat_MIC, family = "binomial")
+
+summary(test)
+plot(test)
+```
+
 # GyrB Protein sequence alignment

### 2018-10-30 notebooks/article_notebook.Rmd
added plot for mechanism per ST and species

@@
 date: '`r format(Sys.time(), "%d.%m.%Y")`'
+always_allow_html: yes
 output:
@@
 
+## Sequence types and resistance mechanisms
+```{r, fig.width=14, fig.height=16}
+# identify which sequence types are represented over three times
+n_ST <- mlst_data[c("id","ST")] %>%
+  group_by(ST) %>%
+  count() %>%
+  filter(n > 3)
+
+# species ST  
+species_st <- mlst_data[c("id","ST")] %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  select(-id) %>%
+  group_by(ST, species) %>%
+  count() %>%
+  filter(ST %in% n_ST$ST)
+
+# prepare micplot data for each ST with n > 3
+micplot_st_df <- mlst_data[c("id", "ST")] %>%
+  left_join(mic_data[c("id", "CIP")], by = "id") %>%
+  group_by(ST, CIP) %>%
+  count() %>%
+  filter(ST %in% n_ST$ST)
+
+# prepare heatmap data frame
+st_heatmap_df <- mlst_data[c("id", "ST")] %>%
+  left_join(mut_report[c("id", "marR")], by = "id") %>%
+  left_join(cols_to_add, by = "id") %>%
+  left_join(acquired_report, by = "id") %>%
+  select(-id) %>%
+  mutate_at(vars(-ST),
+            funs(as.numeric)) %>%
+  group_by(ST) %>%
+  summarise_all(funs(sum, n = n())) %>%
+  ungroup() %>%
+  rename("n" = marR_n) %>%
+  select(-contains("_n")) %>%
+  gather(mechanism, value, -c(ST, n)) %>%
+  mutate(mechanism = sub("_sum", "", mechanism),
+         percent = round(value/n*100),
+         value2 = ifelse(value > 0, 1, 0)) %>%
+  filter(ST %in% n_ST$ST)
+
+# plots
+factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
+factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")
+
+## heatmap
+heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, ST, fill = percent))+
+  geom_tile()+
+  scale_fill_viridis()+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1,
+                                   vjust = 0.3,
+                                   size = 8),
+        axis.title.x = element_blank())
+
+## micplot
+micplot_st <- ggplot(micplot_st_df, aes(ST, CIP))+
+  geom_boxplot(fill = "#67a9cf")+
+  scale_y_continuous(labels = factor_levels,
+                     breaks = factor_breaks,
+                     trans = "log10")+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_text(size = 8),
+        axis.ticks.y = element_blank())+
+  coord_flip()
+
+# n plot
+nplot_st <- ggplot(species_st, aes(ST, n, fill = species))+
+  geom_col(color = "black")+
+  scale_fill_brewer(palette = "Paired")+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_text(size = 8),
+        axis.ticks.y = element_blank(),
+        legend.position = c(0.6, 0.95),
+        legend.title = element_blank())+
+  coord_flip()
+
+cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))
+
+```
+
 # Statistics

### 2018-11-01 notebooks/article_notebook.Rmd
updated gene query

@@
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","oqx","qep")
-mut_genes <- c("gyr","par","marR","soxR","acrR","mcb")
+ac_genes <- c("qnr","oqx","qep","mfp","mcb")
+mut_genes <- c("gyr","par","marR","soxR","acrR","nor","ydh","mdf")
 

### 2018-11-07 notebooks/quast_reference_testing.Rmd
added notebook on quast reference testing

---
title: "Quast Reference Testing"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r, include=FALSE}
# Libraries
pacman::p_load(dplyr, ggplot2, tidyr)
```

```{r, include=FALSE}
# Path to folders
path <- "../data/quast_ref_testing"

# Functions
## searches recursively for all filenames with fastqc.zip
file_names_recursive <- function(filepath) {
  files <- list.files(path = filepath, pattern = "transposed_report.tsv", recursive = TRUE)
  return(files)
}

## Creates clean data frame
create_df <- function(df) {
  df <- df %>%
    rename(ctg_0_bp = X..contigs.....0.bp.,
           ctg_1000_bp = X..contigs.....1000.bp.,
           ctg_5000_bp = X..contigs.....5000.bp.,
           ctg_10000_bp = X..contigs.....10000.bp.,
           ctg_25000_bp = X..contigs.....25000.bp.,
           ctg_50000_bp = X..contigs.....50000.bp.,
           tot_len_0_bp = Total.length.....0.bp.,
           tot_len_1000_bp = Total.length.....1000.bp.,
           tot_len_5000_bp = Total.length.....5000.bp.,
           tot_len_10000_bp = Total.length.....10000.bp.,
           tot_len_25000_bp = Total.length.....25000.bp.,
           tot_len_50000_bp = Total.length.....50000.bp.,
           no_of_contigs = X..contigs,
           no_of_misassemblies = X..misassemblies,
           no_of_misassembled_ctg = X..misassembled.contigs,
           no_of_local_misassemblies = X..local.misassemblies,
           unaligned_contigs = X..unaligned.contigs,
           N_per_100kbp = X..N.s.per.100.kbp,
           mismatches_per_100kbp = X..mismatches.per.100.kbp,
           indels_per_100kbp = X..indels.per.100.kbp,
           no_of_genes = X..genes) %>%
    mutate(Assembly = gsub("(.*?)_.+", "\\1", Assembly),
           reference = sub(".+/(.*?)/transposed_report.tsv", "\\1", ref),
           group = sub("(.*?)/.+/transposed_report.tsv", "\\1", ref),
           short_group = ifelse(group == "hiseq_flex_150bp", "A",
                                ifelse(group == "hiseqr_xt_250bp", "B", "C")))
  return(df)
}

func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

```

```{r}
# Data import and wrangling
refs <- read.table("../data/ecoli_ref.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

filenames <- file_names_recursive(path)

data_list <- lapply(filenames,
                      FUN = function(file) {
                        read.delim(paste0(path, "/", file),
                                   sep = "\t",
                                   header = TRUE,
                                   stringsAsFactors = FALSE)
                      })

data_list <- lapply(data_list, function(df) mutate_all(df, as.character))

names(data_list) <- filenames

df <- bind_rows(data_list, .id = "ref")

df2 <- create_df(df) %>%
  left_join(refs, by = "reference")
```

# Information
## Reference sequences
Below is a table over the references used in the various Quast analyses.
```{r}
refs
```

## Sequencing methodology
```{r}
df2 %>%
  select(group, short_group) %>%
  group_by(short_group) %>%
  summarise_all(funs(func_paste))
```

# Plots
## N50
```{r}
# N50 per reference and sequencing group

ggplot(df2, aes(short_group, as.numeric(N50), fill = short_group))+
  geom_boxplot()+
  labs(y = "N50",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## Number of contigs
```{r}
# Number of contigs
ggplot(df2, aes(short_group, as.numeric(no_of_contigs), fill = short_group))+
  geom_boxplot()+
  labs(y = "Number of contigs",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## Number of misassemblies
```{r}
# Number of misassemblies
ggplot(df2, aes(short_group, as.numeric(no_of_misassemblies), fill = short_group))+
  geom_boxplot()+
  labs(y = "Number of misassemblies",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## Genome fraction
```{r}
ggplot(df2, aes(short_group, as.numeric(Genome.fraction....), fill = short_group))+
  geom_boxplot()+
  labs(y = "Genome fraction (%)",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## N per 100kbp
```{r}
ggplot(df2, aes(short_group, as.numeric(N_per_100kbp), fill = short_group))+
  geom_boxplot()+
  labs(y = "N per 100 kbp",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## Largest contigs
```{r}
ggplot(df2, aes(short_group, as.numeric(Largest.contig), fill = short_group))+
  geom_boxplot()+
  labs(y = "Largets Contig Length",
       x = NULL)+
  guides(fill = FALSE)+
  theme_classic()+
  facet_wrap(~name)
```

## Genome length vs. reference length
```{r}
ggplot(df2, aes(as.numeric(size), as.numeric(Total.length), fill = short_group))+
  geom_jitter(pch = 21, size = 3)+
  labs(y = "Total Length",
       x = "Reference length",
       fill = NULL)+
  theme_classic()
```

## Genome fraction vs reference length
```{r}
ggplot(df2, aes(as.numeric(Genome.fraction....), as.numeric(Total.length), fill = short_group))+
  geom_jitter(pch = 21, size = 3)+
  labs(y = "Total Length",
       x = "Genome Fraction",
       fill = NULL)+
  theme_classic()
```

### 2018-11-07 notebooks/article_notebook.Rmd
minor changes

--- 2018-10-30 notebooks/article_notebook.Rmd
@@
 ```{r Import data, include = FALSE}
-
 # Collapses data frame to unique lines while ignoring NA's
 func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
+
+"%not_in%" <- Negate("%in%")
 
@@
                            header = T,
-                           stringsAsFactors = F)
+                           stringsAsFactors = F) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S"))
 
@@
   left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref)
+  select(id, everything(), -ref) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S"))
 
@@
   left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref)
+  select(id, everything(), -ref) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S"))
 ```
@@
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","oqx","qep")
-mut_genes <- c("gyr","par","marR","soxR","acrR","mcb")
+ac_genes <- c("qnr","oqx","qep","mfp","mcb")
+mut_genes <- c("gyr","par","marR","soxR","acrR","nor","ydh","mdf","sdi")
 
@@
 acquired_report %>%
-  select(-id) %>%
+  select(-c(id ,qepA4)) %>%
   t() %>%
@@
             funs(as.numeric(as.character(.)))) %>%
-  mutate(Present = rowSums(.[,-286])) %>%
+  mutate(Present = rowSums(.[,-282])) %>%
   select(c(Gene, Present)) %>%
@@
 mut_report %>%
-  select(-c(id, marR, nt_change)) %>%
+  select(-c(id, marR, nt_change, mdfA)) %>%
   group_by_all() %>%

### 2018-11-10 notebooks/quast_reference_testing.Rmd
minor changes

@@
 
-## Genome length vs. reference length
-```{r}
-ggplot(df2, aes(as.numeric(size), as.numeric(Total.length), fill = short_group))+
-  geom_jitter(pch = 21, size = 3)+
-  labs(y = "Total Length",
-       x = "Reference length",
-       fill = NULL)+
-  theme_classic()
-```
-
 ## Genome fraction vs reference length
@@
 ggplot(df2, aes(as.numeric(Genome.fraction....), as.numeric(Total.length), fill = short_group))+
-  geom_jitter(pch = 21, size = 3)+
+  geom_point(pch = 21, size = 3)+
   labs(y = "Total Length",

### 2018-11-10 notebooks/article_notebook.Rmd
added figure for MIC combinations

@@
   select(-c(lwr, upr, Total))
+```
+
+## Amount of isolates resistant to specific antimicrobials
+```{r}
+mic_data %>%
+  mutate(SMX_R = ifelse(SMX > 64, 1, 0),
+         TMP_R = ifelse(TMP > 2, 1, 0),
+         CIP_R = ifelse(CIP > 0.06, 1, 0),
+         TET_R = ifelse(TET > 8, 1, 0),
+         NAL_R = ifelse(NAL > 16, 1, 0),
+         CTX_R = ifelse(CTX > 0.25, 1, 0),
+         CHL_R = ifelse(CHL > 16, 1, 0),
+         AMP_R = ifelse(AMP > 8, 1, 0),
+         GEN_R = ifelse(GEN > 2, 1, 0)) %>%
+  select(contains("_R")) %>%
+  gather(AB, value) %>%
+  group_by(AB) %>%
+  summarise_all(funs(sum)) %>%
+  arrange(desc(value)) %>%
+  mutate(AB = sub("_R", "", AB))
 ```
@@
 
+# Phenotypic resistance combinations
+```{r, fig.width=12, fig.height=10}
+mic_dotplot_data <- mic_data %>%
+  mutate(SMX_R = ifelse(SMX > 64, 1, 0),
+         TMP_R = ifelse(TMP > 2, 1, 0),
+         CIP_R = ifelse(CIP > 0.06, 1, 0),
+         TET_R = ifelse(TET > 8, 1, 0),
+         NAL_R = ifelse(NAL > 16, 1, 0),
+         CTX_R = ifelse(CTX > 0.25, 1, 0),
+         CHL_R = ifelse(CHL > 16, 1, 0),
+         AMP_R = ifelse(AMP > 8, 1, 0),
+         GEN_R = ifelse(GEN > 2, 1, 0)) %>%
+  select(contains("_R"), -c(CIP_R, NAL_R)) %>%
+  mutate(None = ifelse(rowSums(.) == 0, 1, 0)) %>%
+  group_by_all() %>%
+  count() %>%
+  ungroup() %>%
+  mutate(id = 1:n()) %>%
+  gather(AM, value, -id, -n) %>%
+  mutate(value = ifelse(AM == "None" & value == 0, NA, value),
+         value = ifelse(AM != "None" & value == 0, NA, value),
+         AM = sub("_R", "", AM)) %>%
+  filter(is.na(value) == FALSE) %>%
+  mutate(AM = factor(AM, ordered = TRUE, levels = c("SMX","TMP","AMP","TET","CHL","CTX","GEN","None")))
+
+mic_nplot_data <- mic_data %>%
+  mutate(SMX_R = ifelse(SMX > 64, 1, 0),
+         TMP_R = ifelse(TMP > 2, 1, 0),
+         CIP_R = ifelse(CIP > 0.06, 1, 0),
+         TET_R = ifelse(TET > 8, 1, 0),
+         NAL_R = ifelse(NAL > 16, 1, 0),
+         CTX_R = ifelse(CTX > 0.25, 1, 0),
+         CHL_R = ifelse(CHL > 16, 1, 0),
+         AMP_R = ifelse(AMP > 8, 1, 0),
+         GEN_R = ifelse(GEN > 2, 1, 0)) %>%
+  select(id, contains("_R"), -c(CIP_R, NAL_R)) %>%
+  mutate(None = ifelse(rowSums(.[,2:8]) == 0, 1, 0)) %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  select(-id) %>%
+  group_by_all() %>%
+  count() %>%
+  group_by_at(vars(SMX_R, TMP_R, TET_R, CTX_R, CHL_R, AMP_R, GEN_R)) %>%
+  { mutate(ungroup(.), group = group_indices(.)) } %>%
+  group_by(group) %>%
+  mutate(group_n = sum(n))
+
+mic_dotplot <- ggplot(mic_dotplot_data, aes(reorder(factor(id), -n), AM, fill = AM)) +
+  geom_point(size = 6, pch = 21)+
+  guides(fill = FALSE)+
+  theme(axis.title = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major.x = element_line())
+
+mic_nplot <- ggplot(mic_nplot_data, aes(reorder(factor(group), -group_n), n, fill = species))+
+  geom_col(color = "black")+
+  scale_fill_brewer(palette = "Paired")+
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        legend.position = c(0.85,0.85),
+        legend.title = element_blank())
+
+cowplot::plot_grid(mic_nplot, mic_dotplot, ncol = 1, align = "v")
+
+```
+
 # Statistics
@@
 
-test <- glm(factor(category) ~ ., data = stat_MIC, family = "binomial")
-
-summary(test)
-plot(test)
+# test <- glm(category ~ ., data = stat_MIC, family = "binomial")
+# 
+# summary(test)
+# plot(test)
 ```

### 2018-11-13 notebooks/article_notebook.Rmd
updated some figures and added informational text

@@
 ## List of genes of interest
+This table lists all the genes of interest in this study, some information about them, and their accession number(s).
 ```{r}
@@
                           grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
-                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9")) %>%
+                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
+                          grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
   filter(test == FALSE,
@@
                           grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
-                          grepl("soxs", cluster, ignore.case = T) == TRUE ~ "soxS",
+                          grepl("soxR", cluster, ignore.case = T) == TRUE ~ "soxR",
                           grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                           grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
-                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR")) %>%
+                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
+                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA")) %>%
   filter(test == FALSE,
@@
                                  Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
-                                 Gene == "soxS" ~ "Transcriptional activator of the superoxide response regulon.",
-                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump."),
+                                 Gene == "soxR" ~ "Transcriptional activator of the superoxide response regulon.",
+                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
+                                 Gene == "mdfA" ~ "Chromosomal efflux pump."),
          Type = "Intrinsic gene")
@@
 ## Percent of isolates with specific mutation combinations in QRDR
+This table presents the different substitution combinations identified in the quinolone resistance determining regions (QRDR) of *gyrA*, *gyrB*, *parC* and *parE*.
 ```{r}
@@
 ## Percent isolates with specific nucleotide changes in QRDR
+This table presents the various nucleotide changes for each amino acid substitution identified in the quinolone resistance determining region (QRDR) of *gyrA*, *gyrB*, *parC* and *parE*.
 ```{r}
@@
 ## Total combinations
+This table presents the total combination of resistance mechanisms, their respective MIC-range for CIP and NAL, how many isolates per group, as well as percentage and confidence intervals.
 ```{r}
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("marR","id")], by = "id") %>%
+  left_join(mut_report[c("marR","mdfA","id")], by = "id") %>%
   left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
@@
 # Figures
-## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
+## Percent isolates with PMQR and/or QRDR mutations in *gyrA*, *gyrB*, *parC* or *parE*
 ```{r}
@@
 ## Total resistance mechanisms per animal species
+This figure presents the percent combination of resistance mechanisms per CIP MIC category.
+The X-axis on the leftmost plot lists all the specific mutations in *gyrA*, *parC* and *parE*, whether *marR* is mutated or not, as well as a summation of all the *qnr* genes and *qepA*.
+The rightmost plot presents how many isolates are present in the given CIP-MIC group (Y-axis in both plots).
 ```{r, warning=F, message=F, fig.height=16, fig.width=20}
@@
          perc = value/n*100)
-# 
-# ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
-#   geom_tile(color = "grey70")+
-#   scale_fill_viridis(option = "A",
-#                      begin = 0.1,
-#                      guide = guide_colorbar(ticks = F,
-#                                             barheight = 0.5,
-#                                             barwidth = 10))+
-#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 7),
-#         axis.text.y = element_text(size = 8),
-#         axis.title.x = element_blank(),
-#         axis.title.y = element_text(size = 10),
-#         legend.title = element_blank(),
-#         legend.position="bottom",
-#         legend.justification="center",
-#         legend.direction="horizontal",
-#         legend.text = element_text(size = 8),
-#         strip.background = element_rect(colour = "black", fill = "grey70"),
-#         strip.text = element_text(size = 10))+
-#   labs(y = "CIP MIC")+
-#   facet_wrap(~species)+
-#   coord_fixed()
 
@@
 ## Correlation plot
+This plot presents the results of a pearson correlation test for each resistance mechanism.
+A positive correlation is represented in yellow and red spectra, a negative correlation represented as blue and light blue, and no correlation in green.
 ```{r}
@@
 ## Sequence types per resistance category
+This plot presents the percent of isolates in each sequence type group that are resistant to multiple antimicrobials, sorted from least resistant at the top and most resistant at the bottom.
+The percentage is based on each ST group, and how many isolates in that group.
 ```{r}
@@
 
-ggplot(ST_MIC, aes(reorder(ST, p), reorder(AB, -p), fill = p))+
+ggplot(ST_MIC, aes(reorder(AB, -p), reorder(ST, -p), fill = p))+
   geom_tile()+
@@
   labs(y = "Antimicrobials")+
-  theme(axis.text.y = element_text(size = 4),
+  theme(axis.text.y = element_text(size = 6),
         axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 8),
         axis.title.x = element_blank(),
-        axis.title.y = element_text(size = 10),
+        axis.title.y = element_blank(),
         legend.position = "left",
@@
         legend.text = element_text(size = 7))+
-  facet_wrap(~ species, scales = "free_x", ncol = 1)
+  facet_wrap(~ species, scales = "free_y", ncol = 2)
 ```
@@
   geom_tile()+
-  scale_fill_viridis()+
+  scale_fill_viridis(option = "A")+
   guides(fill = FALSE)+
@@
 # Phenotypic resistance combinations
+This plot presents the various phenotypic resistance combinations present in the isolates, and how many isolates for each animal species in each group.
 ```{r, fig.width=12, fig.height=10}
@@
   geom_point(size = 6, pch = 21)+
+  scale_fill_viridis(option = "A",
+                     discrete = TRUE)+
   guides(fill = FALSE)+
@@
 
-# Statistics
-## Grouping of MIC values
-```{r}
-stat_MIC <- mic_data[c("id", "CIP")] %>%
-  left_join(total_df, by = "id") %>%
-  left_join(cols_to_add, by = "id") %>%
-  select(-c(gyrA, gyrB, parC, parE)) %>%
-  mutate(category = ifelse(as.numeric(CIP) > 0.25, 1, 0)) %>%
-  select(-c(id, CIP))
-
-# test <- glm(category ~ ., data = stat_MIC, family = "binomial")
-# 
-# summary(test)
-# plot(test)
-```
-
 # GyrB Protein sequence alignment

### 2018-11-15 notebooks/mash_analysis.Rmd
added notebook on mash analysis

---
title: "Quality control - Mash screen"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r}
pacman::p_load(ggplot2, dplyr, tidyr, stringr, purrr, viridis)
```

```{r}
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

scan_max <- function(x) max(scan(text = x, what = "", sep = ",", quiet = TRUE, strip.white = TRUE))

# Identifies filenames in input folder
file_names_mash <- function(filepath) {
  files <- list.files(path = filepath, pattern = "_mash.out")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_mash_data <- function(filepath) {
  files <- file_names_mash(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = FALSE, # no header in files
                          quote = "", # disable quoting, avoid EOS error
                          sep = "\t"
                        )
                      })

  names(data_list) <- files # set name of each df in list 
  data <- bind_rows(data_list, .id = "ref") %>% # bind to data frame
    rename("identity" = V1,
           "shared_hashes" = V2,
           "median_multiplicity" = V3,
           "p_value" = V4,
           "query_id" = V5,
           "query_comment" = V6) %>%
    mutate(p_value = round(p_value, 5))
  return(data)
}

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

filepath <- "../data/mash/"

mash_raw <- get_mash_data(filepath) %>%
  mutate(ref = sub("_L00[0-9]_R1_001.fastq.gz_mash.out", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(-ref) %>%
  select(id, everything())

```

```{r}
mash_results <- mash_raw %>%
  separate(shared_hashes, c("min","max"), sep = "/", remove = FALSE) %>%
  mutate(test = as.numeric(min)/as.numeric(max)) %>%
  select(-c(min, max)) %>%
  filter(test >= 0.1) %>%
  mutate(species = query_comment %>%
           str_replace("\\[.*?\\] ", "") %>%
           sub("^.+_.+\\.[0-9] (.*?), .+", "\\1", .) %>%
           word(1,2),
         phage = grepl("phage", ignore.case = T, species)) %>%
  filter(phage == FALSE) %>%
  {. ->> mash_plot_df} %>%
  select(-phage) %>%
  spread(species, identity)
```

# Identity plot for all isolates

```{r, fig.width=10, fig.height=8}
ggplot(mash_plot_df, aes(id, identity, fill = species)) +
  geom_point(pch = 21, size = 2)+
  geom_hline(yintercept = 0.95, alpha = 0.5)+
  scale_fill_brewer(type = "div", palette = 2)+
  labs(y = "Identity",
       x = "Samples")+
  theme_classic()+
  theme(axis.text.x = element_blank(),
        legend.title = element_blank(),
        axis.ticks.x = element_blank())
```

# Contaminated Isolates
```{r}
test <- mash_plot_df %>%
  filter(species != "Escherichia coli",
         identity >= 0.95) %>%
  select(id)

ids <- test$id

mash_plot_df %>%
  select(-c(phage,shared_hashes,query_id,query_comment,median_multiplicity,test,p_value)) %>%
  filter(id %in% ids) %>%
  mutate(id2 = 1:n()) %>%
  spread(species, identity, fill = NA) %>%
  select(-id2) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  mutate(`Escherichia coli` = sapply(`Escherichia coli`, scan_max)) %>%
  mutate_all(funs(gsub("^$", NA, .)))

```

### 2018-11-15 notebooks/article_notebook.Rmd
minor changes

@@
 
-# Dependencies
-This notebook is using the following packages:
-
-- dplyr
-
-- ggplot2
-
-- purrr
+# Background
+
+## Aims
+
+- Identify potential differences in resistance mechanisms between wild- and production animals.
+
+- Identify potential clones which are species specific.
+
+- Characterize resistance mechanisms in QREC from Norwegian wild- and production animals.
+
+- Identify "main culprit" mechanisms that infer high MIC values for CIP and/or NAL.
+
+- Characterize the population structure of QREC from selected animal species.
+
+## Isolates and Methods
+
+Groups of quinolone resistant *E.
+coli* (QREC) isolates based on animal species, minimum inhibitory concentration (MIC) values to CIP and NAL, and the amount of additional resistances were created, and isolates were randomly selected from these groups (*n* = 285).
+This was done to ensure phenotypic diversity in the isolates.
+Of the 285 isolates, 73 were already sequenced in other projects, and 212 were subjected to whole genome sequencing.
+MIC values were extracted from the internal sample registration system at the Norwegian Veterinary Institute.
+
+The sequences were subjected to quality control with FastQC.
+Mash screen was used to identify contaminant reads in the files.
+Four isolates were excluded from the dataset due to contamination with *Citrobacter* reads.
+This is probably due to contaminated culture during DNA extraction.
+
+To identify resistance genes of interest, the sequences were run through the program "Antimicrobial Resistance Identification By Assembly" (ARIBA), which is included in the Bifrost Pipeline.
+Acquired resistance genes were detected by the use of the ResFinder database, while mutations in intrinsic genes related to quinolone resistance was identified with the MegaRes database.
+Sequence types was also identified by ARIBA, using the Wirth et al. scheme (*E. coli* #1).
+
+## Further work
+
+- Assemble genomes
+
+- Align assemblies
+
+- Land on methodology for phylogenetic analysis (likely core genome SNP - ML)
+
+Results from various analyses are listed below.
+
 ```{r load libraries, message=FALSE, warning=FALSE}
@@
 ac_genes <- c("qnr","oqx","qep","mfp","mcb")
-mut_genes <- c("gyr","par","marR","soxR","acrR","nor","ydh","mdf","sdi")
+mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","mdf","sdi","rpoB")
 
@@
                           grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
+                          grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
                           grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                           grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
-                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA")) %>%
+                          grepl("acrb", ref_name, ignore.case = T) == TRUE ~ "acrB",
+                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA",
+                          grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB")) %>%
   filter(test == FALSE,
@@
   mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
+                                 Gene == "acrB" ~ "Gene that codes for the AcrB efflux pump, part of the AcrAB-TolC efflux system",
                                  Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
@@
                                  Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
+                                 Gene == "marA" ~ "Transcriptional activator og genes involved in the MAR phenotype.",
                                  Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
@@
                                  Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
-                                 Gene == "mdfA" ~ "Chromosomal efflux pump."),
+                                 Gene == "mdfA" ~ "Chromosomal efflux pump.",
+                                 Gene == "rpoB" ~ "Beta-subunit of RNA polymerase, mutations lead to increased expression of the ydhE-pump"),
          Type = "Intrinsic gene")
@@
 mut_report %>%
-  select(-c(id, marR, nt_change, mdfA)) %>%
+  select(gyrA, gyrB, parC, parE) %>%
   group_by_all() %>%
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("marR","mdfA","id")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id") %>%
   left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
@@
     select(gene, species1, species2, p.value) %>%
-    filter(p.value < 0.05)
+    mutate(p.value = round(p.value, 5)) %>%
+    filter(p.value <= 0.05)
   return(df)
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("marR","id")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id") %>%
   left_join(isolate_data[c("id", "species")])
@@
 genes <- c(names(mut_report), names(acquired_report))
-genes <- genes[!genes == "id"]
-
-p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
+genes <- genes[!genes %in% c("id","nt_change")]
+
+cip_n_plot <- ggplot(total_report, aes(reorder(present, as.numeric(CIP)), percent, fill = factor(CIP)))+
   geom_col(color = "black")+
@@
   scale_fill_viridis(discrete = TRUE)+
-  guides(fill = FALSE)+
+  guides(fill=guide_legend(ncol = 10))+
   theme(axis.text.x = element_blank(),
@@
         legend.position = "top",
-        legend.spacing.x = unit(0.27, "cm"))
+        legend.spacing.x = unit(0.2, "cm"),
+        legend.justification = "center")
 
@@
   left_join(acquired_report, by = "id") %>%
+  select(-nt_change) %>%
   gather(gene, value, -id) %>%
@@
 
-p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
+mechanism_dot_plot <- ggplot(dot_report, aes(reorder(present, CIP), gene, fill = gene))+
   geom_point(size = 2, pch = 21)+
@@
 
-plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
+plot_grid(cip_n_plot, mechanism_dot_plot, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
 
@@
             parE_D475E_n,parE_H516Y_n,parE_L416F_n,
-            parE_L488M_n,parE_S458A_n)) %>%
+            parE_L488M_n,parE_S458A_n,marA_n,rpoB_n,
+            mdfA_n,acrB_n)) %>%
   gather(gene, value, -c(species, CIP, marR_n)) %>%
@@
                           gene2 == "marR" ~ 5,
-                          gene2 == "qnr" ~ 6,
-                          gene2 == "qepA4" ~ 7),
+                          gene2 == "marA" ~ 6,
+                          gene2 == "mdfA" ~ 7,
+                          gene2 == "acrB" ~ 8,
+                          gene2 == "rpoB" ~ 9,
+                          gene2 == "qnr" ~ 10,
+                          gene2 == "qepA4" ~ 11),
          perc = value/n*100)
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("marR","id")], by = "id")
+  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id")
 
@@
                           gene1 == "marR" ~ 5,
-                          gene1 == "qnr" ~ 6,
-                          gene1 == "qepA4" ~ 7),
+                          gene1 == "marA" ~ 6,
+                          gene1 == "acrB" ~ 7,
+                          gene1 == "rpoB" ~ 8,
+                          gene1 == "mdfA" ~ 9,
+                          gene1 == "qnr" ~ 10,
+                          gene1 == "qepA4" ~ 11),
            type2 = case_when(gene2 == "gyrA" ~ 1,
@@
                           gene2 == "marR" ~ 5,
-                          gene2 == "qnr" ~ 6,
-                          gene2 == "qepA4" ~ 7))
+                          gene2 == "marA" ~ 6,
+                          gene2 == "acrB" ~ 7,
+                          gene2 == "rpoB" ~ 8,
+                          gene2 == "mdfA" ~ 9,
+                          gene2 == "qnr" ~ 10,
+                          gene2 == "qepA4" ~ 11))
 
@@
 
-# Phenotypic resistance combinations
+## Phenotypic resistance combinations
 This plot presents the various phenotypic resistance combinations present in the isolates, and how many isolates for each animal species in each group.

### 2018-11-20 notebooks/article_notebook.Rmd
adjusted rounding of values to be the same in all tables

@@
   mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,2),
-         lwr = round(get_binCI(Present, Total)[1], 2),
-         upr = round(get_binCI(Present, Total)[2], 2),
+         Percent = round(Present/Total * 100,1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
          "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
@@
 
-## Percent isolates with specific qnr genes
+## Percent isolates with specific PMQR genes
 ```{r}
 acquired_report %>%
-  select(-c(id ,qepA4)) %>%
+  select(-id) %>%
   t() %>%
@@
   rowwise() %>%
-  mutate(Total = 285,
-         Percent = round(Present/Total * 100,2),
-         lwr = round(get_binCI(Present, Total)[1], 2),
-         upr = round(get_binCI(Present, Total)[2], 2),
+  mutate(Total = 281,
+         Percent = round(Present/Total * 100,1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
          "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
-  select(-c(lwr, upr, Total))
+  select(-c(lwr, upr))
 ```
@@
   rowwise() %>%
-  mutate(Total = 285,
-         Percent = round(n/Total * 100,2),
-         lwr = round(get_binCI(n, Total)[1], 2),
-         upr = round(get_binCI(n, Total)[2], 2),
+  mutate(Total = 281,
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1], 1),
+         upr = round(get_binCI(n, Total)[2], 1),
          "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
@@
   rowwise() %>%
-  mutate(Total = 285,
+  mutate(Total = 281,
          Percent = round(n/Total * 100,1),
@@
   mutate(present = ifelse(present == "", "None", present),
-         total = 285,
+         total = 281,
          percent = round(n/total * 100, 1),

### 2018-12-25 notebooks/article_notebook.Rmd
updated phylogenetic tree

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble)
 ```
@@
                        "2016-02-486-2-S",
-                       "2016-02-732-2-S"))
+                       "2016-02-732-2-S",
+                       "2014-01-1741-1-S"))
 
@@
                        "2016-02-486-2-S",
-                       "2016-02-732-2-S"))
-
-mlst_hf150bp <- read.table("../data/mlst/hiseq_flex_150bp_mlst_summarized_results.tsv",
-                           sep = "\t",
-                           header = T,
-                           stringsAsFactors = F)
-
-mlst_hx125bp <- read.table("../data/mlst/hiseq_xt_125bp_mlst_summarized_results.tsv",
-                           sep = "\t",
-                           header = T,
-                           stringsAsFactors = F)
-
-mlst_hrx250bp <- read.table("../data/mlst/hiseqr_xt_250bp_mlst_summarized_results.tsv",
-                           sep = "\t",
-                           header = T,
-                           stringsAsFactors = F)
-
-mlst_data <- rbind(mlst_hf150bp, mlst_hx125bp, mlst_hrx250bp) %>%
-  separate(ST.adk.fumC.gyrB.icd.mdh.purA.recA,
-           into = c("x","ST","adk","fumC","gyrB","icd","mdh","purA","recA"),
-           sep = " ") %>%
-  select(-x) %>%
-  rename("ref" = header) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S"))
+                       "2016-02-732-2-S",
+                       "2014-01-1741-1-S"))
+
+mlst_data <- read.table("../data/mlst_results.txt", sep = "\t", header = TRUE, stringsAsFactors = F) %>%
+  filter(id %not_in% "2014-01-1741-1-S")
 ```
@@
 ac_genes <- c("qnr","oqx","qep","mfp","mcb")
-mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","mdf","sdi","rpoB")
+mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")
 
@@
             funs(as.numeric(as.character(.)))) %>%
-  mutate(Present = rowSums(.[,-282])) %>%
+  mutate(Present = rowSums(.[,-281])) %>%
   select(c(Gene, Present)) %>%
   rowwise() %>%
-  mutate(Total = 281,
+  mutate(Total = 280,
          Percent = round(Present/Total * 100,1),
@@
   rowwise() %>%
-  mutate(Total = 281,
+  mutate(Total = 280,
          Percent = round(n/Total * 100,1),
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","acrB","rpoB")], by = "id") %>%
   left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
@@
   rowwise() %>%
-  mutate(Total = 285,
+  mutate(Total = 280,
          Percent = round(n/Total * 100,1),
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","acrB","rpoB")], by = "id") %>%
   left_join(isolate_data[c("id", "species")])
@@
 ## Percent isolates with PMQR and/or QRDR mutations in *gyrA*, *gyrB*, *parC* or *parE*
-```{r}
+```{r,message = FALSE, warning = FALSE}
 total_report <- mut_report %>%
@@
   mutate(present = ifelse(present == "", "None", present),
-         total = 281,
+         total = 280,
          percent = round(n/total * 100, 1),
@@
 gene_names <- gene_names[-c(1,2)]
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
 
@@
     ) +
-    scale_fill_brewer(palette = "Paired") +
+    scale_fill_manual(values = palette) +
     labs(y = "Percent (%) of isolates") +
@@
     ) +
-    scale_fill_brewer(palette = "Paired") +
+    scale_fill_manual(values = palette) +
     labs(y = "Percent (%) of isolates",
@@
             parE_L488M_n,parE_S458A_n,marA_n,rpoB_n,
-            mdfA_n,acrB_n)) %>%
+            acrB_n)) %>%
   gather(gene, value, -c(species, CIP, marR_n)) %>%
@@
 A positive correlation is represented in yellow and red spectra, a negative correlation represented as blue and light blue, and no correlation in green.
-```{r}
+```{r,message = FALSE, warning = FALSE}
 create_corr_matrix <- function(df) {
@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("id","marR","marA","mdfA","acrB","rpoB")], by = "id")
+  left_join(mut_report[c("id","marR","marA","acrB","rpoB")], by = "id")
 
@@
 
-## Sequence types and resistance mechanisms
-```{r, fig.width=14, fig.height=16}
-# identify which sequence types are represented over three times
-n_ST <- mlst_data[c("id","ST")] %>%
-  group_by(ST) %>%
-  count() %>%
-  filter(n > 3)
-
-# species ST  
-species_st <- mlst_data[c("id","ST")] %>%
-  left_join(isolate_data[c("id", "species")], by = "id") %>%
-  select(-id) %>%
-  group_by(ST, species) %>%
-  count() %>%
-  filter(ST %in% n_ST$ST)
-
-# prepare micplot data for each ST with n > 3
-micplot_st_df <- mlst_data[c("id", "ST")] %>%
-  left_join(mic_data[c("id", "CIP")], by = "id") %>%
-  group_by(ST, CIP) %>%
-  count() %>%
-  filter(ST %in% n_ST$ST)
-
-# prepare heatmap data frame
-st_heatmap_df <- mlst_data[c("id", "ST")] %>%
-  left_join(mut_report[c("id", "marR")], by = "id") %>%
-  left_join(cols_to_add, by = "id") %>%
-  left_join(acquired_report, by = "id") %>%
-  select(-id) %>%
-  mutate_at(vars(-ST),
-            funs(as.numeric)) %>%
-  group_by(ST) %>%
-  summarise_all(funs(sum, n = n())) %>%
-  ungroup() %>%
-  rename("n" = marR_n) %>%
-  select(-contains("_n")) %>%
-  gather(mechanism, value, -c(ST, n)) %>%
-  mutate(mechanism = sub("_sum", "", mechanism),
-         percent = round(value/n*100),
-         value2 = ifelse(value > 0, 1, 0)) %>%
-  filter(ST %in% n_ST$ST)
-
-# plots
-factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
-factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")
-
-## heatmap
-heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, ST, fill = percent))+
-  geom_tile()+
-  scale_fill_viridis(option = "A")+
-  guides(fill = FALSE)+
-  theme(axis.text.x = element_text(angle = 90,
-                                   hjust = 1,
-                                   vjust = 0.3,
-                                   size = 8),
-        axis.title.x = element_blank())
-
-## micplot
-micplot_st <- ggplot(micplot_st_df, aes(ST, CIP))+
-  geom_boxplot(fill = "#67a9cf")+
-  scale_y_continuous(labels = factor_levels,
-                     breaks = factor_breaks,
-                     trans = "log10")+
-  theme(axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 8),
-        axis.ticks.y = element_blank())+
-  coord_flip()
-
-# n plot
-nplot_st <- ggplot(species_st, aes(ST, n, fill = species))+
-  geom_col(color = "black")+
-  scale_fill_brewer(palette = "Paired")+
-  theme(axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.text.x = element_text(size = 8),
-        axis.ticks.y = element_blank(),
-        legend.position = c(0.6, 0.95),
-        legend.title = element_blank())+
-  coord_flip()
-
-cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))
+```{r, fig.width=14, fig.height=16,message = FALSE, warning = FALSE}
+# # identify which sequence types are represented over three times
+# n_ST <- mlst_data[c("id","ST")] %>%
+#   group_by(ST) %>%
+#   count() %>%
+#   filter(n > 3)
+# 
+# # species ST  
+# species_st <- mlst_data[c("id","ST")] %>%
+#   left_join(isolate_data[c("id", "species")], by = "id") %>%
+#   select(-id) %>%
+#   group_by(ST, species) %>%
+#   count() %>%
+#   filter(ST %in% n_ST$ST)
+# 
+# # prepare micplot data for each ST with n > 3
+# micplot_st_df <- mlst_data[c("id", "ST")] %>%
+#   left_join(mic_data[c("id", "CIP")], by = "id") %>%
+#   group_by(ST, CIP) %>%
+#   count() %>%
+#   filter(ST %in% n_ST$ST)
+# 
+# # prepare heatmap data frame
+# st_heatmap_df <- mlst_data[c("id", "ST")] %>%
+#   left_join(mut_report[c("id", "marR")], by = "id") %>%
+#   left_join(cols_to_add, by = "id") %>%
+#   left_join(acquired_report, by = "id") %>%
+#   select(-id) %>%
+#   mutate_at(vars(-ST),
+#             funs(as.numeric)) %>%
+#   group_by(ST) %>%
+#   summarise_all(funs(sum, n = n())) %>%
+#   ungroup() %>%
+#   rename("n" = marR_n) %>%
+#   select(-contains("_n")) %>%
+#   gather(mechanism, value, -c(ST, n)) %>%
+#   mutate(mechanism = sub("_sum", "", mechanism),
+#          percent = round(value/n*100),
+#          value2 = ifelse(value > 0, 1, 0)) %>%
+#   filter(ST %in% n_ST$ST)
+# 
+# # plots
+# factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
+# factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")
+# 
+# ## heatmap
+# heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, factor(ST), fill = percent))+
+#   geom_tile()+
+#   scale_fill_viridis(option = "A")+
+#   guides(fill = FALSE)+
+#   theme(axis.text.x = element_text(angle = 90,
+#                                    hjust = 1,
+#                                    vjust = 0.3,
+#                                    size = 8),
+#         axis.title.x = element_blank())
+# 
+# ## micplot
+# micplot_st <- ggplot(micplot_st_df, aes(factor(ST), CIP))+
+#   geom_boxplot(fill = "#67a9cf")+
+#   scale_y_continuous(labels = factor_levels,
+#                      breaks = factor_breaks,
+#                      trans = "log10")+
+#   theme(axis.title = element_blank(),
+#         axis.text.y = element_blank(),
+#         axis.text.x = element_text(size = 8),
+#         axis.ticks.y = element_blank())+
+#   coord_flip()
+# 
+# # n plot
+# nplot_st <- ggplot(species_st, aes(factor(ST), n, fill = species))+
+#   geom_col(color = "black")+
+#   scale_fill_manual(values = palette) +
+#   theme(axis.title = element_blank(),
+#         axis.text.y = element_blank(),
+#         axis.text.x = element_text(size = 8),
+#         axis.ticks.y = element_blank(),
+#         legend.position = c(0.6, 0.95),
+#         legend.title = element_blank())+
+#   coord_flip()
+# 
+# cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))
 
@@
   geom_col(color = "black")+
-  scale_fill_brewer(palette = "Paired")+
+  scale_fill_manual(values = palette) +
   theme(axis.title.x = element_blank(),
@@
 
-# GyrB Protein sequence alignment
+# Phylogeny
+## Core Genome MLST
+```{r, fig.width=16, fig.height=14, message = FALSE, warning = FALSE}
+cgMLSTdata <- read.table("../data/cgMLST.txt", sep = "\t", row.names=1, colClasses = "factor", header = T)
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+cgMLST_tree <- as.phylo(hclust(daisy(cgMLSTdata, metric="gower")))
+
+ids <- as.data.frame(cgMLST_tree$tip.label)
+names(ids) <- "id"
+
+ids2 <- ids %>%
+  left_join(isolate_data[c("id","species")]) %>%
+  left_join(mlst_data[c("id", "ST")])
+
+cgMLST_tree$tip.label <- ids2$species
+
+group_info <- split(cgMLST_tree$tip.label, f = cgMLST_tree$tip.label)
+cgMLST_tree <- groupOTU(cgMLST_tree, group_info)
+
+cgMLST_tree$tip.label <- ids2$id
+
+p_cgmlst <- ggtree(cgMLST_tree, size = 0.2, color = "#808080", layout = "circular")+
+  geom_tippoint(aes(color = group), size = 2.5)+
+  labs(color = "Species")+
+  scale_color_manual(values = palette)+
+  guides(color = guide_legend(override.aes = list(size = 8)))+
+  theme(legend.position = "right")
+
+palette3 <- c("1-A" = "#8dd3c7",
+              "1-I" = "#80b1d3",
+              "0" = "grey95")
+
+qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
+  mutate_at(vars(-id),
+            funs(ifelse(. == "1", "1-I", "0")))
+
+ac_genes <- acquired_report %>%
+  mutate_at(vars(-id),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
+  select(id, qepA4, qnr) %>%
+  mutate_at(vars(-id),
+            funs(ifelse(. == "1", "1-A", "0")))
+
+total_gene_data <- qr_genes %>%
+  left_join(ac_genes, by = "id") %>%
+  column_to_rownames("id")
+
+fan_tree <- open_tree(p_cgmlst, 9)
+
+rotated_fan_tree <- rotate_tree(fan_tree, 95)
+
+gheatmap(rotated_fan_tree,
+         total_gene_data,
+         offset = 0.005,
+         width = 0.2,
+         colnames_offset_y = 3,
+         colnames_position = "top",
+         font.size = 2.5) +
+  scale_fill_manual(values = palette3,
+                    labels = c("Absent",
+                               "Acquired gene present",
+                               "Intrinsic gene mutated"))
+```
+
 ```{r}
-alignment <- Biostrings::readAAMultipleAlignment("../data/alignment.fasta")
-msaR::msaR(alignment)
-```
+# alignment <- Biostrings::readAAMultipleAlignment("../data/alignment.fasta")
+# msaR::msaR(alignment)
+```

### 2018-12-25 notebooks/assembly_metrics.Rmd
initial commit

---
title: "Assembly Quality Metrics - Quast Results"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r, include = FALSE}
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, magrittr)

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

```

```{r}
path <- "../data/quast/"

file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = ".tsv")
  return(files)
}

create_df <- function(df) {
  df <- df %>%
    rename(ctg_0_bp = X..contigs.....0.bp.,
           ctg_1000_bp = X..contigs.....1000.bp.,
           ctg_5000_bp = X..contigs.....5000.bp.,
           ctg_10000_bp = X..contigs.....10000.bp.,
           ctg_25000_bp = X..contigs.....25000.bp.,
           ctg_50000_bp = X..contigs.....50000.bp.,
           tot_len_0_bp = Total.length.....0.bp.,
           tot_len_1000_bp = Total.length.....1000.bp.,
           tot_len_5000_bp = Total.length.....5000.bp.,
           tot_len_10000_bp = Total.length.....10000.bp.,
           tot_len_25000_bp = Total.length.....25000.bp.,
           tot_len_50000_bp = Total.length.....50000.bp.,
           no_of_contigs = X..contigs,
           no_of_misassemblies = X..misassemblies,
           no_of_misassembled_ctg = X..misassembled.contigs,
           no_of_local_misassemblies = X..local.misassemblies,
           unaligned_contigs = X..unaligned.contigs,
           N_per_100kbp = X..N.s.per.100.kbp,
           mismatches_per_100kbp = X..mismatches.per.100.kbp,
           indels_per_100kbp = X..indels.per.100.kbp,
           genes = X..genes,
           GC = GC....,
           duplication_ratio = Duplication.ratio,
           largest_contig = Largest.contig)
  return(df)
}

```

# Import files and clean data
```{r}
files <- file_names(path)

data_list <- lapply(files,
                    FUN = function(file) {
                      read.delim(paste0(path, "/", file),
                                 sep = "\t",
                                 header = TRUE,
                                 stringsAsFactors = FALSE)
                    })

data_list <- lapply(data_list, function(df) mutate_all(df, as.character))

names(data_list) <- files

df <- bind_rows(data_list)

df_clean <- create_df(df) %>%
  filter(grepl("broken", Assembly) == FALSE) %>%
  filter(Assembly %not_in% c("51-2016-02-486_S31_pilon_spades",
                             "68-2016-17-565_S45_pilon_spades",
                             "52-2016-02-732_S43_pilon_spades",
                             "50-2016-02-428_S19_pilon_spades",
                             "4-2014-01-1741_TAAGGCGA-ACTGCATA_pilon_spades")) %>%
  mutate(ref = Assembly %>%
           gsub("_pilon_spades", "", .)) %>%
  left_join(id_data, by = "ref") %>%
  left_join(isolate_data[c("id","sequencer")], by = "id")

metrics <- df_clean %>%
  select(id, no_of_contigs, largest_contig, GC, N50, NG50, L50, LG50, no_of_misassemblies, duplication_ratio, genes) %>%
  mutate_at(vars(no_of_contigs, largest_contig, N50),
            funs(as.numeric))

metrics

metrics2 <- metrics %>%
  left_join(isolate_data, by = "id")

embed_data <- function(x = mtcars, filename = "file.csv", label= "Get data") {

  # Create encoded Base64 datastream 
  encode_data <- function(x){
    write.csv2(x, "./file.csv")
    enc= sprintf('data:text/csv;base64,%s', openssl::base64_encode(paste0(readLines("./file.csv"), collapse="\n")) )
    unlink("./file.csv")
    return(enc)
  }

  # String result ready to be placed in rmarkdown
  paste0("<a download='", filename, "' href=", encode_data(x), ">", label, "</a>")

}
```

`r embed_data(metrics, filename = "assembly_metrics.csv")`

# Plots

```{r}
ggplot(metrics2, aes(as.numeric(largest_contig), as.numeric(N50)))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()+
  labs(x = "Largest contig size",
       y = "N50")

```

```{r}
ggplot(metrics2, aes(as.numeric(no_of_contigs), as.numeric(N50)))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()+
  labs(x = "Number of contigs",
       y = "N50")
```

### 2019-01-15 figure_script.R
updated with new colours and order

@@ -4,16 +4,15 @@ gene_names <- names(mut_report)
 
 gene_names <- gene_names[-1]
 
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
 ggsave(
   "C:/Users/VI1511/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/percent_mut_per_species.tiff",
   mut_report %>%
-    mutate(
-      ref = gsub("(.*?)\\_.+", "\\1", ref),
-      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
-      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
-    ) %>%
-    arrange(desc(ref)) %>%
-    mutate(id = id) %>%
     left_join(isolate_data, by = "id") %>%
     gather(key, value, gene_names) %>%
     group_by(key, value, species) %>%
@@ -28,6 +27,7 @@ ggsave(
       lwr = round(get_binCI(Present, Total)[1], 1),
       upr = round(get_binCI(Present, Total)[2], 1)
     ) %>%
+    mutate(species = factor(species, levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
     ggplot(aes(key, Percent, fill = species)) +
     geom_col(color = "black",
              position = position_dodge(1)) +
@@ -36,11 +36,8 @@ ggsave(
       width = 0.6,
       position = position_dodge(1)
     ) +
-    scale_x_discrete(limits = c(
-      "gyrA", "gyrB", "parC", "parE",
-      "marR", "soxS", "tolC"
-    )) +
-    scale_fill_brewer(palette = "Paired") +
+    scale_fill_manual(values = palette) +
+    scale_x_discrete(limits = c("gyrA","gyrB","parC","parE","marA","marR","acrB","rpoB")) +
     labs(y = "Percent (%) of isolates") +
     guides(fill = FALSE) +
     theme_classic() +
@@ -65,20 +62,11 @@ ggsave(
 )
 
 
-
 qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
-
 
 ggsave(
[…]

### 2019-01-15 prokka_check.R
initial commit

@@ -1,0 +1,163 @@
+library(stringr)
+library(seqinr)
+library(tidyr)
+library(Biostrings)
+
+gyrA <- readAAStringSet("data/Prokka/DNA_gyrase_GYRA_aa_sequences.fa")
+parC <- readAAStringSet("data/Prokka/Topoisomerase_IV_subunit_A_aa_sequences.fa")
+
+mismatches <- function(query, ref) {
+  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
+                    gapOpening = 3, gapExtension = 1) %>%
+    mismatchTable() %>%
+    mutate(ID=names(query),
+           Pos=PatternStart,
+           Reference_AA=as.character(PatternSubstring),
+           Sample_AA=as.character(SubjectSubstring)) %>%
+    select(ID, Reference_AA, Sample_AA, Pos) %>%
+    filter(Sample_AA != "X")
+}  
+
+paste3 <- function(...,sep=", ") {
+  L <- list(...)
+  L <- lapply(L,function(x) {x[is.na(x)] <- ""; x})
+  ret <-gsub(paste0("(^",sep,"|",sep,"$)"),"",
+             gsub(paste0(sep,sep),sep,
+                  do.call(paste,c(L,list(sep=sep)))))
+  is.na(ret) <- ret==""
+  ret
+}
+
+ariba_flags_gyra <- mut_flags %>%
+  select(-flag) %>%
+  filter(gene == "gyrA") %>%
+  mutate(gyrA_result = ref_ctg_change %>%
+           str_extract_all("\\d+") %>% 
+           map(as.integer) %>%
+           map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
+  mutate(type = case_when(gyrA_result == TRUE ~ "QRDR",
+                          gyrA_result == FALSE & ref_ctg_change != "." ~ "non_QRDR",
+                          gyrA_result == FALSE & ref_ctg_change == "." ~ "wild type")) %>%
+  select(-gyrA_result) %>%
+  group_by(id) %>%
+  mutate(test = 1:n()) %>%
+  ungroup()
+
+mut <- mut_quant %>%
+  select(id, gyrA) %>%
+  mutate(entries = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)})) %>%
+  separate(gyrA, into = as.character(c(1:max(.$entries)))) %>%
+  select(-entries) %>%
[…]

### 2019-01-15 notebooks/article_notebook.Rmd
updated with mlst data and adjusted a few figures

@@
 Sequence types was also identified by ARIBA, using the Wirth et al. scheme (*E. coli* #1).
-
-## Further work
-
-- Assemble genomes
-
-- Align assemblies
-
-- Land on methodology for phylogenetic analysis (likely core genome SNP - ML)
 
@@
 
+## Sequence types
+```{r}
+mlst_data[,c("id","ST")] %>%
+  group_by(ST) %>%
+  count()
+```
+
+## Sequence types per animal species
+```{r}
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(ST, species) %>%
+  count()
+```
+
 ## Significant differences between resistance mechanisms for each animal species
@@
     ) %>%
+    mutate(species = factor(species,
+                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
     ggplot(aes(key, Percent, fill = species)) +
@@
     scale_fill_manual(values = palette) +
+    scale_x_discrete(limits = c("gyrA","gyrB","parC","parE","marA","marR","acrB","rpoB")) +
     labs(y = "Percent (%) of isolates") +
@@
 
-```{r, fig.width=14, fig.height=16,message = FALSE, warning = FALSE}
-# # identify which sequence types are represented over three times
-# n_ST <- mlst_data[c("id","ST")] %>%
-#   group_by(ST) %>%
-#   count() %>%
-#   filter(n > 3)
-# 
-# # species ST  
-# species_st <- mlst_data[c("id","ST")] %>%
-#   left_join(isolate_data[c("id", "species")], by = "id") %>%
-#   select(-id) %>%
-#   group_by(ST, species) %>%
-#   count() %>%
-#   filter(ST %in% n_ST$ST)
-# 
-# # prepare micplot data for each ST with n > 3
-# micplot_st_df <- mlst_data[c("id", "ST")] %>%
-#   left_join(mic_data[c("id", "CIP")], by = "id") %>%
-#   group_by(ST, CIP) %>%
-#   count() %>%
-#   filter(ST %in% n_ST$ST)
-# 
-# # prepare heatmap data frame
-# st_heatmap_df <- mlst_data[c("id", "ST")] %>%
-#   left_join(mut_report[c("id", "marR")], by = "id") %>%
-#   left_join(cols_to_add, by = "id") %>%
-#   left_join(acquired_report, by = "id") %>%
-#   select(-id) %>%
-#   mutate_at(vars(-ST),
-#             funs(as.numeric)) %>%
-#   group_by(ST) %>%
-#   summarise_all(funs(sum, n = n())) %>%
-#   ungroup() %>%
-#   rename("n" = marR_n) %>%
-#   select(-contains("_n")) %>%
-#   gather(mechanism, value, -c(ST, n)) %>%
-#   mutate(mechanism = sub("_sum", "", mechanism),
-#          percent = round(value/n*100),
-#          value2 = ifelse(value > 0, 1, 0)) %>%
-#   filter(ST %in% n_ST$ST)
-# 
-# # plots
-# factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
-# factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")
-# 
-# ## heatmap
-# heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, factor(ST), fill = percent))+
-#   geom_tile()+
-#   scale_fill_viridis(option = "A")+
-#   guides(fill = FALSE)+
-#   theme(axis.text.x = element_text(angle = 90,
-#                                    hjust = 1,
-#                                    vjust = 0.3,
-#                                    size = 8),
-#         axis.title.x = element_blank())
-# 
-# ## micplot
-# micplot_st <- ggplot(micplot_st_df, aes(factor(ST), CIP))+
-#   geom_boxplot(fill = "#67a9cf")+
-#   scale_y_continuous(labels = factor_levels,
-#                      breaks = factor_breaks,
-#                      trans = "log10")+
-#   theme(axis.title = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.text.x = element_text(size = 8),
-#         axis.ticks.y = element_blank())+
-#   coord_flip()
-# 
-# # n plot
-# nplot_st <- ggplot(species_st, aes(factor(ST), n, fill = species))+
-#   geom_col(color = "black")+
-#   scale_fill_manual(values = palette) +
-#   theme(axis.title = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.text.x = element_text(size = 8),
-#         axis.ticks.y = element_blank(),
-#         legend.position = c(0.6, 0.95),
-#         legend.title = element_blank())+
-#   coord_flip()
-# 
-# cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))
-
-```
-
 ## Phenotypic resistance combinations
@@
 
-gheatmap(rotated_fan_tree,
+final_tree <- gheatmap(rotated_fan_tree,
          total_gene_data,
@@
   scale_fill_manual(values = palette3,
-                    labels = c("Absent",
+                    labels = c("Gene/Mutation absent",
                                "Acquired gene present",
                                "Intrinsic gene mutated"))
+final_tree
 ```

### 2019-01-15 notebooks/article_notebook.Rmd
added figure for total resistance mechanisms in all isolates

@@
 # Figures
+
+## Total resistance mechanisms
+```{r}
+total_mechanisms <- mut_report %>%
+  select(-nt_change) %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,11:17]) == 0, 0, 1)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  gather(gene, result, -id) %>%
+  group_by(gene, result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
+  spread(result, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round((Present/Total)*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
+         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))
+
+palette2 <- c("Acquired" = "#ef8a62",
+              "Intrinsic" = "#67a9cf")
+
+ggplot(total_mechanisms, aes(reorder(gene, -Percent), Percent, fill = type)) +
+  geom_col(color = "black") +
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                width = 0.5) +
+  scale_fill_manual(values = palette2) +
+  labs(x = "Genes",
+       y = "Percent (%) Isolates",
+       fill = NULL)
+
+```
+
 ## Percent isolates with PMQR and/or QRDR mutations in *gyrA*, *gyrB*, *parC* or *parE*

### 2019-01-18 notebooks/article_notebook2.Rmd
added new main notebook

---
title: "Article data analysis 2"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr)
```

```{r Import data, include = FALSE}
# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

two_reads <- read.table("../data/no_of_reads.txt",
                          sep = "\t",
                          header = F,
                          stringsAsFactors = F) %>%
  dplyr::rename("ref" = V1,
         "reads" = V2) %>%
  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
  group_by(ref) %>%
  mutate(id = n()) %>%
  {. ->> four_reads} %>%
  filter(id == 2) %>%
  spread(type, reads) %>%
  dplyr::rename("R1-1" = R1,
         "R2-1" = R2) %>%
  mutate("R1-2" = NA,
         "R2-2" = NA) %>%
  select(-id) %>%
  mutate_all(funs(as.character(.)))

four_reads <- four_reads %>%
  filter(id == 4) %>%
  group_by(ref) %>%
  mutate(id = 1:n()) %>%
  spread(id, reads) %>%
  summarise_all(funs(func_paste)) %>%
  dplyr::rename("R1-1" = `1`,
         "R2-1" = `2`,
         "R1-2" = `3`,
         "R2-2" = `4`) %>%
  select(-type) %>%
  mutate_all(funs(as.character(.)))

no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

mlst_data <- read.table("../data/mlst_results.txt", sep = "\t", header = TRUE, stringsAsFactors = F) %>%
  filter(id %not_in% "2014-01-1741-1-S")
```

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx","qep","mfp","mcb")
mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions

func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    dplyr::rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change, ref_nt, ctg_nt) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result),
           nt_change = paste(ref_nt, ctg_nt, sep = "-")) %>%
    mutate(nt_change = ifelse(nt_change == ".-.", "0", nt_change)) %>%
    dplyr::rename("gene" = gene_names) %>%
    select(-c(ref_nt, ctg_nt))
  return(df)
}

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
    select(ref, gene_names, flag, mutation) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, mutation) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut",
           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(id) %>%
    mutate(id2 = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id2, type))
  return(df)
}

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

acquired_flags <- check_flags(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

acquired_filtered <- filter_acquired_table(acquired_table)

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

# ---------------------------------- Tables

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  dplyr::count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  dplyr::count()
```

# Double-check ARIBA results with annotated sequences from Prokka
```{r, warning=FALSE, message=FALSE}
# Functions
mismatches <- function(query, ref) {
  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
                    gapOpening = 3, gapExtension = 1) %>%
    mismatchTable() %>%
    mutate(ID=names(query),
           Pos=PatternStart,
           Reference_AA=as.character(PatternSubstring),
           Sample_AA=as.character(SubjectSubstring)) %>%
    select(ID, Reference_AA, Sample_AA, Pos) %>%
    filter(Sample_AA != "X")
}

# Import sequence data
gyrA <- readAAStringSet("../data/Prokka/DNA_gyrase_GYRA_aa_sequences.fa")
parC <- readAAStringSet("../data/Prokka/Topoisomerase_IV_subunit_A_aa_sequences.fa")
parE <- readAAStringSet("../data/Prokka/DNA_topoisomerase_IV_subunit_B_aa_sequences.fa")

# Identify results from ARIBA
mut_gyrA <- mut_quant %>%
  select(id, gyrA) %>%
  mutate(entries = sapply(strsplit(.$gyrA, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(gyrA, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
gyrA_results <- bind_rows(lapply(seq_along(gyrA[-1]),
                            function(i) mismatches(gyrA[i + 1],
                                                   gyrA[1]))) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "gyrA_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, gyrA_prokka) %>%
  mutate(gyrA_result = gyrA_prokka %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
  filter(gyrA_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_gyrA, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("gyrA_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-gyrA_result)

# Identify results from ARIBA
mut_parC <- mut_quant %>%
  select(id, parC) %>%
  mutate(entries = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)})) %>%
  separate(parC, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parC_results <- bind_rows(lapply(seq_along(parC[-1]),
                                 function(i) mismatches(parC[i + 1],
                                                        parC[1]))) %>%
  mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parC_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parC_prokka) %>%
  mutate(parC_result = parC_prokka %>%
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L))) %>%
  filter(parC_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parC, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parC_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-parC_result)

# Identify results from ARIBA
mut_parE <- mut_quant %>%
  select(id, parE) %>%
  mutate(entries = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
  separate(parE, into = as.character(c(1:max(.$entries)))) %>%
  select(-entries) %>%
  gather(key, value, `1`,`2`) %>%
  filter(is.na(value) == FALSE) %>%
  select(-key) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup()

# Run mutation identification analysis on Prokka sequences and
# match to ARIBA results
parE_results <- bind_rows(lapply(seq_along(parE[-1]),
                                 function(i) mismatches(parE[i + 1],
                                                        parE[1]))) %>%
  #mutate(Pos = Pos + 22) %>%
  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
  select(ID, mutation) %>%
  mutate(ID = sub(".faa", "", ID)) %>%
  dplyr::rename("ref" = ID,
                "parE_prokka" = mutation) %>%
  left_join(id_data, by = "ref") %>%
  select(id, parE_prokka) %>%
  mutate(parE_result = parE_prokka %>%
           str_extract_all("\\d+") %>% 
           map(as.integer) %>% 
           map_lgl(~ any(.x >= 366L & .x <= 523L))) %>%
  filter(parE_result == TRUE) %>%
  group_by(id) %>%
  mutate(test = 1:n()) %>%
  ungroup() %>%
  left_join(mut_parE, by = c("id","test")) %>%
  select(-test) %>%
  dplyr::rename("parE_ariba" = value) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  select(-parE_result)

full_report <- gyrA_results %>%
  left_join(parC_results, by = "id") %>%
  left_join(parE_results, by = "id") %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  mutate(parC_ariba = ifelse(parC_prokka == "S80I" & 
                               parC_ariba == "S58I", "S80I", parC_ariba))

full_report

```

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## List of genes of interest
This table lists all the genes of interest in this study, some information about them, and their accession number(s).
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
                          grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  dplyr::rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxR", cluster, ignore.case = T) == TRUE ~ "soxR",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
                          grepl("acrb", ref_name, ignore.case = T) == TRUE ~ "acrB",
                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA",
                          grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  dplyr::rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "acrB" ~ "Gene that codes for the AcrB efflux pump, part of the AcrAB-TolC efflux system",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "marA" ~ "Transcriptional activator og genes involved in the MAR phenotype.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxR" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
                                 Gene == "mdfA" ~ "Chromosomal efflux pump.",
                                 Gene == "rpoB" ~ "Beta-subunit of RNA polymerase, mutations lead to increased expression of the ydhE-pump"),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

```

## Isolates per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  dplyr::count()

per_species
```

## Sequence types in total
```{r}
tot_mlst <- mlst_data[,c("id","ST")] %>%
  group_by(ST) %>%
  dplyr::count() %>%
  arrange(desc(n))

print(paste0("Total amount of different sequence types: ", nrow(tot_mlst)))

tot_mlst
```

## Sequence types per animal species
```{r}
mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  dplyr::count() %>%
  arrange(desc(n))
```

## Number of sequence types per animal species
```{r}
mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = length(unique(ST))) %>%
  select(species,n) %>%
  summarise_all(funs(func_paste))
```

## Percent isolates with PMQR
```{r}
acquired_report %>%
  select(-id) %>%
  mutate_all(funs(as.numeric(.))) %>%
  mutate(total = rowSums(.),
         result = if_else(total != 0, 1, 0)) %>%
  group_by(result) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
  spread(result, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1))
```

## Percent isolates with specific PMQR genes
```{r}
acquired_report %>%
  select(-id) %>%
  t() %>%
  as.data.frame(.) %>%
  mutate(Gene = rownames(.)) %>%
  mutate_at(vars(-Gene),
            funs(as.numeric(as.character(.)))) %>%
  mutate(Present = rowSums(.[,-281])) %>%
  select(c(Gene, Present)) %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(Present/Total * 100,1))
```

## Percent isolates with mutations in QRDR
```{r}
mut_report %>%
  select(gyrA, gyrB, parC, parE) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
            .funs = funs(as.numeric)) %>%
  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
  select(n, type) %>%
  group_by(type) %>%
  summarise_all(funs(sum)) %>%
  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
  spread(type, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  select(-Total)
```

## Specific mutations in QRDR
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  gather(gene, mut, -ref) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  dplyr::count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut))
```

## Mutation combinations in QRDR
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1))
```

# Figures
## Resistance mechanisms per sequence type
```{r}
genes <- c(names(mut_report), names(acquired_report), "qnr")
genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")]

no_of_mlst <- mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  dplyr::count() %>%
  group_by(ST) %>%
  mutate(group_n = sum(n)) %>%
  filter(group_n > 4)

mlst_mechanisms_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mlst_data[c("id","ST")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,ST)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  group_by(ST, gene, value) %>%
  dplyr::count() %>%
  filter(ST %in% no_of_mlst$ST)

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

p1 <- ggplot(mlst_mechanisms_report, aes(factor(ST), gene, size = n))+
  geom_point()+
  guides(size = FALSE)+
  labs(x = "Sequence Types",
       y = "Genes")+
  scale_y_discrete(limits = c("gyrA","parC","parE","marA","marR","rpoB","qnr")) +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())

p2 <- ggplot(no_of_mlst, aes(factor(ST), n, fill = species))+
  geom_col(color = "black")+
  scale_fill_manual(values = palette)+
  guides(fill = FALSE)+
  theme(axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank())

plot_grid(p2, p1, nrow = 2, align = "v")

```

## Percent resistance mechanisms in total
```{r}
total_mechanisms <- mut_report %>%
  select(-nt_change) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,11:17]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, result, -id) %>%
  group_by(gene, result) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1),
         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))

palette2 <- c("Acquired" = "#8dd3c7",
              "Intrinsic" = "#80b1d3")

ggplot(total_mechanisms, aes(reorder(gene, -Percent), Percent, fill = type)) +
  geom_col(color = "black") +
  scale_fill_manual(values = palette2) +
  labs(x = "Genes",
       y = "Percent (%) Isolates",
       fill = NULL)

```

## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-c(1,2)]

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1)
    ) %>%
    mutate(species = factor(species,
                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    scale_fill_manual(values = palette) +
    scale_x_discrete(limits = c("gyrA","gyrB","parC","parE","marA","marR","acrB","rpoB")) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```

## cgMLST-trees
```{r, fig.width=24, fig.height=22, message = FALSE, warning = FALSE}

palette3 <- c("1-A" = "#8dd3c7",
              "1-I" = "#80b1d3",
              "0" = "grey95")

qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
  mutate_at(vars(-id),
            funs(ifelse(. == "1", "1-I", "0")))

ac_genes <- acquired_report %>%
  mutate_at(vars(-id),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
  select(id, qepA4, qnr) %>%
  mutate_at(vars(-id),
            funs(ifelse(. == "1", "1-A", "0")))

cgMLSTdata <- read.table("../data/cgMLST.txt",
                         sep = "\t",
                         row.names=1,
                         colClasses = "factor",
                         header = T)

broiler_iso <- isolate_data[c("id", "species")] %>%
  filter(species == "Broiler") %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

pig_iso <- isolate_data[c("id", "species")] %>%
  filter(species == "Pig") %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

fox_iso <- isolate_data[c("id", "species")] %>%
  filter(species == "Red fox") %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

bird_iso <- isolate_data[c("id", "species")] %>%
  filter(species == "Wild bird") %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

data_broiler <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% broiler_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_pig <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% pig_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_fox <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% fox_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_bird <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% bird_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

cgMLSTdata_broiler <- read.table("../data/cgMLST_broiler.txt",
                                 sep = "\t",
                                 row.names=1,
                                 colClasses = "factor",
                                 header = T)

cgMLST_tree_broiler <- as.phylo(hclust(daisy(cgMLSTdata_broiler,
                                             metric="gower")))

cgMLSTdata_pig <- read.table("../data/cgMLST_pig.txt",
                             sep = "\t",
                             row.names=1,
                             colClasses = "factor",
                             header = T)

cgMLST_tree_pig <- as.phylo(hclust(daisy(cgMLSTdata_pig,
                                         metric="gower")))

cgMLSTdata_fox <- read.table("../data/cgMLST_fox.txt", 
                             sep = "\t", 
                             row.names=1, 
                             colClasses = "factor", 
                             header = T)

cgMLST_tree_fox <- as.phylo(hclust(daisy(cgMLSTdata_fox,
                                         metric="gower")))

cgMLSTdata_bird <- read.table("../data/cgMLST_bird.txt", 
                              sep = "\t", 
                              row.names=1, 
                              colClasses = "factor", 
                              header = T)

cgMLST_tree_bird <- as.phylo(hclust(daisy(cgMLSTdata_bird,
                                          metric="gower")))

broiler_tree <- ggtree(cgMLST_tree_broiler,
                       layout = "circular",
                       color = "#808080",
                       size = 1) %<+% data_broiler +
  geom_tippoint(size = 4,
                color = "#4575b4")+
  ggtitle("A")+
  theme(plot.title = element_text(size = 40, face = "bold"),
        plot.margin=unit(c(0,0,0,0), "cm"))

pig_tree <- ggtree(cgMLST_tree_pig,
                   layout = "circular",
                   color = "#808080",
                   size = 1) %<+% data_pig +
  geom_tippoint(size = 4,
                color = "#74add1")+
  ggtitle("B")+
  theme(plot.title = element_text(size = 40, face = "bold"),
        plot.margin=unit(c(0,0,0,0), "cm"))

fox_tree <- ggtree(cgMLST_tree_fox,
                   layout = "circular",
                   color = "#808080",
                   size = 1) %<+% data_fox +
  geom_tippoint(size = 4,
                color = "#f46d43")+
  ggtitle("C")+
  theme(plot.title = element_text(size = 40, face = "bold"),
        plot.margin=unit(c(0,0,0,0), "cm"))

bird_tree <- ggtree(cgMLST_tree_bird,
                    layout = "circular",
                    color = "#808080",
                    size = 1) %<+% data_bird +
  geom_tippoint(size = 4,
                color = "#fdae61")+
  ggtitle("D")+
  theme(plot.title = element_text(size = 40, face = "bold"),
        plot.margin=unit(c(0,0,0,0), "cm"))

total_data <- qr_genes %>%
  left_join(ac_genes, by = "id")

broiler_data <- total_data %>%
  filter(id %in% broiler_iso$id) %>%
  column_to_rownames("id")

pig_data <- total_data %>%
  filter(id %in% pig_iso$id) %>%
  column_to_rownames("id")

fox_data <- total_data %>%
  filter(id %in% fox_iso$id) %>%
  column_to_rownames("id")

bird_data <- total_data %>%
  filter(id %in% bird_iso$id) %>%
  column_to_rownames("id")

fan_tree_broiler <- open_tree(broiler_tree, 10)

rotated_fan_tree_broiler <- rotate_tree(fan_tree_broiler, 93)

broiler_annotated_tree <- gheatmap(rotated_fan_tree_broiler,
                                   broiler_data,
                                   offset = 0.01,
                                   width = 0.5,
                                   colnames_offset_y = 0.7,
                                   colnames_position = "top",
                                   font.size = 7) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

fan_tree_pig <- open_tree(pig_tree, 10)

rotated_fan_tree_pig <- rotate_tree(fan_tree_pig, 92)

pig_annotated_tree <- gheatmap(rotated_fan_tree_pig, 
                               pig_data,
                               offset = 0.01, 
                               width = 0.5, 
                               colnames_offset_y = 0.55, 
                               colnames_position = "top", 
                               font.size = 7) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

fan_tree_fox <- open_tree(fox_tree, 10)

rotated_fan_tree_fox <- rotate_tree(fan_tree_fox, 92)

fox_annotated_tree <- gheatmap(rotated_fan_tree_fox, 
                               fox_data, 
                               offset = 0.01, 
                               width = 0.5, 
                               colnames_offset_y = 0.2, 
                               colnames_position = "top", 
                               font.size = 7) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

fan_tree_bird <- open_tree(bird_tree, 10)

rotated_fan_tree_bird <- rotate_tree(fan_tree_bird, 93)

bird_annotated_tree <- gheatmap(rotated_fan_tree_bird, 
                                bird_data, 
                                offset = 0.01, 
                                width = 0.5, 
                                colnames_offset_y = 0.4, 
                                colnames_position = "top", 
                                font.size = 7) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

total_plot <- cowplot::plot_grid(broiler_annotated_tree, 
                                 pig_annotated_tree, 
                                 fox_annotated_tree, 
                                 bird_annotated_tree, 
                                 ncol = 2, 
                                 align = "h")
total_plot

```

## Another color scheme
```{r,fig.width=24, fig.height=22, message = FALSE, warning = FALSE}

qr_genes_alt <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
  mutate(gyrA = ifelse(gyrA == 1, "gyrA_1", gyrA),
         parC = ifelse(parC == 1, "parC_1", parC),
         parE = ifelse(parE == 1, "parE_1", parE),
         rpoB = ifelse(rpoB == 1, "rpoB_1", rpoB),
         marA = ifelse(marA == 1, "marA_1", marA),
         marR = ifelse(marR == 1, "marR_1", marR))

ac_genes_alt <- acquired_report %>%
  mutate_at(vars(-id),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
  select(id, qepA4, qnr) %>%
  mutate(qnr = ifelse(qnr == 1, "qnr_1", qnr),
         qepA4 = ifelse(qepA4 == 1, "qepA4_1", qepA4))

total_data_alt <- qr_genes_alt %>%
  left_join(ac_genes_alt, by = "id")

palette_heatmap <- c("gyrA_1" = "#8dd3c7",
                     "parC_1" = "#ffffb3",
                     "parE_1" = "#bebada",
                     "rpoB_1" = "#fb8072",
                     "marA_1" = "#80b1d3",
                     "marR_1" = "#fdb462",
                     "qepA4_1" = "#b3de69",
                     "qnr_1" = "#fccde5",
                     "0" = "grey95")

broiler_tree <- ggtree(cgMLST_tree_broiler,
                       layout = "circular",
                       color = "#4575b4",
                       size = 1) %<+% data_broiler +
  geom_tippoint(size = 2,
                color = "black")+
  ggtitle("A")+
  theme(plot.title = element_text(size = 20, face = "bold"))

broiler_data <- total_data_alt %>%
  filter(id %in% broiler_iso$id) %>%
  column_to_rownames("id")

broiler_annotated_tree2 <- gheatmap(broiler_tree,
                                   broiler_data,
                                   offset = 0.01,
                                   width = 0.3,
                                   colnames = FALSE)+
  scale_fill_manual(values = palette_heatmap)+
  guides(fill = FALSE)

pig_tree <- ggtree(cgMLST_tree_pig,
                       layout = "circular",
                       color = "#74add1",
                       size = 1) %<+% data_pig +
  geom_tippoint(size = 2,
                color = "black")+
  ggtitle("B")+
  theme(plot.title = element_text(size = 20, face = "bold"))

pig_data <- total_data_alt %>%
  filter(id %in% pig_iso$id) %>%
  column_to_rownames("id")

pig_annotated_tree2 <- gheatmap(pig_tree,
                                   pig_data,
                                   offset = 0.01,
                                   width = 0.3,
                                   colnames = FALSE)+
  scale_fill_manual(values = palette_heatmap)+
  guides(fill = FALSE)

fox_tree <- ggtree(cgMLST_tree_fox,
                       layout = "circular",
                       color = "#f46d43",
                       size = 1) %<+% data_fox +
  geom_tippoint(size = 2,
                color = "black")+
  ggtitle("C")+
  theme(plot.title = element_text(size = 20, face = "bold"))

fox_data <- total_data_alt %>%
  filter(id %in% fox_iso$id) %>%
  column_to_rownames("id")

fox_annotated_tree2 <- gheatmap(fox_tree,
                                   fox_data,
                                   offset = 0.01,
                                   width = 0.3,
                                   colnames = FALSE)+
  scale_fill_manual(values = palette_heatmap)+
  guides(fill = FALSE)

bird_tree <- ggtree(cgMLST_tree_bird,
                       layout = "circular",
                       color = "#fdae61",
                       size = 1) %<+% data_bird +
  geom_tippoint(size = 2,
                color = "black")+
  ggtitle("D")+
  theme(plot.title = element_text(size = 20, face = "bold"))

bird_data <- total_data_alt %>%
  filter(id %in% bird_iso$id) %>%
  column_to_rownames("id")

bird_annotated_tree2 <- gheatmap(bird_tree,
                                   bird_data,
                                   offset = 0.01,
                                   width = 0.3,
                                   colnames = FALSE)+
  scale_fill_manual(values = palette_heatmap)+
  guides(fill = FALSE)

total_plot2 <- cowplot::plot_grid(broiler_annotated_tree2, 
                                 pig_annotated_tree2, 
                                 fox_annotated_tree2, 
                                 bird_annotated_tree2, 
                                 ncol = 2, 
                                 align = "h")
total_plot2
```

### 2019-01-18 notebooks/assembly_metrics.Rmd
minor changes

@@
 
-# Plots
-
+# Mean metrics
 ```{r}
-ggplot(metrics2, aes(as.numeric(largest_contig), as.numeric(N50)))+
-  geom_point()+
-  geom_smooth(method = "lm")+
-  theme_classic()+
-  labs(x = "Largest contig size",
-       y = "N50")
+print(paste0("Mean number of contigs: ", round(mean(metrics2$no_of_contigs), 1)))
+print(paste0("Mean size of largest contigs: ", round(mean(metrics2$largest_contig), 1)))
+print(paste0("Mean N50: ", round(mean(metrics2$N50), 1)))
+print(paste0("Mean NG50: ", round(mean(as.numeric(metrics2$NG50)), 1)))
 
 ```
-
-```{r}
-ggplot(metrics2, aes(as.numeric(no_of_contigs), as.numeric(N50)))+
-  geom_point()+
-  geom_smooth(method = "lm")+
-  theme_classic()+
-  labs(x = "Number of contigs",
-       y = "N50")
-```

### 2019-01-25 notebooks/article_notebook.Rmd
minor changes
minor changes, added figures to article notebooks

@@
   left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
-  left_join(mut_report[c("id","marR","marA","acrB","rpoB")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","rpoB")], by = "id") %>%
   left_join(isolate_data[c("id", "species")])
@@
   group_by_all() %>%
-  count() %>%
+  dplyr::count() %>%
   ungroup() %>%
@@
 
+test <- hclust(daisy(cgMLSTdata, metric="gower"))
+
 ids <- as.data.frame(cgMLST_tree$tip.label)
@@
 ```{r}
-# alignment <- Biostrings::readAAMultipleAlignment("../data/alignment.fasta")
-# msaR::msaR(alignment)
-```
+alignment <- Biostrings::readAAMultipleAlignment("../data/alignment.fasta")
+msaR::msaR(alignment)
+```

### 2019-01-25 notebooks/article_notebook2.Rmd
minor changes, added figures to article notebooks

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr, scales)
 ```
@@
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","oqx","qep","mfp","mcb")
-mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")
+ac_genes <- c("qnr","oqx","qep","aac6")
+mut_genes <- c("gyr","par","marR","marA","sox","rpoB","rob","gad")
 
@@
 mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
+  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
   left_join(id_data, by = "ref") %>%
@@
 ```{r}
-
-acquired_ref_names <- acquired_data$ref_name
-acquired_ref <- c()
-
-for (i in ac_genes) {
-  for (j in acquired_ref_names) {
-    if (grepl(i,j, ignore.case = T) == TRUE) {
-      acquired_ref <- c(acquired_ref, j)
-    }
-  }
-}
-
-acquired_ref <- unique(acquired_ref)
-
-acquired_ref_genes <- acquired_data %>%
-  mutate(test = duplicated(ref_name),
-         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
-                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
-                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
-                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
-                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
-                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
-                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
-                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
-                          grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
-  filter(test == FALSE,
-         ref_name %in% acquired_ref_names) %>%
-  select(ref_name, gene) %>%
-  mutate(compl = complete.cases(.)) %>%
-  filter(compl == TRUE) %>%
-  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
-  group_by(gene) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(gene, ref_name2) %>%
-  dplyr::rename("Accession number(s)" = ref_name2,
-         "Gene" = gene) %>%
-  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
-         Type = "Acquired")
-
-df_ref_names <- mut_data$cluster
-ref_names <- c()
-
-for (i in mut_genes) {
-  for (j in df_ref_names) {
-    if (grepl(i,j, ignore.case = T) == TRUE) {
-      ref_names <- c(ref_names, j)
-    }
-  }
-}
-
-ref_names <- unique(ref_names)
-
-intrinsic_ref_genes <- mut_data %>%
-  mutate(test = duplicated(cluster),
-         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
-                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
-                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
-                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
-                          grepl("soxR", cluster, ignore.case = T) == TRUE ~ "soxR",
-                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
-                          grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
-                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
-                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
-                          grepl("acrb", ref_name, ignore.case = T) == TRUE ~ "acrB",
-                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA",
-                          grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB")) %>%
-  filter(test == FALSE,
-         cluster %in% ref_names) %>%
-  select(ref_name, gene, free_text) %>%
-  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
-  select(-ref_name) %>%
-  group_by(gene) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(gene, ref_name2) %>%
-  dplyr::rename("Accession number(s)" = ref_name2,
-         "Gene" = gene) %>%
-  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
-                                 Gene == "acrB" ~ "Gene that codes for the AcrB efflux pump, part of the AcrAB-TolC efflux system",
-                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
-                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
-                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
-                                 Gene == "marA" ~ "Transcriptional activator og genes involved in the MAR phenotype.",
-                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
-                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
-                                 Gene == "soxR" ~ "Transcriptional activator of the superoxide response regulon.",
-                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
-                                 Gene == "mdfA" ~ "Chromosomal efflux pump.",
-                                 Gene == "rpoB" ~ "Beta-subunit of RNA polymerase, mutations lead to increased expression of the ydhE-pump"),
-         Type = "Intrinsic gene")
-
-ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)
-
-ref_genes
+# 
+# acquired_ref_names <- acquired_data$ref_name
+# acquired_ref <- c()
+# 
+# for (i in ac_genes) {
+#   for (j in acquired_ref_names) {
+#     if (grepl(i,j, ignore.case = T) == TRUE) {
+#       acquired_ref <- c(acquired_ref, j)
+#     }
+#   }
+# }
+# 
+# acquired_ref <- unique(acquired_ref)
+# 
+# acquired_ref_genes <- acquired_data %>%
+#   mutate(test = duplicated(ref_name),
+#          gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
+#                           grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
+#                           grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
+#                           grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
+#                           grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
+#                           grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
+#                           grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
+#                           grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
+#                           grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
+#   filter(test == FALSE,
+#          ref_name %in% acquired_ref_names) %>%
+#   select(ref_name, gene) %>%
+#   mutate(compl = complete.cases(.)) %>%
+#   filter(compl == TRUE) %>%
+#   mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
+#   group_by(gene) %>%
+#   summarise_all(funs(func_paste)) %>%
+#   select(gene, ref_name2) %>%
+#   dplyr::rename("Accession number(s)" = ref_name2,
+#          "Gene" = gene) %>%
+#   mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
+#          Type = "Acquired")
+#   
+# df_ref_names <- mut_data$cluster
+# ref_names <- c()
+# 
+# for (i in mut_genes) {
+#   for (j in df_ref_names) {
+#     if (grepl(i,j, ignore.case = T) == TRUE) {
+#       ref_names <- c(ref_names, j)
+#     }
+#   }
+# }
+# 
+# ref_names <- unique(ref_names)
+# 
+# intrinsic_ref_genes <- mut_data %>%
+#   mutate(test = duplicated(cluster),
+#          gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
+#                           grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
+#                           grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
+#                           grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
+#                           grepl("soxS", cluster, ignore.case = T) == TRUE ~ "soxS",
+#                           grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
+#                           grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
+#                           grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB",
+#                           grepl("rob", ref_name, ignore.case = T) == TRUE ~ "robA")) %>%
+#   filter(test == FALSE,
+#          cluster %in% ref_names) %>%
+#   select(ref_name, gene, free_text) %>%
+#   mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
+#   select(-ref_name) %>%
+#   group_by(gene) %>%
+#   summarise_all(funs(func_paste)) %>%
+#   select(gene, ref_name2) %>%
+#   dplyr::rename("Accession number(s)" = ref_name2,
+#          "Gene" = gene)
+#   # mutate(Description = case_when(Gene == "gyrA" ~ ,
+#   #                                Gene == "gyrB" ~ ,
+#   #                                Gene == "marR" ~ ,
+#   #                                Gene == "marA" ~ ,
+#   #                                Gene == "parC" ~ ,
+#   #                                Gene == "parE" ~ ,
+#   #                                Gene == "soxS" ~ "Transcriptional activator of superoxide response regulon",
+#   #                                Gene == "rpoB" ~ "RNA polymerase subunit B",
+#   #                                Gene == "robA" ~ ),
+#   #        Type = "Intrinsic gene")
+#                 
+# ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)
+# 
+# ref_genes
 
@@
   arrange(desc(n))
+```
+
+## Most abundant (n > 3) sequence types per animal species
+```{r}
+mlst_n <- mlst_data[,c("id","ST")] %>%
+  group_by(ST) %>%
+  dplyr::count() %>%
+  filter(n > 3)
+
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  spread(ST, species) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-id) %>%
+  gather(ST, species) %>%
+  filter(ST %in% mlst_n$ST)
 ```
@@
   spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1))
+```
+
+## Percent isolates with qnr family genes
+```{r}
+acquired_report %>%
+  select(-c(qepA4, id)) %>%
+  mutate_all(funs(as.numeric)) %>%
+  dplyr::rename("qnrA" = qnrA1) %>%
+  mutate(qnrB = ifelse(rowSums(.[,2:4]) >= 1, 1, 0),
+         qnrS = ifelse(rowSums(.[,5:7]) >= 1, 1, 0)) %>%
+  select(qnrA, qnrB, qnrS) %>%
+  gather(gene, value) %>%
+  group_by(gene, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 0, "Absent", "Present")) %>%
+  spread(value, n) %>%
   mutate(Total = Present + Absent,
@@
 
+## qnr-genes per animal species
+```{r}
+qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
+
+acquired_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, qnr_genes) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  arrange(key, desc(Percent))
+```
+
 ## Percent isolates with mutations in QRDR
@@
   select(-Total)
+```
+
+## Percent mutations in each intrinsic gene
+```{r}
+mut_report %>%
+  select(-nt_change) %>%
+  gather(gene, value, -id) %>%
+  group_by(gene, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1)) %>%
+  arrange(desc(Percent))
 ```
@@
   select(-c(id, flag)) %>%
-  gather(gene, mut, -ref) %>%
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S",
+                       "2014-01-1741-1-S")) %>%
+  gather(gene, mut, -id) %>%
   mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
@@
   separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
   select(-mut) %>%
@@
          Percent = round(n/Total * 100,1)) %>%
-  dplyr::rename("Mutation" = mut))
+  dplyr::rename("Mutation" = mut) %>%
+  arrange(gene, desc(Percent)))
+```
+
+## Specific QRDR mutations per animal species
+```{r}
+suppressWarnings(clean_mut_data %>%
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  summarise_all(funs(func_paste2)) %>%
+  dplyr::select(-c(id, flag)) %>%
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S",
+                       "2014-01-1741-1-S")) %>%
+  gather(gene, mut, -id) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
+  select(-mut) %>%
+  dplyr::rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  group_by(gene, mut, species) %>%
+  dplyr::count() %>%
+  rowwise() %>%
+  mutate(Total = case_when(species == "Broiler" ~ 87,
+                           species == "Pig" ~ 75,
+                           species == "Red fox" ~ 52,
+                           species == "Wild bird" ~ 66),
+         Percent = round(n/Total * 100,1)) %>%
+  dplyr::rename("Mutation" = mut) %>%
+  arrange(gene, desc(Percent)))
 ```
@@
   mutate(Total = 280,
-         Percent = round(n/Total * 100,1))
+         Percent = round(n/Total * 100,1)) %>%
+  arrange(desc(Percent))
+```
+
+## Mutations in intrinsic genes per animal species
+```{r}
+gene_names <- names(mut_report)
+
+gene_names <- gene_names[-c(1,2)]
+
+mut_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, gene_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  dplyr::rename("Gene" = key,
+                "Species" = species) %>%
+  dplyr::select(Gene, Species, Present, Total, Percent) %>%
+  arrange(Gene, desc(Percent))
+```
+
+## Mechanisms in isolates without QRDR mutations
+```{r}
+mut_report %>%
+  filter(gyrA == 0,
+         parC == 0,
+         parE == 0) %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  select(-c(nt_change, id)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  dplyr::select(-c(gyrA,gyrB,parC,parE))
+```
+
+## Total combinations of resistance mechanisms
+```{r}
+mut_report %>%
+  select(-c(nt_change, gyrA, parC, parE)) %>%
+  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
+  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
+  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  arrange(desc(CIP))
+```
+
+## Chi Squared test for significant differences in resistance mechanisms
+```{r}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    unlist() %>% unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
+do_chisq_test <- function(df) {
+  df <- df %>%
+    dplyr::slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
+          dplyr::filter(Var2 < Var1) %>% t) %>%
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    dplyr::do(data_frame(gene = .$key,
+                  test = dplyr::first(.$test),
+                  species1 = dplyr::first(.$species),
+                  species2 = dplyr::last(.$species),
+                  data = list(matrix(c(.$Absent,
+                                       .$Present),
+                                     ncol = 2)))) %>%
+    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
+    mutate(p.value = map_dbl(chi_test, "p.value"),
+           dupl = duplicated(test)) %>%
+    dplyr::filter(dupl == FALSE) %>%
+    ungroup() %>%
+    select(gene, species1, species2, p.value) %>%
+    mutate(p.value = round(p.value, 5)) %>%
+    dplyr::filter(p.value <= 0.05)
+  return(df)
+}
+
+total_df <- acquired_report %>%
+  left_join(mut_report, by = "id") %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  select(-nt_change)
+
+mechanism_per_species <- total_df %>%
+  select(id, species, everything()) %>%
+  select(-id) %>%
+  gather(key, value, -species) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  select(key, species, Present, Absent) %>%
+  split(f = .$key)
+
+intrinsic_names <- names(mut_report)
+intrinsic_names <- intrinsic_names[-c(1,2)]
+
+intrinsic_gene_perc <- mut_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, intrinsic_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  select(key, species, Percent) %>%
+  dplyr::rename("gene" = key)
+
+acquired_names <- names(acquired_report)
+acquired_names <- acquired_names[-1]
+
+acquired_genes_perc <- acquired_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, acquired_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  select(key, species, Percent) %>%
+  dplyr::rename("gene" = key)
+
+total_perc <- rbind(intrinsic_gene_perc, acquired_genes_perc)
+
+p_values_total <- suppressWarnings(lapply(mechanism_per_species, function(x) do_chisq_test(x))) %>%
+  bind_rows() %>%
+  left_join(total_perc, by = c("gene", c("species1" = "species"))) %>%
+  dplyr::rename("Percent species 1" = Percent) %>%
+  left_join(total_perc, by = c("gene", c("species2" = "species"))) %>%
+  dplyr::rename("Percent species 2" = Percent) %>%
+  dplyr::select(gene, species1, `Percent species 1`, species2, `Percent species 2`, p.value)
+
+p_values_total
+
+```
+
+## Resistant isolates per antimicrobial per species
+```{r}
+percent_am_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round((Present/Total)*100, 1))
+
+percent_am_res
+```
+
+## Significant differences between the amount of resistant isolates
+```{r}
+cip_nal_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  filter(compound %in% c("CIP","NAL")) %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  select(compound, species, Present, Absent) %>%
+  dplyr::rename("key" = compound) %>%
+  split(f = .$key)
+
+p_values_res <- suppressWarnings(lapply(cip_nal_res, function(x) do_chisq_test(x))) %>%
+  bind_rows() %>%
+  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species1" = "species",
+                                                                       "gene" = "compound")) %>%
+  dplyr::rename("Percent_species1" = Percent) %>%
+  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species2" = "species",
+                                                                       "gene" = "compound")) %>%
+  dplyr::rename("Percent_species2" = Percent,
+                "Compound" = gene) %>%
+  select(Compound, species1, Percent_species1, species2, Percent_species2, p.value)
+
+p_values_res
+
+rest_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  filter(compound %not_in% c("CIP","NAL")) %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  select(compound, species, Present, Absent) %>%
+  dplyr::rename("key" = compound) %>%
+  split(f = .$key)
+
+p_values_res_rest <- suppressWarnings(lapply(rest_res, function(x) do_chisq_test(x))) %>%
+  bind_rows() %>%
+  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species1" = "species",
+                                                                       "gene" = "compound")) %>%
+  dplyr::rename("Percent_species1" = Percent) %>%
+  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species2" = "species",
+                                                                       "gene" = "compound")) %>%
+  dplyr::rename("Percent_species2" = Percent,
+                "Compound" = gene) %>%
+  select(Compound, species1, Percent_species1, species2, Percent_species2, p.value)
+
+p_values_res_rest
 ```
@@
 # Figures
+
+## Sequence types per animal group
+```{r}
+palette4 <- c("Livestock" = "#80b1d3",
+              "Wild" = "#fb8072")
+
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(animal_group = ifelse(species %in% c("Broiler","Pig"), "Livestock", "Wild")) %>%
+  group_by(ST, animal_group) %>%
+  dplyr::count() %>%
+  group_by(ST) %>%
+  mutate(total = dplyr::first(n) + dplyr::last(n)) %>%
+  filter(total > 4) %>%
+  ggplot(aes(factor(ST), n, fill = animal_group))+
+  geom_col(color = "black")+
+  labs(x = "Sequence Types")+
+  scale_fill_manual(values = palette4)+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.45),
+        legend.title = element_blank())
+```
+
 ## Resistance mechanisms per sequence type
@@
   mutate(group_n = sum(n)) %>%
-  filter(group_n > 4)
+  filter(group_n > 3)
 
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
@@
        y = "Genes")+
-  scale_y_discrete(limits = c("gyrA","parC","parE","marA","marR","rpoB","qnr")) +
   theme(axis.text.x = element_text(size = 10),
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,11:17]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
@@
          Percent = round((Present/Total)*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
          type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))
@@
   geom_col(color = "black") +
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                width = 0.5)+
   scale_fill_manual(values = palette2) +
@@
       Total = Present + Absent,
-      Percent = round(Present / Total * 100, 1)
+      Percent = round(Present / Total * 100, 1),
+      lwr = round(get_binCI(Present, Total)[1], 1),
+      upr = round(get_binCI(Present, Total)[2], 1)
     ) %>%
@@
                             levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
-    ggplot(aes(key, Percent, fill = species)) +
+    ggplot(aes(reorder(key, -Percent), Percent, fill = species)) +
     geom_col(color = "black",
              position = position_dodge(1)) +
+    geom_errorbar(
+      aes(ymin = lwr, ymax = upr),
+      width = 0.6,
+      position = position_dodge(1)
+    ) +
     scale_fill_manual(values = palette) +
-    scale_x_discrete(limits = c("gyrA","gyrB","parC","parE","marA","marR","acrB","rpoB")) +
     labs(y = "Percent (%) of isolates") +
@@
     facet_wrap(~ species)
+```
+
+## Percent QRDR mutations per animal species
+```{r}
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+suppressWarnings(clean_mut_data %>%
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  summarise_all(funs(func_paste2)) %>%
+  dplyr::select(-c(id, flag)) %>%
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  filter(id %not_in% c("2016-17-565-1-S",
+                       "2016-02-428-2-S",
+                       "2016-02-486-2-S",
+                       "2016-02-732-2-S",
+                       "2014-01-1741-1-S")) %>%
+  gather(gene, mut, -id) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
+  select(-mut) %>%
+  dplyr::rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  group_by(gene, mut, species) %>%
+  dplyr::count() %>%
+  rowwise() %>%
+  mutate(Total = case_when(species == "Broiler" ~ 87,
+                           species == "Pig" ~ 75,
+                           species == "Red fox" ~ 52,
+                           species == "Wild bird" ~ 66),
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1], 1),
+         upr = round(get_binCI(n, Total)[2], 1)) %>%
+  dplyr::rename("Mutation" = mut)) %>%
+  ggplot(aes(reorder(Mutation, -Percent), Percent, fill = species))+
+  geom_col(color = "black", position = position_dodge(0.9))+
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                position = position_dodge(0.9),
+                width = 0.5)+
+  scale_fill_manual(values = palette)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
+        axis.title.x = element_blank())+
+  facet_wrap(~ gene, scale = "free_x")
+
 ```
@@
 
+## Correlation plot for resistance mechanisms
+```{r, fig.height=8, fig.width=7, warning=FALSE}
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+total_df <- acquired_report %>%
+  left_join(mut_report, by = "id") %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  select(-nt_change) %>%
+  mutate_at(vars(-c(id,species)),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,3:9]) > 0, 1, 0)) %>%
+  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  split(., f = .$species) %>%
+  lapply(., function(x) select(x, -species))
+
+corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
+  bind_rows(.id = "species") %>%
+  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
+           gene2 = gsub("(.*?)_.+", "\\1", var2),
+           var1 = gsub(".+_(.*?)", "\\1", var1),
+           var2 = gsub(".+_(.*?)", "\\1", var2),
+           type1 = case_when(gene1 == "gyrA" ~ 1,
+                          gene1 == "gyrB" ~ 2,
+                          gene1 == "parC" ~ 3,
+                          gene1 == "parE" ~ 4,
+                          gene1 == "marR" ~ 5,
+                          gene1 == "marA" ~ 6,
+                          gene1 == "rpoB" ~ 7,
+                          gene1 == "soxS" ~ 8,
+                          gene1 == "robA" ~ 9,
+                          gene1 == "gadX" ~ 10,
+                          gene1 == "qnr" ~ 11,
+                          gene1 == "qepA4" ~ 12),
+           type2 = case_when(gene2 == "gyrA" ~ 1,
+                          gene2 == "gyrB" ~ 2,
+                          gene2 == "parC" ~ 3,
+                          gene2 == "parE" ~ 4,
+                          gene2 == "marR" ~ 5,
+                          gene2 == "marA" ~ 6,
+                          gene2 == "rpoB" ~ 7,
+                          gene2 == "soxS" ~ 8,
+                          gene2 == "robA" ~ 9,
+                          gene2 == "gadX" ~ 10,
+                          gene2 == "qnr" ~ 11,
+                          gene2 == "qepA4" ~ 12))
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradientn(colors=cols,
+                       na.value="grey95",
+                       limits=c(-1, 1),
+                       guide=guide_colourbar(ticks=T,nbin=10,
+                                             barheight=.5,label=T, 
+                                             barwidth=10))+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 8),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank(),
+        legend.position="bottom",
+        legend.justification="center",
+        legend.direction="horizontal")+
+  coord_fixed()+
+  facet_wrap(~species)
+```
+
+## MIC-plots
+```{r}
+total_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  dplyr::select(-nt_change) %>%
+  mutate_at(vars(-id),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-ref) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(present,CIP) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(present = ifelse(present == "", "None", present),
+         total = 280,
+         percent = round(n/total * 100, 1),
+         lwr = round(get_binCI(n, total)[1], 1),
+         upr = round(get_binCI(n, total)[2], 1)) %>%
+  ungroup() %>%
+  mutate(entries = sapply(strsplit(.$present, ","),
+                          FUN = function(x) {length(x)}))
+
+genes <- c(names(mut_report), names(acquired_report), "qnr")
+genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")]
+
+n_plot <- ggplot(total_report, aes(reorder(present, entries), percent, fill = factor(CIP)))+
+  geom_col(color = "black")+
+  labs(y = "Percent (%) isolates")+
+  scale_fill_viridis(discrete = TRUE)+
+  guides(fill=guide_legend(ncol = 10))+
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        axis.text.y = element_text(size = 8),
+        axis.title.x = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_blank(),
+        legend.position = "top",
+        legend.spacing.x = unit(0.2, "cm"),
+        legend.justification = "center")
+
+dot_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  select(-nt_change) %>%
+  mutate_at(vars(-id),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  select(-c(ref, id)) %>%
+  ungroup() %>%
+  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
+  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
+         present = ifelse(present == "", "None", present)) %>%
+  gather(gene, value, -c(present,CIP)) %>%
+  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
+         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
+  na.omit() %>%
+  mutate(entries = sapply(strsplit(.$present, ","),
+                          FUN = function(x) {length(x)}))
+
+dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
+  geom_point(size = 2)+
+  scale_y_discrete(limits = c("gyrA","parC","parE","gadX","marA","marR","rpoB","soxS","qepA4","qnr"))+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.title = element_blank(),
+        axis.ticks.x = element_blank())
+
+plot_grid(n_plot, dot_plot, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
+```
+
+## Additional resistance levels per animal species
+```{r}
+mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round((Present/Total)*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
+         species = factor(species, levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
+  ggplot(aes(reorder(compound, -Percent), Percent, fill = species))+
+  geom_col(color = "black")+
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                width = 0.6)+
+  guides(fill = FALSE)+
+  scale_fill_manual(values = palette)+
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))+
+  facet_wrap(~species)
+```
+
+## CIP/NAL MIC-values per animal species
+```{r}
+mic_data[,c("id","CIP")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(CIP,species) %>%
+  dplyr::count() %>%
+  ggplot(aes(factor(CIP), n, fill = species))+
+  geom_col(color = "black", position = position_fill())+
+  labs(y = "Percent isolates",
+       x = "CIP MIC")+
+  scale_fill_manual(values = palette)+
+  scale_y_continuous(labels = percent_format())+
+  theme(legend.title = element_blank())
+
+mic_data[,c("id","NAL")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(NAL,species) %>%
+  dplyr::count() %>%
+  ggplot(aes(factor(NAL), n, fill = species))+
+  geom_col(color = "black", position = position_fill())+
+  labs(y = "Percent isolates",
+       x = "NAL MIC")+
+  scale_fill_manual(values = palette)+
+  scale_y_continuous(labels = percent_format())+
+  theme(legend.title = element_blank())
+
+```
+
 ## cgMLST-trees
-```{r, fig.width=24, fig.height=22, message = FALSE, warning = FALSE}
-
-palette3 <- c("1-A" = "#8dd3c7",
+```{r, fig.width=28, fig.height=28, message = FALSE, warning = FALSE}
+
+palette3 <- c("1-A" = "#fc8d59",
               "1-I" = "#80b1d3",
@@
 
-qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
+qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR","gadX","soxS")] %>%
   mutate_at(vars(-id),
@@
                 color = "#4575b4")+
-  ggtitle("A")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
+  geom_tiplab2(aes(label = ST),
+               offset = 0.01,
+               size = 8)+
+  ggtitle("Broiler")+
+  theme(plot.title = element_text(size = 40))
 
@@
                 color = "#74add1")+
-  ggtitle("B")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
+    geom_tiplab2(aes(label = ST),
+               offset = 0.01,
+               size = 8)+
+  ggtitle("Pig")+
+  theme(plot.title = element_text(size = 40))
 
@@
                 color = "#f46d43")+
-  ggtitle("C")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
+    geom_tiplab2(aes(label = ST),
+               offset = 0.01,
+               size = 8)+
+  ggtitle("Red fox")+
+  theme(plot.title = element_text(size = 40))
 
@@
                 color = "#fdae61")+
-  ggtitle("D")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
+    geom_tiplab2(aes(label = ST),
+               offset = 0.01,
+               size = 8)+
+  ggtitle("Wild bird")+
+  theme(plot.title = element_text(size = 40))
 
@@
                                    broiler_data,
-                                   offset = 0.01,
+                                   offset = 0.06,
                                    width = 0.5,
@@
                                    colnames_position = "top",
-                                   font.size = 7) +
+                                   font.size = 8,
+                                   color = "grey90") +
   scale_fill_manual(values = palette3) +
@@
                                pig_data,
-                               offset = 0.01, 
+                               offset = 0.06, 
                                width = 0.5, 
@@
                                colnames_position = "top", 
-                               font.size = 7) +
+                               font.size = 8,
+                               color = "grey90") +
   scale_fill_manual(values = palette3) +
@@
                                fox_data, 
-                               offset = 0.01, 
+                               offset = 0.06, 
                                width = 0.5, 
@@
                                 bird_data, 
-                                offset = 0.01, 
+                                offset = 0.06, 
                                 width = 0.5, 
@@
                                 colnames_position = "top", 
-                                font.size = 7) +
+                                font.size = 8,
+                                color = "grey90") +
   scale_fill_manual(values = palette3) +
@@
 
-total_plot <- cowplot::plot_grid(broiler_annotated_tree, 
-                                 pig_annotated_tree, 
-                                 fox_annotated_tree, 
-                                 bird_annotated_tree, 
-                                 ncol = 2, 
-                                 align = "h")
-total_plot
-
-```
-
-## Another color scheme
-```{r,fig.width=24, fig.height=22, message = FALSE, warning = FALSE}
-
-qr_genes_alt <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
-  mutate(gyrA = ifelse(gyrA == 1, "gyrA_1", gyrA),
-         parC = ifelse(parC == 1, "parC_1", parC),
-         parE = ifelse(parE == 1, "parE_1", parE),
-         rpoB = ifelse(rpoB == 1, "rpoB_1", rpoB),
-         marA = ifelse(marA == 1, "marA_1", marA),
-         marR = ifelse(marR == 1, "marR_1", marR))
-
-ac_genes_alt <- acquired_report %>%
-  mutate_at(vars(-id),
-            funs(as.numeric)) %>%
-  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
-  select(id, qepA4, qnr) %>%
-  mutate(qnr = ifelse(qnr == 1, "qnr_1", qnr),
-         qepA4 = ifelse(qepA4 == 1, "qepA4_1", qepA4))
-
-total_data_alt <- qr_genes_alt %>%
-  left_join(ac_genes_alt, by = "id")
-
-palette_heatmap <- c("gyrA_1" = "#8dd3c7",
-                     "parC_1" = "#ffffb3",
-                     "parE_1" = "#bebada",
-                     "rpoB_1" = "#fb8072",
-                     "marA_1" = "#80b1d3",
-                     "marR_1" = "#fdb462",
-                     "qepA4_1" = "#b3de69",
-                     "qnr_1" = "#fccde5",
-                     "0" = "grey95")
-
-broiler_tree <- ggtree(cgMLST_tree_broiler,
-                       layout = "circular",
-                       color = "#4575b4",
-                       size = 1) %<+% data_broiler +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("A")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-broiler_data <- total_data_alt %>%
-  filter(id %in% broiler_iso$id) %>%
-  column_to_rownames("id")
-
-broiler_annotated_tree2 <- gheatmap(broiler_tree,
-                                   broiler_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-pig_tree <- ggtree(cgMLST_tree_pig,
-                       layout = "circular",
-                       color = "#74add1",
-                       size = 1) %<+% data_pig +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("B")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-pig_data <- total_data_alt %>%
-  filter(id %in% pig_iso$id) %>%
-  column_to_rownames("id")
-
-pig_annotated_tree2 <- gheatmap(pig_tree,
-                                   pig_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-fox_tree <- ggtree(cgMLST_tree_fox,
-                       layout = "circular",
-                       color = "#f46d43",
-                       size = 1) %<+% data_fox +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("C")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-fox_data <- total_data_alt %>%
-  filter(id %in% fox_iso$id) %>%
-  column_to_rownames("id")
-
-fox_annotated_tree2 <- gheatmap(fox_tree,
-                                   fox_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-bird_tree <- ggtree(cgMLST_tree_bird,
-                       layout = "circular",
-                       color = "#fdae61",
-                       size = 1) %<+% data_bird +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("D")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-bird_data <- total_data_alt %>%
-  filter(id %in% bird_iso$id) %>%
-  column_to_rownames("id")
-
-bird_annotated_tree2 <- gheatmap(bird_tree,
-                                   bird_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-total_plot2 <- cowplot::plot_grid(broiler_annotated_tree2, 
-                                 pig_annotated_tree2, 
-                                 fox_annotated_tree2, 
-                                 bird_annotated_tree2, 
-                                 ncol = 2, 
-                                 align = "h")
-total_plot2
-```
+broiler_annotated_tree
+pig_annotated_tree
+fox_annotated_tree
+bird_annotated_tree
+
+```
+
+# Tree-stats
+```{r, fig.width=10, fig.height=9}
+print(paste0("Broiler median distance: ",
+             round(median(cophenetic.phylo(cgMLST_tree_broiler)), 2)))
+
+print(paste0("Pig median distance: ",
+             round(median(cophenetic.phylo(cgMLST_tree_pig)), 2)))
+
+print(paste0("Red fox median distance: ",
+             round(median(cophenetic.phylo(cgMLST_tree_fox)), 2)))
+
+print(paste0("Wild bird median distance: ",
+             round(median(cophenetic.phylo(cgMLST_tree_bird)), 2)))
+
+par(mfrow = c(2,2))
+
+hist(cophenetic.phylo(cgMLST_tree_broiler),
+     breaks = 20,
+     main = "Broiler distance frequency",
+     xlab = "Distance",
+     ylim = c(0,6000))
+
+hist(cophenetic.phylo(cgMLST_tree_pig),
+     breaks = 20,
+     main = "Pig distance frequency",
+     xlab = "Distance",
+     ylim = c(0,6000))
+
+hist(cophenetic.phylo(cgMLST_tree_fox),
+     breaks = 20,
+     main = "Red fox distance frequency",
+     xlab = "Distance",
+     ylim = c(0,6000))
+
+hist(cophenetic.phylo(cgMLST_tree_bird),
+     breaks = 20,
+     main = "Wild bird distance frequency",
+     xlab = "Distance",
+     ylim = c(0,6000))
+
+```

### 2019-02-07 notebooks/article_notebook2.Rmd
minor changes

@@
 
+## Specific mutations in non-QRDR intrinsic genes
+These mutations are the most prominent in each gene where mutations were identified (n > 2).
+The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
+```{r, message=FALSE, warning=FALSE}
+mut_filtered %>%
+  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
+  select(id, gene, mut) %>%
+  mutate(entries = sapply(strsplit(.$mut, ","),
+                          FUN = function(x) {length(x)})) %>%
+  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
+  mutate(value = str_trim(value)) %>%
+  select(id, gene, value) %>%
+  filter(value != 0,
+         is.na(value) == FALSE) %>%
+  group_by(gene, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  filter(n > 2) %>%
+  mutate(percent = case_when(
+    gene == "gadX" ~ 
+      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
+    gene == "marA" ~ 
+      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
+    gene == "marR" ~ 
+      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
+    gene == "rpoB" ~ 
+      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
+    gene == "soxS" ~ 
+      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
+  arrange(gene, desc(percent)) %>%
+  dplyr::rename("Gene" = gene,
+                "Mutation" = value,
+                "Percent" = percent)
+
+```
+
 ## Mechanisms in isolates without QRDR mutations
@@
 ```{r}
-AddCol <- function(df, col_name) {
-  # split rows by delimeters
-  string_to_proc <- df %>% select(!!col_name) %>%
-    unlist() %>% str_split(regex("\\, |\\,")) 
-  # find unique entries
-  unique_strings <- string_to_proc %>%
-    unlist() %>% unique()
-  # construct names of the new columns
-  cols_names <- paste(col_name, unique_strings, sep = "_")
-  # construct 0/1-content columns for each unique entry
-  cols_content <- sapply(function(i) {
-    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
-                             X = string_to_proc)))
-  }, X = seq_along(unique_strings))
-  res <- data.frame(cols_content)
-  names(res) <- cols_names
-  return(res)
-}
-
 do_chisq_test <- function(df) {
@@
 
+## Resistance mechanisms per sequence type
+```{r}
+mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  gather(gene, value, -ST) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  filter(Present != 0) %>%
+  mutate(Total = Present + Absent) %>%
+  filter(Total > 2) %>%
+  rowwise() %>%
+  mutate(Percent = round(Present/Total*100, 1)) %>%
+  ungroup() %>%
+  group_by(ST) %>%
+  mutate(no_of_mechanisms = n())
+```
+
+## Resistance mechanisms per sequence type per animal species
+```{r}
+mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  gather(gene, value, -c(ST, species)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  filter(Present != 0) %>%
+  mutate(Total = Present + Absent) %>%
+  filter(Total > 2) %>%
+  rowwise() %>%
+  mutate(Percent = round(Present/Total*100, 1)) %>%
+  ungroup() %>%
+  group_by(ST, species) %>%
+  mutate(no_of_mechanisms = n()) %>%
+  arrange(ST, species)
+```
+
 ## Significant differences between the amount of resistant isolates
@@
 p_values_res_rest
+```
+
+## Correlation between specific mutations and qnr-genes
+```{r}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    base::unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    base::unlist() %>% base::unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(base::unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor(method = "spearman") %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+col_proc <- c("gyrA","gyrB","parC","parE")
+
+cols_to_add <- sapply(function(i) {AddCol(df = mut_quant, col_name = col_proc[i])},
+                      X = seq_along(col_proc)) %>%
+  bind_cols() %>%
+  select(-c(gyrA_0, parC_0, gyrB_0, parE_0)) %>%
+  mutate_all(funs(as.numeric)) %>%
+  {. ->> tot_mut_report} %>%
+  select(gyrA_S83L, gyrA_D87N, parC_S80I) %>%
+  mutate(gyrA_S83L_D87N = ifelse(gyrA_S83L == 1 & gyrA_D87N == 1, 1, 0),
+         gyrA2_parC1 = ifelse(gyrA_S83L == 1 & gyrA_D87N == 1 & parC_S80I == 1, 1, 0))
+
+test <- mut_quant %>%
+  cbind(tot_mut_report) %>%
+  select(-contains("mut")) %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(-id),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,21:27]) == 0, 0, 1)) %>%
+  select(-id)
+
+test2 <- create_corr_matrix(test)
+
+cor.test(test$gyrA_S83L, test$qnr)
 ```
@@
 ```{r, fig.height=8, fig.width=7, warning=FALSE}
-create_corr_matrix <- function(df) {
-  df <- df %>%
-    as.matrix %>%
-    cor %>%
-    as.data.frame %>%
-    rownames_to_column(var = "var1") %>%
-    gather(var2, value, -var1) %>%
-    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
-           var2 = gsub("(.*?)_R$", "\\1", var2))
-  return(df)
-}
-
 total_df <- acquired_report %>%
@@
   left_join(isolate_data[c("id","species")], by = "id") %>%
-  select(-nt_change) %>%
-  mutate_at(vars(-c(id,species)),
-            funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,3:9]) > 0, 1, 0)) %>%
-  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  select(-c(nt_change, id)) %>%
+  mutate_at(vars(-species),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,2:8]) > 0, 1, 0)) %>%
+  {. ->> total_resistance_report} %>%
   split(., f = .$species) %>%
@@
                           gene1 == "gadX" ~ 10,
-                          gene1 == "qnr" ~ 11,
-                          gene1 == "qepA4" ~ 12),
+                          gene1 == "qnrA1" ~ 11,
+                          gene1 == "qnrB19" ~ 12,
+                          gene1 == "qnrB6" ~ 13,
+                          gene1 == "qnrB60" ~ 14,
+                          gene1 == "qnrS1" ~ 15,
+                          gene1 == "qnrS2" ~ 16,
+                          gene1 == "qnrS4" ~ 17,
+                          gene1 == "qnr" ~ 18,
+                          gene1 == "qepA4" ~ 19),
            type2 = case_when(gene2 == "gyrA" ~ 1,
@@
                           gene2 == "gadX" ~ 10,
-                          gene2 == "qnr" ~ 11,
-                          gene2 == "qepA4" ~ 12))
+                          gene2 == "qnrA1" ~ 11,
+                          gene2 == "qnrB19" ~ 12,
+                          gene2 == "qnrB6" ~ 13,
+                          gene2 == "qnrB60" ~ 14,
+                          gene2 == "qnrS1" ~ 15,
+                          gene2 == "qnrS2" ~ 16,
+                          gene2 == "qnrS4" ~ 17,
+                          gene2 == "qnr" ~ 18,
+                          gene2 == "qepA4" ~ 19))
 
@@
   geom_tile(color = "white")+
-  scale_fill_gradientn(colors=cols,
-                       na.value="grey95",
-                       limits=c(-1, 1),
-                       guide=guide_colourbar(ticks=T,nbin=10,
-                                             barheight=.5,label=T, 
-                                             barwidth=10))+
+  scale_fill_gradient(low = "black",
+                      high = "white",
+                      na.value = "grey95")+
   labs(fill = NULL)+
@@
         axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
-        axis.title = element_blank(),
-        legend.position="bottom",
-        legend.justification="center",
-        legend.direction="horizontal")+
+        axis.title = element_blank())+
   coord_fixed()+
@@
 
-## cgMLST-trees
-```{r, fig.width=28, fig.height=28, message = FALSE, warning = FALSE}
-
+# Dendrograms
+
+## Number of NA's per isolate in cgMLST allele matrix
+The red outliers have been excluded from the cgMLST analysis due to too many missing alleles
+```{r}
+cgMLSTdata <- read.table("../data/cgMLST.txt",
+                         sep = "\t",
+                         colClasses = "factor",
+                         header = T)
+
+cgMLSTdata[cgMLSTdata == "NaN"] <- NA
+
+cgMLSTdata2 <- cgMLSTdata %>%
+  mutate(n_NA = apply(., 1, function(x) sum(is.na(x))),
+         group = 1)
+
+ggplot(cgMLSTdata2, aes(factor(group), n_NA)) +
+  geom_boxplot(outlier.size = 2) +
+  geom_point(data = cgMLSTdata2[cgMLSTdata2$n_NA > 500,], aes(factor(group), n_NA),
+             pch = 21,
+             size = 2,
+             fill = "red") +
+  labs(x = NULL,
+       y = "Number of NA's per isolate") +
+  theme_classic() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank())
+
+```
+
+## cgMLST-tree total
+```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
 palette3 <- c("1-A" = "#fc8d59",
               "1-I" = "#80b1d3",
-              "0" = "grey95")
+              "0" = "grey95",
+              "0.03" = "#440154FF",
+              "0.06" = "#482878FF",
+              "0.12" = "#3E4A89FF",
+              "0.25" = "#31688EFF",
+              "0.5" = "#26828EFF",
+              "1" = "#1F9E89FF",
+              "2" = "#35B779FF",
+              "4" = "#6DCD59FF",
+              "8" = "#B4DE2CFF",
+              "16" = "#FDE725FF")
 
@@
 
+total_data <- qr_genes %>%
+  left_join(ac_genes, by = "id")
+
+total_data_id <- total_data %>%
+  column_to_rownames("id")
+
+metadata_total <- mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id", "species")], by = "id")
+
+cgMLSTdata_complete <- cgMLSTdata2 %>%
+  filter(n_NA < 500) %>%
+  select(-n_NA) %>%
+  column_to_rownames("Genome")
+
+total_tree <- as.phylo(hclust(daisy(cgMLSTdata_complete, metric = "gower"), method = "average"))
+
+total_tree_annotated <- ggtree(total_tree,
+                               layout = "circular",
+                               color = "#808080",
+                               size = 0.5) %<+% metadata_total +
+                               geom_tippoint(aes(color = species),
+                               size = 3) +
+                               geom_tiplab2(aes(label = ST),
+                               offset = 0.01,
+                               size = 4) +
+  scale_color_manual(values = palette)
+
+total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)
+
+gheatmap(total_tree_opened,
+         total_data_id,
+         offset = 0.04,
+         width = 0.3,
+         colnames_offset_y = 2.5,
+         colnames_position = "top",
+         font.size = 5) +
+  scale_fill_manual(values = palette3,
+                    labels = c("Gene/Mutation absent",
+                               "Acquired gene present",
+                               "Intrinsic gene mutated")) +
+  theme(legend.position = c(0.9,0.9),
+        legend.text = element_text(size = 20))
+
+```
+
+## Roary tree
+```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
+roary_tree <- read.tree("../data/Roary/accessory_binary_genes.fa.newick")
+
+tree_refs <- roary_tree$tip.label
+
+id_data2 <- id_data %>%
+  mutate(test = ref %in% tree_refs) %>%
+  filter(test == TRUE)
+
+id_df <- left_join(data.frame(ref = tree_refs), id_data2, by = "ref") %>%
+  mutate(id = ifelse(is.na(id) == TRUE, ref, id))
+
+roary_tree$tip.label <- id_df$id
+
+total_data2 <- total_data %>%
+  filter(id %in% id_df$id) %>%
+  column_to_rownames("id")
+
+metadata_total2 <- metadata_total %>%
+  filter(id %in% id_df$id)
+
+roary_tree_colored <- ggtree(roary_tree,
+       layout = "circular",
+       color = "#808080",
+       size = 1) %<+% metadata_total2 +
+  geom_tiplab2(aes(label = ST, color = species),
+               offset = 0.01,
+               size = 6,
+               align = TRUE) +
+  scale_color_manual(values = palette)
+
+total_tree_opened2 <- rotate_tree(open_tree(roary_tree_colored, 8), 93)
+
+gheatmap(total_tree_opened2,
+         total_data2,
+         offset = 0.065,
+         width = 0.3,
+         colnames_offset_y = 1.9,
+         colnames_position = "top",
+         font.size = 5) +
+  scale_fill_manual(values = palette3,
+                    labels = c("Gene/Mutation absent",
+                               "Acquired gene present",
+                               "Intrinsic gene mutated")) +
+  theme(legend.position = c(0.9,0.9),
+        legend.text = element_text(size = 20))
+
+```
+
+## cgMLST-tree per animal species
+```{r, fig.width=28, fig.height=28, message = FALSE, warning = FALSE}
+
+palette3 <- c("1-A" = "#fc8d59",
+              "1-I" = "#80b1d3",
+              "0" = "grey95")
+
+qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR","gadX","soxS")] %>%
+  mutate_at(vars(-id),
+            funs(ifelse(. == "1", "1-I", "0")))
+
+ac_genes <- acquired_report %>%
+  mutate_at(vars(-id),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
+  select(id, qepA4, qnr) %>%
+  mutate_at(vars(-id),
+            funs(ifelse(. == "1", "1-A", "0")))
+
 cgMLSTdata <- read.table("../data/cgMLST.txt",
                          sep = "\t",
-                         row.names=1,
                          colClasses = "factor",
                          header = T)
+
+cgMLSTdata[cgMLSTdata == "NaN"] <- NA
+
+cgMLSTdata2 <- cgMLSTdata %>%
+  mutate(n_NA = apply(., 1, function(x) sum(is.na(x)))) %>%
+  filter(n_NA < 500) %>%
+  select(-n_NA) %>%
+  column_to_rownames("Genome")
 
@@
 
-data_broiler <- cgMLSTdata %>%
+rawdata_broiler <- cgMLSTdata2 %>%
   rownames_to_column("id") %>%
@@
   select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_pig <- cgMLSTdata %>%
+  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
+  {. ->> data_broiler } %>%
+  select(-ST)
+
+rawdata_pig <- cgMLSTdata2 %>%
   rownames_to_column("id") %>%
@@
   select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_fox <- cgMLSTdata %>%
+  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
+  {. ->> data_pig } %>%
+  select(-ST)
+
+rawdata_fox <- cgMLSTdata2 %>%
   rownames_to_column("id") %>%
@@
   select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_bird <- cgMLSTdata %>%
+  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
+  {. ->> data_fox } %>%
+  select(-ST)
+
+rawdata_bird <- cgMLSTdata2 %>%
   rownames_to_column("id") %>%
@@
   select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
+  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
+  {. ->> data_bird } %>%
+  select(-ST)
+
+write.table(rawdata_broiler, "../data/cgMLST_broiler.txt", sep = "\t", row.names = FALSE)
+write.table(rawdata_pig, "../data/cgMLST_pig.txt", sep = "\t", row.names = FALSE)
+write.table(rawdata_fox, "../data/cgMLST_fox.txt", sep = "\t", row.names = FALSE)
+write.table(rawdata_bird, "../data/cgMLST_bird.txt", sep = "\t", row.names = FALSE)
 
@@
 
+cgMLSTdata_broiler[cgMLSTdata_broiler == "NaN"] <- NA
+
 cgMLST_tree_broiler <- as.phylo(hclust(daisy(cgMLSTdata_broiler,
-                                             metric="gower")))
+                                             metric="gower"), method = "average"))
 
@@
 
+cgMLSTdata_pig[cgMLSTdata_pig == "NaN"] <- NA
+
 cgMLST_tree_pig <- as.phylo(hclust(daisy(cgMLSTdata_pig,
-                                         metric="gower")))
+                                         metric="gower"), method = "average"))
 
@@
 
+cgMLSTdata_fox[cgMLSTdata_fox == "NaN"] <- NA
+
 cgMLST_tree_fox <- as.phylo(hclust(daisy(cgMLSTdata_fox,
-                                         metric="gower")))
+                                         metric="gower"), method = "average"))
 
@@
 
+cgMLSTdata_bird[cgMLSTdata_bird == "NaN"] <- NA
+
 cgMLST_tree_bird <- as.phylo(hclust(daisy(cgMLSTdata_bird,
-                                          metric="gower")))
+                                          metric="gower"), method = "average"))
 
@@
 ```{r, fig.width=10, fig.height=9}
+dist_broiler <- cophenetic.phylo(cgMLST_tree_broiler)
+dist_broiler[dist_broiler > 0.95] <- NA
+
+dist_pig <- cophenetic.phylo(cgMLST_tree_pig)
+dist_pig[dist_pig > 0.95] <- NA
+
+dist_fox <- cophenetic.phylo(cgMLST_tree_fox)
+dist_fox[dist_fox > 0.95] <- NA
+
+dist_bird <- cophenetic.phylo(cgMLST_tree_bird)
+dist_bird[dist_bird > 0.95] <- NA
+
 print(paste0("Broiler median distance: ",
-             round(median(cophenetic.phylo(cgMLST_tree_broiler)), 2)))
+             round(median(dist_broiler, na.rm = TRUE), 2)))
 
 print(paste0("Pig median distance: ",
-             round(median(cophenetic.phylo(cgMLST_tree_pig)), 2)))
+             round(median(dist_pig, na.rm = TRUE), 2)))
 
 print(paste0("Red fox median distance: ",
-             round(median(cophenetic.phylo(cgMLST_tree_fox)), 2)))
+             round(median(dist_fox, na.rm = TRUE), 2)))
 
 print(paste0("Wild bird median distance: ",
-             round(median(cophenetic.phylo(cgMLST_tree_bird)), 2)))
+             round(median(dist_bird, na.rm = TRUE), 2)))
 
@@
 
-hist(cophenetic.phylo(cgMLST_tree_broiler),
+x1 <- hist(dist_broiler,
      breaks = 20,
+     plot = FALSE)
+
+x1$density = x1$counts/sum(x1$counts)*100
+
+plot(x1,
+     freq = FALSE,
      main = "Broiler distance frequency",
      xlab = "Distance",
-     ylim = c(0,6000))
-
-hist(cophenetic.phylo(cgMLST_tree_pig),
+     ylim = c(0, 50))
+
+x2 <- hist(dist_pig,
      breaks = 20,
+     plot = FALSE)
+
+x2$density = x2$counts/sum(x2$counts)*100
+
+plot(x2,
+     freq = FALSE,
      main = "Pig distance frequency",
      xlab = "Distance",
-     ylim = c(0,6000))
-
-hist(cophenetic.phylo(cgMLST_tree_fox),
+     ylim = c(0, 50))
+
+x3 <- hist(dist_fox,
      breaks = 20,
+     plot = FALSE)
+
+x3$density = x3$counts/sum(x3$counts)*100
+
+plot(x3,
+     freq = FALSE,
      main = "Red fox distance frequency",
      xlab = "Distance",
-     ylim = c(0,6000))
-
-hist(cophenetic.phylo(cgMLST_tree_bird),
+     ylim = c(0, 50))
+
+x4 <- hist(dist_bird,
      breaks = 20,
+     plot = FALSE)
+
+x4$density = x4$counts/sum(x4$counts)*100
+
+plot(x4,
+     freq = FALSE,
      main = "Wild bird distance frequency",
      xlab = "Distance",
-     ylim = c(0,6000))
-
-```
+     ylim = c(0, 50))
+
+```

### 2019-02-07 notebooks/assembly_metrics.Rmd
added boxplots on summary metrics to verify similarity between isolates from animal species and sequencer

@@
 ```{r, include = FALSE}
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, magrittr)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, scales)
 
@@
   df <- df %>%
-    rename(ctg_0_bp = X..contigs.....0.bp.,
+    dplyr::rename(ctg_0_bp = X..contigs.....0.bp.,
            ctg_1000_bp = X..contigs.....1000.bp.,
@@
 ```
+
+# Summarised assembly stats
+## Summary stats per animal species
+```{r}
+p1 <- ggplot(metrics2, aes(species, N50)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p2 <- ggplot(metrics2, aes(species, no_of_contigs)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "No. of contigs") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p3 <- ggplot(metrics2, aes(species, largest_contig)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "Largest contig") +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p4 <- ggplot(metrics2, aes(species, as.numeric(no_of_misassemblies))) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "No. of misassemblies") +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+plot_grid(p1, p2, p3, p4, ncol = 2, nrow = 2, align = "v")
+```
+
+## Summary stats per sequencing method
+```{r}
+p5 <- ggplot(metrics2, aes(sequencer, N50)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p6 <- ggplot(metrics2, aes(sequencer, no_of_contigs)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "No. of contigs") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p7 <- ggplot(metrics2, aes(sequencer, largest_contig)) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "Largest contig") +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+p8 <- ggplot(metrics2, aes(sequencer, as.numeric(no_of_misassemblies))) +
+  stat_boxplot(geom = "errorbar", width = 0.4) +
+  geom_boxplot() +
+  labs(y = "No. of misassemblies") +
+  scale_y_continuous(labels = comma) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(size = 8))
+
+plot_grid(p5, p6, p7, p8, ncol = 2, nrow = 2, align = "v")
+```

### 2019-02-07 notebooks/distance_calc_test.Rmd
notebook on distance calculation testing for dendrograms

---
title: "Distance calculation testing"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r}
pacman::p_load(dplyr, ggtree, ape, tibble, cluster)
```

```{r}
testdata_tree <- read.delim("../data/test_data_cgMLST_funksjon.txt", sep = "\t", header = TRUE, stringsAsFactors = TRUE) %>%
  select(-X) %>%
  column_to_rownames("ref")

print("Raw data without NA's:")
testdata_tree

test_tree <- as.phylo(hclust(daisy(testdata_tree, metric="gower")))

print("Daisy function calculation on raw data without NAs:")
daisy(testdata_tree, metric = "gower")

print("Distances calculated on the raw data without NA's with cophenetic.phylo:")
cophenetic.phylo(test_tree)

ggtree(test_tree)+
  geom_tiplab()

testdata_tree_NA1 <- testdata_tree %>%
  mutate(g4 = ifelse(g1 == 6, NA, g4))

row.names(testdata_tree_NA1) <- c("A","B","C","D","E")

print("Raw data with NA's:")
testdata_tree_NA1

test_tree2 <- as.phylo(hclust(daisy(testdata_tree_NA1, metric="gower")))

print("Daisy function calculation on raw data with one NA:")
daisy(testdata_tree_NA1, metric = "gower")

print("Distances calculated on the raw data with one NA with cophenetic.phylo:")
cophenetic.phylo(test_tree2)

ggtree(test_tree2)+
  geom_tiplab()

```

# dist vs. daisy
```{r}
print("Dataset without NA's:")

x <-
  as.data.frame(matrix(
    c(
      "5",
      "8",
      "13",
      "10",
      "15",
      "6",
      "8",
      "13",
      "10",
      "15",
      "7",
      "9",
      "13",
      "10",
      "15",
      "4",
      "12",
      "16",
      "10",
      "15",
      "4",
      "12",
      "20",
      "10",
      "15"
    ),
    ncol = 5,
    byrow = T
  ))

x

test1 <- as.phylo(hclust(dist(x, method = "manhattan")))
test2 <- as.phylo(hclust(daisy(x, metric = "gower")))

print("Test with dist:")
cophenetic.phylo(test1)
ggtree(test1) +
  geom_tiplab()

print("Test with daisy:")
cophenetic.phylo(test2)
ggtree(test2) +
  geom_tiplab()

print("Dataset with NA's:")
# Med NA'er
x_na <-
  as.data.frame(matrix(
    c(
      "5",
      "8",
      "13",
      "10",
      "15",
      "6",
      "8",
      NA,
      "10",
      "15",
      "7",
      "9",
      "13",
      "10",
      "15",
      "4",
      "12",
      "16",
      "10",
      "15",
      "4",
      "12",
      "20",
      "10",
      "15"
    ),
    ncol = 5,
    byrow = T
  ))

x_na

test1_na <- as.phylo(hclust(dist(x_na)))
test2_na <- as.phylo(hclust(daisy(x_na, metric = "gower")))

print("Test with dist:")
cophenetic.phylo(test1_na)
ggtree(test1_na) +
  geom_tiplab()

print("Test with daisy:")
cophenetic.phylo(test2_na)
ggtree(test2_na) +
  geom_tiplab()

```

### 2019-03-15 notebooks/article_notebook2.Rmd
removed clustering analysis

--- 2019-01-18 notebooks/article_notebook2.Rmd
@@
 # Libraries
+# WARNING: This will install packages on your system if not present in library path!
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, ape, cluster, tibble, Biostrings, seqinr, stringr, scales, reprex)
 ```
@@
 ```{r Import data, include = FALSE}
+# Vector of excluded samples
+ex_samples <- c(
+  "2016-17-565-1-S",
+  "2016-02-428-2-S",
+  "2016-02-486-2-S",
+  "2016-02-732-2-S",
+  "2014-01-1741-1-S"
+  )
+
 # Collapses data frame to unique lines while ignoring NA's
@@
                            stringsAsFactors = F) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+  filter(id %not_in% ex_samples)
 
@@
 
-no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
+no_of_reads <- as.data.frame(rbind(as.matrix(two_reads),
+                                   as.matrix(four_reads))) %>%
   left_join(id_data, by = "ref") %>%
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+  filter(id %not_in% ex_samples)
 
@@
                            stringsAsFactors = F) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-mlst_data <- read.table("../data/mlst_results.txt", sep = "\t", header = TRUE, stringsAsFactors = F) %>%
+  filter(id %not_in% ex_samples)
+
+mlst_data <- read.table("../data/mlst_results.txt",
+                        sep = "\t",
+                        header = TRUE,
+                        stringsAsFactors = F) %>%
   filter(id %not_in% "2014-01-1741-1-S")
@@
 
-```{r ARIBA analysis, include = FALSE}
+# ARIBA reports analyses
+This script filters results from ARIBA reports for genes of interest
+```{r ARIBA analysis}
+# Report directories
 megares_report_loc <- "../data/ariba/megares_results"
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","oqx","qep","mfp","mcb")
-mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")
-
-# adjust parameters for filtering
-if (length(ac_genes) == 1) {
-  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
-    ac_genes <- "ALL"
-    } else {
-      ac_genes <- ac_genes
-    }
-}
-
-if (length(mut_genes) == 1) {  
-  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
-    mut_genes <- "ALL"
-    } else {
-      mut_genes <- mut_genes
-    }
-}
-# -------------------------------------- Libraries
-
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
-               forcats,purrr,stringr,kableExtra,
-               knitr,IRdisplay,reprex,svglite)
-
-# -------------------------------------- Functions
-
-func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")
+
+# Genes of interest. These genes are partially matched to 
+# the generated gene names from the "fix_gene_names" function
+ac_genes <- c("qnr","oqx","qep","aac6")
+mut_genes <- c("gyr","par","marR","marA",
+               "sox","rpoB","rob","gad")
+
+# -------------------------- Functions ----------------------------
+
+# Paste unique cell values together, separate with ","
+func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
+                                 collapse = ",")
 
 # Confidence interval function
-get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
-                                                c("lwr", "upr")))
+get_binCI <- function(x, n) as.numeric(
+  setNames(
+    binom.test(x,n)$conf.int*100, 
+    c("lwr", "upr")
+    )
+  )
 
@@
 file_names <- function(filepath) {
-  files <- list.files(path = filepath, pattern = "amr_report.tsv")
+  files <- list.files(path = filepath, 
+                      pattern = "amr_report.tsv")
   return(files)
@@
 
-# Import ariba data from report.tsv from chosen database used in ariba
+# Import ariba data from report.tsv identified with file_names
 get_ariba_data <- function(filepath) {
+  # Identify file names in filepath
   files <- file_names(filepath)
 
+  # Import data
   data_list <- lapply(files,
@@
 
+  # Set file names in list
   names(data_list) <- files
-  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
+
+  # Bind to data frame and convert columns to character
+  data <- bind_rows(
+    lapply(
+      data_list,
+      function(x) map(x, as.character)
+      ),
+    .id = "ref")
+
   return(data)
@@
 
-# Corrects the gene names found in the "cluster" column
+# Corrects gene names 
 fix_gene_names <- function(df) {
+  # Identify unique entries in ref_name column
   genes <- unique(df$ref_name)
+
+  # Strip away excess characters
   new_names <- gsub("^(.*?)\\..*", "\\1", genes)
@@
 
+  # make all genes lower case, with last letter 
+  # upper case if four characters long
   gene_names <- c()
+
   for (i in new_names) {
-    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
+    p <- paste(
+      tolower(
+        substring(i, 1,3)
+        ),
+      substring(i, 4),
+      sep = "",
+      collapse = " ")
     gene_names <- c(gene_names,p)
   }
+
+  # Create data frame for matching names below
   df2 <- data.frame(genes,gene_names) %>%
@@
 
+  # Left join to match with ref_name column
+  # and strip ref column for file suffix
   df <- df %>%
@@
 
-# Function that handles megares data and returns a data frame with information on whether a gene
-# is mutated or not. Includes control for QRDR in gyrA.
+# Function that handles megares data and returns a data 
+# frame with information on whether a gene is mutated 
+# or not
 create_mut_table <- function(df) {
@@
 
-# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
+# Corrects for mutations in QRDR for gyrA, gyrB, parC and parE genes.
 # The function returns columns of 1/0 values for whether or not the 
@@
 # parE: AA 366 - 523
-# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
+# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
 fix_gyr_par_results <- function(df) {
   df <- df %>%
-    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
+    # control for mutation within QRDR for gyrA
+    mutate(gyrA_result = mut %>% 
              str_extract_all("\\d+") %>% # from reprex package
              map(as.integer) %>% # converts all to integer
-             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
-           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
+             # returns TRUE/FALSE whether value is within range or not
+             map_lgl(~ any(.x >= 67L & .x <= 106L)), 
+           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result),
            gyrA_result = as.integer(gyrA_result),
-           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
+           # control for mutation within QRDR for gyrB
+           gyrB_result = mut %>%
              str_extract_all("\\d+") %>%
@@
            gyrB_result = as.integer(gyrB_result),
-           parC_result = mut %>% # control for mutation within QRDR for parC
+           # control for mutation within QRDR for parC
+           parC_result = mut %>% 
              str_extract_all("\\d+") %>% 
@@
            parC_result = as.integer(parC_result),
-           parE_result = mut %>% # control for mutation within QRDR for parE
+           # control for mutation within QRDR for parE
+           parE_result = mut %>% 
              str_extract_all("\\d+") %>% 
@@
            parE_result = as.integer(parE_result)) %>%
+    # mergeresults from all four genes into one column
     mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
@@
                                       gene == "parE" ~ parE_result)) %>%
-    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
+    # merge results from other genes with the four genes
+    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"),
+                                 result_gyr_par, result)) %>%
     select(-c(gyrA_result,
@@
     mutate(id = 1:n()) %>%
+    # Filter flags
     filter(flag %in% flag_selection) %>%
@@
 
-# Function that returns a filtered dataframe based on the string "genes"
+# Function that returns a filtered dataframe based 
+# on the vector "mut_genes"
 filter_mut_table <- function(df) {
+  # Identify unique genes in column
   report_genes <- unique(df$gene)
   grep_genes <- c()
+
+  # match names from name vector to column entries
   for (gene in mut_genes) {
@@
   }
+
+  # Filter data frame based on values in grep_genes
   df <- df %>%
@@
 
-# Function that returns a filtered dataframe based on the string "acquired_genes"
+# Function that returns a filtered dataframe based
+# on the string "ac_genes"
 filter_acquired_table <- function(df) {
+  # Identify unique genes in column
   report_genes <- unique(df$gene)
   grep_genes <- c()
+
+  # match names from vector to column entries
   for (gene in ac_genes) {
@@
   df <- df %>%
+    # Count Presence/Absence of each gene
     group_by(gene, result_total) %>%
@@
     ungroup() %>%
-    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
+    # Convert binary to "Present" and "Absent"
+    mutate(result_total = if_else(result_total == 1, 
+                                  "Present", "Absent")) %>%
+    # Spread to separate Present and Absent columns
     spread(result_total, n, fill = 0) %>%
     rowwise() %>%
+    # Calculate percentages
     mutate(Total = Present + Absent,
@@
   df1 <- df %>%
+    # Filter out all other genes
     filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
-    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
+    select(-nt_change) %>%
+    # Count how many mutations in each gene per row
+    mutate(entries = sapply(strsplit(.$mut, ","),
+                            FUN = function(x) {length(x)})) %>%
+    # Separate each mutation to own column
     separate(mut, into = as.character(c(1:max(.$entries)))) %>%
     select(-entries) %>%
+    # Gather all mutations together into one column,
+    # to get one row per mutation
     gather(id, mut, -c(ref, gene, result, type))
 
+  # Correct for QRDR mutations in the four genes
   df2 <- fix_gyr_par_results(df1)
 
+  # Generate final table
   df3 <- df2 %>%
+    # Filter out mutations outide QRDR
     mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
     filter(test == 1) %>%
+    # Create own column for each gene
     spread(gene, mut) %>%
     group_by(ref) %>%
+    # Summarise all columns based on unique values
     summarise_all(funs(func_paste)) %>%
+    # Remove "0" values for correct counts below
     mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
               .funs = function(x) ifelse(x == "0", "", x)) %>%
-    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
-           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
-           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
-           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
+    # Count mutations in each column entry
+    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","),
+                             FUN = function(x) {length(x)}),
+           mut_gyrB = sapply(strsplit(.$gyrB, ","),
+                             FUN = function(x) {length(x)}),
+           mut_parC = sapply(strsplit(.$parC, ","),
+                             FUN = function(x) {length(x)}),
+           mut_parE = sapply(strsplit(.$parE, ","),
+                             FUN = function(x) {length(x)})) %>%
     select(-c(type, id, result_total, test)) %>%
+    # Reinsert "0" values 
     mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
@@
 
-# -------------------------------------- Analysis
+# -------------------------- Analysis ---------------------------
 # Import data
@@
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+  # Filters out unwanted samples
+  filter(id %not_in% ex_samples)
 
@@
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-# Wrangle
+  filter(id %not_in% ex_samples)
+
 mut_table <- create_mut_table(clean_mut_data)
+
 mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
@@
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+  filter(id %not_in% ex_samples)
 
 mut_filtered <- filter_mut_table(mut_table_fixed)
+
 acquired_table <- create_acquired_table(clean_acquired_data) %>%
@@
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+  filter(id %not_in% ex_samples)
 
@@
 
-# Stats
-if (length(mut_genes) == 1) {
-  if (mut_genes == "ALL") {
-    mut_stats <- calc_stats(mut_table_fixed)
-    mut_report <- create_mut_report(mut_table_fixed)
-    }
-} else {
 mut_stats <- calc_stats(mut_filtered)
 mut_report <- create_mut_report(mut_filtered)
-}
-
-if (length(ac_genes) == 1) {
-  if (ac_genes == "ALL") {
-    acquired_stats <- calc_stats(acquired_table)
-    acquired_report <- create_acquired_report(acquired_table)
-  }
-} else {
+
 acquired_stats <- calc_stats(acquired_filtered)
 acquired_report <- create_acquired_report(acquired_filtered)
-}
 
 mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
+  # Correct for erroneous entry from ARIBA
+  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
   left_join(id_data, by = "ref") %>%
   select(id, everything(), -ref) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-# ---------------------------------- Tables
+  filter(id %not_in% ex_samples)
 
@@
 
-# Double-check ARIBA results with annotated sequences from Prokka
+# Tables
+## Isolate data
+This table presents the isolates and their accompanying metadata.
+```{r}
+isolate_data_complete <- isolate_data %>%
+  # combine with read data
+  left_join(no_of_reads, by = "id")
+
+isolate_data_complete
+```
+
+## Isolates per species
+This table presents how many isolates were included per species.
+```{r}
+per_species <- isolate_data %>%
+  group_by(species) %>%
+  dplyr::count()
+
+per_species
+```
+
+## Sequence types in total
+```{r}
+tot_mlst <- mlst_data[,c("id","ST")] %>%
+  group_by(ST) %>%
+  dplyr::count() %>%
+  arrange(desc(n))
+
+print(paste0("Total amount of different sequence types: ", nrow(tot_mlst)))
+
+tot_mlst
+```
+
+## Sequence types per animal species
+This table presents all the different sequence types identified in each animal species, arranged in a descending order.
+```{r}
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(ST, species) %>%
+  dplyr::count() %>%
+  arrange(desc(n))
+```
+
+## Most abundant sequence types per animal species
+This table presents the most abundant sequence types identified in each animal species (n > 3).
+```{r}
+# Identify which STs have more than three observations
+mlst_n <- mlst_data[,c("id","ST")] %>%
+  group_by(ST) %>%
+  dplyr::count() %>%
+  # filter out all STs with less than four observations
+  filter(n > 3)
+
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  spread(ST, species) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-id) %>%
+  gather(ST, species) %>%
+  # filter data frame on the STs identified above
+  filter(ST %in% mlst_n$ST)
+```
+
+## Number of sequence types per animal species
+This table presents the amount of different sequence types that were identified in each animal species in total.
+```{r}
+mlst_data[,c("id","ST")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(species) %>%
+  # count how many STs per group
+  mutate(n = length(unique(ST))) %>%
+  select(species,n) %>%
+  summarise_all(funs(func_paste))
+```
+
+## Percent isolates with PMQR
+This table presents the percentage of isolates with identified PMQR determinants in total.
+```{r}
+acquired_report %>%
+  select(-id) %>%
+  mutate_all(funs(as.numeric(.))) %>%
+  # presence/absence of qnr genes in total
+  mutate(total = rowSums(.),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  # count groups
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1))
+```
+
+## Percent isolates with qnr family genes
+```{r}
+acquired_report %>%
+  select(-c(qepA4, id)) %>%
+  mutate_all(funs(as.numeric)) %>%
+  dplyr::rename("qnrA" = qnrA1) %>%
+  # identify presence/absence for family groups 
+  # from the gene subgroups
+  mutate(qnrB = ifelse(rowSums(.[,2:4]) >= 1, 1, 0),
+         qnrS = ifelse(rowSums(.[,5:7]) >= 1, 1, 0)) %>%
+  select(qnrA, qnrB, qnrS) %>%
+  gather(gene, value) %>%
+  group_by(gene, value) %>%
+  # count groups
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 0, "Absent", "Present")) %>%
+  spread(value, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1))
+```
+
+## Percent isolates with specific PMQR genes
+```{r}
+acquired_report %>%
+  select(-id) %>%
+  # transpose table
+  t() %>%
+  as.data.frame(.) %>%
+  mutate(Gene = rownames(.)) %>%
+  mutate_at(vars(-Gene),
+            funs(as.numeric(as.character(.)))) %>%
+  mutate(Present = rowSums(.[,-281])) %>%
+  select(c(Gene, Present)) %>%
+  rowwise() %>%
+  mutate(Total = 280,
+         Percent = round(Present/Total * 100,1))
+```
+
+## qnr-genes per animal species
+This table presents the specific qnr-genes identified in each animal species.
+```{r}
+qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")
+
+acquired_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, qnr_genes) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  arrange(key, desc(Percent))
+```
+
+## Percent isolates with mutations in QRDR
+This table presents the percent of isolates with mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
+```{r}
+mut_report %>%
+  select(gyrA, gyrB, parC, parE) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
+            .funs = funs(as.numeric)) %>%
+  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
+  select(n, type) %>%
+  group_by(type) %>%
+  summarise_all(funs(sum)) %>%
+  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
+  spread(type, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1)) %>%
+  select(-Total)
+```
+
+## Percent mutations in each intrinsic gene
+This table presents the percent mutations in all intrinsic _E. coli_ genes included in this study.
+```{r}
+mut_report %>%
+  select(-nt_change) %>%
+  gather(gene, value, -id) %>%
+  group_by(gene, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1)) %>%
+  arrange(desc(Percent))
+```
+
+## Specific mutations in QRDR
+This table presents the specific mutations identified in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
+```{r, warning=FALSE, message=FALSE}
+clean_mut_data %>%
+  # correct non-mutated entries
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
+                                 "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  # filter quality flags and genes
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  # create columns for each gene
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  # summarise for each sample
+  summarise_all(funs(func_paste2)) %>%
+  select(-c(id, flag)) %>%
+  # set correct sample id's
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  # filter out excluded samples
+  filter(id %not_in% ex_samples) %>%
+  gather(gene, mut, -id) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
+                      0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  # create one row per mutation
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  # correct erroneous result from ARIBA
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
+                            "S80I", aa_change)) %>%
+  select(-mut) %>%
+  dplyr::rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  # fix results for the four genes
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  group_by(gene, mut, nt_change) %>%
+  # count groups
+  dplyr::count() %>%
+  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
+  rowwise() %>%
+  mutate(Total = 280,
+         Percent = round(n/Total * 100,1)) %>%
+  dplyr::rename("Mutation" = mut) %>%
+  arrange(gene, desc(Percent))
+```
+
+## Specific QRDR mutations per animal species
+```{r, warning=FALSE, message=FALSE}
+clean_mut_data %>%
+  # correct non-mutated entries
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
+                                 "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  # filter quality flags and genes
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  # create columns for each gene
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  summarise_all(funs(func_paste2)) %>%
+  dplyr::select(-c(id, flag)) %>%
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  # filter out excluded samples
+  filter(id %not_in% ex_samples) %>%
+  gather(gene, mut, -id) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
+                      0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  # correct errouneous result from ARIBA
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
+                            "S80I", aa_change)) %>%
+  select(-mut) %>%
+  dplyr::rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  # fix results for the four genes
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  group_by(gene, mut, species) %>%
+  dplyr::count() %>%
+  rowwise() %>%
+  # set total amount of isolates per animal species
+  mutate(Total = case_when(species == "Broiler" ~ 87,
+                           species == "Pig" ~ 75,
+                           species == "Red fox" ~ 52,
+                           species == "Wild bird" ~ 66),
+         Percent = round(n/Total * 100,1)) %>%
+  dplyr::rename("Mutation" = mut) %>%
+  arrange(gene, desc(Percent))
+```
+
+## Mutation combinations in QRDR
+This table presents all the observed combinations of mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
+```{r}
+mut_quant %>%
+  group_by(gyrA, gyrB, parC, parE) %>%
+  # count combinations across columns
+  dplyr::count() %>%
+  rowwise() %>%
+  mutate(Total = 280,
+         Percent = round(n/Total * 100,1)) %>%
+  arrange(desc(Percent))
+```
+
+## Mutations in intrinsic genes per animal species
+This table presents the percentage of isolates with mutations for each intrinsic gene.
+```{r}
+# Identify gene names
+gene_names <- names(mut_report)
+gene_names <- gene_names[-c(1,2)]
+
+mut_report %>%
+  left_join(isolate_data, by = "id") %>%
+  # use gene names to specify which columns to gather
+  gather(key, value, gene_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  dplyr::rename("Gene" = key,
+                "Species" = species) %>%
+  dplyr::select(Gene, Species, Present, Total, Percent) %>%
+  arrange(Gene, desc(Percent))
+```
+
+## Specific mutations in non-QRDR intrinsic genes
+These mutations are the most prominent in each gene where mutations were identified (n > 2).
+The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
+```{r, message=FALSE, warning=FALSE}
+mut_filtered %>%
+  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
+  select(id, gene, mut) %>%
+  mutate(entries = sapply(strsplit(.$mut, ","),
+                          FUN = function(x) {length(x)})) %>%
+  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
+  mutate(value = str_trim(value)) %>%
+  select(id, gene, value) %>%
+  filter(value != 0,
+         is.na(value) == FALSE) %>%
+  group_by(gene, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  filter(n > 2) %>%
+  mutate(percent = case_when(
+    gene == "gadX" ~ 
+      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
+    gene == "marA" ~ 
+      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
+    gene == "marR" ~ 
+      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
+    gene == "rpoB" ~ 
+      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
+    gene == "soxS" ~ 
+      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
+  arrange(gene, desc(percent)) %>%
+  dplyr::rename("Gene" = gene,
+                "Mutation" = value,
+                "Percent" = percent)
+
+```
+
+## Mechanisms in isolates without QRDR mutations
+This table presents the mechanisms identified in isolates with no mutations in either of _gyrA_, _gyrB_, _parC_ and/or _parE_.
+```{r}
+specific_ac_genes <- acquired_report %>%
+  gather(PMQR, value, -id) %>%
+  filter(value == 1) %>%
+  select(-value)
+
+mut_report %>%
+  # all gyrB were 0
+  filter(gyrA == 0,
+         parC == 0,
+         parE == 0) %>%
+  left_join(specific_ac_genes, by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  dplyr::select(-c(gyrA,gyrB,parC,parE))
+```
+
+## Total combinations of resistance mechanisms
+This table presents all combinations of resistance mechanisms included in this study, and their respective MIC.
+```{r}
+mut_report %>%
+  select(-c(nt_change, gyrA, parC, parE)) %>%
+  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
+            funs(as.numeric(.))) %>%
+  # summarise qnr-genes
+  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
+  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
+  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  arrange(desc(CIP))
+```
+
+## Significant differences in resistance mechanisms
+This table presents the significant (p < 0.05) differences in resistance mechanisms between species, identified with chi-squared tests.
+```{r, message=FALSE, warning=FALSE}
+# function for doing chi-squared tests on multiple entries in a list
+do_chisq_test <- function(df) {
+  df <- df %>%
+    # create columns with all possible comparisons
+    dplyr::slice(expand.grid(1:length(unique(species)),
+                             1:length(unique(species))) %>%
+                   rev %>% 
+          dplyr::filter(Var2 < Var1) %>% t) %>%
+    # create test ID
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    # specify values for chi-squred tests
+    dplyr::do(data_frame(gene = .$key,
+                  test = dplyr::first(.$test),
+                  species1 = dplyr::first(.$species),
+                  species2 = dplyr::last(.$species),
+                  data = list(matrix(c(.$Absent,
+                                       .$Present),
+                                     ncol = 2)))) %>%
+    # do chi-squared tests
+    mutate(chi_test = map(data, chisq.test,
+                          correct = FALSE)) %>%
+    mutate(p.value = map_dbl(chi_test, "p.value"),
+           dupl = duplicated(test)) %>%
+    # filter out duplicated tests
+    dplyr::filter(dupl == FALSE) %>%
+    ungroup() %>%
+    select(gene, species1, species2, p.value) %>%
+    # round off p-value column
+    mutate(p.value = round(p.value, 5)) %>%
+    dplyr::filter(p.value <= 0.05)
+  return(df)
+}
+
+# create total mechanism table
+total_df <- acquired_report %>%
+  left_join(mut_report, by = "id") %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  select(-nt_change)
+
+# create list of genes with total presence/absence values
+# from each animal species
+mechanism_per_species <- total_df %>%
+  select(id, species, everything()) %>%
+  select(-id) %>%
+  gather(key, value, -species) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  select(key, species, Present, Absent) %>%
+  split(f = .$key)
+
+# Get intrinsic gene names
+intrinsic_names <- names(mut_report)
+intrinsic_names <- intrinsic_names[-c(1,2)]
+
+# create percentage table over intrinsic genes
+intrinsic_gene_perc <- mut_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, intrinsic_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  select(key, species, Percent) %>%
+  dplyr::rename("gene" = key)
+
+# Get acquired gene names
+acquired_names <- names(acquired_report)
+acquired_names <- acquired_names[-1]
+
+# create percentage table over acquired genes
+acquired_genes_perc <- acquired_report %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, acquired_names) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  select(key, species, Percent) %>%
+  dplyr::rename("gene" = key)
+
+# create total percentage table
+total_perc <- rbind(intrinsic_gene_perc, acquired_genes_perc)
+
+# run chi-squared tests
+p_values_total <- lapply(mechanism_per_species, function(x) do_chisq_test(x)) %>%
+  # bind results
+  bind_rows() %>%
+  # specify percentages for species 1
+  left_join(total_perc, by = c("gene", c("species1" = "species"))) %>%
+  dplyr::rename("Percent species 1" = Percent) %>%
+  # specify percentages for species 2
+  left_join(total_perc, by = c("gene", c("species2" = "species"))) %>%
+  dplyr::rename("Percent species 2" = Percent) %>%
+  dplyr::select(gene,
+                species1,
+                `Percent species 1`,
+                species2,
+                `Percent species 2`,
+                p.value)
+
+p_values_total
+
+```
+
+## Percent resistance towards all tested compounds
+This table presents the percent resistance towards the compounds sulfamethoxazole (SMX), trimethoprim (TMP), ciprofloxacin (CIP), tetracycline (TET), nalidixic acid (NAL), cefotaxime (CTX), chloramphenicol (CHL), ampicillin (AMP), and gentamicin (GEN).
+```{r}
+percent_am_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  # set resistant/nonresistant based on MIC cutoffs
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  # calculate percentages
+  mutate(Total = Present + Absent,
+         Percent = round((Present/Total)*100, 1))
+
+percent_am_res
+```
+
+## Resistance mechanisms per sequence type
+This table presents the total resistance mechanisms identified in each sequence type (n > 2).
+```{r}
+mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  gather(gene, value, -ST) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  filter(Present != 0) %>%
+  mutate(Total = Present + Absent) %>%
+  # filter out all STs observed less than three times
+  filter(Total > 2) %>%
+  rowwise() %>%
+  mutate(Percent = round(Present/Total*100, 1)) %>%
+  ungroup() %>%
+  group_by(ST) %>%
+  mutate(no_of_mechanisms = n())
+```
+
+## Resistance mechanisms per sequence type per animal species
+This table presents each resistance mechanism identified in each sequence type (n > 2) per animal species.
+```{r}
+mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  gather(gene, value, -c(ST, species)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  filter(Present != 0) %>%
+  mutate(Total = Present + Absent) %>%
+  # filter out all STs observed less than three times
+  filter(Total > 2) %>%
+  rowwise() %>%
+  mutate(Percent = round(Present/Total*100, 1)) %>%
+  ungroup() %>%
+  group_by(ST, species) %>%
+  mutate(no_of_mechanisms = n()) %>%
+  arrange(ST, species)
+```
+
+## Mutations in intrinsic genes for isolates with CIP MIC < 0.12
+This table presents the mutations identified in isolates with a MIC-value for ciprofloxacin below the ECOFF. Note that all mutations in all genes are listed here, not just the QRDR-located ones. 
+```{r}
+cip_mic_sens <- mic_data %>%
+  select(id, CIP, NAL) %>%
+  filter(CIP < 0.12)
+
+acquired_genes <- acquired_report %>%
+  gather(gene, value, -id) %>%
+  filter(value == 1) %>%
+  select(-value) %>%
+  dplyr::rename("PMQR" = gene)
+
+mut_table_fixed %>%
+  filter(id %in% cip_mic_sens$id,
+         gene %in% gene_names) %>%
+  select(id, gene, mut) %>%
+  spread(gene, mut) %>%
+  left_join(acquired_genes, by = "id") %>%
+  left_join(mic_data[,c("id", "CIP", "NAL")], by = "id") %>%
+  left_join(isolate_data[,c("id","species","method")], by = "id")
+```
+
+## Significant differences between the amount of resistant isolates
+These tables present the significant (p < 0.05) differences in isolates resistant to quinolones (table 1), and the rest of the tested compounds (table 2) for each animal species.
+```{r, message=FALSE, warning=FALSE}
+# create table over amount of resistant and sensitive isolates for CIP and NAL
+cip_nal_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  select(compound, species, Present, Absent) %>%
+  dplyr::rename("key" = compound) %>%
+  split(f = .$key)
+
+# do chi-squared test
+p_values_res <- lapply(cip_nal_res, function(x) do_chisq_test(x)) %>%
+  bind_rows() %>%
+  # insert percentages for species 1
+  left_join(percent_am_res[,c("species","compound","Percent")],
+            by = c("species1" = "species",
+                   "gene" = "compound")) %>%
+  dplyr::rename("Percent_species1" = Percent) %>%
+  # insert percentages for species 2
+  left_join(percent_am_res[,c("species","compound","Percent")],
+            by = c("species2" = "species",
+                   "gene" = "compound")) %>%
+  dplyr::rename("Percent_species2" = Percent,
+                "Compound" = gene) %>%
+  select(Compound,
+         species1,
+         Percent_species1,
+         species2,
+         Percent_species2,
+         p.value)
+
+p_values_res
+
+# dataset for all other compounds
+rest_res <- mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  select(compound, species, Present, Absent) %>%
+  dplyr::rename("key" = compound) %>%
+  split(f = .$key)
+
+p_values_res_rest <- lapply(rest_res, function(x) do_chisq_test(x)) %>%
+  bind_rows() %>%
+  # insert percentages for species 1
+  left_join(percent_am_res[,c("species","compound","Percent")], 
+            by = c("species1" = "species",
+                   "gene" = "compound")) %>%
+  dplyr::rename("Percent_species1" = Percent) %>%
+  # insert percentages for species 2
+  left_join(percent_am_res[,c("species","compound","Percent")],
+            by = c("species2" = "species",
+                   "gene" = "compound")) %>%
+  dplyr::rename("Percent_species2" = Percent,
+                "Compound" = gene) %>%
+  select(Compound,
+         species1,
+         Percent_species1,
+         species2,
+         Percent_species2,
+         p.value)
+
+p_values_res_rest
+```
+
+## Correlation between specific mutations and qnr-genes
+```{r}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    base::unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    base::unlist() %>% base::unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(base::unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
+col_proc <- c("gyrA","gyrB","parC","parE")
+
+tot_mut_report <- sapply(function(i) {AddCol(df = mut_quant, col_name = col_proc[i])},
+                      X = seq_along(col_proc)) %>%
+  bind_cols() %>%
+  select(-c(gyrA_0, parC_0, gyrB_0, parE_0)) %>%
+  mutate_all(funs(as.numeric))
+
+test <- mut_quant %>%
+  cbind(tot_mut_report) %>%
+  select(-contains("mut")) %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  left_join(acquired_report, by = "id") %>%
+  mutate_at(vars(-id),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,21:27]) == 0, 0, 1)) %>%
+  select(-id)
+
+cor.test(test$gyrA_S83L, test$qnr)
+```
+
+## Double-check ARIBA results
+This chunk double-checks the mutations found in _gyrA_, _parC_ and _parE_ against an internal mutation analysis script.
+The sequences are taken from Prokka, and reference wild-type versions of the genes are used to identify mismatches with the sample sequences.
 ```{r, warning=FALSE, message=FALSE}
@@
   select(id, gyrA_prokka) %>%
-  mutate(gyrA_result = gyrA_prokka %>% # control for mutation within QRDR for gyrA
+  # control for mutation within QRDR for gyrA
+  mutate(gyrA_result = gyrA_prokka %>%
              str_extract_all("\\d+") %>% # from reprex package
@@
   dplyr::rename("gyrA_ariba" = value) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S")) %>%
+  filter(id %not_in% ex_samples) %>%
   select(-gyrA_result)
@@
   select(id, parC) %>%
-  mutate(entries = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)})) %>%
+  mutate(entries = sapply(strsplit(.$parC, ","),
+                          FUN = function(x) {length(x)})) %>%
   separate(parC, into = as.character(c(1:max(.$entries)))) %>%
@@
   dplyr::rename("parC_ariba" = value) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S")) %>%
+  filter(id %not_in% ex_samples) %>%
   select(-parC_result)
@@
   select(id, parE) %>%
-  mutate(entries = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
+  mutate(entries = sapply(strsplit(.$parE, ","),
+                          FUN = function(x) {length(x)})) %>%
   separate(parE, into = as.character(c(1:max(.$entries)))) %>%
@@
   dplyr::rename("parE_ariba" = value) %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S")) %>%
+  filter(id %not_in% ex_samples) %>%
   select(-parE_result)
@@
 
-# Tables
-## Isolate data
-```{r}
-isolate_data_complete <- isolate_data %>%
-  left_join(no_of_reads, by = "id")
-
-isolate_data_complete
-```
-
-## List of genes of interest
-This table lists all the genes of interest in this study, some information about them, and their accession number(s).
-```{r}
-
-acquired_ref_names <- acquired_data$ref_name
-acquired_ref <- c()
-
-for (i in ac_genes) {
-  for (j in acquired_ref_names) {
-    if (grepl(i,j, ignore.case = T) == TRUE) {
-      acquired_ref <- c(acquired_ref, j)
-    }
-  }
-}
-
-acquired_ref <- unique(acquired_ref)
-
-acquired_ref_genes <- acquired_data %>%
-  mutate(test = duplicated(ref_name),
-         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
-                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
-                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
-                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
-                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
-                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
-                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
-                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
-                          grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
-  filter(test == FALSE,
-         ref_name %in% acquired_ref_names) %>%
-  select(ref_name, gene) %>%
-  mutate(compl = complete.cases(.)) %>%
-  filter(compl == TRUE) %>%
-  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
-  group_by(gene) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(gene, ref_name2) %>%
-  dplyr::rename("Accession number(s)" = ref_name2,
-         "Gene" = gene) %>%
-  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
-         Type = "Acquired")
-
-df_ref_names <- mut_data$cluster
-ref_names <- c()
-
-for (i in mut_genes) {
-  for (j in df_ref_names) {
-    if (grepl(i,j, ignore.case = T) == TRUE) {
-      ref_names <- c(ref_names, j)
-    }
-  }
-}
-
-ref_names <- unique(ref_names)
-
-intrinsic_ref_genes <- mut_data %>%
-  mutate(test = duplicated(cluster),
-         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
-                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
-                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
-                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
-                          grepl("soxR", cluster, ignore.case = T) == TRUE ~ "soxR",
-                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
-                          grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
-                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
-                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
-                          grepl("acrb", ref_name, ignore.case = T) == TRUE ~ "acrB",
-                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA",
-                          grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB")) %>%
-  filter(test == FALSE,
-         cluster %in% ref_names) %>%
-  select(ref_name, gene, free_text) %>%
-  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
-  select(-ref_name) %>%
-  group_by(gene) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(gene, ref_name2) %>%
-  dplyr::rename("Accession number(s)" = ref_name2,
-         "Gene" = gene) %>%
-  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
-                                 Gene == "acrB" ~ "Gene that codes for the AcrB efflux pump, part of the AcrAB-TolC efflux system",
-                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
-                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
-                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
-                                 Gene == "marA" ~ "Transcriptional activator og genes involved in the MAR phenotype.",
-                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
-                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
-                                 Gene == "soxR" ~ "Transcriptional activator of the superoxide response regulon.",
-                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
-                                 Gene == "mdfA" ~ "Chromosomal efflux pump.",
-                                 Gene == "rpoB" ~ "Beta-subunit of RNA polymerase, mutations lead to increased expression of the ydhE-pump"),
-         Type = "Intrinsic gene")
-
-ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)
-
-ref_genes
-
-```
-
-## Isolates per species
-```{r}
-per_species <- isolate_data %>%
-  group_by(species) %>%
-  dplyr::count()
-
-per_species
-```
-
-## Sequence types in total
-```{r}
-tot_mlst <- mlst_data[,c("id","ST")] %>%
-  group_by(ST) %>%
-  dplyr::count() %>%
-  arrange(desc(n))
-
-print(paste0("Total amount of different sequence types: ", nrow(tot_mlst)))
-
-tot_mlst
-```
-
-## Sequence types per animal species
-```{r}
+# Figures
+
+## Sequence types per animal group
+```{r}
+palette4 <- c("Livestock" = "#80b1d3",
+              "Wild" = "#fb8072")
+
 mlst_data[,c("id","ST")] %>%
   left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(ST, species) %>%
-  dplyr::count() %>%
-  arrange(desc(n))
-```
-
-## Number of sequence types per animal species
-```{r}
-mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(species) %>%
-  mutate(n = length(unique(ST))) %>%
-  select(species,n) %>%
-  summarise_all(funs(func_paste))
-```
-
-## Percent isolates with PMQR
-```{r}
-acquired_report %>%
-  select(-id) %>%
-  mutate_all(funs(as.numeric(.))) %>%
-  mutate(total = rowSums(.),
-         result = if_else(total != 0, 1, 0)) %>%
-  group_by(result) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
-  spread(result, n) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1))
-```
-
-## Percent isolates with specific PMQR genes
-```{r}
-acquired_report %>%
-  select(-id) %>%
-  t() %>%
-  as.data.frame(.) %>%
-  mutate(Gene = rownames(.)) %>%
-  mutate_at(vars(-Gene),
-            funs(as.numeric(as.character(.)))) %>%
-  mutate(Present = rowSums(.[,-281])) %>%
-  select(c(Gene, Present)) %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(Present/Total * 100,1))
-```
-
-## Percent isolates with mutations in QRDR
-```{r}
-mut_report %>%
-  select(gyrA, gyrB, parC, parE) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
-            .funs = funs(as.numeric)) %>%
-  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
-  select(n, type) %>%
-  group_by(type) %>%
-  summarise_all(funs(sum)) %>%
-  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
-  spread(type, n) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1)) %>%
-  select(-Total)
-```
-
-## Specific mutations in QRDR
-```{r}
-suppressWarnings(clean_mut_data %>%
-  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
-         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
-         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
-         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
-         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
-         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
-  select(ref, gene_names, flag, mutation) %>%
-  filter(flag %in% flag_selection,
-         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
-  mutate(id = 1:n()) %>%
-  spread(gene_names, mutation) %>%
-  group_by(ref) %>%
-  summarise_all(funs(func_paste2)) %>%
-  select(-c(id, flag)) %>%
-  gather(gene, mut, -ref) %>%
-  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
-         result = ifelse(mut != 0, 1, 0),
-         result = as.integer(result),
-         type = "mut") %>%
-  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
-  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
-  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
-  select(-mut) %>%
-  dplyr::rename("mut" = aa_change) %>%
-  filter(is.na(nt_change) == FALSE) %>%
-  fix_gyr_par_results(.) %>%
-  filter(result_total == 1) %>%
-  group_by(gene, mut, nt_change) %>%
-  dplyr::count() %>%
-  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(n/Total * 100,1)) %>%
-  dplyr::rename("Mutation" = mut))
-```
-
-## Mutation combinations in QRDR
-```{r}
-mut_quant %>%
-  group_by(gyrA, gyrB, parC, parE) %>%
-  dplyr::count() %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(n/Total * 100,1))
-```
-
-# Figures
+  mutate(animal_group = ifelse(species %in% c("Broiler","Pig"), "Livestock", "Wild")) %>%
+  group_by(ST, animal_group) %>%
+  dplyr::count() %>%
+  group_by(ST) %>%
+  mutate(total = dplyr::first(n) + dplyr::last(n)) %>%
+  filter(total > 4) %>%
+  ggplot(aes(factor(ST), n, fill = animal_group))+
+  geom_col(color = "black")+
+  labs(x = "Sequence Types")+
+  scale_fill_manual(values = palette4)+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.45),
+        legend.title = element_blank())
+```
+
 ## Resistance mechanisms per sequence type
@@
 
+## Percent QRDR mutations per animal species
+```{r}
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+suppressWarnings(clean_mut_data %>%
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  summarise_all(funs(func_paste2)) %>%
+  dplyr::select(-c(id, flag)) %>%
+  left_join(id_data, by = "ref") %>%
+  dplyr::select(id, everything(), -ref) %>%
+  filter(id %not_in% ex_samples) %>%
+  gather(gene, mut, -id) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
+  select(-mut) %>%
+  dplyr::rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  group_by(gene, mut, species) %>%
+  dplyr::count() %>%
+  rowwise() %>%
+  mutate(Total = case_when(species == "Broiler" ~ 87,
+                           species == "Pig" ~ 75,
+                           species == "Red fox" ~ 52,
+                           species == "Wild bird" ~ 66),
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1], 1),
+         upr = round(get_binCI(n, Total)[2], 1)) %>%
+  dplyr::rename("Mutation" = mut)) %>%
+  ggplot(aes(reorder(Mutation, -Percent), Percent, fill = species))+
+  geom_col(color = "black", position = position_dodge(0.9))+
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                position = position_dodge(0.9),
+                width = 0.5)+
+  scale_fill_manual(values = palette)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
+        axis.title.x = element_blank())+
+  facet_wrap(~ gene, scale = "free_x")
+
+```
+
 ## Percent QREC isolates with Qnr genes per species
 ```{r}
-qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
+qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")
 
@@
 
-## cgMLST-trees
-```{r, fig.width=24, fig.height=22, message = FALSE, warning = FALSE}
-
-palette3 <- c("1-A" = "#8dd3c7",
-              "1-I" = "#80b1d3",
-              "0" = "grey95")
-
-qr_genes <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
+## Correlation plot for resistance mechanisms
+```{r, fig.height=8, fig.width=7, warning=FALSE}
+total_df <- acquired_report %>%
+  left_join(mut_report, by = "id") %>%
+  left_join(isolate_data[c("id","species")], by = "id") %>%
+  select(-c(nt_change, id)) %>%
+  mutate_at(vars(-species),
+            funs(as.numeric)) %>%
+  mutate(qnr = ifelse(rowSums(.[,2:8]) > 0, 1, 0)) %>%
+  {. ->> total_resistance_report} %>%
+  split(., f = .$species) %>%
+  lapply(., function(x) select(x, -species))
+
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor(method = "spearman") %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
+  bind_rows(.id = "species") %>%
+  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
+           gene2 = gsub("(.*?)_.+", "\\1", var2),
+           var1 = gsub(".+_(.*?)", "\\1", var1),
+           var2 = gsub(".+_(.*?)", "\\1", var2),
+           type1 = case_when(gene1 == "gyrA" ~ 1,
+                          gene1 == "gyrB" ~ 2,
+                          gene1 == "parC" ~ 3,
+                          gene1 == "parE" ~ 4,
+                          gene1 == "marR" ~ 5,
+                          gene1 == "marA" ~ 6,
+                          gene1 == "rpoB" ~ 7,
+                          gene1 == "soxS" ~ 8,
+                          gene1 == "robA" ~ 9,
+                          gene1 == "gadX" ~ 10,
+                          gene1 == "qnrA1" ~ 11,
+                          gene1 == "qnrB19" ~ 12,
+                          gene1 == "qnrB6" ~ 13,
+                          gene1 == "qnrB60" ~ 14,
+                          gene1 == "qnrS1" ~ 15,
+                          gene1 == "qnrS2" ~ 16,
+                          gene1 == "qnrS4" ~ 17,
+                          gene1 == "qnr" ~ 18,
+                          gene1 == "qepA4" ~ 19),
+           type2 = case_when(gene2 == "gyrA" ~ 1,
+                          gene2 == "gyrB" ~ 2,
+                          gene2 == "parC" ~ 3,
+                          gene2 == "parE" ~ 4,
+                          gene2 == "marR" ~ 5,
+                          gene2 == "marA" ~ 6,
+                          gene2 == "rpoB" ~ 7,
+                          gene2 == "soxS" ~ 8,
+                          gene2 == "robA" ~ 9,
+                          gene2 == "gadX" ~ 10,
+                          gene2 == "qnrA1" ~ 11,
+                          gene2 == "qnrB19" ~ 12,
+                          gene2 == "qnrB6" ~ 13,
+                          gene2 == "qnrB60" ~ 14,
+                          gene2 == "qnrS1" ~ 15,
+                          gene2 == "qnrS2" ~ 16,
+                          gene2 == "qnrS4" ~ 17,
+                          gene2 == "qnr" ~ 18,
+                          gene2 == "qepA4" ~ 19))
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradient(low = "black",
+                      high = "white",
+                      na.value = "grey95")+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 8),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank())+
+  coord_fixed()+
+  facet_wrap(~species)
+```
+
+## MIC-plots
+```{r}
+total_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  dplyr::select(-nt_change) %>%
   mutate_at(vars(-id),
-            funs(ifelse(. == "1", "1-I", "0")))
-
-ac_genes <- acquired_report %>%
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-ref) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(present,CIP) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(present = ifelse(present == "", "None", present),
+         total = 280,
+         percent = round(n/total * 100, 1),
+         lwr = round(get_binCI(n, total)[1], 1),
+         upr = round(get_binCI(n, total)[2], 1)) %>%
+  ungroup() %>%
+  mutate(entries = sapply(strsplit(.$present, ","),
+                          FUN = function(x) {length(x)}))
+
+genes <- c(names(mut_report), names(acquired_report), "qnr")
+genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")]
+
+n_plot <- ggplot(total_report, aes(reorder(present, entries), percent, fill = factor(CIP)))+
+  geom_col(color = "black")+
+  labs(y = "Percent (%) isolates")+
+  scale_fill_viridis(discrete = TRUE)+
+  guides(fill=guide_legend(ncol = 10))+
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        axis.text.y = element_text(size = 8),
+        axis.title.x = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_blank(),
+        legend.position = "top",
+        legend.spacing.x = unit(0.2, "cm"),
+        legend.justification = "center")
+
+dot_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  select(-nt_change) %>%
   mutate_at(vars(-id),
-            funs(as.numeric)) %>%
-  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
-  select(id, qepA4, qnr) %>%
-  mutate_at(vars(-id),
-            funs(ifelse(. == "1", "1-A", "0")))
-
-cgMLSTdata <- read.table("../data/cgMLST.txt",
-                         sep = "\t",
-                         row.names=1,
-                         colClasses = "factor",
-                         header = T)
-
-broiler_iso <- isolate_data[c("id", "species")] %>%
-  filter(species == "Broiler") %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-pig_iso <- isolate_data[c("id", "species")] %>%
-  filter(species == "Pig") %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-fox_iso <- isolate_data[c("id", "species")] %>%
-  filter(species == "Red fox") %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-bird_iso <- isolate_data[c("id", "species")] %>%
-  filter(species == "Wild bird") %>%
-  filter(id %not_in% c("2016-17-565-1-S",
-                       "2016-02-428-2-S",
-                       "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
-
-data_broiler <- cgMLSTdata %>%
-  rownames_to_column("id") %>%
-  filter(id %in% broiler_iso$id) %>%
-  dplyr::rename("Genome" = `id`) %>%
-  select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_pig <- cgMLSTdata %>%
-  rownames_to_column("id") %>%
-  filter(id %in% pig_iso$id) %>%
-  dplyr::rename("Genome" = `id`) %>%
-  select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_fox <- cgMLSTdata %>%
-  rownames_to_column("id") %>%
-  filter(id %in% fox_iso$id) %>%
-  dplyr::rename("Genome" = `id`) %>%
-  select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-data_bird <- cgMLSTdata %>%
-  rownames_to_column("id") %>%
-  filter(id %in% bird_iso$id) %>%
-  dplyr::rename("Genome" = `id`) %>%
-  select(Genome, everything()) %>%
-  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))
-
-cgMLSTdata_broiler <- read.table("../data/cgMLST_broiler.txt",
-                                 sep = "\t",
-                                 row.names=1,
-                                 colClasses = "factor",
-                                 header = T)
-
-cgMLST_tree_broiler <- as.phylo(hclust(daisy(cgMLSTdata_broiler,
-                                             metric="gower")))
-
-cgMLSTdata_pig <- read.table("../data/cgMLST_pig.txt",
-                             sep = "\t",
-                             row.names=1,
-                             colClasses = "factor",
-                             header = T)
-
-cgMLST_tree_pig <- as.phylo(hclust(daisy(cgMLSTdata_pig,
-                                         metric="gower")))
-
-cgMLSTdata_fox <- read.table("../data/cgMLST_fox.txt", 
-                             sep = "\t", 
-                             row.names=1, 
-                             colClasses = "factor", 
-                             header = T)
-
-cgMLST_tree_fox <- as.phylo(hclust(daisy(cgMLSTdata_fox,
-                                         metric="gower")))
-
-cgMLSTdata_bird <- read.table("../data/cgMLST_bird.txt", 
-                              sep = "\t", 
-                              row.names=1, 
-                              colClasses = "factor", 
-                              header = T)
-
-cgMLST_tree_bird <- as.phylo(hclust(daisy(cgMLSTdata_bird,
-                                          metric="gower")))
-
-broiler_tree <- ggtree(cgMLST_tree_broiler,
-                       layout = "circular",
-                       color = "#808080",
-                       size = 1) %<+% data_broiler +
-  geom_tippoint(size = 4,
-                color = "#4575b4")+
-  ggtitle("A")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
-
-pig_tree <- ggtree(cgMLST_tree_pig,
-                   layout = "circular",
-                   color = "#808080",
-                   size = 1) %<+% data_pig +
-  geom_tippoint(size = 4,
-                color = "#74add1")+
-  ggtitle("B")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
-
-fox_tree <- ggtree(cgMLST_tree_fox,
-                   layout = "circular",
-                   color = "#808080",
-                   size = 1) %<+% data_fox +
-  geom_tippoint(size = 4,
-                color = "#f46d43")+
-  ggtitle("C")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
-
-bird_tree <- ggtree(cgMLST_tree_bird,
-                    layout = "circular",
-                    color = "#808080",
-                    size = 1) %<+% data_bird +
-  geom_tippoint(size = 4,
-                color = "#fdae61")+
-  ggtitle("D")+
-  theme(plot.title = element_text(size = 40, face = "bold"),
-        plot.margin=unit(c(0,0,0,0), "cm"))
-
-total_data <- qr_genes %>%
-  left_join(ac_genes, by = "id")
-
-broiler_data <- total_data %>%
-  filter(id %in% broiler_iso$id) %>%
-  column_to_rownames("id")
-
-pig_data <- total_data %>%
-  filter(id %in% pig_iso$id) %>%
-  column_to_rownames("id")
-
-fox_data <- total_data %>%
-  filter(id %in% fox_iso$id) %>%
-  column_to_rownames("id")
-
-bird_data <- total_data %>%
-  filter(id %in% bird_iso$id) %>%
-  column_to_rownames("id")
-
-fan_tree_broiler <- open_tree(broiler_tree, 10)
-
-rotated_fan_tree_broiler <- rotate_tree(fan_tree_broiler, 93)
-
-broiler_annotated_tree <- gheatmap(rotated_fan_tree_broiler,
-                                   broiler_data,
-                                   offset = 0.01,
-                                   width = 0.5,
-                                   colnames_offset_y = 0.7,
-                                   colnames_position = "top",
-                                   font.size = 7) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-fan_tree_pig <- open_tree(pig_tree, 10)
-
-rotated_fan_tree_pig <- rotate_tree(fan_tree_pig, 92)
-
-pig_annotated_tree <- gheatmap(rotated_fan_tree_pig, 
-                               pig_data,
-                               offset = 0.01, 
-                               width = 0.5, 
-                               colnames_offset_y = 0.55, 
-                               colnames_position = "top", 
-                               font.size = 7) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-fan_tree_fox <- open_tree(fox_tree, 10)
-
-rotated_fan_tree_fox <- rotate_tree(fan_tree_fox, 92)
-
-fox_annotated_tree <- gheatmap(rotated_fan_tree_fox, 
-                               fox_data, 
-                               offset = 0.01, 
-                               width = 0.5, 
-                               colnames_offset_y = 0.2, 
-                               colnames_position = "top", 
-                               font.size = 7) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-fan_tree_bird <- open_tree(bird_tree, 10)
-
-rotated_fan_tree_bird <- rotate_tree(fan_tree_bird, 93)
-
-bird_annotated_tree <- gheatmap(rotated_fan_tree_bird, 
-                                bird_data, 
-                                offset = 0.01, 
-                                width = 0.5, 
-                                colnames_offset_y = 0.4, 
-                                colnames_position = "top", 
-                                font.size = 7) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-total_plot <- cowplot::plot_grid(broiler_annotated_tree, 
-                                 pig_annotated_tree, 
-                                 fox_annotated_tree, 
-                                 bird_annotated_tree, 
-                                 ncol = 2, 
-                                 align = "h")
-total_plot
-
-```
-
-## Another color scheme
-```{r,fig.width=24, fig.height=22, message = FALSE, warning = FALSE}
-
-qr_genes_alt <- mut_report[c("id","gyrA","parC","parE","rpoB","marA","marR")] %>%
-  mutate(gyrA = ifelse(gyrA == 1, "gyrA_1", gyrA),
-         parC = ifelse(parC == 1, "parC_1", parC),
-         parE = ifelse(parE == 1, "parE_1", parE),
-         rpoB = ifelse(rpoB == 1, "rpoB_1", rpoB),
-         marA = ifelse(marA == 1, "marA_1", marA),
-         marR = ifelse(marR == 1, "marR_1", marR))
-
-ac_genes_alt <- acquired_report %>%
-  mutate_at(vars(-id),
-            funs(as.numeric)) %>%
-  mutate(qnr = ifelse(rowSums(.[,3:9]) != 0 , 1, 0)) %>%
-  select(id, qepA4, qnr) %>%
-  mutate(qnr = ifelse(qnr == 1, "qnr_1", qnr),
-         qepA4 = ifelse(qepA4 == 1, "qepA4_1", qepA4))
-
-total_data_alt <- qr_genes_alt %>%
-  left_join(ac_genes_alt, by = "id")
-
-palette_heatmap <- c("gyrA_1" = "#8dd3c7",
-                     "parC_1" = "#ffffb3",
-                     "parE_1" = "#bebada",
-                     "rpoB_1" = "#fb8072",
-                     "marA_1" = "#80b1d3",
-                     "marR_1" = "#fdb462",
-                     "qepA4_1" = "#b3de69",
-                     "qnr_1" = "#fccde5",
-                     "0" = "grey95")
-
-broiler_tree <- ggtree(cgMLST_tree_broiler,
-                       layout = "circular",
-                       color = "#4575b4",
-                       size = 1) %<+% data_broiler +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("A")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-broiler_data <- total_data_alt %>%
-  filter(id %in% broiler_iso$id) %>%
-  column_to_rownames("id")
-
-broiler_annotated_tree2 <- gheatmap(broiler_tree,
-                                   broiler_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-pig_tree <- ggtree(cgMLST_tree_pig,
-                       layout = "circular",
-                       color = "#74add1",
-                       size = 1) %<+% data_pig +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("B")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-pig_data <- total_data_alt %>%
-  filter(id %in% pig_iso$id) %>%
-  column_to_rownames("id")
-
-pig_annotated_tree2 <- gheatmap(pig_tree,
-                                   pig_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-fox_tree <- ggtree(cgMLST_tree_fox,
-                       layout = "circular",
-                       color = "#f46d43",
-                       size = 1) %<+% data_fox +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("C")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-fox_data <- total_data_alt %>%
-  filter(id %in% fox_iso$id) %>%
-  column_to_rownames("id")
-
-fox_annotated_tree2 <- gheatmap(fox_tree,
-                                   fox_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-bird_tree <- ggtree(cgMLST_tree_bird,
-                       layout = "circular",
-                       color = "#fdae61",
-                       size = 1) %<+% data_bird +
-  geom_tippoint(size = 2,
-                color = "black")+
-  ggtitle("D")+
-  theme(plot.title = element_text(size = 20, face = "bold"))
-
-bird_data <- total_data_alt %>%
-  filter(id %in% bird_iso$id) %>%
-  column_to_rownames("id")
-
-bird_annotated_tree2 <- gheatmap(bird_tree,
-                                   bird_data,
-                                   offset = 0.01,
-                                   width = 0.3,
-                                   colnames = FALSE)+
-  scale_fill_manual(values = palette_heatmap)+
-  guides(fill = FALSE)
-
-total_plot2 <- cowplot::plot_grid(broiler_annotated_tree2, 
-                                 pig_annotated_tree2, 
-                                 fox_annotated_tree2, 
-                                 bird_annotated_tree2, 
-                                 ncol = 2, 
-                                 align = "h")
-total_plot2
-```
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  select(-c(ref, id)) %>%
+  ungroup() %>%
+  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
+  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
+         present = ifelse(present == "", "None", present)) %>%
+  gather(gene, value, -c(present,CIP)) %>%
+  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
+         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
+  na.omit() %>%
+  mutate(entries = sapply(strsplit(.$present, ","),
+                          FUN = function(x) {length(x)}))
+
+dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
+  geom_point(size = 2)+
+  scale_y_discrete(limits = c("gyrA","parC","parE","gadX","marA","marR","rpoB","soxS","qepA4","qnr"))+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.title = element_blank(),
+        axis.ticks.x = element_blank())
+
+plot_grid(n_plot, dot_plot, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
+```
+
+## Additional resistance levels per animal species
+```{r}
+mic_data %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
+         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
+         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
+         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
+         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
+         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
+         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
+         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
+         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
+  select(species,contains("_R")) %>%
+  gather(compound, value, -species) %>%
+  mutate(compound = sub("_R", "", compound)) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round((Present/Total)*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1),
+         species = factor(species, levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
+  ggplot(aes(reorder(compound, -Percent), Percent, fill = species))+
+  geom_col(color = "black")+
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                width = 0.6)+
+  guides(fill = FALSE)+
+  scale_fill_manual(values = palette)+
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))+
+  facet_wrap(~species)
+```
+
+## CIP/NAL MIC-values per animal species
+```{r}
+mic_data[,c("id","CIP")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(CIP,species) %>%
+  dplyr::count() %>%
+  ggplot(aes(factor(CIP), n, fill = species))+
+  geom_col(color = "black", position = position_fill())+
+  labs(y = "Percent isolates",
+       x = "CIP MIC")+
+  scale_fill_manual(values = palette)+
+  scale_y_continuous(labels = percent_format())+
+  theme(legend.title = element_blank())
+
+mic_data[,c("id","NAL")] %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(NAL,species) %>%
+  dplyr::count() %>%
+  ggplot(aes(factor(NAL), n, fill = species))+
+  geom_col(color = "black", position = position_fill())+
+  labs(y = "Percent isolates",
+       x = "NAL MIC")+
+  scale_fill_manual(values = palette)+
+  scale_y_continuous(labels = percent_format())+
+  theme(legend.title = element_blank())
+
+```

### 2019-03-15 notebooks/Clustering analysis.Rmd
initial commit

---
title: "Clustering analysis and statistics"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r libraries, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(cluster)
library(Biostrings)
library(ape)
library(tibble)
library(cowplot)
```

# cgMLST Schema and quality control
```{r import data}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

# Isolate data
id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

# cgMLST data
cgmlst_stats <- read.table("../data/chewbbaca/complete_results/mdata_stats.tsv",
                         sep = "\t",
                         header = TRUE,
                         stringsAsFactors = FALSE)

cgMLST_allele_data <- read.table("../data/chewbbaca/complete_results/cgMLST.tsv",
                              sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE) %>%
  mutate(FILE = sub("_pilon_spades.fasta", "", FILE)) %>%
  dplyr::rename("ref" = FILE) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% ex_samples)

```

```{r Number of missing loci}
boxplot(cgmlst_stats$number.of.missing.data,
        ylab = "Number of missing loci")
```

# Total dendrogram of all isolates
```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
cgMLST_data <- read.table("../data/chewbbaca/complete_results/clean_cgMLST_data.txt",
                   sep = "\t",
                   header = TRUE,
                   colClasses = "factor")

total_tree <- as.phylo(hclust(daisy(cgMLST_data, metric = "gower"), method = "average"))

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE)

tree_heatmap_data <- read.table("../data/chewbbaca/complete_results/tree_heatmap_data.txt",
                            sep = "\t",
                            header = TRUE)

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

palette2 <- c("1-A" = "#fc8d59",
              "1-I" = "#80b1d3",
              "0" = "grey95")

total_tree_annotated <- ggtree(total_tree,
                               layout = "circular",
                               color = "#808080",
                               size = 0.5) %<+% tree_metadata +
                               geom_tippoint(aes(color = species),
                               size = 3) +
                               geom_tiplab2(aes(label = ST),
                               offset = 0.01,
                               size = 4) +
  scale_color_manual(values = palette)

total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)

gheatmap(total_tree_opened,
         tree_heatmap_data,
         offset = 0.04,
         width = 0.3,
         colnames_offset_y = 2.5,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette2,
                    labels = c("Gene/Mutation absent",
                               "Acquired gene present",
                               "Intrinsic gene mutated")) +
  theme(legend.position = c(0.9,0.9),
        legend.text = element_text(size = 20))
```

```{r, fig.height=25, fig.width=25, message=FALSE, warning=FALSE}

metadata_species <- read.table("../data/chewbbaca/complete_results/metadata_species_specific_trees.txt",
                               sep = "\t",
                               header = TRUE)

palette3 <- c("0" = "grey95",
              # gyrA
              "S83L" = "#c6dbef",
              "S83A" = "#9ecae1",
              "D87N" = "#6baed6",
              "D87H" = "#4292c6",
              "D87Y" = "#2171b5",
              "D87G" = "#08519c",
              "S83L, D87N" = "#08306b",
              # parC
              "S80I" = "#a1d99b",
              "S80R" = "#74c476",
              "S57T" = "#41ab5d",
              "A56T, S80I" = "#238b45",
              "S80I, E84V" = "#005a32",
              # parE
              "D475E" = "#fdd0a2",
              "D463N" = "#fdae6b",
              "S458A" = "#fd8d3c",
              "L416F" = "#f16913",
              "H516Y" = "#d94801",
              "L488M, A512T" = "#8c2d04",
              # PMQR
              "qnrA1" = "#762a83",
              "qnrB19" = "#ef8a62",
              "qnrS1" = "#a6dba0",
              "qnrS2" = "#5aae61",
              "qnrS4" = "#1b7837",
              "qepA4" = "#80cdc1",
              # num
              "1" = "#440154FF",
              "2" = "#471164FF",
              "3" = "#481F70FF",
              "4" = "#472D7BFF",
              "5" = "#443A83FF",
              "6" = "#404688FF",
              "7" = "#3B528BFF",
              "8" = "#365D8DFF",
              "9" = "#31688EFF",
              "10" = "#2C728EFF",
              "11" = "#287C8EFF",
              "12" = "#24868EFF",
              "13" = "#21908CFF",
              "14" = "#1F9A8AFF",
              "15" = "#20A486FF",
              "16" = "#27AD81FF",
              "17" = "#35B779FF",
              "18" = "#47C16EFF",
              "19" = "#5DC863FF",
              "20" = "#75D054FF",
              "21" = "#8FD744FF",
              "22" = "#AADC32FF",
              "23" = "#C7E020FF",
              "24" = "#E3E418FF",
              "25" = "#FDE725FF")

# Broiler tree
broiler_cgMLST <- read.table("../data/chewbbaca/complete_results/broiler_cgmlst.txt",
                   sep = "\t",
                   header = TRUE,
                   colClasses = "factor")

broiler_tree <- as.phylo(hclust(daisy(broiler_cgMLST, metric = "gower"), method = "average"))

broiler_tree_annotated <- ggtree(broiler_tree,
                                 layout = "circular",
                                 color = "#808080",
                                 size = 0.5) %<+% tree_metadata +
  geom_tippoint(color = "#4575b4",
                size = 4) +
  geom_tiplab2(aes(label = ST),
               offset = 0.01,
               size = 6)

broiler_tree_opened <- rotate_tree(open_tree(broiler_tree_annotated, 10), 92.5)

broiler_complete <- gheatmap(broiler_tree_opened,
         metadata_species,
         offset = 0.09,
         width = 0.5,
         colnames_offset_y = 0.75,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

# Pig tree
pig_cgMLST <- read.table("../data/chewbbaca/complete_results/pig_cgmlst.txt",
                   sep = "\t",
                   header = TRUE,
                   colClasses = "factor")

pig_tree <- as.phylo(hclust(daisy(pig_cgMLST, metric = "gower"), method = "average"))

pig_tree_annotated <- ggtree(pig_tree,
                                 layout = "circular",
                                 color = "#808080",
                                 size = 0.5) %<+% tree_metadata +
  geom_tippoint(color = "#74add1",
                size = 4) +
  geom_tiplab2(aes(label = ST),
               offset = 0.01,
               size = 5.5)

pig_tree_opened <- rotate_tree(open_tree(pig_tree_annotated, 10), 92.5)

pig_complete <- gheatmap(pig_tree_opened,
         metadata_species,
         offset = 0.09,
         width = 0.5,
         colnames_offset_y = 0.53,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

# Red fox tree
fox_cgMLST <- read.table("../data/chewbbaca/complete_results/fox_cgmlst.txt",
                   sep = "\t",
                   header = TRUE,
                   colClasses = "factor")

fox_tree <- as.phylo(hclust(daisy(fox_cgMLST, metric = "gower"), method = "average"))

fox_tree_annotated <- ggtree(fox_tree,
                                 layout = "circular",
                                 color = "#808080",
                                 size = 0.5) %<+% tree_metadata +
  geom_tippoint(color = "#f46d43",
                size = 4) +
  geom_tiplab2(aes(label = ST),
               offset = 0.01,
               size = 5.5)

fox_tree_opened <- rotate_tree(open_tree(fox_tree_annotated, 10), 92)

fox_complete <- gheatmap(fox_tree_opened,
         metadata_species,
         offset = 0.09,
         width = 0.5,
         colnames_offset_y = 0.17,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

# Wild bird tree
bird_cgMLST <- read.table("../data/chewbbaca/complete_results/bird_cgmlst.txt",
                   sep = "\t",
                   header = TRUE,
                   colClasses = "factor")

bird_tree <- as.phylo(hclust(daisy(bird_cgMLST, metric = "gower"), method = "average"))

bird_tree_annotated <- ggtree(bird_tree,
                                 layout = "circular",
                                 color = "#808080",
                                 size = 0.5) %<+% tree_metadata +
  geom_tippoint(color = "#f46d43",
                size = 4) +
  geom_tiplab2(aes(label = ST),
               offset = 0.01,
               size = 5.5)

bird_tree_opened <- rotate_tree(open_tree(bird_tree_annotated, 10), 92.5)

bird_complete <- gheatmap(bird_tree_opened,
         metadata_species,
         offset = 0.09,
         width = 0.5,
         colnames_offset_y = 0.38,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3) +
  guides(fill = FALSE)

# Total plot
tot_plot <- plot_grid(broiler_complete,
          pig_complete,
          fox_complete,
          bird_complete,
          nrow = 2,
          ncol = 2,
          align = "h")
tot_plot

```

```{r}
# calculates distances from cgMLST data and returns
# the lower triangular of the matrix, with (diag = FALSE)
# or without (diag = TRUE) the diagonal
lower_tri_dist_calc <- function(data, diag = TRUE, data_frame = FALSE) {
  dist <- as.matrix(daisy(data, metric = "gower"))
  if (diag == TRUE) {
    dist[upper.tri(dist, diag = TRUE)] <- NA
  } else {
    dist[upper.tri(dist, diag = FALSE)] <- NA
  }
  dist_vec <- as.vector(as.matrix(dist))
  dist_compl <- dist_vec[!is.na(dist_vec)]

  if (data_frame == TRUE) {
    dist_compl <- as.data.frame(as.matrix(dist))
  }
  return(dist_compl)
}

broiler_dist <- lower_tri_dist_calc(broiler_cgMLST)
pig_dist <- lower_tri_dist_calc(pig_cgMLST)
fox_dist <- lower_tri_dist_calc(fox_cgMLST)
bird_dist <- lower_tri_dist_calc(bird_cgMLST)

# Distance plots
par(mfrow = c(2,2))

x1 <- hist(broiler_dist,
     breaks = 20,
     plot = FALSE)

x1$density = x1$counts/sum(x1$counts)*100

plot(x1,
     freq = FALSE,
     main = "Broiler distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x2 <- hist(pig_dist,
     breaks = 20,
     plot = FALSE)

x2$density = x2$counts/sum(x2$counts)*100

plot(x2,
     freq = FALSE,
     main = "Pig distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x3 <- hist(fox_dist,
     breaks = 20,
     plot = FALSE)

x3$density = x3$counts/sum(x3$counts)*100

plot(x3,
     freq = FALSE,
     main = "Red fox distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x4 <- hist(bird_dist,
     breaks = 20,
     plot = FALSE)

x4$density = x4$counts/sum(x4$counts)*100

plot(x4,
     freq = FALSE,
     main = "Wild bird distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

```

# Statistics
```{r, warning=FALSE, message=FALSE}
alleles <- cgMLST_allele_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         new_id = case_when(species == "Broiler" ~ paste("BR", n, sep = "_"),
                            species == "Pig" ~ paste("PI", n, sep = "_"),
                            species == "Red fox" ~ paste("RF", n, sep = "_"),
                            species == "Wild bird" ~ paste("WB", n, sep = "_"))) %>%
  ungroup() %>%
  select(new_id, everything(), -c(species, id)) %>%
  mutate_all(funs(as.factor)) %>%
  column_to_rownames("new_id")

run_df <- lower_tri_dist_calc(alleles, data_frame = TRUE) %>%
  rownames_to_column("id") %>%
  gather(isolate1, value, -id) %>%
  dplyr::rename("isolate2" = id) %>%
  mutate(group1 = substr(isolate1, start = 1, stop = 2),
         group2 = substr(isolate2, start = 1, stop = 2)) %>%
  select(isolate1, group1, isolate2, group2, value) %>%
  filter(is.na(value) == FALSE,
         group1 == group2) %>%
  arrange(group1, group2)

below_calc <- function(df, run, dist_value = 0.1) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(n = value < dist_value) %>%
    group_by(group1, group2, n) %>%
    count() %>%
    ungroup() %>%
    mutate(n = ifelse(n == TRUE, "Below", "Above")) %>%
    spread(n, nn) %>%
    mutate(Below = ifelse(is.na(Below) == TRUE, 0, Below)) %>%
    mutate(Percent_below = round(Below / (Above + Below) * 100, 2),
           iter = run)
  return(df)
}

below_calc2 <- function(df, run, dist_value = 0.1) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(n = n()) %>%
    sample_n(min(.$n)) %>%
    select(-n) %>%
    mutate(n = value < dist_value) %>%
    group_by(group1, group2, n) %>%
    count() %>%
    ungroup() %>%
    mutate(n = ifelse(n == TRUE, "Below", "Above")) %>%
    spread(n, nn) %>%
    mutate(Below = ifelse(is.na(Below) == TRUE, 0, Below)) %>%
    mutate(Percent_below = round(Below / (Above + Below) * 100, 2),
           iter = run)
  return(df)
}

above_calc <- function(df, run, dist_value = 0.9) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(n = value > dist_value) %>%
    group_by(group1, group2, n) %>%
    count() %>%
    ungroup() %>%
    mutate(n = ifelse(n == TRUE, "Above", "Below")) %>%
    spread(n, nn) %>%
    mutate(Above = ifelse(is.na(Above) == TRUE, 0, Above)) %>%
    mutate(Percent_above = round(Above / (Above + Below) * 100, 2),
           iter = run)
  return(df)
}

above_calc2 <- function(df, run, dist_value = 0.9) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(n = n()) %>%
    sample_n(min(.$n)) %>%
    select(-n) %>%
    mutate(n = value > dist_value) %>%
    group_by(group1, group2, n) %>%
    count() %>%
    ungroup() %>%
    mutate(n = ifelse(n == TRUE, "Above", "Below")) %>%
    spread(n, nn) %>%
    mutate(Above = ifelse(is.na(Above) == TRUE, 0, Above)) %>%
    mutate(Percent_above = round(Above / (Above + Below) * 100, 2),
           iter = run)
  return(df)
}

mean_calc <- function(df, run) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(metric = round(mean(value), 2)) %>%
    select(group1, group2, metric) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(iter = run,
           metric = as.numeric(metric))
}

median_calc <- function(df, run) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(metric = round(median(value), 2)) %>%
    select(group1, group2, metric) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(iter = run,
           metric = as.numeric(metric))
}

quantile_calc <- function(df, run) {
  df <- df %>%
    group_by(group1, group2) %>%
    mutate(quant_0 = quantile(value)[1],
           quant_25 = quantile(value)[2],
           quant_50 = quantile(value)[3],
           quant_75 = quantile(value)[4],
           quant_100 = quantile(value)[5]) %>%
    select(group1, group2, contains("quant")) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(vars(contains("quant")),
              funs(as.numeric)) %>%
    mutate(iter = run)
}

iterate_samples <- function(df, run, method = "mean", dist_value) {

  df <- df %>%
    mutate(value = sample(value))

    if (method == "mean") {
      df <- mean_calc(df, run)
    }

    if (method == "below") {
     df <- below_calc2(df, run, dist_value)
    }

    if (method == "above") {
     df <- above_calc2(df, run, dist_value)
    }

    if (method == "median") {
      df <- median_calc(df, run)
    }

    if (method == "quantile") {
      df <- quantile_calc(df, run)
    }

    return(df)
}

# run iterations on selected data frame
run_iterations <- function(df, runs = 1000, seed = 10, method = "mean", dist_value) {
  # set seed
  set.seed(seed)

  # create output list
  output <- list()

  # create expected values
    if (method == "mean") {
      orig <- mean_calc(df, 0)
    }

    if (method == "below") {
     orig <- below_calc(df, 0, dist_value)
    }

   if (method == "above") {
     orig <- above_calc(df, 0, dist_value)
    }

    if (method == "median") {
      orig <- median_calc(df, 0)
    }

    if (method == "quantile") {
      orig <- quantile_calc(df, 0)
    }

  output <- c(output, list(orig))

  # run iterations
  for (i in 1:runs) {
    result <- iterate_samples(df,
                              i,
                              method = method,
                              dist_value = dist_value)
    output <- c(output, list(result))
  }
  return(output)
}

# create distribution plots from iteration data
create_plot <- function(df, group_one, column, exp_value, title, binwidth = 0.2) {

  df %>%
    filter(group1 == group_one) %>%
    ggplot(aes(!! sym(column))) +
    geom_histogram(binwidth = 0.2, color = "black",
                   fill = "grey80") +
    geom_segment(aes(x = exp_value, xend = exp_value, y = 70, yend = 0),
                 arrow = arrow(length = unit(0.5, "cm"))) +
    annotate("text", x = exp_value, y = 75, label = "Expected value", angle = 90, size = 3, hjust = 0) +
    ggtitle(title)
}

# Run functions
complete_data_below <- run_iterations(run_df,
                                runs = 1000,
                                method = "below",
                                dist_value = 0.1) %>%
  bind_rows()

complete_data_above <- run_iterations(run_df,
                                runs = 1000,
                                method = "above",
                                dist_value = 0.8) %>%
  bind_rows()

# Plots

p1 <- ggplot(complete_data_below, aes(Percent_below)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
                fill = "grey80") +
  stat_function(fun = dnorm,
                args = with(complete_data_below, c(mean = mean(Percent_below), sd = sd(Percent_below))),
                color = "red") +
  geom_segment(aes(x = complete_data_below$Percent_below[1],
                   xend = complete_data_below$Percent_below[1], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#4575b4") +
  geom_segment(aes(x = complete_data_below$Percent_below[2],
                   xend = complete_data_below$Percent_below[2], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#74add1") +
  geom_segment(aes(x = complete_data_below$Percent_below[3],
                   xend = complete_data_below$Percent_below[3], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#f46d43") +
  geom_segment(aes(x = complete_data_below$Percent_below[4],
                   xend = complete_data_below$Percent_below[4], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#fdae61") +
  geom_vline(xintercept = mean(complete_data_below$Percent_below),
             size = 1)

p2 <- ggplot(complete_data_above, aes(Percent_above)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
                fill = "grey80") +
  stat_function(fun = dnorm,
                args = with(complete_data_above, c(mean = mean(Percent_above), sd = sd(Percent_above))),
                color = "red") +
  geom_segment(aes(x = complete_data_above$Percent_above[1],
                   xend = complete_data_above$Percent_above[1], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#4575b4") +
  geom_segment(aes(x = complete_data_above$Percent_above[2],
                   xend = complete_data_above$Percent_above[2], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#74add1") +
  geom_segment(aes(x = complete_data_above$Percent_above[3],
                   xend = complete_data_above$Percent_above[3], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#f46d43") +
  geom_segment(aes(x = complete_data_above$Percent_above[4],
                   xend = complete_data_above$Percent_above[4], y = 0.15, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#fdae61") +
  geom_vline(xintercept = mean(complete_data_above$Percent_above),
             size = 1)

plot_grid(p1, p2, ncol = 2, nrow = 1, align = "v")

mean_data_below <- complete_data_below %>%
  mutate(mean = mean(Percent_below),
         dist_from_mean = mean - Percent_below)

mp1 <- ggplot(mean_data_below, aes(dist_from_mean))+
  geom_histogram(binwidth = 0.15,
                 color = "black") +
  geom_segment(aes(x = mean_data_below$dist_from_mean[1],
                   xend = mean_data_below$dist_from_mean[1], y = 100, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#4575b4") +
  geom_segment(aes(x = mean_data_below$dist_from_mean[2],
                   xend = mean_data_below$dist_from_mean[2], y = 100, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#74add1") +
  geom_segment(aes(x = mean_data_below$dist_from_mean[3],
                   xend = mean_data_below$dist_from_mean[3], y = 100, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#f46d43") +
  geom_segment(aes(x = mean_data_below$dist_from_mean[4],
                   xend = mean_data_below$dist_from_mean[4], y = 100, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#fdae61")

mean_data_above <- complete_data_above %>%
  mutate(mean = mean(Percent_above),
         dist_from_mean = mean - Percent_above)

mp2 <- ggplot(mean_data_above, aes(dist_from_mean))+
  geom_histogram(binwidth = 0.15,
                 color = "black") +
  geom_segment(aes(x = mean_data_above$dist_from_mean[1],
                   xend = mean_data_above$dist_from_mean[1], y = 50, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#4575b4") +
  geom_segment(aes(x = mean_data_above$dist_from_mean[2],
                   xend = mean_data_above$dist_from_mean[2], y = 50, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#74add1") +
  geom_segment(aes(x = mean_data_above$dist_from_mean[3],
                   xend = mean_data_above$dist_from_mean[3], y = 50, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#f46d43") +
  geom_segment(aes(x = mean_data_above$dist_from_mean[4],
                   xend = mean_data_above$dist_from_mean[4], y = 50, yend = 0),
               arrow = arrow(length = unit(0.4, "cm"),
                             type = "closed"),
               color = "#fdae61")

plot_grid(mp1, mp2, ncol = 2, nrow = 1, align = "v")

chsq_data_below <- complete_data_below[1:4,] %>%
  mutate(Total = Above + Below)

br_pi_csq <- chsq_data_below[1:2,] %>%
  select(Total, Below) %>%
  as.matrix

br_rf_csq <- chsq_data_below[c(1,3),] %>%
  select(Total, Below) %>%
  as.matrix

br_wb_csq <- chsq_data_below[c(1,4),] %>%
  select(Total, Below) %>%
  as.matrix

pi_rf_csq <- chsq_data_below[c(2,3),] %>%
  select(Total, Below) %>%
  as.matrix

pi_wb_csq <- chsq_data_below[c(2,4),] %>%
  select(Total, Below) %>%
  as.matrix

rf_wb_csq <- chsq_data_below[c(3,4),] %>%
  select(Total, Below) %>%
  as.matrix

chisq.test(br_pi_csq, correct = FALSE)
chisq.test(br_rf_csq, correct = FALSE)
chisq.test(br_wb_csq, correct = FALSE)
chisq.test(pi_rf_csq, correct = FALSE)
chisq.test(pi_wb_csq, correct = FALSE)
chisq.test(rf_wb_csq, correct = FALSE)

```

### 2019-03-27 notebooks/Clustering analysis.Rmd
minor changes

@@
 library(cowplot)
+library(knitr)
+library(png)
 ```
 
-# cgMLST Schema and quality control
+# Core genome MLST schema and quality control
+The _E. coli_ scheme from Enterobase [1] was used with chewBBACA [2] to get the cgMLST profile for each isolate.
+The number of missing loci in total (sum of LNF, NIPH, PLOT, ALM and ASM from chewBBACA) is plotted below for all isolates.
+
 ```{r import data}
@@
 
+detailed_cgmlst_stats <- read.table("../data/chewbbaca/results_statistics.tsv",
+                                    sep = "\t",
+                                    header = TRUE,
+                                    stringsAsFactors = FALSE)
+
 cgMLST_allele_data <- read.table("../data/chewbbaca/complete_results/cgMLST.tsv",
@@
 
-```{r Number of missing loci}
+```{r cgMLST stats}
 boxplot(cgmlst_stats$number.of.missing.data,
@@
 
+Below follows the detailed results for all loci in the cgMLST scheme.
+EXC = excact matches, INF = novel allele, LNF = locus not found, NIPH = paralogous locus, PLOT = possible loci on tip of contig, ALM = allele 20 % larger than reference, ASM = allele 20 % smaller than reference.
+The numbers below each boxplot denote the mean.
+
+```{r}
+test <- detailed_cgmlst_stats %>%
+  gather(metric, value, -Genome) %>%
+  group_by(metric) %>%
+  mutate(mean_val = round(mean(value), 1))
+
+ggplot(test, aes(metric, value)) +
+  stat_boxplot(geom = "errorbar", width = 0.5) +
+  geom_boxplot() +
+  geom_text(data = subset(test,
+                          Genome == "2006-01-2184_S30_pilon_spades.fasta"),
+            aes(label = mean_val),
+            y = -60,
+            size = 3.8) +
+  scale_x_discrete(limits = c("EXC",
+                              "INF",
+                              "LNF",
+                              "NIPH",
+                              "PLOT",
+                              "ALM",
+                              "ASM")) +
+  labs(y = "Number of loci") +
+  theme(axis.title.x = element_blank(),
+        axis.ticks.x = element_blank())
+```
+
 # Total dendrogram of all isolates
+Genetic distances were calculated from the allele data from the cgMLST report, and a dendrogram was drawn from these distances.
+The annotations are based on data from the ARIBA resistance gene analysis.
 ```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
@@
                             sep = "\t",
-                            header = TRUE)
+                            header = TRUE) %>%
+  select(-gadX)
 
@@
 
+# Dendrograms per animal species
 ```{r, fig.height=25, fig.width=25, message=FALSE, warning=FALSE}
@@
               "1" = "#440154FF",
-              "2" = "#471164FF",
-              "3" = "#481F70FF",
-              "4" = "#472D7BFF",
-              "5" = "#443A83FF",
-              "6" = "#404688FF",
-              "7" = "#3B528BFF",
-              "8" = "#365D8DFF",
-              "9" = "#31688EFF",
-              "10" = "#2C728EFF",
-              "11" = "#287C8EFF",
-              "12" = "#24868EFF",
-              "13" = "#21908CFF",
-              "14" = "#1F9A8AFF",
-              "15" = "#20A486FF",
-              "16" = "#27AD81FF",
-              "17" = "#35B779FF",
-              "18" = "#47C16EFF",
-              "19" = "#5DC863FF",
-              "20" = "#75D054FF",
-              "21" = "#8FD744FF",
-              "22" = "#AADC32FF",
-              "23" = "#C7E020FF",
-              "24" = "#E3E418FF",
-              "25" = "#FDE725FF")
+              "2" = "#3B528BFF",
+              "3" = "#21908CFF",
+              "4" = "#5DC863FF",
+              "5" = "#FDE725FF")
 
@@
          metadata_species,
-         offset = 0.09,
+         offset = 0.1,
          width = 0.5,
@@
                                  size = 0.5) %<+% tree_metadata +
-  geom_tippoint(color = "#f46d43",
+  geom_tippoint(color = "#fdae61",
                 size = 4) +
@@
           align = "h")
-tot_plot
 
+include_graphics("../figures/species_dendrogram.png")
 ```
+
+# Statistics
+## Distribution of calculated distances per animal species
+Below follows a plot of the distribution of the lower triangular distances for each animal species.
+A bimodal distribution is present in all four plots.
+The filtering values of < 0.1 and > 0.8 is used for downstream statistics to acertain whether the broiler isolates are more homogenous than the isolates from other species.
 
@@
 
-# Statistics
-```{r, warning=FALSE, message=FALSE}
+## Distribution of iterated distances
+Distribution of percentages of distances < 0.1 or > 0.8.
+The distribution is based on 1000 iterations of the distance data.
+Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird.
+The mean value is denoted as a vertical black line.
+```{r, warning=FALSE, message=FALSE, dpi=300}
+# generate data frame from allele data
 alleles <- cgMLST_allele_data %>%
@@
 
+# generate data frame for iteration analysis
 run_df <- lower_tri_dist_calc(alleles, data_frame = TRUE) %>%
@@
 
-below_calc <- function(df, run, dist_value = 0.1) {
+# functions used in iteration analyses
+
+## Calculates the sum of observations lower than dist_value1 and
+## bigger than dist_value2, and calculates the percentage of
+## observed values for each group
+without <- function(df, run, dist_value1 = 0.1, dist_value2 = 0.8) {
   df <- df %>%
     group_by(group1, group2) %>%
-    mutate(n = value < dist_value) %>%
+    mutate(n = value < dist_value1 | value > dist_value2) %>%
+    group_by(group1, group2, n) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(n = ifelse(n == TRUE, "Within", "Without")) %>%
+    spread(n, nn) %>%
+    mutate(Percent = round(Within / (Within + Without) * 100, 2),
+           iter = run)
+  return(df)
+}
+
+## Calculates the sum of observations between dist_value1 and dist_value2
+## and calculates the percentage of observed values for each group
+within <- function(df, run, dist_value1 = 0.1, dist_value2 = 0.8) {
+  df <- df %>%
+    group_by(group1, group2) %>%
+    mutate(n = value > dist_value1 & value < dist_value2) %>%
+    group_by(group1, group2, n) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(n = ifelse(n == TRUE, "Within", "Without")) %>%
+    spread(n, nn) %>%
+    mutate(Percent = round(Within / (Within + Without) * 100, 2),
+           iter = run)
+  return(df)
+}
+
+## Calculates the sum of observations below dist_value1
+## for each group
+below_calc <- function(df, run, dist_value1 = 0.1) {
+  df <- df %>%
+    group_by(group1, group2) %>%
+    mutate(n = value < dist_value1) %>%
     group_by(group1, group2, n) %>%
@@
 
-below_calc2 <- function(df, run, dist_value = 0.1) {
+## Calculates the sum of observations above dist_value1
+## for each group
+above_calc <- function(df, run, dist_value1 = 0.9) {
   df <- df %>%
     group_by(group1, group2) %>%
-    mutate(n = n()) %>%
-    sample_n(min(.$n)) %>%
-    select(-n) %>%
-    mutate(n = value < dist_value) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Below", "Above")) %>%
-    spread(n, nn) %>%
-    mutate(Below = ifelse(is.na(Below) == TRUE, 0, Below)) %>%
-    mutate(Percent_below = round(Below / (Above + Below) * 100, 2),
-           iter = run)
-  return(df)
-}
-
-above_calc <- function(df, run, dist_value = 0.9) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = value > dist_value) %>%
+    mutate(n = value > dist_value1) %>%
     group_by(group1, group2, n) %>%
@@
 
-above_calc2 <- function(df, run, dist_value = 0.9) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = n()) %>%
-    sample_n(min(.$n)) %>%
-    select(-n) %>%
-    mutate(n = value > dist_value) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Above", "Below")) %>%
-    spread(n, nn) %>%
-    mutate(Above = ifelse(is.na(Above) == TRUE, 0, Above)) %>%
-    mutate(Percent_above = round(Above / (Above + Below) * 100, 2),
-           iter = run)
-  return(df)
-}
-
+## Calculates the mean of observations for each group
 mean_calc <- function(df, run) {
@@
 
+## Calculates the median of observations for each group
 median_calc <- function(df, run) {
@@
 
+## Calculates the quantiles of the observations for each group
 quantile_calc <- function(df, run) {
@@
 
-iterate_samples <- function(df, run, method = "mean", dist_value) {
+# Iteration function
+iterate_samples <- function(df, run, method = "mean", dist_value1, dist_value2) {
 
+  # randomize the value column, regardless of groups
   df <- df %>%
@@
 
+  # run specified function
   if (method == "mean") {
@@
   if (method == "below") {
-     df <- below_calc2(df, run, dist_value)
+   df <- below_calc(df, run, dist_value1)
     }
@@
   if (method == "above") {
-     df <- above_calc2(df, run, dist_value)
+   df <- above_calc(df, run, dist_value1)
+    }
+
+  if (method == "within") {
+    df <- within(df, run, dist_value1, dist_value2)
+    }
+
+  if (method == "without") {
+    df <- without(df, run, dist_value1, dist_value2)
     }
@@
 # run iterations on selected data frame
-run_iterations <- function(df, runs = 1000, seed = 10, method = "mean", dist_value) {
+run_iterations <- function(df, runs = 1000, seed = 10, method = "mean", dist_value1, dist_value2) {
   # set seed
@@
   if (method == "below") {
-     orig <- below_calc(df, 0, dist_value)
+   orig <- below_calc(df, 0, dist_value1)
     }
@@
   if (method == "above") {
-     orig <- above_calc(df, 0, dist_value)
+   orig <- above_calc(df, 0, dist_value1)
+    }
+
+  if (method == "within") {
+    orig <- within(df, 0, dist_value1, dist_value2)
+    }
+
+  if (method == "without") {
+    orig <- without(df, 0, dist_value1, dist_value2)
     }
@@
                               method = method,
-                              dist_value = dist_value)
+                              dist_value1 = dist_value1,
+                              dist_value2 = dist_value2)
     output <- c(output, list(result))
@@
 
-# create distribution plots from iteration data
-create_plot <- function(df, group_one, column, exp_value, title, binwidth = 0.2) {
+# Run functions
+summary_stats <- run_df %>%
+  mutate(total_mean = round(mean(value)*100, 3)) %>%
+  group_by(group1) %>%
+  mutate(mean = round(mean(value)*100, 3),
+         var = round(var(value)*100, 3)) %>%
+  select(group1, total_mean, mean, var) %>%
+  summarise_all(funs(func_paste))
 
-  df %>%
-    filter(group1 == group_one) %>%
-    ggplot(aes(!! sym(column))) +
-    geom_histogram(binwidth = 0.2, color = "black",
-                   fill = "grey80") +
-    geom_segment(aes(x = exp_value, xend = exp_value, y = 70, yend = 0),
-                 arrow = arrow(length = unit(0.5, "cm"))) +
-    annotate("text", x = exp_value, y = 75, label = "Expected value", angle = 90, size = 3, hjust = 0) +
-    ggtitle(title)
-}
+summary_stats2 <- run_df %>%
+  mutate(total_mean = round(mean(value), 3)) %>%
+  group_by(group1) %>%
+  mutate(mean = round(mean(value), 3),
+         var = round(var(value), 3)) %>%
+  select(group1, total_mean, mean, var) %>%
+  summarise_all(funs(func_paste))
 
-# Run functions
-complete_data_below <- run_iterations(run_df,
+complete_data <- run_iterations(run_df,
+                                runs = 1000,
+                                method = "without",
+                                dist_value1 = 0.1,
+                                dist_value2 = 0.8) %>%
+  bind_rows() %>%
+  mutate(Percent = as.numeric(Percent))
+
+lower_data <- run_iterations(run_df,
                              runs = 1000,
                              method = "below",
-                                dist_value = 0.1) %>%
-  bind_rows()
-
-complete_data_above <- run_iterations(run_df,
-                                runs = 1000,
-                                method = "above",
-                                dist_value = 0.8) %>%
-  bind_rows()
+                             dist_value1 = 0.1) %>%
+  bind_rows() %>%
+  mutate(Percent_below = as.numeric(Percent_below))
 
@@
 
-p1 <- ggplot(complete_data_below, aes(Percent_below)) +
+ggplot(complete_data, aes(Percent)) +
   geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
@@
   stat_function(fun = dnorm,
-                args = with(complete_data_below, c(mean = mean(Percent_below), sd = sd(Percent_below))),
+                args = with(complete_data, c(mean = mean(Percent), sd = sd(Percent))),
                 color = "red") +
-  geom_segment(aes(x = complete_data_below$Percent_below[1],
-                   xend = complete_data_below$Percent_below[1], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = complete_data$Percent[1],
+                   xend = complete_data$Percent[1], y = 0.08, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#4575b4") +
-  geom_segment(aes(x = complete_data_below$Percent_below[2],
-                   xend = complete_data_below$Percent_below[2], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = complete_data$Percent[2],
+                   xend = complete_data$Percent[2], y = 0.08, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#74add1") +
-  geom_segment(aes(x = complete_data_below$Percent_below[3],
-                   xend = complete_data_below$Percent_below[3], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = complete_data$Percent[3],
+                   xend = complete_data$Percent[3], y = 0.08, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#f46d43") +
-  geom_segment(aes(x = complete_data_below$Percent_below[4],
-                   xend = complete_data_below$Percent_below[4], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = complete_data$Percent[4],
+                   xend = complete_data$Percent[4], y = 0.08, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#fdae61") +
-  geom_vline(xintercept = mean(complete_data_below$Percent_below),
-             size = 1)
+  geom_text(label = "p < 0.001",
+            x = 83.30,
+            y = 0.16,
+            angle = 90) +
+  geom_vline(xintercept = as.numeric(summary_stats$mean[1]),
+             size = 1,
+             color = "#4575b4") +
+  geom_vline(xintercept = as.numeric(summary_stats$mean[2]),
+             size = 1,
+             color = "#74add1") +
+  geom_vline(xintercept = as.numeric(summary_stats$mean[3]),
+             size = 1,
+             color = "#f46d43") +
+  geom_vline(xintercept = as.numeric(summary_stats$mean[4]),
+             size = 1,
+             color = "#fdae61") +
+  labs(x = "Percent (%) distances < 0.1 or > 0.8",
+       y = "Density") +
+  ggtitle(label = "Distribution of percentages",
+          subtitle = "Number of iterations: 1000") +
+  theme(plot.title = element_text(hjust = 0))
 
-p2 <- ggplot(complete_data_above, aes(Percent_above)) +
+ggplot(lower_data, aes(Percent_below)) +
   geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
@@
   stat_function(fun = dnorm,
-                args = with(complete_data_above, c(mean = mean(Percent_above), sd = sd(Percent_above))),
+                args = with(lower_data, c(mean = mean(Percent_below), sd = sd(Percent_below))),
                 color = "red") +
-  geom_segment(aes(x = complete_data_above$Percent_above[1],
-                   xend = complete_data_above$Percent_above[1], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = lower_data$Percent_below[1],
+                   xend = lower_data$Percent_below[1], y = 0.15, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#4575b4") +
-  geom_segment(aes(x = complete_data_above$Percent_above[2],
-                   xend = complete_data_above$Percent_above[2], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = lower_data$Percent_below[2],
+                   xend = lower_data$Percent_below[2], y = 0.15, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#74add1") +
-  geom_segment(aes(x = complete_data_above$Percent_above[3],
-                   xend = complete_data_above$Percent_above[3], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = lower_data$Percent_below[3],
+                   xend = lower_data$Percent_below[3], y = 0.15, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#f46d43") +
-  geom_segment(aes(x = complete_data_above$Percent_above[4],
-                   xend = complete_data_above$Percent_above[4], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
+  geom_segment(aes(x = lower_data$Percent_below[4],
+                   xend = lower_data$Percent_below[4], y = 0.15, yend = 0),
+               arrow = arrow(length = unit(0.3, "cm"),
                              type = "closed"),
                color = "#fdae61") +
-  geom_vline(xintercept = mean(complete_data_above$Percent_above),
-             size = 1)
+  geom_text(label = "p < 0.001",
+            x = 5.87,
+            y = 0.32,
+            angle = 90) +
+  geom_vline(xintercept = mean(lower_data$Percent_below),
+             size = 1) +
+  labs(x = "Percent (%) distances < 0.1",
+       y = "Density") +
+  ggtitle(label = "Distribution of Percentages",
+          subtitle = "Number of iterations: 1000") +
+  theme(plot.title = element_text(hjust = 0))
 
-plot_grid(p1, p2, ncol = 2, nrow = 1, align = "v")
-
-mean_data_below <- complete_data_below %>%
-  mutate(mean = mean(Percent_below),
-         dist_from_mean = mean - Percent_below)
-
-mp1 <- ggplot(mean_data_below, aes(dist_from_mean))+
-  geom_histogram(binwidth = 0.15,
-                 color = "black") +
-  geom_segment(aes(x = mean_data_below$dist_from_mean[1],
-                   xend = mean_data_below$dist_from_mean[1], y = 100, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#4575b4") +
-  geom_segment(aes(x = mean_data_below$dist_from_mean[2],
-                   xend = mean_data_below$dist_from_mean[2], y = 100, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#74add1") +
-  geom_segment(aes(x = mean_data_below$dist_from_mean[3],
-                   xend = mean_data_below$dist_from_mean[3], y = 100, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#f46d43") +
-  geom_segment(aes(x = mean_data_below$dist_from_mean[4],
-                   xend = mean_data_below$dist_from_mean[4], y = 100, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#fdae61")
-
-mean_data_above <- complete_data_above %>%
-  mutate(mean = mean(Percent_above),
-         dist_from_mean = mean - Percent_above)
-
-mp2 <- ggplot(mean_data_above, aes(dist_from_mean))+
-  geom_histogram(binwidth = 0.15,
-                 color = "black") +
-  geom_segment(aes(x = mean_data_above$dist_from_mean[1],
-                   xend = mean_data_above$dist_from_mean[1], y = 50, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#4575b4") +
-  geom_segment(aes(x = mean_data_above$dist_from_mean[2],
-                   xend = mean_data_above$dist_from_mean[2], y = 50, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#74add1") +
-  geom_segment(aes(x = mean_data_above$dist_from_mean[3],
-                   xend = mean_data_above$dist_from_mean[3], y = 50, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#f46d43") +
-  geom_segment(aes(x = mean_data_above$dist_from_mean[4],
-                   xend = mean_data_above$dist_from_mean[4], y = 50, yend = 0),
-               arrow = arrow(length = unit(0.4, "cm"),
-                             type = "closed"),
-               color = "#fdae61")
-
-plot_grid(mp1, mp2, ncol = 2, nrow = 1, align = "v")
-
-chsq_data_below <- complete_data_below[1:4,] %>%
-  mutate(Total = Above + Below)
-
-br_pi_csq <- chsq_data_below[1:2,] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-br_rf_csq <- chsq_data_below[c(1,3),] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-br_wb_csq <- chsq_data_below[c(1,4),] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-pi_rf_csq <- chsq_data_below[c(2,3),] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-pi_wb_csq <- chsq_data_below[c(2,4),] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-rf_wb_csq <- chsq_data_below[c(3,4),] %>%
-  select(Total, Below) %>%
-  as.matrix
-
-chisq.test(br_pi_csq, correct = FALSE)
-chisq.test(br_rf_csq, correct = FALSE)
-chisq.test(br_wb_csq, correct = FALSE)
-chisq.test(pi_rf_csq, correct = FALSE)
-chisq.test(pi_wb_csq, correct = FALSE)
-chisq.test(rf_wb_csq, correct = FALSE)
+summary_stats2
 
 ```
+
+# References
+[1] Alikhan N-F, Zhou Z, Sergeant MJ, Achtman M.
+A genomic overview of the population structure of Salmonella.
+Casadesús J, editor.
+PLOS Genet [Internet].
+2018 Apr 5;14(4):e1007261.
+Available from: https://dx.plos.org/10.1371/journal.pgen.1007261
+
+[2] Machado MP, Silva M, Carriço JA, Silva DN, Santos S, Ramirez M, et al. chewBBACA: A complete suite for gene-by-gene schema creation and strain identification. Microb Genomics. 2018;1–7.

### 2019-03-27 notebooks/article_notebook2.Rmd
minor changes

@@
 # Libraries
-# WARNING: This will install packages on your system if not present in library path!
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, ape, cluster, tibble, Biostrings, seqinr, stringr, scales, reprex)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(cowplot)
+library(viridis)
+library(purrr)
+library(tibble)
+library(stringr)
+library(Biostrings)
+library(seqinr)
+library(scales)
+library(gridExtra)
 ```
@@
 mut_genes <- c("gyr","par","marR","marA",
-               "sox","rpoB","rob","gad")
+               "sox","rpoB","rob")
 
@@
     mutate(gene_names = as.character(gene_names),
-           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))
+           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref)) %>%
+    mutate(gene_names = ifelse(gene_names == "soxS", "soxR", gene_names))
 
@@
 
+## Percent mutations in each intrinsic gene per animal species
+```{r}
+mut_report %>%
+  select(-nt_change) %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  gather(gene, value, -c(id, species)) %>%
+  group_by(gene, species, value) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,1)) %>%
+  arrange(gene)
+```
+
 ## Specific mutations in QRDR
@@
   filter(n > 2) %>%
-  mutate(percent = case_when(
-    gene == "gadX" ~ 
-      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
+  mutate(Total = case_when(
     gene == "marA" ~
-      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
+      length(which(mut_report[, "marA"] == 1)),
     gene == "marR" ~
-      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
+      length(which(mut_report[, "marR"] == 1)),
     gene == "rpoB" ~
-      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
-    gene == "soxS" ~ 
-      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
+      length(which(mut_report[, "rpoB"] == 1)),
+    gene == "soxR" ~
+      length(which(mut_report[, "soxR"] == 1))),
+         percent = case_when(
+    gene == "marA" ~ 
+      round(n / Total*100, 1),
+    gene == "marR" ~ 
+      round(n / Total*100, 1),
+    gene == "rpoB" ~ 
+      round(n / Total*100, 1),
+    gene == "soxR" ~ 
+      round(n / Total*100, 1))) %>%
   arrange(gene, desc(percent)) %>%
   dplyr::rename("Gene" = gene,
-                "Mutation" = value,
+                "Substitution" = value,
                 "Percent" = percent)
@@
   # summarise qnr-genes
-  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
   left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
@@
 
-## Double-check ARIBA results
-This chunk double-checks the mutations found in _gyrA_, _parC_ and _parE_ against an internal mutation analysis script.
-The sequences are taken from Prokka, and reference wild-type versions of the genes are used to identify mismatches with the sample sequences.
-```{r, warning=FALSE, message=FALSE}
+```{r, include=FALSE}
 # Functions
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
+  mutate(qnr = ifelse(rowSums(.[,13:18]) == 0, 0, 1)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
@@
     mutate(species = factor(species,
-                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
-    ggplot(aes(reorder(key, -Percent), Percent, fill = species)) +
+                            levels = c("Broiler","Pig","Wild bird","Red fox")),
+           key = paste0(toupper(substr(key, 1, 1)),
+                        substr(key, 2, nchar(key))),
+           group = if_else(key %in% c("GyrA","GyrB","ParC","ParE"), 1, 2)) %>%
+  ggplot(aes(key, Percent, fill = species)) +
     geom_col(color = "black",
@@
     scale_fill_manual(values = palette) +
+    scale_x_discrete(labels = c("GyrA","GyrB","ParC","ParE",
+                                "MarA","MarR","SoxR","RpoB",
+                                "RobA")) +
     labs(y = "Percent (%) of isolates") +
@@
              position = position_dodge(0.9)) +
+    geom_errorbar(
+      aes(ymin = lwr, ymax = upr),
+      width = 0.6,
+      position = position_dodge(0.9)
+    ) +
     scale_fill_manual(values = palette) +
@@
       axis.text = element_text(size = 12),
+      axis.text.x = element_text(face = "italic"),
       axis.title.y = element_text(size = 14),
@@
                           gene1 == "rpoB" ~ 7,
-                          gene1 == "soxS" ~ 8,
+                          gene1 == "soxR" ~ 8,
                           gene1 == "robA" ~ 9,
-                          gene1 == "gadX" ~ 10,
-                          gene1 == "qnrA1" ~ 11,
-                          gene1 == "qnrB19" ~ 12,
-                          gene1 == "qnrB6" ~ 13,
-                          gene1 == "qnrB60" ~ 14,
-                          gene1 == "qnrS1" ~ 15,
-                          gene1 == "qnrS2" ~ 16,
-                          gene1 == "qnrS4" ~ 17,
-                          gene1 == "qnr" ~ 18,
-                          gene1 == "qepA4" ~ 19),
+                          gene1 == "qnrA1" ~ 10,
+                          gene1 == "qnrB19" ~ 11,
+                          gene1 == "qnrB6" ~ 12,
+                          gene1 == "qnrB60" ~ 13,
+                          gene1 == "qnrS1" ~ 14,
+                          gene1 == "qnrS2" ~ 15,
+                          gene1 == "qnrS4" ~ 16,
+                          gene1 == "qnr" ~ 17,
+                          gene1 == "qepA4" ~ 18),
            type2 = case_when(gene2 == "gyrA" ~ 1,
@@
                           gene2 == "rpoB" ~ 7,
-                          gene2 == "soxS" ~ 8,
+                          gene2 == "soxR" ~ 8,
                           gene2 == "robA" ~ 9,
-                          gene2 == "gadX" ~ 10,
-                          gene2 == "qnrA1" ~ 11,
-                          gene2 == "qnrB19" ~ 12,
-                          gene2 == "qnrB6" ~ 13,
-                          gene2 == "qnrB60" ~ 14,
-                          gene2 == "qnrS1" ~ 15,
-                          gene2 == "qnrS2" ~ 16,
-                          gene2 == "qnrS4" ~ 17,
-                          gene2 == "qnr" ~ 18,
-                          gene2 == "qepA4" ~ 19))
+                          gene2 == "qnrA1" ~ 10,
+                          gene2 == "qnrB19" ~ 11,
+                          gene2 == "qnrB6" ~ 12,
+                          gene2 == "qnrB60" ~ 13,
+                          gene2 == "qnrS1" ~ 14,
+                          gene2 == "qnrS2" ~ 15,
+                          gene2 == "qnrS4" ~ 16,
+                          gene2 == "qnr" ~ 17,
+                          gene2 == "qepA4" ~ 18))
 
@@
 
+## Mechanisms per animal species
+```{r, fig.height=10, fig.width=15, warning=FALSE, message=FALSE}
+AddCol <- function(df, col_name) {
+  # split rows by delimeters
+  string_to_proc <- df %>% select(!!col_name) %>%
+    unlist() %>% str_split(regex("\\, |\\,")) 
+  # find unique entries
+  unique_strings <- string_to_proc %>%
+    unlist() %>% unique()
+  # construct names of the new columns
+  cols_names <- paste(col_name, unique_strings, sep = "_")
+  # construct 0/1-content columns for each unique entry
+  cols_content <- sapply(function(i) {
+    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
+                             X = string_to_proc)))
+  }, X = seq_along(unique_strings))
+  res <- data.frame(cols_content)
+  names(res) <- cols_names
+  return(res)
+}
+
+total_df <- acquired_report %>%
+  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
+  left_join(mut_report[c("id","marR","marA","rpoB","soxR","robA")], by = "id") %>%
+  left_join(isolate_data[c("id", "species")])
+
+names(total_df) <- c("id","qepA4","qnrA1","qnrB19",
+                     "qnrB6","qnrB60","qnrS1","qnrS2",
+                     "qnrS4","gyrA","gyrB","parC",
+                     "parE","MarR","MarA","RpoB",
+                     "SoxR","RobA","species")
+
+col_proc <- c("gyrA","gyrB","parC","parE")
+
+id <- total_df$id
+
+cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
+                      X = seq_along(col_proc)) %>%
+  bind_cols() %>%
+  cbind(id) %>%
+  gather(key, value, -id) %>%
+  filter(grepl("(.*?)_0", key) == FALSE) %>%
+  spread(key, value)
+
+total_df2 <- total_df %>%
+  select(-c(gyrA, gyrB, parC, parE)) %>%
+  left_join(cols_to_add, by = "id") %>%
+  select(id, species, everything()) %>%
+  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
+            funs(as.numeric(.))) %>%
+  mutate(qnr = ifelse(rowSums(.[,4:10]) == 0, 0, 1)) %>%
+  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  select(-id) %>%
+  mutate_at(vars(-species, CIP),
+            funs(as.numeric(as.character(.)))) %>%
+  group_by(species, CIP) %>%
+  summarise_all(funs(sum, n = n())) %>%
+  dplyr::rename("n" = MarR_n) %>%
+  select(-contains("_n")) %>%
+  gather(gene, value, -c(species, CIP, n)) %>%
+  mutate(gene2 = sub("(.*?)_(.*?)_sum", "\\1", gene),
+         gene2 = sub("(.*?)_sum", "\\1", gene2),
+         gene = sub("(.*?)_(.*?)_sum", "\\2", gene),
+         gene = sub("(.*?)_sum", "\\1", gene),
+         type = case_when(gene2 == "gyrA" ~ 1,
+                          gene2 == "gyrB" ~ 2,
+                          gene2 == "parC" ~ 3,
+                          gene2 == "parE" ~ 4,
+                          gene2 == "MarR" ~ 5,
+                          gene2 == "MarA" ~ 6,
+                          gene2 == "RpoB" ~ 7,
+                          gene2 == "SoxR" ~ 8,
+                          gene2 == "RobA" ~ 9,
+                          gene2 == "qnr" ~ 10,
+                          gene2 == "qepA4" ~ 11),
+         perc = value/n*100)
+
+p1 <- ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
+  geom_tile()+
+  scale_fill_viridis(option = "A",
+                     begin = 0.1,
+                     guide = guide_colorbar(barheight = 10,
+                                            barwidth = 0.8,
+                                            label.position = "left",
+                                            label.hjust = 0.5,
+                                            label.vjust = 0.3,
+                                            ticks = TRUE))+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        strip.background = element_rect(colour = "black", fill = "grey70"),
+        strip.text = element_text(size = 12),
+        legend.position = "left",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 10),
+        plot.margin = unit(c(0,0,0,0), "cm"))+
+  labs(y = "CIP MIC")+
+  facet_wrap(~species, ncol = 1)+
+  coord_fixed()
+
+nplot_data <- total_df %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(species, CIP) %>%
+  dplyr::count() %>%
+  ungroup()
+
+n_names <- c("Broiler" = "n",
+             "Pig" = "n",
+             "Red fox" = "n",
+             "Wild bird" = "n")
+
+p2 <- ggplot(nplot_data, aes(factor(CIP), n))+
+  geom_col()+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_text(size = 10),
+        axis.ticks.y = element_blank(),
+        strip.text = element_text(size = 12, face = "italic"),
+        strip.background = element_rect(colour = "black", fill = "grey70"),
+        plot.margin = unit(c(0,0,-10,0), "cm"))+
+  facet_wrap(~species, ncol = 1, labeller = as_labeller(n_names))+
+  coord_flip()
+
+p1g <- ggplotGrob(p1)
+p2g <- ggplotGrob(p2)
+
+g <- cbind(p1g, p2g, size = "first")
+
+grid.arrange(g)
+```
+
 ## MIC-plots
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
   select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
@@
             funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
+  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
   select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
@@
   geom_point(size = 2)+
-  scale_y_discrete(limits = c("gyrA","parC","parE","gadX","marA","marR","rpoB","soxS","qepA4","qnr"))+
+  scale_y_discrete(limits = c("gyrA","parC","parE","marA","marR","rpoB","soxR","qepA4","qnr"))+
   guides(fill = FALSE)+

### 2019-03-27 figure_script.R

@@ -1,9 +1,266 @@
 ## Script for generating figures for the article
 
+## Total dendrogram
+cgMLST_data <- read.table("data/chewbbaca/complete_results/clean_cgMLST_data.txt",
+                          sep = "\t",
+                          header = TRUE,
+                          colClasses = "factor")
+
+total_tree <- as.phylo(hclust(daisy(cgMLST_data, metric = "gower"), method = "average"))
+
+
+tree_metadata <- read.table("data/chewbbaca/complete_results/tree_metadata.txt",
+                            sep = "\t",
+                            header = TRUE)
+
+tree_heatmap_data <- read.table("data/chewbbaca/complete_results/tree_heatmap_data.txt",
+                                sep = "\t",
+                                header = TRUE) %>%
+  select(-gadX)
+
+names(tree_heatmap_data) <- c("GyrA","ParC","ParE","RpoB","MarA","MarR","SoxR","qepA4","qnr")
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+palette2 <- c("1-A" = "#fc8d59",
+              "1-I" = "#80b1d3",
+              "0" = "grey95")
+
+total_tree_annotated <- ggtree(total_tree,
+                               layout = "circular",
+                               color = "#808080",
+                               size = 0.5) %<+% tree_metadata +
+  geom_tippoint(aes(color = species),
+                size = 1.5) +
+  geom_tiplab2(aes(label = ST),
+               offset = 0.01,
+               size = 2.5) +
+  scale_color_manual(values = palette)
+
+total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)
+
+ggsave(
+  "C:/Users/vi1511/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/total_dendrogram.tiff",
+  gheatmap(
+    total_tree_opened,
+    tree_heatmap_data,
+    offset = 0.04,
+    width = 0.3,
+    colnames_offset_y = 2.5,
+    colnames_position = "top",
+    font.size = 3
+  ) +
+    scale_fill_manual(
+      values = palette2,
+      labels = c(
+        "Gene/Mutation absent",
+        "Acquired gene present",
+        "Intrinsic gene mutated"
+      )
+    ) +
+    theme(
+      legend.position = c(0.9, 0.9),
+      legend.text = element_text(size = 10)
+    ),
[…]

### 2019-03-27 notebooks/Clustering analysis.Rmd
removed erroneous mean values from plot

@@
 summary_stats <- run_df %>%
-  mutate(total_mean = round(mean(value)*100, 3)) %>%
-  group_by(group1) %>%
-  mutate(mean = round(mean(value)*100, 3),
-         var = round(var(value)*100, 3)) %>%
-  select(group1, total_mean, mean, var) %>%
-  summarise_all(funs(func_paste))
-
-summary_stats2 <- run_df %>%
   mutate(total_mean = round(mean(value), 3)) %>%
@@
   mutate(Percent_below = as.numeric(Percent_below))
+
+summary_stats
 
@@
             angle = 90) +
-  geom_vline(xintercept = as.numeric(summary_stats$mean[1]),
-             size = 1,
-             color = "#4575b4") +
-  geom_vline(xintercept = as.numeric(summary_stats$mean[2]),
-             size = 1,
-             color = "#74add1") +
-  geom_vline(xintercept = as.numeric(summary_stats$mean[3]),
-             size = 1,
-             color = "#f46d43") +
-  geom_vline(xintercept = as.numeric(summary_stats$mean[4]),
-             size = 1,
-             color = "#fdae61") +
   labs(x = "Percent (%) distances < 0.1 or > 0.8",
@@
 
-summary_stats2
-
 ```

### 2019-03-28 figure_script.R
fixed erroneous x-axis

@@ -298,7 +298,7 @@ ggsave(
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
-   scale_x_discrete(labels = c("GyrA","GyrB","ParC","ParE",
+   scale_x_discrete(limits = c("GyrA","GyrB","ParC","ParE",
                                "MarA","MarR","SoxR","RpoB",
                                "RobA")) +
    labs(y = "Percent (%) of isolates") +

### 2019-03-28 notebooks/article_notebook2.Rmd
changed x-axis on barplot figure

@@
     scale_fill_manual(values = palette) +
-    scale_x_discrete(labels = c("GyrA","GyrB","ParC","ParE",
+    scale_x_discrete(limits = c("GyrA","GyrB","ParC","ParE",
                                 "MarA","MarR","SoxR","RpoB",

### 2019-04-24 scripts/ARIBA_analysis.R
initial commit

@@ -1,0 +1,460 @@
+# Libraries
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(purrr)
+
+# Report directories
+megares_report_loc <- "../data/ariba/megares_results"
+resfinder_report_loc <- "../data/ariba/resfinder_results"
+
+# Genes of interest. These genes are partially matched to 
+# the generated gene names from the "fix_gene_names" function
+ac_genes <- c("qnr","oqx","qep","aac6")
+mut_genes <- c("gyr","par","marR","marA",
+               "sox","rpoB","rob")
+
+# -------------------------- Functions ----------------------------
+
+# Paste unique cell values together, separate with ","
+func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
+                                 collapse = ",")
+
+
+# Confidence interval function
+get_binCI <- function(x, n) as.numeric(
+  setNames(
+    binom.test(x,n)$conf.int*100, 
+    c("lwr", "upr")
+  )
+)
+
+
+# Identifies filenames in input folder
+file_names <- function(filepath) {
+  files <- list.files(path = filepath, 
+                      pattern = "amr_report.tsv")
+  return(files)
+}
+
+
+# Import ariba data from report.tsv identified with file_names
+get_ariba_data <- function(filepath) {
+  # Identify file names in filepath
+  files <- file_names(filepath) 
+  
+  # Import data
+  data_list <- lapply(files,
+                      FUN = function(file) {
+                        read.delim(
+                          paste0(filepath, "/", file),
+                          stringsAsFactors = F,
+                          header = TRUE,
+                          sep = "\t"
+                        )
+                      })
+  
+  # Set file names in list
+  names(data_list) <- files
+  
+  # Bind to data frame and convert columns to character
+  data <- bind_rows(
+    lapply(
+      data_list,
+      function(x) map(x, as.character)
+    ),
+    .id = "ref")
+  
+  return(data)
+}
+
+
+# Corrects gene names 
+fix_gene_names <- function(df) {
+  # Identify unique entries in ref_name column
+  genes <- unique(df$ref_name)
+  
+  # Strip away excess characters
+  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
+  new_names <- gsub("_", "", new_names, fixed = T)
[…]

### 2019-04-24 scripts/chisq_tests.R

@@ -1,0 +1,124 @@
+# Chi-squared tests
+
+# function for doing chi-squared tests on multiple entries in a list
+do_chisq_test <- function(df) {
+  df <- df %>%
+    # create columns with all possible comparisons
+    dplyr::slice(expand.grid(1:length(unique(species)),
+                             1:length(unique(species))) %>%
+                   rev %>% 
+                   dplyr::filter(Var2 < Var1) %>% t) %>%
+    # create test ID
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    # specify values for chi-squred tests
+    dplyr::do(data_frame(gene = .$key,
+                         test = dplyr::first(.$test),
+                         species1 = dplyr::first(.$species),
+                         species2 = dplyr::last(.$species),
+                         data = list(matrix(c(.$Absent,
+                                              .$Present),
+                                            ncol = 2)))) %>%
+    # do chi-squared tests
+    mutate(chi_test = map(data, chisq.test,
+                          correct = FALSE)) %>%
+    mutate(p.value = round(map_dbl(chi_test, "p.value"),3),
+           df = map_dbl(chi_test, "parameter"),
+           x_sq = round(map_dbl(chi_test, "statistic"),2),
+           dupl = duplicated(test)) %>%
+    # filter out duplicated tests
+    dplyr::filter(dupl == FALSE) %>%
+    ungroup() %>%
+    rowwise() %>%
+    mutate(n = sum(unlist(data))) %>%
+    select(gene, species1, species2, n, df, x_sq, p.value) %>%
+    # round off p-value column
+    mutate(p.value = round(p.value, 5)) %>%
+    dplyr::filter(p.value <= 0.05)
+  return(df)
+}
+
+# create total mechanism table
+total_df <- acquired_report %>%
+  left_join(mut_report, by = "id") %>%
+  left_join(isolate_data[c("id", "species")], by = "id") %>%
+  select(-nt_change)
+
+# create list of genes with total presence/absence values
+# from each animal species
+mechanism_per_species <- total_df %>%
+  select(id, species, everything()) %>%
+  select(-id) %>%
+  gather(key, value, -species) %>%
+  group_by_all() %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
[…]

### 2019-04-24 notebooks/Clustering analysis.Rmd
minor changes

@@
 library(knitr)
-library(png)
+library(distanceR)
 ```
@@
   "2016-17-565-1-S",
+  "2016-02-415-2-S",
   "2016-02-428-2-S",
@@
                          header = TRUE,
-                         stringsAsFactors = FALSE)
+                         stringsAsFactors = FALSE) %>%
+  filter(FILE != "2016-02-415_S208_pilon_spades.fasta")
 
@@
                                     header = TRUE,
-                                    stringsAsFactors = FALSE)
+                                    stringsAsFactors = FALSE) %>%
+  filter(Genome != "2016-02-415_S208_pilon_spades.fasta")
 
@@
 
+names(tree_heatmap_data) <- c("GyrA","ParC","ParE","RpoB","MarA","MarR","SoxR","qepA4","qnr")
+
 palette <- c("Broiler" = "#4575b4",
@@
                                sep = "\t",
-                               header = TRUE)
+                               header = TRUE) %>%
+  select(gyrA, parC, parE, PMQR) %>%
+  dplyr::rename("GyrA" = gyrA,
+         "ParC" = parC,
+         "ParE" = parE)
 
@@
               # PMQR
-              "qnrA1" = "#762a83",
-              "qnrB19" = "#ef8a62",
-              "qnrS1" = "#a6dba0",
-              "qnrS2" = "#5aae61",
-              "qnrS4" = "#1b7837",
-              "qepA4" = "#80cdc1",
-              # num
-              "1" = "#440154FF",
-              "2" = "#3B528BFF",
-              "3" = "#21908CFF",
-              "4" = "#5DC863FF",
-              "5" = "#FDE725FF")
+              "qnrA1" = "#dadaeb",
+              "qnrB19" = "#bcbddc",
+              "qnrS1" = "#9e9ac8",
+              "qnrS2" = "#807dba",
+              "qnrS4" = "#6a51a3",
+              "qepA4" = "#4a1486")
 
 # Broiler tree
-broiler_cgMLST <- read.table("../data/chewbbaca/complete_results/broiler_cgmlst.txt",
-                   sep = "\t",
-                   header = TRUE,
-                   colClasses = "factor")
 
-broiler_tree <- as.phylo(hclust(daisy(broiler_cgMLST, metric = "gower"), method = "average"))
+broiler_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_broiler.tsv")
 
@@
   geom_tiplab2(aes(label = ST),
-               offset = 0.01,
+               offset = 0.02,
                size = 6)
 
-broiler_tree_opened <- rotate_tree(open_tree(broiler_tree_annotated, 10), 92.5)
-
-broiler_complete <- gheatmap(broiler_tree_opened,
+broiler_complete <- gheatmap(broiler_tree_annotated,
          metadata_species,
@@
          width = 0.5,
-         colnames_offset_y = 0.75,
-         colnames_position = "top",
-         font.size = 5) +
+         colnames = FALSE) +
   scale_fill_manual(values = palette3) +
@@
 # Pig tree
-pig_cgMLST <- read.table("../data/chewbbaca/complete_results/pig_cgmlst.txt",
-                   sep = "\t",
-                   header = TRUE,
-                   colClasses = "factor")
-
-pig_tree <- as.phylo(hclust(daisy(pig_cgMLST, metric = "gower"), method = "average"))
+pig_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_pig.tsv")
 
@@
   geom_tiplab2(aes(label = ST),
-               offset = 0.01,
+               offset = 0.02,
                size = 5.5)
 
-pig_tree_opened <- rotate_tree(open_tree(pig_tree_annotated, 10), 92.5)
-
-pig_complete <- gheatmap(pig_tree_opened,
+pig_complete <- gheatmap(pig_tree_annotated,
          metadata_species,
@@
          width = 0.5,
-         colnames_offset_y = 0.53,
-         colnames_position = "top",
-         font.size = 5) +
+         colnames = FALSE) +
   scale_fill_manual(values = palette3) +
@@
 # Red fox tree
-fox_cgMLST <- read.table("../data/chewbbaca/complete_results/fox_cgmlst.txt",
-                   sep = "\t",
-                   header = TRUE,
-                   colClasses = "factor")
-
-fox_tree <- as.phylo(hclust(daisy(fox_cgMLST, metric = "gower"), method = "average"))
+fox_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_fox.tsv")
 
@@
   geom_tiplab2(aes(label = ST),
-               offset = 0.01,
+               offset = 0.02,
                size = 5.5)
 
-fox_tree_opened <- rotate_tree(open_tree(fox_tree_annotated, 10), 92)
-
-fox_complete <- gheatmap(fox_tree_opened,
+fox_complete <- gheatmap(fox_tree_annotated,
          metadata_species,
@@
          width = 0.5,
-         colnames_offset_y = 0.17,
-         colnames_position = "top",
-         font.size = 5) +
+         colnames = FALSE) +
   scale_fill_manual(values = palette3) +
@@
 # Wild bird tree
-bird_cgMLST <- read.table("../data/chewbbaca/complete_results/bird_cgmlst.txt",
-                   sep = "\t",
-                   header = TRUE,
-                   colClasses = "factor")
-
-bird_tree <- as.phylo(hclust(daisy(bird_cgMLST, metric = "gower"), method = "average"))
+bird_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_bird.tsv")
 
@@
   geom_tiplab2(aes(label = ST),
-               offset = 0.01,
+               offset = 0.02,
                size = 5.5)
 
-bird_tree_opened <- rotate_tree(open_tree(bird_tree_annotated, 10), 92.5)
-
-bird_complete <- gheatmap(bird_tree_opened,
+bird_complete <- gheatmap(bird_tree_annotated,
          metadata_species,
@@
          width = 0.5,
-         colnames_offset_y = 0.38,
-         colnames_position = "top",
-         font.size = 5) +
+         colnames = FALSE) +
   scale_fill_manual(values = palette3) +
@@
           ncol = 2,
-          align = "h")
+          align = "h",
+          labels = c("Broiler","Pig","Red fox","Wild bird"))
 
-include_graphics("../figures/species_dendrogram.png")
+include_graphics("../figures/species_dendrogram2.png")
+```
+
+# Minimum spanning tree
+A minimum spanning tree calculated from the cgMLST data with GrapeTree [3] using MSTreeV2.
+Number of allele differences are denoted as numbers on branches between nodes.
+The branches have been collapsed on ten differences or less.
+The number inside each node is the sequence type based on the Warwick 7-gene MLST scheme [4].
+```{r, fig.width=15, fig.height=12}
+include_graphics("../figures/GrapeTree_cgMLST.png")
 ```
@@
 }
+
+broiler_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_broiler.tsv",
+                           sep = "\t",
+                           header = TRUE,
+                           colClasses = "factor",
+                           row.names = 1)
+
+pig_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_pig.tsv",
+                           sep = "\t",
+                           header = TRUE,
+                           colClasses = "factor",
+                           row.names = 1)
+
+fox_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_fox.tsv",
+                           sep = "\t",
+                           header = TRUE,
+                           colClasses = "factor",
+                           row.names = 1)
+
+bird_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_bird.tsv",
+                           sep = "\t",
+                           header = TRUE,
+                           colClasses = "factor",
+                           row.names = 1)
 
@@
             angle = 90) +
+  geom_vline(xintercept = mean(complete_data$Percent),
+             size = 1) +
   labs(x = "Percent (%) distances < 0.1 or > 0.8",
@@
 # References
-[1] Alikhan N-F, Zhou Z, Sergeant MJ, Achtman M.
+[1] Alikhan N-F et al.
 A genomic overview of the population structure of Salmonella.
@@
 
-[2] Machado MP, Silva M, Carriço JA, Silva DN, Santos S, Ramirez M, et al. chewBBACA: A complete suite for gene-by-gene schema creation and strain identification. Microb Genomics. 2018;1–7.
+[2] Machado MP et al. chewBBACA: A complete suite for gene-by-gene schema creation and strain identification. Microb Genomics. 2018;1–7.
+
+[3] Z Zhou et al. (2018) “GrapeTree: Visualization of core genomic relationships among 100,000 bacterial pathogens”, Genome Res. doi: https://doi.org/10.1101/gr.232397.117
+
+[4] Wirth T et al. (2006) "Sex and virulence in Escherichia coli : an evolutionary perspective", Molecular Microbiology (2006) 60 (5), 1136–1151, doi: 10.1111/j.1365-2958.2006.05172.x

### 2019-04-24 notebooks/article_notebook2.Rmd
removed unneccessary tables

--- 2019-03-15 notebooks/article_notebook2.Rmd
@@
 ---
-title: "Article data analysis 2"
+title: "Additional Tables"
 author: "Håkon Kaspersen"
@@
 
+# Libraries
 ```{r load libraries, message=FALSE, warning=FALSE}
 # Libraries
-# WARNING: This will install packages on your system if not present in library path!
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, ape, cluster, tibble, Biostrings, seqinr, stringr, scales, reprex)
+library(dplyr)
+library(tidyr)
+library(kableExtra)
 ```
@@
   "2016-17-565-1-S",
+  "2016-02-415-2-S",
   "2016-02-428-2-S",
   "2016-02-486-2-S",
-  "2016-02-732-2-S",
-  "2014-01-1741-1-S"
+  "2016-02-732-2-S"
   )
@@
                         header = TRUE,
-                        stringsAsFactors = F) %>%
-  filter(id %not_in% "2014-01-1741-1-S")
+                        stringsAsFactors = F)
 ```
 
-# ARIBA reports analyses
-This script filters results from ARIBA reports for genes of interest
-```{r ARIBA analysis}
-# Report directories
-megares_report_loc <- "../data/ariba/megares_results"
-resfinder_report_loc <- "../data/ariba/resfinder_results"
+```{r message=FALSE}
+source("../scripts/ARIBA_analysis.R")
 
-# Genes of interest. These genes are partially matched to 
-# the generated gene names from the "fix_gene_names" function
-ac_genes <- c("qnr","oqx","qep","aac6")
-mut_genes <- c("gyr","par","marR","marA",
-               "sox","rpoB","rob","gad")
-
-# -------------------------- Functions ----------------------------
-
-# Paste unique cell values together, separate with ","
-func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
-                                 collapse = ",")
-
-# Confidence interval function
-get_binCI <- function(x, n) as.numeric(
-  setNames(
-    binom.test(x,n)$conf.int*100, 
-    c("lwr", "upr")
-    )
-  )
-
-# Identifies filenames in input folder
-file_names <- function(filepath) {
-  files <- list.files(path = filepath, 
-                      pattern = "amr_report.tsv")
-  return(files)
-}
-
-# Import ariba data from report.tsv identified with file_names
-get_ariba_data <- function(filepath) {
-  # Identify file names in filepath
-  files <- file_names(filepath)
-
-  # Import data
-  data_list <- lapply(files,
-                      FUN = function(file) {
-                        read.delim(
-                          paste0(filepath, "/", file),
-                          stringsAsFactors = F,
-                          header = TRUE,
-                          sep = "\t"
-                        )
-                      })
-
-  # Set file names in list
-  names(data_list) <- files
-
-  # Bind to data frame and convert columns to character
-  data <- bind_rows(
-    lapply(
-      data_list,
-      function(x) map(x, as.character)
-      ),
-    .id = "ref")
-
-  return(data)
-}
-
-# Corrects gene names 
-fix_gene_names <- function(df) {
-  # Identify unique entries in ref_name column
-  genes <- unique(df$ref_name)
-
-  # Strip away excess characters
-  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
-  new_names <- gsub("_", "", new_names, fixed = T)
-  new_names <- gsub("-", "", new_names, fixed = T)
-
-  # make all genes lower case, with last letter 
-  # upper case if four characters long
-  gene_names <- c()
-
-  for (i in new_names) {
-    p <- paste(
-      tolower(
-        substring(i, 1,3)
-        ),
-      substring(i, 4),
-      sep = "",
-      collapse = " ")
-    gene_names <- c(gene_names,p)
-  }
-
-  # Create data frame for matching names below
-  df2 <- data.frame(genes,gene_names) %>%
-    mutate(genes = as.character(genes)) %>%
-    dplyr::rename(ref_name = genes)
-
-  # Left join to match with ref_name column
-  # and strip ref column for file suffix
-  df <- df %>%
-    left_join(df2, by = "ref_name") %>%
-    mutate(gene_names = as.character(gene_names),
-           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))
-
-  return(df)
-}
-
-# Function for selecting genes of interest and filtering the 
-# columns in the data frame on the resulting vector
-select_genes <- function(df, gene_string) {
-  names_df <- names(df)
-  x <- "ref"
-  for (i in names_df) {
-    for (j in gene_string) {
-      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
-        x <- c(x, i)
-      }
-    }
-  }
-  df <- df %>%
-    select(one_of(x))
-  return(df)
-}
-
-# Allowed flags from the ARIBA report
-flag_selection <- c("19","27","147","155","403",
-                    "411","915","923","787","795",
-                    "531","539","659","667","787","795")
-
-# Function that returns info on flag selection
-check_flags <- function(df) {
-  df <- df %>%
-    select(ref, gene_names, flag, ref_ctg_change, ref_nt, ctg_nt) %>%
-    mutate(flag_result = flag %in% flag_selection,
-           flag_result = as.integer(flag_result),
-           nt_change = paste(ref_nt, ctg_nt, sep = "-")) %>%
-    mutate(nt_change = ifelse(nt_change == ".-.", "0", nt_change)) %>%
-    dplyr::rename("gene" = gene_names) %>%
-    select(-c(ref_nt, ctg_nt))
-  return(df)
-}
-
-# Function that handles megares data and returns a data 
-# frame with information on whether a gene is mutated 
-# or not
-create_mut_table <- function(df) {
-  df <- df %>%
-    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
-           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
-           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
-           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
-           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
-           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
-    select(ref, gene_names, flag, mutation) %>%
-    filter(flag %in% flag_selection) %>%
-    mutate(id = 1:n()) %>%
-    spread(gene_names, mutation) %>%
-    group_by(ref) %>%
-    summarise_all(funs(func_paste)) %>%
-    select(-c(id, flag)) %>%
-    gather(gene, mut, -ref) %>%
-    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
-           result = ifelse(mut != 0, 1, 0),
-           result = as.integer(result),
-           type = "mut",
-           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
-           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
-  return(df)
-}
-
-# Corrects for mutations in QRDR for gyrA, gyrB, parC and parE genes.
-# The function returns columns of 1/0 values for whether or not the 
-# mutations reported by ARIBA is within the QRDR in the gene:
-# gyrA: AA 67 - 106
-# gyrB: AA 333 - 481
-# ParC: AA 51 - 170
-# parE: AA 366 - 523
-# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
-fix_gyr_par_results <- function(df) {
-  df <- df %>%
-    # control for mutation within QRDR for gyrA
-    mutate(gyrA_result = mut %>% 
-             str_extract_all("\\d+") %>% # from reprex package
-             map(as.integer) %>% # converts all to integer
-             # returns TRUE/FALSE whether value is within range or not
-             map_lgl(~ any(.x >= 67L & .x <= 106L)), 
-           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result),
-           gyrA_result = as.integer(gyrA_result),
-           # control for mutation within QRDR for gyrB
-           gyrB_result = mut %>%
-             str_extract_all("\\d+") %>%
-             map(as.integer) %>%
-             map_lgl(~ any(.x >= 333L & .x <= 481L)),
-           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
-           gyrB_result = as.integer(gyrB_result),
-           # control for mutation within QRDR for parC
-           parC_result = mut %>% 
-             str_extract_all("\\d+") %>% 
-             map(as.integer) %>% 
-             map_lgl(~ any(.x >= 51L & .x <= 170L)),
-           parC_result = if_else(gene != "parC", NA, parC_result),
-           parC_result = as.integer(parC_result),
-           # control for mutation within QRDR for parE
-           parE_result = mut %>% 
-             str_extract_all("\\d+") %>% 
-             map(as.integer) %>% 
-             map_lgl(~ any(.x >= 366L & .x <= 523L)),
-           parE_result = if_else(gene != "parE", NA, parE_result),
-           parE_result = as.integer(parE_result)) %>%
-    # mergeresults from all four genes into one column
-    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
-                                      gene == "gyrB" ~ gyrB_result,
-                                      gene == "parC" ~ parC_result,
-                                      gene == "parE" ~ parE_result)) %>%
-    # merge results from other genes with the four genes
-    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"),
-                                 result_gyr_par, result)) %>%
-    select(-c(gyrA_result,
-              gyrB_result,
-              parC_result,
-              parE_result,
-              result_gyr_par,
-              result))
-  return(df)
-}
-
-# Function that handles resfinder data and returns a data frame with 
-# presence/absence for acquired genes                           
-create_acquired_table <- function(df) {
-  df <- df %>%
-    select(ref, gene_names, flag, ref_ctg_change) %>%
-    mutate(id = 1:n()) %>%
-    # Filter flags
-    filter(flag %in% flag_selection) %>%
-    spread(gene_names, ref_ctg_change) %>%
-    group_by(ref) %>%
-    summarise_all(funs(func_paste)) %>%
-    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
-    select(-c(id, flag)) %>%
-    gather(gene, result,-ref) %>%
-    mutate(result_total = ifelse(result == "", 0, 1),
-           result_total = as.character(result_total),
-           type = "gene") %>%
-    select(-result)
-  return(df)
-}
-
-# Function that returns a filtered dataframe based 
-# on the vector "mut_genes"
-filter_mut_table <- function(df) {
-  # Identify unique genes in column
-  report_genes <- unique(df$gene)
-  grep_genes <- c()
-
-  # match names from name vector to column entries
-  for (gene in mut_genes) {
-    for (g in report_genes) {
-      if (grepl(gene, g, ignore.case = T) == TRUE) {
-        grep_genes <- c(grep_genes, g)
-      }
-    }
-  }
-
-  # Filter data frame based on values in grep_genes
-  df <- df %>%
-    filter(gene %in% grep_genes) %>%
-    mutate(gene = as.character(gene))
-  return(df)
-}
-
-# Function that returns a filtered dataframe based
-# on the string "ac_genes"
-filter_acquired_table <- function(df) {
-  # Identify unique genes in column
-  report_genes <- unique(df$gene)
-  grep_genes <- c()
-
-  # match names from vector to column entries
-  for (gene in ac_genes) {
-    for (g in report_genes) {
-      if (grepl(gene, g, ignore.case = T) == TRUE) {
-        grep_genes <- c(grep_genes, g)
-      }
-    }
-  }
-
-  # Filter data frame based on values in grep_genes
-  df <- df %>%
-    filter(gene %in% grep_genes) %>%
-    mutate(gene = gsub("_", "", gene),
-           gene = as.character(gene))
-  return(df)
-}
-
-# calculates percentage of present/absent mutations and genes
-calc_stats <- function(df) {
-  df <- df %>%
-    # Count Presence/Absence of each gene
-    group_by(gene, result_total) %>%
-    dplyr::count() %>%
-    ungroup() %>%
-    # Convert binary to "Present" and "Absent"
-    mutate(result_total = if_else(result_total == 1, 
-                                  "Present", "Absent")) %>%
-    # Spread to separate Present and Absent columns
-    spread(result_total, n, fill = 0) %>%
-    rowwise() %>%
-    # Calculate percentages
-    mutate(Total = Present + Absent,
-           Percent = round(Present/Total*100, 1))
-  return(df)
-}
-
-# calculates how many mutations are present in the genes
-# gyrA, gyrB, parC and parE
-calc_no_of_mut <- function(df) {
-  df1 <- df %>%
-    # Filter out all other genes
-    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
-    select(-nt_change) %>%
-    # Count how many mutations in each gene per row
-    mutate(entries = sapply(strsplit(.$mut, ","),
-                            FUN = function(x) {length(x)})) %>%
-    # Separate each mutation to own column
-    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
-    select(-entries) %>%
-    # Gather all mutations together into one column,
-    # to get one row per mutation
-    gather(id, mut, -c(ref, gene, result, type))
-
-  # Correct for QRDR mutations in the four genes
-  df2 <- fix_gyr_par_results(df1)
-
-  # Generate final table
-  df3 <- df2 %>%
-    # Filter out mutations outide QRDR
-    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
-    filter(test == 1) %>%
-    # Create own column for each gene
-    spread(gene, mut) %>%
-    group_by(ref) %>%
-    # Summarise all columns based on unique values
-    summarise_all(funs(func_paste)) %>%
-    # Remove "0" values for correct counts below
-    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
-              .funs = function(x) ifelse(x == "0", "", x)) %>%
-    # Count mutations in each column entry
-    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","),
-                             FUN = function(x) {length(x)}),
-           mut_gyrB = sapply(strsplit(.$gyrB, ","),
-                             FUN = function(x) {length(x)}),
-           mut_parC = sapply(strsplit(.$parC, ","),
-                             FUN = function(x) {length(x)}),
-           mut_parE = sapply(strsplit(.$parE, ","),
-                             FUN = function(x) {length(x)})) %>%
-    select(-c(type, id, result_total, test)) %>%
-    # Reinsert "0" values 
-    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
-              .funs = function(x) ifelse(x == "", "0", x))
-  return(df3)
-}
-
-# creates a data frame with one row per sample, and 1/0 results 
-# for mutations in respective genes in columns
-create_mut_report <- function(df) {
-  df <- df %>%
-    select(-mut) %>%
-    group_by(id) %>%
-    mutate(id2 = 1:n()) %>%
-    spread(gene, result_total) %>%
-    summarise_all(funs(func_paste)) %>%
-    select(-c(id2, type))
-  return(df)
-}
-
-# creates a data frame with one row per sample, and 1/0 results
-# for acquired genes in columns
-create_acquired_report <- function(df) {
-  df <- df %>%
-    group_by(id) %>%
-    mutate(id2 = 1:n()) %>%
-    spread(gene, result_total) %>%
-    summarise_all(funs(func_paste)) %>%
-    select(-c(id2, type))
-  return(df)
-}
-
-# -------------------------- Analysis ---------------------------
-# Import data
-mut_data <- get_ariba_data(megares_report_loc)
-acquired_data <- get_ariba_data(resfinder_report_loc)
-
-# Clean data
-clean_mut_data <- fix_gene_names(mut_data)
-clean_acquired_data <- fix_gene_names(acquired_data)
-
-# Check flags
-mut_flags <- check_flags(clean_mut_data) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  # Filters out unwanted samples
-  filter(id %not_in% ex_samples)
-
-acquired_flags <- check_flags(clean_acquired_data) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-
-mut_table <- create_mut_table(clean_mut_data)
-
-mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-
-mut_filtered <- filter_mut_table(mut_table_fixed)
-
-acquired_table <- create_acquired_table(clean_acquired_data) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-
-acquired_filtered <- filter_acquired_table(acquired_table)
-
-mut_stats <- calc_stats(mut_filtered)
-mut_report <- create_mut_report(mut_filtered)
-
-acquired_stats <- calc_stats(acquired_filtered)
-acquired_report <- create_acquired_report(acquired_filtered)
-
-mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
-  # Correct for erroneous entry from ARIBA
-  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-
-# number of mutations table
-mut_no <- mut_quant %>%
-  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
-  dplyr::count()
-
-# number of isolates per mutation combination
-mut_comb <- mut_quant %>%
-  group_by(gyrA, gyrB, parC, parE) %>%
-  dplyr::count()
 ```
@@
 # Tables
-## Isolate data
-This table presents the isolates and their accompanying metadata.
+
+## Group listing
+This table presents the groups created in the isolate selection.
+The "Additional Resistances" column contains how many additional antimicrobials the isolates in that group is resistant to, other than CIP and/or NAL.
 ```{r}
-isolate_data_complete <- isolate_data %>%
-  # combine with read data
-  left_join(no_of_reads, by = "id")
-
-isolate_data_complete
+mic_data %>%
+  group_by(antres, CIP, NAL) %>%
+  count() %>%
+  dplyr::rename("Additional Resistances" = antres) %>%
+  ungroup() %>%
+  mutate(Group = 1:n()) %>%
+  dplyr::select(Group, everything()) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped","hover"))
 ```
 
-## Isolates per species
-This table presents how many isolates were included per species.
-```{r}
-per_species <- isolate_data %>%
-  group_by(species) %>%
-  dplyr::count()
-
-per_species
-```
-
-## Sequence types in total
-```{r}
-tot_mlst <- mlst_data[,c("id","ST")] %>%
-  group_by(ST) %>%
-  dplyr::count() %>%
-  arrange(desc(n))
-
-print(paste0("Total amount of different sequence types: ", nrow(tot_mlst)))
-
-tot_mlst
-```
-
-## Sequence types per animal species
-This table presents all the different sequence types identified in each animal species, arranged in a descending order.
-```{r}
-mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(ST, species) %>%
-  dplyr::count() %>%
-  arrange(desc(n))
-```
-
-## Most abundant sequence types per animal species
-This table presents the most abundant sequence types identified in each animal species (n > 3).
-```{r}
-# Identify which STs have more than three observations
-mlst_n <- mlst_data[,c("id","ST")] %>%
-  group_by(ST) %>%
-  dplyr::count() %>%
-  # filter out all STs with less than four observations
-  filter(n > 3)
-
-mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  spread(ST, species) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(-id) %>%
-  gather(ST, species) %>%
-  # filter data frame on the STs identified above
-  filter(ST %in% mlst_n$ST)
-```
-
-## Number of sequence types per animal species
-This table presents the amount of different sequence types that were identified in each animal species in total.
-```{r}
-mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(species) %>%
-  # count how many STs per group
-  mutate(n = length(unique(ST))) %>%
-  select(species,n) %>%
-  summarise_all(funs(func_paste))
-```
-
-## Percent isolates with PMQR
-This table presents the percentage of isolates with identified PMQR determinants in total.
-```{r}
-acquired_report %>%
-  select(-id) %>%
-  mutate_all(funs(as.numeric(.))) %>%
-  # presence/absence of qnr genes in total
-  mutate(total = rowSums(.),
-         result = if_else(total != 0, 1, 0)) %>%
-  group_by(result) %>%
-  # count groups
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
-  spread(result, n) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1))
-```
-
-## Percent isolates with qnr family genes
-```{r}
-acquired_report %>%
-  select(-c(qepA4, id)) %>%
-  mutate_all(funs(as.numeric)) %>%
-  dplyr::rename("qnrA" = qnrA1) %>%
-  # identify presence/absence for family groups 
-  # from the gene subgroups
-  mutate(qnrB = ifelse(rowSums(.[,2:4]) >= 1, 1, 0),
-         qnrS = ifelse(rowSums(.[,5:7]) >= 1, 1, 0)) %>%
-  select(qnrA, qnrB, qnrS) %>%
-  gather(gene, value) %>%
-  group_by(gene, value) %>%
-  # count groups
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 0, "Absent", "Present")) %>%
-  spread(value, n) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1))
-```
-
-## Percent isolates with specific PMQR genes
-```{r}
-acquired_report %>%
-  select(-id) %>%
-  # transpose table
-  t() %>%
-  as.data.frame(.) %>%
-  mutate(Gene = rownames(.)) %>%
-  mutate_at(vars(-Gene),
-            funs(as.numeric(as.character(.)))) %>%
-  mutate(Present = rowSums(.[,-281])) %>%
-  select(c(Gene, Present)) %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(Present/Total * 100,1))
-```
-
-## qnr-genes per animal species
-This table presents the specific qnr-genes identified in each animal species.
-```{r}
-qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")
-
-acquired_report %>%
-  left_join(isolate_data, by = "id") %>%
-  gather(key, value, qnr_genes) %>%
-  group_by(key, value, species) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present / Total * 100, 1)) %>%
-  arrange(key, desc(Percent))
-```
-
-## Percent isolates with mutations in QRDR
-This table presents the percent of isolates with mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
-```{r}
-mut_report %>%
-  select(gyrA, gyrB, parC, parE) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
-            .funs = funs(as.numeric)) %>%
-  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
-  select(n, type) %>%
-  group_by(type) %>%
-  summarise_all(funs(sum)) %>%
-  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
-  spread(type, n) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1)) %>%
-  select(-Total)
-```
-
-## Percent mutations in each intrinsic gene
-This table presents the percent mutations in all intrinsic _E. coli_ genes included in this study.
-```{r}
-mut_report %>%
-  select(-nt_change) %>%
-  gather(gene, value, -id) %>%
-  group_by(gene, value) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total * 100,1)) %>%
-  arrange(desc(Percent))
-```
-
-## Specific mutations in QRDR
-This table presents the specific mutations identified in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
-```{r, warning=FALSE, message=FALSE}
-clean_mut_data %>%
-  # correct non-mutated entries
-  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
-                                 "", ref_ctg_change),
-         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
-         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
-         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
-         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
-         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
-  select(ref, gene_names, flag, mutation) %>%
-  # filter quality flags and genes
-  filter(flag %in% flag_selection,
-         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
-  mutate(id = 1:n()) %>%
-  # create columns for each gene
-  spread(gene_names, mutation) %>%
-  group_by(ref) %>%
-  # summarise for each sample
-  summarise_all(funs(func_paste2)) %>%
-  select(-c(id, flag)) %>%
-  # set correct sample id's
-  left_join(id_data, by = "ref") %>%
-  dplyr::select(id, everything(), -ref) %>%
-  # filter out excluded samples
-  filter(id %not_in% ex_samples) %>%
-  gather(gene, mut, -id) %>%
-  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
-                      0, mut),
-         result = ifelse(mut != 0, 1, 0),
-         result = as.integer(result),
-         type = "mut") %>%
-  # create one row per mutation
-  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
-  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
-  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
-  # correct erroneous result from ARIBA
-  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
-                            "S80I", aa_change)) %>%
-  select(-mut) %>%
-  dplyr::rename("mut" = aa_change) %>%
-  filter(is.na(nt_change) == FALSE) %>%
-  # fix results for the four genes
-  fix_gyr_par_results(.) %>%
-  filter(result_total == 1) %>%
-  group_by(gene, mut, nt_change) %>%
-  # count groups
-  dplyr::count() %>%
-  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(n/Total * 100,1)) %>%
-  dplyr::rename("Mutation" = mut) %>%
-  arrange(gene, desc(Percent))
-```
-
-## Specific QRDR mutations per animal species
-```{r, warning=FALSE, message=FALSE}
-clean_mut_data %>%
-  # correct non-mutated entries
-  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".",
-                                 "", ref_ctg_change),
-         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
-         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
-         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
-         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
-         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
-  select(ref, gene_names, flag, mutation) %>%
-  # filter quality flags and genes
-  filter(flag %in% flag_selection,
-         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
-  mutate(id = 1:n()) %>%
-  # create columns for each gene
-  spread(gene_names, mutation) %>%
-  group_by(ref) %>%
-  summarise_all(funs(func_paste2)) %>%
-  dplyr::select(-c(id, flag)) %>%
-  left_join(id_data, by = "ref") %>%
-  dplyr::select(id, everything(), -ref) %>%
-  # filter out excluded samples
-  filter(id %not_in% ex_samples) %>%
-  gather(gene, mut, -id) %>%
-  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE,
-                      0, mut),
-         result = ifelse(mut != 0, 1, 0),
-         result = as.integer(result),
-         type = "mut") %>%
-  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
-  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
-  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
-  # correct errouneous result from ARIBA
-  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I",
-                            "S80I", aa_change)) %>%
-  select(-mut) %>%
-  dplyr::rename("mut" = aa_change) %>%
-  filter(is.na(nt_change) == FALSE) %>%
-  # fix results for the four genes
-  fix_gyr_par_results(.) %>%
-  filter(result_total == 1) %>%
-  left_join(isolate_data[c("id","species")], by = "id") %>%
-  group_by(gene, mut, species) %>%
-  dplyr::count() %>%
-  rowwise() %>%
-  # set total amount of isolates per animal species
-  mutate(Total = case_when(species == "Broiler" ~ 87,
-                           species == "Pig" ~ 75,
-                           species == "Red fox" ~ 52,
-                           species == "Wild bird" ~ 66),
-         Percent = round(n/Total * 100,1)) %>%
-  dplyr::rename("Mutation" = mut) %>%
-  arrange(gene, desc(Percent))
-```
-
-## Mutation combinations in QRDR
-This table presents all the observed combinations of mutations in the QRDR of _gyrA_, _gyrB_, _parC_ and _parE_.
-```{r}
-mut_quant %>%
-  group_by(gyrA, gyrB, parC, parE) %>%
-  # count combinations across columns
-  dplyr::count() %>%
-  rowwise() %>%
-  mutate(Total = 280,
-         Percent = round(n/Total * 100,1)) %>%
-  arrange(desc(Percent))
-```
-
-## Mutations in intrinsic genes per animal species
-This table presents the percentage of isolates with mutations for each intrinsic gene.
-```{r}
-# Identify gene names
-gene_names <- names(mut_report)
-gene_names <- gene_names[-c(1,2)]
-
-mut_report %>%
-  left_join(isolate_data, by = "id") %>%
-  # use gene names to specify which columns to gather
-  gather(key, value, gene_names) %>%
-  group_by(key, value, species) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present / Total * 100, 1)) %>%
-  dplyr::rename("Gene" = key,
-                "Species" = species) %>%
-  dplyr::select(Gene, Species, Present, Total, Percent) %>%
-  arrange(Gene, desc(Percent))
-```
-
-## Specific mutations in non-QRDR intrinsic genes
-These mutations are the most prominent in each gene where mutations were identified (n > 2).
-The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
-```{r, message=FALSE, warning=FALSE}
-mut_filtered %>%
-  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
-  select(id, gene, mut) %>%
-  mutate(entries = sapply(strsplit(.$mut, ","),
-                          FUN = function(x) {length(x)})) %>%
-  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
-  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
-  mutate(value = str_trim(value)) %>%
-  select(id, gene, value) %>%
-  filter(value != 0,
-         is.na(value) == FALSE) %>%
-  group_by(gene, value) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  filter(n > 2) %>%
-  mutate(percent = case_when(
-    gene == "gadX" ~ 
-      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
-    gene == "marA" ~ 
-      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
-    gene == "marR" ~ 
-      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
-    gene == "rpoB" ~ 
-      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
-    gene == "soxS" ~ 
-      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
-  arrange(gene, desc(percent)) %>%
-  dplyr::rename("Gene" = gene,
-                "Mutation" = value,
-                "Percent" = percent)
-
-```
-
-## Mechanisms in isolates without QRDR mutations
-This table presents the mechanisms identified in isolates with no mutations in either of _gyrA_, _gyrB_, _parC_ and/or _parE_.
+## Mechanisms in isolates without substitutions in QRDR
+This table presents the mechanisms identified in isolates with no amino acid substitutions in either of GyrA, GyrB, ParC and/or ParE.
 ```{r}
@@
   ungroup() %>%
-  dplyr::select(-c(gyrA,gyrB,parC,parE))
+  dplyr::select(-c(gyrA,gyrB,parC,parE)) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
 ```
 
-## Total combinations of resistance mechanisms
-This table presents all combinations of resistance mechanisms included in this study, and their respective MIC.
-```{r}
-mut_report %>%
-  select(-c(nt_change, gyrA, parC, parE)) %>%
-  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
-  left_join(acquired_report, by = "id") %>%
-  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
-            funs(as.numeric(.))) %>%
-  # summarise qnr-genes
-  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
-  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
-  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  arrange(desc(CIP))
-```
-
-## Significant differences in resistance mechanisms
-This table presents the significant (p < 0.05) differences in resistance mechanisms between species, identified with chi-squared tests.
-```{r, message=FALSE, warning=FALSE}
-# function for doing chi-squared tests on multiple entries in a list
-do_chisq_test <- function(df) {
-  df <- df %>%
-    # create columns with all possible comparisons
-    dplyr::slice(expand.grid(1:length(unique(species)),
-                             1:length(unique(species))) %>%
-                   rev %>% 
-          dplyr::filter(Var2 < Var1) %>% t) %>%
-    # create test ID
-    mutate(test = rep(1:(n()/2), each = 2)) %>%
-    group_by(test) %>%
-    # specify values for chi-squred tests
-    dplyr::do(data_frame(gene = .$key,
-                  test = dplyr::first(.$test),
-                  species1 = dplyr::first(.$species),
-                  species2 = dplyr::last(.$species),
-                  data = list(matrix(c(.$Absent,
-                                       .$Present),
-                                     ncol = 2)))) %>%
-    # do chi-squared tests
-    mutate(chi_test = map(data, chisq.test,
-                          correct = FALSE)) %>%
-    mutate(p.value = map_dbl(chi_test, "p.value"),
-           dupl = duplicated(test)) %>%
-    # filter out duplicated tests
-    dplyr::filter(dupl == FALSE) %>%
-    ungroup() %>%
-    select(gene, species1, species2, p.value) %>%
-    # round off p-value column
-    mutate(p.value = round(p.value, 5)) %>%
-    dplyr::filter(p.value <= 0.05)
-  return(df)
-}
-
-# create total mechanism table
-total_df <- acquired_report %>%
-  left_join(mut_report, by = "id") %>%
-  left_join(isolate_data[c("id", "species")], by = "id") %>%
-  select(-nt_change)
-
-# create list of genes with total presence/absence values
-# from each animal species
-mechanism_per_species <- total_df %>%
-  select(id, species, everything()) %>%
-  select(-id) %>%
-  gather(key, value, -species) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  select(key, species, Present, Absent) %>%
-  split(f = .$key)
-
-# Get intrinsic gene names
-intrinsic_names <- names(mut_report)
-intrinsic_names <- intrinsic_names[-c(1,2)]
-
-# create percentage table over intrinsic genes
-intrinsic_gene_perc <- mut_report %>%
-  left_join(isolate_data, by = "id") %>%
-  gather(key, value, intrinsic_names) %>%
-  group_by(key, value, species) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present / Total * 100, 1)) %>%
-  select(key, species, Percent) %>%
-  dplyr::rename("gene" = key)
-
-# Get acquired gene names
-acquired_names <- names(acquired_report)
-acquired_names <- acquired_names[-1]
-
-# create percentage table over acquired genes
-acquired_genes_perc <- acquired_report %>%
-  left_join(isolate_data, by = "id") %>%
-  gather(key, value, acquired_names) %>%
-  group_by(key, value, species) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present / Total * 100, 1)) %>%
-  select(key, species, Percent) %>%
-  dplyr::rename("gene" = key)
-
-# create total percentage table
-total_perc <- rbind(intrinsic_gene_perc, acquired_genes_perc)
-
-# run chi-squared tests
-p_values_total <- lapply(mechanism_per_species, function(x) do_chisq_test(x)) %>%
-  # bind results
-  bind_rows() %>%
-  # specify percentages for species 1
-  left_join(total_perc, by = c("gene", c("species1" = "species"))) %>%
-  dplyr::rename("Percent species 1" = Percent) %>%
-  # specify percentages for species 2
-  left_join(total_perc, by = c("gene", c("species2" = "species"))) %>%
-  dplyr::rename("Percent species 2" = Percent) %>%
-  dplyr::select(gene,
-                species1,
-                `Percent species 1`,
-                species2,
-                `Percent species 2`,
-                p.value)
-
-p_values_total
-
-```
-
-## Percent resistance towards all tested compounds
-This table presents the percent resistance towards the compounds sulfamethoxazole (SMX), trimethoprim (TMP), ciprofloxacin (CIP), tetracycline (TET), nalidixic acid (NAL), cefotaxime (CTX), chloramphenicol (CHL), ampicillin (AMP), and gentamicin (GEN).
-```{r}
-percent_am_res <- mic_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  # set resistant/nonresistant based on MIC cutoffs
-  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
-         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
-         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
-         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
-         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
-         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
-         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
-         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
-         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
-  select(species,contains("_R")) %>%
-  gather(compound, value, -species) %>%
-  mutate(compound = sub("_R", "", compound)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  # calculate percentages
-  mutate(Total = Present + Absent,
-         Percent = round((Present/Total)*100, 1))
-
-percent_am_res
-```
-
-## Resistance mechanisms per sequence type
-This table presents the total resistance mechanisms identified in each sequence type (n > 2).
-```{r}
-mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  left_join(mlst_data[,c("id","ST")], by = "id") %>%
-  select(-c(nt_change, id)) %>%
-  gather(gene, value, -ST) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  filter(Present != 0) %>%
-  mutate(Total = Present + Absent) %>%
-  # filter out all STs observed less than three times
-  filter(Total > 2) %>%
-  rowwise() %>%
-  mutate(Percent = round(Present/Total*100, 1)) %>%
-  ungroup() %>%
-  group_by(ST) %>%
-  mutate(no_of_mechanisms = n())
-```
-
-## Resistance mechanisms per sequence type per animal species
-This table presents each resistance mechanism identified in each sequence type (n > 2) per animal species.
-```{r}
-mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  left_join(mlst_data[,c("id","ST")], by = "id") %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  select(-c(nt_change, id)) %>%
-  gather(gene, value, -c(ST, species)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  filter(Present != 0) %>%
-  mutate(Total = Present + Absent) %>%
-  # filter out all STs observed less than three times
-  filter(Total > 2) %>%
-  rowwise() %>%
-  mutate(Percent = round(Present/Total*100, 1)) %>%
-  ungroup() %>%
-  group_by(ST, species) %>%
-  mutate(no_of_mechanisms = n()) %>%
-  arrange(ST, species)
-```
-
-## Mutations in intrinsic genes for isolates with CIP MIC < 0.12
-This table presents the mutations identified in isolates with a MIC-value for ciprofloxacin below the ECOFF. Note that all mutations in all genes are listed here, not just the QRDR-located ones. 
+## Substitutions in intrinsic genes for isolates with CIP MIC < 0.12
+This table presents the substitutions identified in isolates with a MIC-value for ciprofloxacin below the ECOFF.
+Note that all substitutions in all genes are listed here, not just the QRDR-located ones.
 ```{r}
@@
 
+gene_names <- names(mut_report)
+gene_names <- gene_names[-c(1,2)]
+
 mut_table_fixed %>%
@@
   left_join(mic_data[,c("id", "CIP", "NAL")], by = "id") %>%
-  left_join(isolate_data[,c("id","species","method")], by = "id")
+  left_join(isolate_data[,c("id","species","method")], by = "id") %>%
+  dplyr::select(-c(id, PMQR, method)) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
 ```
-
-## Significant differences between the amount of resistant isolates
-These tables present the significant (p < 0.05) differences in isolates resistant to quinolones (table 1), and the rest of the tested compounds (table 2) for each animal species.
-```{r, message=FALSE, warning=FALSE}
-# create table over amount of resistant and sensitive isolates for CIP and NAL
-cip_nal_res <- mic_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  mutate(CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
-         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0)) %>%
-  select(species,contains("_R")) %>%
-  gather(compound, value, -species) %>%
-  mutate(compound = sub("_R", "", compound)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
-  spread(value, n, fill = 0) %>%
-  select(compound, species, Present, Absent) %>%
-  dplyr::rename("key" = compound) %>%
-  split(f = .$key)
-
-# do chi-squared test
-p_values_res <- lapply(cip_nal_res, function(x) do_chisq_test(x)) %>%
-  bind_rows() %>%
-  # insert percentages for species 1
-  left_join(percent_am_res[,c("species","compound","Percent")],
-            by = c("species1" = "species",
-                   "gene" = "compound")) %>%
-  dplyr::rename("Percent_species1" = Percent) %>%
-  # insert percentages for species 2
-  left_join(percent_am_res[,c("species","compound","Percent")],
-            by = c("species2" = "species",
-                   "gene" = "compound")) %>%
-  dplyr::rename("Percent_species2" = Percent,
-                "Compound" = gene) %>%
-  select(Compound,
-         species1,
-         Percent_species1,
-         species2,
-         Percent_species2,
-         p.value)
-
-p_values_res
-
-# dataset for all other compounds
-rest_res <- mic_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
-         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
-         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
-         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
-         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
-         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
-         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
-  select(species,contains("_R")) %>%
-  gather(compound, value, -species) %>%
-  mutate(compound = sub("_R", "", compound)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
-  spread(value, n, fill = 0) %>%
-  select(compound, species, Present, Absent) %>%
-  dplyr::rename("key" = compound) %>%
-  split(f = .$key)
-
-p_values_res_rest <- lapply(rest_res, function(x) do_chisq_test(x)) %>%
-  bind_rows() %>%
-  # insert percentages for species 1
-  left_join(percent_am_res[,c("species","compound","Percent")], 
-            by = c("species1" = "species",
-                   "gene" = "compound")) %>%
-  dplyr::rename("Percent_species1" = Percent) %>%
-  # insert percentages for species 2
-  left_join(percent_am_res[,c("species","compound","Percent")],
-            by = c("species2" = "species",
-                   "gene" = "compound")) %>%
-  dplyr::rename("Percent_species2" = Percent,
-                "Compound" = gene) %>%
-  select(Compound,
-         species1,
-         Percent_species1,
-         species2,
-         Percent_species2,
-         p.value)
-
-p_values_res_rest
-```
-
-## Correlation between specific mutations and qnr-genes
-```{r}
-AddCol <- function(df, col_name) {
-  # split rows by delimeters
-  string_to_proc <- df %>% select(!!col_name) %>%
-    base::unlist() %>% str_split(regex("\\, |\\,")) 
-  # find unique entries
-  unique_strings <- string_to_proc %>%
-    base::unlist() %>% base::unique()
-  # construct names of the new columns
-  cols_names <- paste(col_name, unique_strings, sep = "_")
-  # construct 0/1-content columns for each unique entry
-  cols_content <- sapply(function(i) {
-    as.integer(base::unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
-                             X = string_to_proc)))
-  }, X = seq_along(unique_strings))
-  res <- data.frame(cols_content)
-  names(res) <- cols_names
-  return(res)
-}
-
-col_proc <- c("gyrA","gyrB","parC","parE")
-
-tot_mut_report <- sapply(function(i) {AddCol(df = mut_quant, col_name = col_proc[i])},
-                      X = seq_along(col_proc)) %>%
-  bind_cols() %>%
-  select(-c(gyrA_0, parC_0, gyrB_0, parE_0)) %>%
-  mutate_all(funs(as.numeric))
-
-test <- mut_quant %>%
-  cbind(tot_mut_report) %>%
-  select(-contains("mut")) %>%
-  select(-c(gyrA, gyrB, parC, parE)) %>%
-  left_join(acquired_report, by = "id") %>%
-  mutate_at(vars(-id),
-            funs(as.numeric)) %>%
-  mutate(qnr = ifelse(rowSums(.[,21:27]) == 0, 0, 1)) %>%
-  select(-id)
-
-cor.test(test$gyrA_S83L, test$qnr)
-```
-
-## Double-check ARIBA results
-This chunk double-checks the mutations found in _gyrA_, _parC_ and _parE_ against an internal mutation analysis script.
-The sequences are taken from Prokka, and reference wild-type versions of the genes are used to identify mismatches with the sample sequences.
-```{r, warning=FALSE, message=FALSE}
-# Functions
-mismatches <- function(query, ref) {
-  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
-                    gapOpening = 3, gapExtension = 1) %>%
-    mismatchTable() %>%
-    mutate(ID=names(query),
-           Pos=PatternStart,
-           Reference_AA=as.character(PatternSubstring),
-           Sample_AA=as.character(SubjectSubstring)) %>%
-    select(ID, Reference_AA, Sample_AA, Pos) %>%
-    filter(Sample_AA != "X")
-}
-
-# Import sequence data
-gyrA <- readAAStringSet("../data/Prokka/DNA_gyrase_GYRA_aa_sequences.fa")
-parC <- readAAStringSet("../data/Prokka/Topoisomerase_IV_subunit_A_aa_sequences.fa")
-parE <- readAAStringSet("../data/Prokka/DNA_topoisomerase_IV_subunit_B_aa_sequences.fa")
-
-# Identify results from ARIBA
-mut_gyrA <- mut_quant %>%
-  select(id, gyrA) %>%
-  mutate(entries = sapply(strsplit(.$gyrA, ","),
-                          FUN = function(x) {length(x)})) %>%
-  separate(gyrA, into = as.character(c(1:max(.$entries)))) %>%
-  select(-entries) %>%
-  gather(key, value, `1`,`2`) %>%
-  filter(is.na(value) == FALSE) %>%
-  select(-key) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup()
-
-# Run mutation identification analysis on Prokka sequences and
-# match to ARIBA results
-gyrA_results <- bind_rows(lapply(seq_along(gyrA[-1]),
-                            function(i) mismatches(gyrA[i + 1],
-                                                   gyrA[1]))) %>%
-  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
-  select(ID, mutation) %>%
-  mutate(ID = sub(".faa", "", ID)) %>%
-  dplyr::rename("ref" = ID,
-                "gyrA_prokka" = mutation) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, gyrA_prokka) %>%
-  # control for mutation within QRDR for gyrA
-  mutate(gyrA_result = gyrA_prokka %>%
-             str_extract_all("\\d+") %>% # from reprex package
-             map(as.integer) %>% # converts all to integer
-             map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
-  filter(gyrA_result == TRUE) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup() %>%
-  left_join(mut_gyrA, by = c("id","test")) %>%
-  select(-test) %>%
-  dplyr::rename("gyrA_ariba" = value) %>%
-  filter(id %not_in% ex_samples) %>%
-  select(-gyrA_result)
-
-# Identify results from ARIBA
-mut_parC <- mut_quant %>%
-  select(id, parC) %>%
-  mutate(entries = sapply(strsplit(.$parC, ","),
-                          FUN = function(x) {length(x)})) %>%
-  separate(parC, into = as.character(c(1:max(.$entries)))) %>%
-  select(-entries) %>%
-  gather(key, value, `1`,`2`) %>%
-  filter(is.na(value) == FALSE) %>%
-  select(-key) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup()
-
-# Run mutation identification analysis on Prokka sequences and
-# match to ARIBA results
-parC_results <- bind_rows(lapply(seq_along(parC[-1]),
-                                 function(i) mismatches(parC[i + 1],
-                                                        parC[1]))) %>%
-  mutate(Pos = Pos + 22) %>%
-  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
-  select(ID, mutation) %>%
-  mutate(ID = sub(".faa", "", ID)) %>%
-  dplyr::rename("ref" = ID,
-                "parC_prokka" = mutation) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, parC_prokka) %>%
-  mutate(parC_result = parC_prokka %>%
-             str_extract_all("\\d+") %>% 
-             map(as.integer) %>% 
-             map_lgl(~ any(.x >= 51L & .x <= 170L))) %>%
-  filter(parC_result == TRUE) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup() %>%
-  left_join(mut_parC, by = c("id","test")) %>%
-  select(-test) %>%
-  dplyr::rename("parC_ariba" = value) %>%
-  filter(id %not_in% ex_samples) %>%
-  select(-parC_result)
-
-# Identify results from ARIBA
-mut_parE <- mut_quant %>%
-  select(id, parE) %>%
-  mutate(entries = sapply(strsplit(.$parE, ","),
-                          FUN = function(x) {length(x)})) %>%
-  separate(parE, into = as.character(c(1:max(.$entries)))) %>%
-  select(-entries) %>%
-  gather(key, value, `1`,`2`) %>%
-  filter(is.na(value) == FALSE) %>%
-  select(-key) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup()
-
-# Run mutation identification analysis on Prokka sequences and
-# match to ARIBA results
-parE_results <- bind_rows(lapply(seq_along(parE[-1]),
-                                 function(i) mismatches(parE[i + 1],
-                                                        parE[1]))) %>%
-  #mutate(Pos = Pos + 22) %>%
-  mutate(mutation = paste0(Reference_AA, Pos, Sample_AA)) %>%
-  select(ID, mutation) %>%
-  mutate(ID = sub(".faa", "", ID)) %>%
-  dplyr::rename("ref" = ID,
-                "parE_prokka" = mutation) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, parE_prokka) %>%
-  mutate(parE_result = parE_prokka %>%
-           str_extract_all("\\d+") %>% 
-           map(as.integer) %>% 
-           map_lgl(~ any(.x >= 366L & .x <= 523L))) %>%
-  filter(parE_result == TRUE) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup() %>%
-  left_join(mut_parE, by = c("id","test")) %>%
-  select(-test) %>%
-  dplyr::rename("parE_ariba" = value) %>%
-  filter(id %not_in% ex_samples) %>%
-  select(-parE_result)
-
-full_report <- gyrA_results %>%
-  left_join(parC_results, by = "id") %>%
-  left_join(parE_results, by = "id") %>%
-  group_by(id) %>%
-  summarise_all(funs(func_paste)) %>%
-  mutate(parC_ariba = ifelse(parC_prokka == "S80I" & 
-                               parC_ariba == "S58I", "S80I", parC_ariba))
-
-full_report
-
-```
-
-# Figures
-
-## Sequence types per animal group
-```{r}
-palette4 <- c("Livestock" = "#80b1d3",
-              "Wild" = "#fb8072")
-
-mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  mutate(animal_group = ifelse(species %in% c("Broiler","Pig"), "Livestock", "Wild")) %>%
-  group_by(ST, animal_group) %>%
-  dplyr::count() %>%
-  group_by(ST) %>%
-  mutate(total = dplyr::first(n) + dplyr::last(n)) %>%
-  filter(total > 4) %>%
-  ggplot(aes(factor(ST), n, fill = animal_group))+
-  geom_col(color = "black")+
-  labs(x = "Sequence Types")+
-  scale_fill_manual(values = palette4)+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.45),
-        legend.title = element_blank())
-```
-
-## Resistance mechanisms per sequence type
-```{r}
-genes <- c(names(mut_report), names(acquired_report), "qnr")
-genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")]
-
-no_of_mlst <- mlst_data[,c("id","ST")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(ST, species) %>%
-  dplyr::count() %>%
-  group_by(ST) %>%
-  mutate(group_n = sum(n)) %>%
-  filter(group_n > 3)
-
-mlst_mechanisms_report <- mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
-            funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
-  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
-  gather(gene, value, -id) %>%
-  mutate(present = ifelse(value == 1, gene, NA),
-         ref = 1:n()) %>%
-  spread(gene, value) %>%
-  group_by(id) %>%
-  summarise_all(funs(func_paste)) %>%
-  left_join(mlst_data[c("id","ST")], by = "id") %>%
-  select(-c(ref, id)) %>%
-  ungroup() %>%
-  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
-  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
-         present = ifelse(present == "", "None", present)) %>%
-  gather(gene, value, -c(present,ST)) %>%
-  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
-         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
-  na.omit() %>%
-  group_by(ST, gene, value) %>%
-  dplyr::count() %>%
-  filter(ST %in% no_of_mlst$ST)
-
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
-
-p1 <- ggplot(mlst_mechanisms_report, aes(factor(ST), gene, size = n))+
-  geom_point()+
-  guides(size = FALSE)+
-  labs(x = "Sequence Types",
-       y = "Genes")+
-  theme(axis.text.x = element_text(size = 10),
-        axis.text.y = element_text(size = 10),
-        axis.ticks.x = element_blank())
-
-p2 <- ggplot(no_of_mlst, aes(factor(ST), n, fill = species))+
-  geom_col(color = "black")+
-  scale_fill_manual(values = palette)+
-  guides(fill = FALSE)+
-  theme(axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.text.x = element_blank())
-
-plot_grid(p2, p1, nrow = 2, align = "v")
-
-```
-
-## Percent resistance mechanisms in total
-```{r}
-total_mechanisms <- mut_report %>%
-  select(-nt_change) %>%
-  left_join(acquired_report, by = "id") %>%
-  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
-            funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
-  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
-  gather(gene, result, -id) %>%
-  group_by(gene, result) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
-  spread(result, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round((Present/Total)*100, 1),
-         lwr = round(get_binCI(Present, Total)[1], 1),
-         upr = round(get_binCI(Present, Total)[2], 1),
-         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))
-
-palette2 <- c("Acquired" = "#8dd3c7",
-              "Intrinsic" = "#80b1d3")
-
-ggplot(total_mechanisms, aes(reorder(gene, -Percent), Percent, fill = type)) +
-  geom_col(color = "black") +
-  geom_errorbar(aes(ymin = lwr, ymax = upr),
-                width = 0.5)+
-  scale_fill_manual(values = palette2) +
-  labs(x = "Genes",
-       y = "Percent (%) Isolates",
-       fill = NULL)
-
-```
-
-## Percent QREC isolates with mutations in intrinsic genes
-```{r}
-gene_names <- names(mut_report)
-
-gene_names <- gene_names[-c(1,2)]
-
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
-
-mut_report %>%
-    left_join(isolate_data, by = "id") %>%
-    gather(key, value, gene_names) %>%
-    group_by(key, value, species) %>%
-    dplyr::count() %>%
-    ungroup() %>%
-    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-    spread(value, n, fill = 0) %>%
-    rowwise() %>%
-    mutate(
-      Total = Present + Absent,
-      Percent = round(Present / Total * 100, 1),
-      lwr = round(get_binCI(Present, Total)[1], 1),
-      upr = round(get_binCI(Present, Total)[2], 1)
-    ) %>%
-    mutate(species = factor(species,
-                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
-    ggplot(aes(reorder(key, -Percent), Percent, fill = species)) +
-    geom_col(color = "black",
-             position = position_dodge(1)) +
-    geom_errorbar(
-      aes(ymin = lwr, ymax = upr),
-      width = 0.6,
-      position = position_dodge(1)
-    ) +
-    scale_fill_manual(values = palette) +
-    labs(y = "Percent (%) of isolates") +
-    guides(fill = FALSE) +
-    theme_classic() +
-    theme(
-      axis.text.x = element_text(
-        angle = 90,
-        size = 12,
-        hjust = 1,
-        vjust = 0.4
-      ),
-      axis.title.x = element_blank(),
-      axis.text.y = element_text(size = 12),
-      axis.title.y = element_text(size = 14),
-      strip.text = element_text(size = 12)
-    ) +
-    facet_wrap(~ species)
-```
-
-## Percent QRDR mutations per animal species
-```{r}
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
-
-suppressWarnings(clean_mut_data %>%
-  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
-         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
-         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
-         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
-         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
-         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
-  select(ref, gene_names, flag, mutation) %>%
-  filter(flag %in% flag_selection,
-         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
-  mutate(id = 1:n()) %>%
-  spread(gene_names, mutation) %>%
-  group_by(ref) %>%
-  summarise_all(funs(func_paste2)) %>%
-  dplyr::select(-c(id, flag)) %>%
-  left_join(id_data, by = "ref") %>%
-  dplyr::select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples) %>%
-  gather(gene, mut, -id) %>%
-  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
-         result = ifelse(mut != 0, 1, 0),
-         result = as.integer(result),
-         type = "mut") %>%
-  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
-  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
-  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
-  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
-  select(-mut) %>%
-  dplyr::rename("mut" = aa_change) %>%
-  filter(is.na(nt_change) == FALSE) %>%
-  fix_gyr_par_results(.) %>%
-  filter(result_total == 1) %>%
-  left_join(isolate_data[c("id","species")], by = "id") %>%
-  group_by(gene, mut, species) %>%
-  dplyr::count() %>%
-  rowwise() %>%
-  mutate(Total = case_when(species == "Broiler" ~ 87,
-                           species == "Pig" ~ 75,
-                           species == "Red fox" ~ 52,
-                           species == "Wild bird" ~ 66),
-         Percent = round(n/Total * 100,1),
-         lwr = round(get_binCI(n, Total)[1], 1),
-         upr = round(get_binCI(n, Total)[2], 1)) %>%
-  dplyr::rename("Mutation" = mut)) %>%
-  ggplot(aes(reorder(Mutation, -Percent), Percent, fill = species))+
-  geom_col(color = "black", position = position_dodge(0.9))+
-  geom_errorbar(aes(ymin = lwr, ymax = upr),
-                position = position_dodge(0.9),
-                width = 0.5)+
-  scale_fill_manual(values = palette)+
-  guides(fill = FALSE)+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
-        axis.title.x = element_blank())+
-  facet_wrap(~ gene, scale = "free_x")
-
-```
-
-## Percent QREC isolates with Qnr genes per species
-```{r}
-qnr_genes <- c("qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")
-
-acquired_report %>%
-    left_join(isolate_data, by = "id") %>%
-    gather(key, value, qnr_genes) %>%
-    group_by(key, value, species) %>%
-    dplyr::count() %>%
-    ungroup() %>%
-    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-    spread(value, n, fill = 0) %>%
-    rowwise() %>%
-    mutate(
-      Total = Present + Absent,
-      Percent = round(Present / Total * 100, 1)
-    ) %>%
-    ggplot(aes(key, Percent, fill = species)) +
-    geom_col(color = "black",
-             position = position_dodge(0.9)) +
-    scale_fill_manual(values = palette) +
-    labs(y = "Percent (%) of isolates",
-         fill = NULL) +
-    theme_classic() +
-    theme(
-      axis.text = element_text(size = 12),
-      axis.title.y = element_text(size = 14),
-      legend.text = element_text(size = 12),
-      axis.title.x = element_blank(),
-      legend.justification = c(0, 1),
-      legend.position = c(0.82, 0.97)
-    )
-```
-
-## Correlation plot for resistance mechanisms
-```{r, fig.height=8, fig.width=7, warning=FALSE}
-total_df <- acquired_report %>%
-  left_join(mut_report, by = "id") %>%
-  left_join(isolate_data[c("id","species")], by = "id") %>%
-  select(-c(nt_change, id)) %>%
-  mutate_at(vars(-species),
-            funs(as.numeric)) %>%
-  mutate(qnr = ifelse(rowSums(.[,2:8]) > 0, 1, 0)) %>%
-  {. ->> total_resistance_report} %>%
-  split(., f = .$species) %>%
-  lapply(., function(x) select(x, -species))
-
-create_corr_matrix <- function(df) {
-  df <- df %>%
-    as.matrix %>%
-    cor(method = "spearman") %>%
-    as.data.frame %>%
-    rownames_to_column(var = "var1") %>%
-    gather(var2, value, -var1) %>%
-    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
-           var2 = gsub("(.*?)_R$", "\\1", var2))
-  return(df)
-}
-
-corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
-  bind_rows(.id = "species") %>%
-  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
-           gene2 = gsub("(.*?)_.+", "\\1", var2),
-           var1 = gsub(".+_(.*?)", "\\1", var1),
-           var2 = gsub(".+_(.*?)", "\\1", var2),
-           type1 = case_when(gene1 == "gyrA" ~ 1,
-                          gene1 == "gyrB" ~ 2,
-                          gene1 == "parC" ~ 3,
-                          gene1 == "parE" ~ 4,
-                          gene1 == "marR" ~ 5,
-                          gene1 == "marA" ~ 6,
-                          gene1 == "rpoB" ~ 7,
-                          gene1 == "soxS" ~ 8,
-                          gene1 == "robA" ~ 9,
-                          gene1 == "gadX" ~ 10,
-                          gene1 == "qnrA1" ~ 11,
-                          gene1 == "qnrB19" ~ 12,
-                          gene1 == "qnrB6" ~ 13,
-                          gene1 == "qnrB60" ~ 14,
-                          gene1 == "qnrS1" ~ 15,
-                          gene1 == "qnrS2" ~ 16,
-                          gene1 == "qnrS4" ~ 17,
-                          gene1 == "qnr" ~ 18,
-                          gene1 == "qepA4" ~ 19),
-           type2 = case_when(gene2 == "gyrA" ~ 1,
-                          gene2 == "gyrB" ~ 2,
-                          gene2 == "parC" ~ 3,
-                          gene2 == "parE" ~ 4,
-                          gene2 == "marR" ~ 5,
-                          gene2 == "marA" ~ 6,
-                          gene2 == "rpoB" ~ 7,
-                          gene2 == "soxS" ~ 8,
-                          gene2 == "robA" ~ 9,
-                          gene2 == "gadX" ~ 10,
-                          gene2 == "qnrA1" ~ 11,
-                          gene2 == "qnrB19" ~ 12,
-                          gene2 == "qnrB6" ~ 13,
-                          gene2 == "qnrB60" ~ 14,
-                          gene2 == "qnrS1" ~ 15,
-                          gene2 == "qnrS2" ~ 16,
-                          gene2 == "qnrS4" ~ 17,
-                          gene2 == "qnr" ~ 18,
-                          gene2 == "qepA4" ~ 19))
-
-cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
-          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
-
-ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
-  geom_tile(color = "white")+
-  scale_fill_gradient(low = "black",
-                      high = "white",
-                      na.value = "grey95")+
-  labs(fill = NULL)+
-  theme_classic()+
-  theme(axis.text = element_text(size = 8),
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
-        axis.title = element_blank())+
-  coord_fixed()+
-  facet_wrap(~species)
-```
-
-## MIC-plots
-```{r}
-total_report <- mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  dplyr::select(-nt_change) %>%
-  mutate_at(vars(-id),
-            funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
-  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
-  gather(gene, value, -id) %>%
-  mutate(present = ifelse(value == 1, gene, NA),
-         ref = 1:n()) %>%
-  spread(gene, value) %>%
-  group_by(id) %>%
-  summarise_all(funs(func_paste)) %>%
-  select(-ref) %>%
-  left_join(mic_data[c("id","CIP")], by = "id") %>%
-  group_by(present,CIP) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  rowwise() %>%
-  mutate(present = ifelse(present == "", "None", present),
-         total = 280,
-         percent = round(n/total * 100, 1),
-         lwr = round(get_binCI(n, total)[1], 1),
-         upr = round(get_binCI(n, total)[2], 1)) %>%
-  ungroup() %>%
-  mutate(entries = sapply(strsplit(.$present, ","),
-                          FUN = function(x) {length(x)}))
-
-genes <- c(names(mut_report), names(acquired_report), "qnr")
-genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrS1","qnrS2","qnrS4")]
-
-n_plot <- ggplot(total_report, aes(reorder(present, entries), percent, fill = factor(CIP)))+
-  geom_col(color = "black")+
-  labs(y = "Percent (%) isolates")+
-  scale_fill_viridis(discrete = TRUE)+
-  guides(fill=guide_legend(ncol = 10))+
-  theme(axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.title.y = element_text(size = 10),
-        axis.text.y = element_text(size = 8),
-        axis.title.x = element_blank(),
-        legend.text = element_text(size = 8),
-        legend.title = element_blank(),
-        legend.position = "top",
-        legend.spacing.x = unit(0.2, "cm"),
-        legend.justification = "center")
-
-dot_report <- mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  select(-nt_change) %>%
-  mutate_at(vars(-id),
-            funs(as.numeric(.))) %>%
-  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
-  select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
-  gather(gene, value, -id) %>%
-  mutate(present = ifelse(value == 1, gene, NA),
-         ref = 1:n()) %>%
-  spread(gene, value) %>%
-  group_by(id) %>%
-  summarise_all(funs(func_paste)) %>%
-  left_join(mic_data[c("id","CIP")], by = "id") %>%
-  select(-c(ref, id)) %>%
-  ungroup() %>%
-  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
-  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
-         present = ifelse(present == "", "None", present)) %>%
-  gather(gene, value, -c(present,CIP)) %>%
-  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
-         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
-  na.omit() %>%
-  mutate(entries = sapply(strsplit(.$present, ","),
-                          FUN = function(x) {length(x)}))
-
-dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
-  geom_point(size = 2)+
-  scale_y_discrete(limits = c("gyrA","parC","parE","gadX","marA","marR","rpoB","soxS","qepA4","qnr"))+
-  guides(fill = FALSE)+
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_text(size = 8),
-        axis.title = element_blank(),
-        axis.ticks.x = element_blank())
-
-plot_grid(n_plot, dot_plot, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
-```
-
-## Additional resistance levels per animal species
-```{r}
-mic_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
-         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
-         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
-         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
-         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
-         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
-         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
-         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
-         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
-  select(species,contains("_R")) %>%
-  gather(compound, value, -species) %>%
-  mutate(compound = sub("_R", "", compound)) %>%
-  group_by_all() %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round((Present/Total)*100, 1),
-         lwr = round(get_binCI(Present, Total)[1], 1),
-         upr = round(get_binCI(Present, Total)[2], 1),
-         species = factor(species, levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
-  ggplot(aes(reorder(compound, -Percent), Percent, fill = species))+
-  geom_col(color = "black")+
-  geom_errorbar(aes(ymin = lwr, ymax = upr),
-                width = 0.6)+
-  guides(fill = FALSE)+
-  scale_fill_manual(values = palette)+
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))+
-  facet_wrap(~species)
-```
-
-## CIP/NAL MIC-values per animal species
-```{r}
-mic_data[,c("id","CIP")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(CIP,species) %>%
-  dplyr::count() %>%
-  ggplot(aes(factor(CIP), n, fill = species))+
-  geom_col(color = "black", position = position_fill())+
-  labs(y = "Percent isolates",
-       x = "CIP MIC")+
-  scale_fill_manual(values = palette)+
-  scale_y_continuous(labels = percent_format())+
-  theme(legend.title = element_blank())
-
-mic_data[,c("id","NAL")] %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(NAL,species) %>%
-  dplyr::count() %>%
-  ggplot(aes(factor(NAL), n, fill = species))+
-  geom_col(color = "black", position = position_fill())+
-  labs(y = "Percent isolates",
-       x = "NAL MIC")+
-  scale_fill_manual(values = palette)+
-  scale_y_continuous(labels = percent_format())+
-  theme(legend.title = element_blank())
-
-```

### 2019-04-24 notebooks/assembly_metrics.Rmd
minor changes

@@
                        "2016-02-486-2-S",
-                       "2016-02-732-2-S",
-                       "2014-01-1741-1-S"))
+                       "2016-02-732-2-S"))
 
@@
                              "52-2016-02-732_S43_pilon_spades",
-                             "50-2016-02-428_S19_pilon_spades",
-                             "4-2014-01-1741_TAAGGCGA-ACTGCATA_pilon_spades")) %>%
+                             "50-2016-02-428_S19_pilon_spades")) %>%
   mutate(ref = Assembly %>%
@@
 ```{r}
-print(paste0("Mean number of contigs: ", round(mean(metrics2$no_of_contigs), 1)))
-print(paste0("Mean size of largest contigs: ", round(mean(metrics2$largest_contig), 1)))
-print(paste0("Mean N50: ", round(mean(metrics2$N50), 1)))
-print(paste0("Mean NG50: ", round(mean(as.numeric(metrics2$NG50)), 1)))
+metrics3 <- filter(metrics2, id != "2014-01-1741-1-S")
+
+print(paste0("Mean number of contigs: ", round(mean(metrics3$no_of_contigs), 1)))
+print(paste0("Mean size of largest contigs: ", round(mean(metrics3$largest_contig), 1)))
+print(paste0("Mean N50: ", round(mean(metrics3$N50), 1)))
+print(paste0("Mean NG50: ", round(mean(as.numeric(metrics3$NG50)), 1)))
 
@@
 ```{r}
-p1 <- ggplot(metrics2, aes(species, N50)) +
+p1 <- ggplot(metrics3, aes(species, N50)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p2 <- ggplot(metrics2, aes(species, no_of_contigs)) +
+p2 <- ggplot(metrics3, aes(species, no_of_contigs)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p3 <- ggplot(metrics2, aes(species, largest_contig)) +
+p3 <- ggplot(metrics3, aes(species, largest_contig)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p4 <- ggplot(metrics2, aes(species, as.numeric(no_of_misassemblies))) +
+p4 <- ggplot(metrics3, aes(species, as.numeric(no_of_misassemblies))) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 ```{r}
-p5 <- ggplot(metrics2, aes(sequencer, N50)) +
+p5 <- ggplot(metrics3, aes(sequencer, N50)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p6 <- ggplot(metrics2, aes(sequencer, no_of_contigs)) +
+p6 <- ggplot(metrics3, aes(sequencer, no_of_contigs)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p7 <- ggplot(metrics2, aes(sequencer, largest_contig)) +
+p7 <- ggplot(metrics3, aes(sequencer, largest_contig)) +
   stat_boxplot(geom = "errorbar", width = 0.4) +
@@
 
-p8 <- ggplot(metrics2, aes(sequencer, as.numeric(no_of_misassemblies))) +
+p8 <- ggplot(metrics3, aes(sequencer, as.numeric(no_of_misassemblies))) +
   stat_boxplot(geom = "errorbar", width = 0.4) +

### 2019-04-24 figure_script.R
updated colors and text

@@ -43,7 +43,7 @@ total_tree_annotated <- ggtree(total_tree,
 total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)
 
 ggsave(
-  "C:/Users/vi1511/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/total_dendrogram.tiff",
+  "C:/Users/vi1511.VETINST/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/total_dendrogram.tiff",
   gheatmap(
     total_tree_opened,
     tree_heatmap_data,
@@ -77,7 +77,11 @@ ggsave(
 
 metadata_species <- read.table("data/chewbbaca/complete_results/metadata_species_specific_trees.txt",
                                sep = "\t",
-                               header = TRUE)
+                               header = TRUE) %>%
+  select(gyrA, parC, parE, PMQR) %>%
+  dplyr::rename("GyrA" = gyrA,
+                "ParC" = parC,
+                "ParE" = parE)
 
 palette3 <- c("0" = "grey95",
               # gyrA
@@ -102,139 +106,99 @@ palette3 <- c("0" = "grey95",
               "H516Y" = "#d94801",
               "L488M, A512T" = "#8c2d04",
               # PMQR
-              "qnrA1" = "#762a83",
-              "qnrB19" = "#ef8a62",
-              "qnrS1" = "#a6dba0",
-              "qnrS2" = "#5aae61",
-              "qnrS4" = "#1b7837",
-              "qepA4" = "#80cdc1",
-              # num
-              "1" = "#440154FF",
-              "2" = "#3B528BFF",
-              "3" = "#21908CFF",
-              "4" = "#5DC863FF",
-              "5" = "#FDE725FF")
+              "qnrA1" = "#dadaeb",
+              "qnrB19" = "#bcbddc",
+              "qnrS1" = "#9e9ac8",
+              "qnrS2" = "#807dba",
+              "qnrS4" = "#6a51a3",
+              "qepA4" = "#4a1486")
 
 # Broiler tree
-broiler_cgMLST <- read.table("data/chewbbaca/complete_results/broiler_cgmlst.txt",
-                             sep = "\t",
-                             header = TRUE,
-                             colClasses = "factor")
 
-broiler_tree <- as.phylo(hclust(daisy(broiler_cgMLST, metric = "gower"), method = "average"))
[…]

### 2019-04-24 alignment_script.R
removed unneccessary scripts

@@ -1,56 +1,0 @@
-# Testing import of gyrB sequences from .gz files created by ARIBA
-library(Biostrings)
-library(msa)
-library(DECIPHER)
-
-# File locations
-path <- "data/ariba/megares_full/"
-
-## Functions
-# Identifies filenames in input folder
-file_names <- function(filepath) {
-  files <- list.files(path = filepath,
-                      pattern = "assembled_genes.fa.gz",
-                      recursive = TRUE)
-  return(files)
-}
-
-# Import data from assembled_genes.fa.gz files created by ariba
-get_seq_data <- function(filepath) {
-  files <- file_names(filepath)
-  
-  data_list <- lapply(files,
-                      FUN = function(file) {
-                        readDNAStringSet(paste0(filepath, "/", file))
-                      })
-  names(data_list) <- files
-  return(data_list)
-}
-
-# Filter out specific gene of interest
-filter_gyrB <- function(fasta_file, gene) {
-  gyrB_pos <- which(grepl(gene, names(fasta_file), ignore.case = TRUE) == TRUE)
-  gyrB_seq <- fasta_file[gyrB_pos]
-  return(gyrB_seq)
-}
-
-# Append sequences to fasta file
-append_fasta_file <- function(list) {
-  i <- 1
-  for (x in list) {
-    write.fasta(x, names(list[i]), "data/sequences.fasta", open = "a")
-    i <- i + 1
-  }
-}
-
-## Run functions
-sequences <- get_seq_data(path)
-gyrB_sequences <- lapply(sequences, function(x) filter_gyrB(x, "gyrb"))
-gyrB_ref <- readDNAStringSet("data/gyrB_ref.fasta")
-write.fasta(gyrB_ref, "gyrB_ref", "data/sequences.fasta", open = "w")
-append_fasta_file(gyrB_sequences)
-
-sequences_gyrB <- readDNAStringSet("data/sequences.fasta")
-trans_seqs <- Biostrings::translate(sequences_gyrB)
-alignment <- AlignSeqs(trans_seqs, iterations = 0, refinements = 0)
-writeXStringSet(alignment, "data/alignment.fasta")

### 2019-04-24 ariba_script.R

@@ -1,464 +1,0 @@
-#!/usr/bin/env Rscript
-args <- commandArgs(trailingOnly = TRUE)
-
-megares_report_loc <- "data/ariba/megares_results"
-resfinder_report_loc <- "data/ariba/resfinder_results"
-ac_genes <- c("qnr","aac","oqx")
-mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")
-
-# adjust parameters for filtering
-if (length(ac_genes) == 1) {
-  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
-    ac_genes <- "ALL"
-    } else {
-      ac_genes <- ac_genes
-    }
-}
-
-if (length(mut_genes) == 1) {  
-  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
-    mut_genes <- "ALL"
-    } else {
-      mut_genes <- mut_genes
-    }
-}
-# -------------------------------------- Libraries
-
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
-               forcats,purrr,stringr,kableExtra,
-               knitr,IRdisplay,reprex,svglite)
-
-# -------------------------------------- Functions
-# Collapses data frame to unique lines while ignoring NA's
-func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
-
-# Confidence interval function
-get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
-                                                c("lwr", "upr")))
-
-# Identifies filenames in input folder
-file_names <- function(filepath) {
-  files <- list.files(path = filepath, pattern = "amr_report.tsv")
-  return(files)
-}
-
-# Import ariba data from report.tsv from chosen database used in ariba
-get_ariba_data <- function(filepath) {
-  files <- file_names(filepath)
-  
-  data_list <- lapply(files,
-                      FUN = function(file) {
-                        read.delim(
-                          paste0(filepath, "/", file),
-                          stringsAsFactors = F,
-                          header = TRUE,
-                          sep = "\t"
-                        )
-                      })
-  
-  names(data_list) <- files
-  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
[…]

### 2019-04-24 prokka_check.R

@@ -1,163 +1,0 @@
-library(stringr)
-library(seqinr)
-library(tidyr)
-library(Biostrings)
-
-gyrA <- readAAStringSet("data/Prokka/DNA_gyrase_GYRA_aa_sequences.fa")
-parC <- readAAStringSet("data/Prokka/Topoisomerase_IV_subunit_A_aa_sequences.fa")
-
-mismatches <- function(query, ref) {
-  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
-                    gapOpening = 3, gapExtension = 1) %>%
-    mismatchTable() %>%
-    mutate(ID=names(query),
-           Pos=PatternStart,
-           Reference_AA=as.character(PatternSubstring),
-           Sample_AA=as.character(SubjectSubstring)) %>%
-    select(ID, Reference_AA, Sample_AA, Pos) %>%
-    filter(Sample_AA != "X")
-}  
-
-paste3 <- function(...,sep=", ") {
-  L <- list(...)
-  L <- lapply(L,function(x) {x[is.na(x)] <- ""; x})
-  ret <-gsub(paste0("(^",sep,"|",sep,"$)"),"",
-             gsub(paste0(sep,sep),sep,
-                  do.call(paste,c(L,list(sep=sep)))))
-  is.na(ret) <- ret==""
-  ret
-}
-
-ariba_flags_gyra <- mut_flags %>%
-  select(-flag) %>%
-  filter(gene == "gyrA") %>%
-  mutate(gyrA_result = ref_ctg_change %>%
-           str_extract_all("\\d+") %>% 
-           map(as.integer) %>%
-           map_lgl(~ any(.x >= 67L & .x <= 106L))) %>%
-  mutate(type = case_when(gyrA_result == TRUE ~ "QRDR",
-                          gyrA_result == FALSE & ref_ctg_change != "." ~ "non_QRDR",
-                          gyrA_result == FALSE & ref_ctg_change == "." ~ "wild type")) %>%
-  select(-gyrA_result) %>%
-  group_by(id) %>%
-  mutate(test = 1:n()) %>%
-  ungroup()
-
-mut <- mut_quant %>%
-  select(id, gyrA) %>%
-  mutate(entries = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)})) %>%
-  separate(gyrA, into = as.character(c(1:max(.$entries)))) %>%
-  select(-entries) %>%
[…]

### 2019-04-24 functions.R
removed unneccessary scripts

@@ -1,12 +1,0 @@
-# Functions
-
-## Save plots
-save_plots <- function(filename, plot, dpi) {
-  ggsave(filename,
-         plot,
-         device = "tiff",
-         units = "cm",
-         height = 20,
-         width = 25,
-         dpi = dpi)
-}

### 2019-04-24 notebooks/article_notebook2.Rmd
removed some headers

@@
 
-# Libraries
-```{r load libraries, message=FALSE, warning=FALSE}
+```{r include=FALSE}
 # Libraries
@@
 
-```{r message=FALSE}
+```{r include=FALSE}
 source("../scripts/ARIBA_analysis.R")
-
 ```
 
-# Tables
-
-## Group listing
+# Group listing
 This table presents the groups created in the isolate selection.
@@
 
-## Mechanisms in isolates without substitutions in QRDR
+# Mechanisms in isolates without substitutions in QRDR
 This table presents the mechanisms identified in isolates with no amino acid substitutions in either of GyrA, GyrB, ParC and/or ParE.
@@
 
-## Substitutions in intrinsic genes for isolates with CIP MIC < 0.12
+# Substitutions in intrinsic genes for isolates with CIP MIC < 0.12
 This table presents the substitutions identified in isolates with a MIC-value for ciprofloxacin below the ECOFF.

### 2019-05-05 figure_script.R
changed node size in legend

@@ -1,5 +1,13 @@
 ## Script for generating figures for the article
 
+library(ggplot2)
+library(dplyr)
+library(ggtree)
+library(tibble)
+library(vegan)
+library(cluster)
+library(ape)
+
 ## Total dendrogram
 cgMLST_data <- read.table("data/chewbbaca/complete_results/clean_cgMLST_data.txt",
                           sep = "\t",
@@ -15,10 +23,8 @@ tree_metadata <- read.table("data/chewbbaca/complete_results/tree_metadata.txt",
 
 tree_heatmap_data <- read.table("data/chewbbaca/complete_results/tree_heatmap_data.txt",
                                 sep = "\t",
-                                header = TRUE) %>%
-  select(-gadX)
+                                header = TRUE)
 
-names(tree_heatmap_data) <- c("GyrA","ParC","ParE","RpoB","MarA","MarR","SoxR","qepA4","qnr")
 
 palette <- c("Broiler" = "#4575b4",
              "Pig" = "#74add1",
@@ -37,34 +43,39 @@ total_tree_annotated <- ggtree(total_tree,
                 size = 1.5) +
   geom_tiplab2(aes(label = ST),
                offset = 0.01,
-               size = 2.5) +
+               size = 2.1) +
   scale_color_manual(values = palette)
 
-total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)
+total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 94)
+
+total_tree_heatmap <- gheatmap(
+  total_tree_opened,
+  tree_heatmap_data,
+  offset = 0.04,
+  width = 0.3,
+  colnames_offset_y = 2.5,
+  colnames_position = "top",
+  font.size = 3
+) +
+  scale_fill_manual(
+    values = palette2,
+    labels = c(
+      "Gene/Mutation absent",
+      "Acquired gene present",
+      "Intrinsic gene mutated"
+    )
+  ) +
+  theme(
+    legend.position = c(0.9, 0.9),
+    legend.text = element_text(size = 10)
+  ) +
+  
+  guides(colour = guide_legend(override.aes = list(size=5)))
+
 
 ggsave(
   "C:/Users/vi1511.VETINST/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/total_dendrogram.tiff",
-  gheatmap(
-    total_tree_opened,
-    tree_heatmap_data,
-    offset = 0.04,
[…]

### 2019-05-05 notebooks/Clustering analysis.Rmd
removed some figures and changed headings

@@
 library(knitr)
+library(vegan)
 library(distanceR)
@@
 
-# Total dendrogram of all isolates
-Genetic distances were calculated from the allele data from the cgMLST report, and a dendrogram was drawn from these distances.
-The annotations are based on data from the ARIBA resistance gene analysis.
-```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
-cgMLST_data <- read.table("../data/chewbbaca/complete_results/clean_cgMLST_data.txt",
-                   sep = "\t",
-                   header = TRUE,
-                   colClasses = "factor")
-
-total_tree <- as.phylo(hclust(daisy(cgMLST_data, metric = "gower"), method = "average"))
-
-tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
-                            sep = "\t",
-                            header = TRUE)
-
-tree_heatmap_data <- read.table("../data/chewbbaca/complete_results/tree_heatmap_data.txt",
-                            sep = "\t",
-                            header = TRUE) %>%
-  select(-gadX)
-
-names(tree_heatmap_data) <- c("GyrA","ParC","ParE","RpoB","MarA","MarR","SoxR","qepA4","qnr")
-
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
-
-palette2 <- c("1-A" = "#fc8d59",
-              "1-I" = "#80b1d3",
-              "0" = "grey95")
-
-total_tree_annotated <- ggtree(total_tree,
-                               layout = "circular",
-                               color = "#808080",
-                               size = 0.5) %<+% tree_metadata +
-                               geom_tippoint(aes(color = species),
-                               size = 3) +
-                               geom_tiplab2(aes(label = ST),
-                               offset = 0.01,
-                               size = 4) +
-  scale_color_manual(values = palette)
-
-total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)
-
-gheatmap(total_tree_opened,
-         tree_heatmap_data,
-         offset = 0.04,
-         width = 0.3,
-         colnames_offset_y = 2.5,
-         colnames_position = "top",
-         font.size = 5) +
-  scale_fill_manual(values = palette2,
-                    labels = c("Gene/Mutation absent",
-                               "Acquired gene present",
-                               "Intrinsic gene mutated")) +
-  theme(legend.position = c(0.9,0.9),
-        legend.text = element_text(size = 20))
-```
-
-# Dendrograms per animal species
-```{r, fig.height=25, fig.width=25, message=FALSE, warning=FALSE}
-
-metadata_species <- read.table("../data/chewbbaca/complete_results/metadata_species_specific_trees.txt",
-                               sep = "\t",
-                               header = TRUE) %>%
-  select(gyrA, parC, parE, PMQR) %>%
-  dplyr::rename("GyrA" = gyrA,
-         "ParC" = parC,
-         "ParE" = parE)
-
-palette3 <- c("0" = "grey95",
-              # gyrA
-              "S83L" = "#c6dbef",
-              "S83A" = "#9ecae1",
-              "D87N" = "#6baed6",
-              "D87H" = "#4292c6",
-              "D87Y" = "#2171b5",
-              "D87G" = "#08519c",
-              "S83L, D87N" = "#08306b",
-              # parC
-              "S80I" = "#a1d99b",
-              "S80R" = "#74c476",
-              "S57T" = "#41ab5d",
-              "A56T, S80I" = "#238b45",
-              "S80I, E84V" = "#005a32",
-              # parE
-              "D475E" = "#fdd0a2",
-              "D463N" = "#fdae6b",
-              "S458A" = "#fd8d3c",
-              "L416F" = "#f16913",
-              "H516Y" = "#d94801",
-              "L488M, A512T" = "#8c2d04",
-              # PMQR
-              "qnrA1" = "#dadaeb",
-              "qnrB19" = "#bcbddc",
-              "qnrS1" = "#9e9ac8",
-              "qnrS2" = "#807dba",
-              "qnrS4" = "#6a51a3",
-              "qepA4" = "#4a1486")
-
-# Broiler tree
-
-broiler_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_broiler.tsv")
-
-broiler_tree_annotated <- ggtree(broiler_tree,
-                                 layout = "circular",
-                                 color = "#808080",
-                                 size = 0.5) %<+% tree_metadata +
-  geom_tippoint(color = "#4575b4",
-                size = 4) +
-  geom_tiplab2(aes(label = ST),
-               offset = 0.02,
-               size = 6)
-
-broiler_complete <- gheatmap(broiler_tree_annotated,
-         metadata_species,
-         offset = 0.1,
-         width = 0.5,
-         colnames = FALSE) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-# Pig tree
-pig_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_pig.tsv")
-
-pig_tree_annotated <- ggtree(pig_tree,
-                                 layout = "circular",
-                                 color = "#808080",
-                                 size = 0.5) %<+% tree_metadata +
-  geom_tippoint(color = "#74add1",
-                size = 4) +
-  geom_tiplab2(aes(label = ST),
-               offset = 0.02,
-               size = 5.5)
-
-pig_complete <- gheatmap(pig_tree_annotated,
-         metadata_species,
-         offset = 0.09,
-         width = 0.5,
-         colnames = FALSE) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-# Red fox tree
-fox_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_fox.tsv")
-
-fox_tree_annotated <- ggtree(fox_tree,
-                                 layout = "circular",
-                                 color = "#808080",
-                                 size = 0.5) %<+% tree_metadata +
-  geom_tippoint(color = "#f46d43",
-                size = 4) +
-  geom_tiplab2(aes(label = ST),
-               offset = 0.02,
-               size = 5.5)
-
-fox_complete <- gheatmap(fox_tree_annotated,
-         metadata_species,
-         offset = 0.09,
-         width = 0.5,
-         colnames = FALSE) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-# Wild bird tree
-bird_tree <- calc_tree("../data/chewbbaca/complete_results/cgMLST_bird.tsv")
-
-bird_tree_annotated <- ggtree(bird_tree,
-                                 layout = "circular",
-                                 color = "#808080",
-                                 size = 0.5) %<+% tree_metadata +
-  geom_tippoint(color = "#fdae61",
-                size = 4) +
-  geom_tiplab2(aes(label = ST),
-               offset = 0.02,
-               size = 5.5)
-
-bird_complete <- gheatmap(bird_tree_annotated,
-         metadata_species,
-         offset = 0.09,
-         width = 0.5,
-         colnames = FALSE) +
-  scale_fill_manual(values = palette3) +
-  guides(fill = FALSE)
-
-# Total plot
-tot_plot <- plot_grid(broiler_complete,
-          pig_complete,
-          fox_complete,
-          bird_complete,
-          nrow = 2,
-          ncol = 2,
-          align = "h",
-          labels = c("Broiler","Pig","Red fox","Wild bird"))
-
-include_graphics("../figures/species_dendrogram2.png")
-```
-
 # Minimum spanning tree
@@
 
-# Statistics
-## Distribution of calculated distances per animal species
-Below follows a plot of the distribution of the lower triangular distances for each animal species.
-A bimodal distribution is present in all four plots.
-The filtering values of < 0.1 and > 0.8 is used for downstream statistics to acertain whether the broiler isolates are more homogenous than the isolates from other species.
-
+# Allele distance distribution
 ```{r}
@@
 
-## Distribution of iterated distances
+# Distribution of iterated distances
 Distribution of percentages of distances < 0.1 or > 0.8.
@@
 
+# NMDS Analysis of resistance mechanisms
+Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types.
+```{r}
+res_mechanisms <- read.table("../data/MDS_data.txt",
+                       sep = "\t",
+                       header = TRUE,
+                       stringsAsFactors = FALSE) %>%
+  column_to_rownames("id")
+
+bin_dist <- dist(res_mechanisms,
+                 method = "binary")
+
+nmds_df <- metaMDS(bin_dist,
+                   trymax = 200,
+                   trace = FALSE) # silence the output
+
+print(paste0("Stress: ", round(nmds_df$stress, 3)))
+
+stressplot(nmds_df)
+
+```
+
 # References

### 2019-05-05 scripts/ARIBA_analysis.R
minor changes
changed user on commit

@@ -6,8 +6,8 @@ library(tibble)
 library(purrr)
 
 # Report directories
-megares_report_loc <- "../data/ariba/megares_results"
-resfinder_report_loc <- "../data/ariba/resfinder_results"
+megares_report_loc <- "../data/ariba/megares_results/"
+resfinder_report_loc <- "../data/ariba/resfinder_results/"
 
 # Genes of interest. These genes are partially matched to 
 # the generated gene names from the "fix_gene_names" function

### 2019-05-05 scripts/chisq_tests.R
minor changes

@@ -122,3 +122,49 @@ p_values_total <- lapply(mechanism_per_species, function(x) do_chisq_test(x)) %>
 p_values_total %>%
   kable() %>%
   kable_styling(bootstrap_options = c("striped", "hover"))
+
+
+
+## PMQR determinants
+test <- acquired_report %>%
+  mutate_at(vars(-id),
+            funs(as.numeric)) %>%
+  mutate(PMQR = ifelse(rowSums(.[,2:9]) > 0, 1, 0)) %>%
+  select(id, PMQR) %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(species, PMQR) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(PMQR = ifelse(PMQR == 1, "Present", "Absent")) %>%
+  spread(PMQR, n, fill = 0) %>%
+  dplyr::slice(expand.grid(1:length(unique(species)),
+                           1:length(unique(species))) %>%
+                 rev %>% 
+                 dplyr::filter(Var2 < Var1) %>% t) %>%
+  # create test ID
+  mutate(test = rep(1:(n()/2), each = 2)) %>%
+  group_by(test) %>%
+  # specify values for chi-squred tests
+  dplyr::do(data_frame(test = dplyr::first(.$test),
+                       species1 = dplyr::first(.$species),
+                       species2 = dplyr::last(.$species),
+                       data = list(matrix(c(.$Absent,
+                                            .$Present),
+                                          ncol = 2)))) %>%
+  # do chi-squared tests
+  mutate(chi_test = map(data, chisq.test,
+                        correct = FALSE)) %>%
+  mutate(p.value = round(map_dbl(chi_test, "p.value"),3),
+         df = map_dbl(chi_test, "parameter"),
+         x_sq = round(map_dbl(chi_test, "statistic"),2),
+         dupl = duplicated(test)) %>%
+  # filter out duplicated tests
+  dplyr::filter(dupl == FALSE) %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(n = sum(unlist(data))) %>%
+  select(species1, species2, n, df, x_sq, p.value) %>%
+  # round off p-value column
+  mutate(p.value = round(p.value, 5)) %>%
+  dplyr::filter(p.value <= 0.05)
+

### 2019-05-05 notebooks/fastqc_analysis.Rmd

---
title: "Fastqc Analysis of Raw Reads"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This analysis depends on the following packages:

- fastqcr

- dplyr

- ggplot2

- tidyr

- viridis

- ggsci

- scales

```{r, include = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(fastqcr,dplyr,ggplot2,tidyr,viridis,ggsci,scales)
```

```{r, include = FALSE}
input_dir <- "C:/Users/VI1511/OneDrive - Veterinærinstituttet/R/R_Projects/qrec_wgs1/data/fastqc"
```

# Functions
```{r}
# Function that lists all the file names in the input folder
file_names <- function(filepath, folder) {
  files <- list.files(path = paste0(filepath, "/", folder), pattern = "_fastqc.zip")
  return(files)
}

# Function that searches recursively for all filenames with fastqc.zip
file_names_recursive <- function(filepath) {
  files <- list.files(path = filepath, pattern = "_fastqc.zip", recursive = TRUE)
  return(files)
}

# Lists folders in the input folder
folder_names <- function(filepath) {
  folders <- list.files(path = filepath)
  return(folders)
}

# Identifies the names of the files and groups them accoring to their respective folders
get_grouping_variable <- function(path, folder) {
  files <- file_names(path, folder)
  df <- data.frame(files = files, group = folder, stringsAsFactors = FALSE)
  df$files <- gsub("(.*?)_fastqc.zip", "\\1", df$files)
  colnames(df) <- c("ref","group")
  return(df)
}

# Creates a data frame with grouping information of the reports
create_group_df <- function(path) {
  folders <- folder_names(path)
  if (length(folders) > 1) {
    df <- lapply(folders, function(folder) get_grouping_variable(path, folder))
    df <- bind_rows(df)
  } else {
    df <- get_grouping_variable(path, folders)
  }
  return(df)
}

# Function that filter out duplicated counts
filter_counts <- function(df) {
  df <- df %>%
    mutate(dupl = duplicated(Count)) %>%
    filter(dupl == FALSE)
  return(df)
}

# Function that prepares the sequence length data for plotting
prepare_seq_len_data <- function(list) {
  x <- split(list$sequence_length_distribution, list$sequence_length_distribution$group)
  x <- lapply(x, filter_counts)
  x <- bind_rows(x)
  return(x)
}

# Function that imports and wrangles fastqc data
get_fastqc_data <- function(filepath) {
  folders <- folder_names(filepath)

  get_files <- file_names_recursive(filepath)

  data_list <- lapply(get_files,
                      FUN = function(file) {
                        qc_read(paste0(filepath, "/", file),
                                modules = "all",
                                verbose = FALSE)
                      })
  names(data_list) <- gsub("(.*?)/(.*?)_fastqc.zip", "\\2", get_files)
  data_list <- purrr::transpose(data_list)

  data_list$sequence_length_distribution <- NULL
  data_list$kmer_content <- NULL

  list_names <- names(data_list)
  list_numbers <- 1:length(list_names)
  for (i in list_numbers) {
    assign(list_names[i], bind_rows(data_list[[i]], .id = "ref"))
  }
  df_list <- list(summary,
                  basic_statistics,
                  per_base_sequence_quality,
                  per_tile_sequence_quality,
                  per_sequence_quality_scores,
                  per_base_sequence_content,
                  per_sequence_gc_content,
                  per_base_n_content,
                  sequence_duplication_levels,
                  overrepresented_sequences,
                  adapter_content,
                  total_deduplicated_percentage)
  names(df_list) <- list_names
  df_list$basic_statistics <- df_list$basic_statistics %>%
    spread(Measure,Value) %>%
    left_join(group_df, by = "ref")
  return(df_list)
}
```

# Grouping
This function creates a data frame with the grouping information of the files, based on the folder names that hold them. Each file is specified a group name based on the folder they are in. This information is used to specify each group of files in the plots below.

```{r}
group_df <- create_group_df(input_dir)
```

# Import data
This function imports the data from the location specified.
```{r}
df_list <- get_fastqc_data(input_dir)
```

# Plotting

## Adapter content
This plot presents the adapter content for all sequences in each group. Each adapter type is specified by color. The X-axis represents the read positions, where the left side represents the start of the read and the right side the end.

```{r, fig.height=8, fig.width=12}
df_list$adapter_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  # get data ready for boxplot by gathering the columns into two columns
  gather(key,
  value,-c(ref,
  Position,
  group,
  seqlen)) %>%
  # plotting
  ggplot(aes(factor(
  Position,
  levels = unique(Position),
  ordered = TRUE
  ), value, color = key)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(
  x = "Position in Read",
  y = "Percent (%) Adapter Content",
  color = NULL,
  title = "Adapter content"
  ) +
  scale_colour_jama() +
  scale_y_continuous(limits = c(0, 20)) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.title = element_text(size = 14),
  axis.ticks.x = element_blank(),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom"
  ) +
  # creates separate windows for each group
  facet_wrap( ~ group, scales = "free_x", dir = "v")
```

## Per base sequence content
This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  gather(key, value,-c(ref, Base, group, seqlen)) %>%
  ggplot(aes(factor(
  Base, levels = unique(Base), ordered = TRUE
  ), value, color = key)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(
  x = "Position in Read",
  y = "Percent (%)",
  color = NULL,
  title = "Per base sequence content"
  ) +
  theme_classic() +
  scale_color_jama() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom"
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x", dir = "v")
```

## Sequence duplication levels
This plot presents the duplication levels for all reads in each group. The X-axis contain "bins" of different duplication levels, and the number is a representation of how many of the same read is present. For example, the bin "3" represents the reads that have triplicates in the file.

```{r, fig.height=8, fig.width=12}
df_list$sequence_duplication_levels %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  gather(key, value,-c(ref, `Duplication Level`, group, seqlen)) %>%
  ggplot(aes(
  factor(
  `Duplication Level`,
  levels = unique(`Duplication Level`),
  ordered = TRUE
  ),
  value,
  fill = key
  )) +
  geom_boxplot(outlier.size = 0.5) +
  scale_fill_manual(values = c("#ef8a62",
  "#67a9cf")) +
  theme_classic() +
  labs(x = "Duplication Level",
  y = "Percent (%) of Sequences",
  fill = NULL,
  title = "Sequence duplication levels") +
  theme(
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 12),
  legend.position = "bottom",
  axis.text.x = element_text(
  size = 12,
  angle = 90,
  hjust = 1,
  vjust = 0.4
  )
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x")
```

## Sequence quality scores
This plot presents the sequence quality scores for all reads in each group. The X-axis represent the phred-score for each read, and the y-axis represent the amount of reads per phred-score. Higher score is better.

```{r, fig.height=8, fig.width=12}
df_list$per_sequence_quality_scores %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(Quality), Count, fill = factor(Quality))) +
  geom_boxplot(outlier.size = 0.5) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Quality",
  y = "Number of reads",
  title = "Per sequence quality scores") +
  guides(fill = FALSE) +
  theme_classic() +
  theme(
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group)
```

## Per sequence GC-content
This plot presents the % GC-content for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$per_sequence_gc_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(`GC Content`), Count)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(x = "GC content (%)",
  y = "Number of reads",
  title = "Per sequence GC content") +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = as.character(seq(
  from = 0, to = 100, by = 10
  ))) +
  theme_classic() +
  theme(
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x")
```

## Per base N-content
This plot presents the per base N content for all reads in each group. N's are added to the read instead of a base if the sequencer is unable to do the base call with sufficient confidence. As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
```{r, fig.height=8, fig.width=12}
df_list$per_base_n_content %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  ggplot(aes(factor(
  Base, levels = unique(Base), ordered = TRUE
  ), `N-Count`)) +
  geom_boxplot(fill = "#e6e6e6",
  outlier.size = 0.5) +
  labs(x = "Position in read",
  title = "Per base N content") +
  guides(fill = FALSE) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12)
  ) +
  # creates separate windows for each group
  facet_wrap(~ group, scales = "free_x", dir = "v")
```

## Total deduplicated percentage
This plot presents the total deduplicated percentage for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$total_deduplicated_percentage %>%
  gather(key = "ref", value = `1`) %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length",
  perc = `1`) %>%
  mutate(perc = as.numeric(perc)) %>%
  ggplot(aes(factor(group), perc)) +
  geom_boxplot(fill = "#e6e6e6",
  outlier.size = 0.5) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "Group",
  y = "Total percentage of deduplicated reads",
  title = "Total deduplicated percentage") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
  axis.title = element_text(size = 14))
```

## Per base sequence quality
This plot presents the per base mean sequence quality for all reads in each group. The green line depicts the "good quality" limit of the phred-score 28. As before, the X-axis represents the position in each read.
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_quality %>%
  # import information from another data frame from same list,
  # to get grouping information etc.
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename(seqlen = "Sequence length") %>%
  mutate(Base = factor(Base,
  levels = unique(Base),
  ordered = TRUE)) %>%
  group_by(group) %>%
  mutate(xmax = length(unique(Base)) + 1) %>%
  ungroup() %>%
  ggplot(aes(Base,
  Mean)) +
  geom_boxplot(outlier.size = 0.4,
  fill = "#7f7f7f") +
  geom_hline(aes(yintercept = 28),
  color = "green") +
  labs(x = "Position in read",
  y = "Sequence quality",
  title = "Per base mean sequence quality") +
  scale_y_continuous(limits = c(0, 42)) +
  theme_classic() +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title = element_text(size = 14),
  strip.text = element_text(size = 14),
  axis.text.y = element_text(size = 12)
  ) +
  # creates separate windows for each group
  facet_wrap( ~ group, scales = "free_x", dir = "v")
```

## Size of genome per sequencing method
```{r}
df_list$basic_statistics %>%
  left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
  rename("tot_seq" = `Total Sequences`) %>%
  mutate(tot_seq = as.numeric(tot_seq)) %>%
  ggplot(aes(group.x, tot_seq))+
  geom_boxplot()+
  labs(y = "Total Sequences")+
  theme_classic()+
  theme(axis.title.x = element_blank())
```

### 2019-05-05 notebooks/mash_analysis.Rmd

@@
 
-mash_plot_df %>%
+test2 <- mash_plot_df %>%
   select(-c(phage,shared_hashes,query_id,query_comment,median_multiplicity,test,p_value)) %>%

### 2019-05-05 notebooks/Clustering analysis.Rmd
minor changes to headers and removed pacman library handling

@@
 
-```{r libraries, include = FALSE}
+```{r libraries, message=FALSE, warning=FALSE}
 library(dplyr)
@@
 ```
-
-# Core genome MLST schema and quality control
-The _E. coli_ scheme from Enterobase [1] was used with chewBBACA [2] to get the cgMLST profile for each isolate.
-The number of missing loci in total (sum of LNF, NIPH, PLOT, ALM and ASM from chewBBACA) is plotted below for all isolates.
 
@@
 ```
+
+# Core genome MLST schema and quality control
+The _E. coli_ scheme from Enterobase [1] was used with chewBBACA [2] to get the cgMLST profile for each isolate.
+The number of missing loci in total (sum of LNF, NIPH, PLOT, ALM and ASM from chewBBACA) is plotted below for all isolates.
 

### 2019-05-05 notebooks/mash_analysis.Rmd
minor changes to headers and removed pacman library handling

@@
 ---
-title: "Quality control - Mash screen"
+title: "Mash screen results"
 author: "Håkon Kaspersen"
@@
 
-```{r}
-pacman::p_load(ggplot2, dplyr, tidyr, stringr, purrr, viridis)
+```{r, warning=FALSE, message=FALSE}
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(purrr)
+library(viridis)
 ```

### 2019-05-05 notebooks/article_notebook2.Rmd
added gyrB substitutions

@@
 
-```{r include=FALSE}
+```{r message = FALSE, warning=FALSE}
 # Libraries
@@
 
-```{r Import data, include = FALSE}
+```{r Import data, message=FALSE, warning=FALSE}
 # Vector of excluded samples
@@
 ```
+
+# Substitutions in GyrB outside QRDR
+```{r}
+mut_filtered %>%
+  filter(gene == "gyrB") %>%
+  group_by(mut) %>%
+  count() %>%
+  filter(mut != "0") %>%
+  dplyr::rename("AA Substitution" = mut) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
+```

### 2019-05-05 notebooks/assembly_metrics.Rmd

@@
 
+```{r, message=FALSE, warning=FALSE}
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(cowplot)
+library(scales)
+```
+
 ```{r, include = FALSE}
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, scales)
-
 "%not_in%" <- Negate("%in%")
@@
 
-# Import files and clean data
-```{r}
+```{r, message=FALSE, warning=FALSE}
 files <- file_names(path)
@@
 metrics <- df_clean %>%
-  select(id, no_of_contigs, largest_contig, GC, N50, NG50, L50, LG50, no_of_misassemblies, duplication_ratio, genes) %>%
-  mutate_at(vars(no_of_contigs, largest_contig, N50),
+  select(id,
+         no_of_contigs,
+         largest_contig,
+         GC,
+         N50,
+         NG50,
+         L50,
+         LG50,
+         no_of_misassemblies,
+         duplication_ratio,
+         genes) %>%
+  mutate_at(vars(no_of_contigs,
+                 largest_contig,
+                 N50),
             funs(as.numeric))
@@
 
-embed_data <- function(x = mtcars, filename = "file.csv", label= "Get data") {
+embed_data <- function(x = mtcars,
+                       filename = "file.csv",
+                       label= "Get data") {
 
@@
     write.csv2(x, "./file.csv")
-    enc= sprintf('data:text/csv;base64,%s', openssl::base64_encode(paste0(readLines("./file.csv"), collapse="\n")) )
+    enc= sprintf('data:text/csv;base64,%s',
+                 openssl::base64_encode(
+                   paste0(
+                     readLines("./file.csv"),
+                     collapse="\n")
+                   )
+                 )
     unlink("./file.csv")

### 2019-05-05 notebooks/fastqc_analysis.Rmd

@@
 ---
-title: "Fastqc Analysis of Raw Reads"
+title: "Fastqc Analysis of raw reads"
 author: "Håkon Kaspersen"
@@
 
-# Dependencies
-This analysis depends on the following packages:
-
-- fastqcr
-
-- dplyr
-
-- ggplot2
-
-- tidyr
-
-- viridis
-
-- ggsci
-
-- scales
-
-```{r, include = FALSE}
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load(fastqcr,dplyr,ggplot2,tidyr,viridis,ggsci,scales)
+```{r, message=FALSE, warning=FALSE}
+library(fastqcr)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(viridis)
+library(ggsci)
+library(scales)
 ```
@@
 
-# Functions
 ```{r}

### 2019-05-05 notebooks/Clustering analysis.Rmd

@@
 # Allele distance distribution
-```{r}
+Basic statistics are presented in the table.
+```{r, message=FALSE, warning=FALSE}
 # calculates distances from cgMLST data and returns
@@
 
-```
-
-# Distribution of iterated distances
-Distribution of percentages of distances < 0.1 or > 0.8.
-The distribution is based on 1000 iterations of the distance data.
-Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird.
-The mean value is denoted as a vertical black line.
-```{r, warning=FALSE, message=FALSE, dpi=300}
 # generate data frame from allele data
@@
 
+run_df %>%
+  mutate(total_mean = round(mean(value), 3)) %>%
+  group_by(group1) %>%
+  mutate(mean = round(mean(value), 3),
+         var = round(var(value), 3)) %>%
+  select(group1, total_mean, mean, var) %>%
+  summarise_all(funs(func_paste))
+```
+
+# Distribution of iterated distances
+Distribution of percentages of distances < 0.1 or > 0.8.
+The distribution is based on 1000 iterations of the distance data.
+Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird.
+The mean value is denoted as a vertical black line.
+```{r, warning=FALSE, message=FALSE, dpi=300}
 # functions used in iteration analyses
@@
 # Run functions
-summary_stats <- run_df %>%
-  mutate(total_mean = round(mean(value), 3)) %>%
-  group_by(group1) %>%
-  mutate(mean = round(mean(value), 3),
-         var = round(var(value), 3)) %>%
-  select(group1, total_mean, mean, var) %>%
-  summarise_all(funs(func_paste))
 
@@
   mutate(Percent_below = as.numeric(Percent_below))
-
-summary_stats
 

### 2019-05-06 notebooks/mash_analysis.Rmd
fixed error

@@
 # Contaminated Isolates
-```{r}
+```{r, warning=FALSE, message=FALSE}
 test <- mash_plot_df %>%
@@
 
-test2 <- mash_plot_df %>%
+mash_plot_df %>%
   select(-c(phage,shared_hashes,query_id,query_comment,median_multiplicity,test,p_value)) %>%

### 2019-05-21 figure_script.R
minor changes

@@ -370,7 +370,7 @@ plot_data <- as.data.frame(nmds_df$points) %>%
   mutate(ST = sub("_.+","",ID),
          ST = factor(ST, levels = c("10","58","117","155","162","355")))
 
-nmds_plot <- ggplot(plot_data, 
+ggplot(plot_data, 
        aes(x = MDS1,
            y = MDS2,
            fill = ST)) +

### 2019-05-22 notebooks/Clustering analysis.Rmd
added tables and minor fixes

@@
 library(tibble)
-library(cowplot)
 library(knitr)
@@
                            header = T,
-                           stringsAsFactors = F) %>%
+                           stringsAsFactors = F)
   filter(id %not_in% ex_samples)
@@
   filter(id %not_in% ex_samples)
-
 ```

### 2019-05-22 notebooks/article_notebook2.Rmd
added tables and minor fixes

@@
 ```
+
+# R and R-Packages
+R version 3.6.0 [1] was used in the present study. The following packages were used:
+
+| Package       | Version | Reference |
+|---------------|---------|-----------|
+| dplyr         | 0.8.0.1 |   [2]     |
+| tidyr         | 0.8.3   |   [3]     |
+| ggplot2       | 3.1.1   |   [4]     |
+| purrr         | 0.3.2   |   [5]     |
+| tibble        | 2.1.1   |   [6]     |
+| Biostrings    | 2.50.2  |   [7]     |
+| vegan         | 2.5.4   |   [8]     |
+| stringr       | 1.4.0   |   [9]     |
+| cluster       | 2.0.8   |   [10]    |
+| ape           | 5.3     |   [11]    |
+| ggtree        | 1.14.6  |   [12]    |
+| RColorBrewer  | 1.1-2   |   [13]    |
+| cowplot       | 0.9.4   |   [14]    |
+| knitr         | 1.22    |   [15]    |
+| kableExtra    | 1.1.0   |   [16]    |
+| ggsci         | 2.9     |   [17]    |
+| fastqcr       | 0.1.2   |   [18]    |
+| scales        | 1.0.0   |   [19]    |
+| viridis       | 0.5.1   |   [20]    |
+
+# References
+[1] RCoreTeam. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/. 2018.
+
+[2] Wickham H, François R, Henry L, Müller K. dplyr: A Grammar of Data Manipulation [Internet]. 2019. Available from: https://cran.r-project.org/package=dplyr
+
+[3] Wickham H, Henry L. tidyr: Easily Tidy Data with “spread()” and “gather()” Functions [Internet]. 2019. Available from: https://cran.r-project.org/package=tidyr
+
+[4] Wickham H, Chang W, Henry L, Pedersen TL, Takahashi K, Wilke C, et al.
+ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics [Internet].
+2019.
+Available from: https://cran.r-project.org/package=ggplot2
+
+[5] Henry L, Wickham H. purrr: Functional Programming Tools [Internet]. 2019. Available from: https://cran.r-project.org/package=purrr
+
+[6] Müller K, Wickham H. tibble: Simple Data Frames [Internet]. 2019. Available from: https://cran.r-project.org/package=tibble
+
+[7] Pagès H, Aboyoun P, Gentleman R, DebRoy S. Biostrings: Efficient manipulation of biological strings. 2019.
+
+[8] Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, et al. vegan: Community Ecology Package [Internet]. 2019. Available from: https://cran.r-project.org/package=vegan
+
+[9] Wickham H. stringr: Simple, Consistent Wrappers for Common String Operations [Internet]. 2019. Available from: https://cran.r-project.org/package=stringr
+
+[10] Maechler M, Rousseeuw P, Struyf A, Hubert M.
+cluster: “Finding Groups in Data”: Cluster Analysis Extended Rousseeuw et al.
+[Internet].
+2019.
+Available from: https://cran.r-project.org/package=cluster
+
+[11] Paradis E, Blomberg S, Bolker B, Brown J, Claude J, Cuong HS, et al. ape: Analyses of Phylogenetics and Evolution [Internet]. 2019. Available from: https://cran.r-project.org/package=ape
+
+[12] Yu G, Smith DK, Zhu H, Guan Y, Lam TT-Y.
+Ggtree: a package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
+McInerny G, editor.
+Methods Ecol Evol [Internet].
+2017 Jan;8(1):28–36.
+Available from: http://doi.wiley.com/10.1111/2041-210X.12628
+
+[13] Neuwirth E. RColorBrewer: ColorBrewer Palettes [Internet]. 2014. Available from: https://cran.r-project.org/package=RColorBrewer
+
+[14] Wilke CO. cowplot: Streamlined Plot Theme and Plot Annotations for “ggplot2” [Internet]. 2019. Available from: https://cran.r-project.org/package=cowplot
+
+[15] Xie Y. knitr: A General-Purpose Package for Dynamic Report Generation in R [Internet]. 2019. Available from: https://cran.r-project.org/package=knitr
+
+[16] Zhu H. kableExtra: Construct Complex Table with “kable” and Pipe Syntax [Internet]. 2019. Available from: https://cran.r-project.org/package=kableExtra
+
+[17] Xiao N. ggsci: Scientific Journal and Sci-Fi Themed Color Palettes for “ggplot2” [Internet]. 2018. Available from: https://cran.r-project.org/package=ggsci
+
+[18] Kassambara A. fastqcr: Quality Control of Sequencing Data [Internet]. 2019. Available from: https://cran.r-project.org/package=fastqcr
+
+[19] Wickham H. scales: Scale Functions for Visualization [Internet]. 2018. Available from: https://cran.r-project.org/package=scales
+
+[20] Garnier S. viridis: Default Color Maps from “matplotlib” [Internet]. 2018. Available from: https://cran.r-project.org/package=viridis

### 2019-06-11 scripts/ARIBA_analysis.R
minor changes

@@ -17,6 +17,8 @@ mut_genes <- c("gyr","par","marR","marA",
 
 # -------------------------- Functions ----------------------------
 
+"%not_in%" <- Negate("%in%")
+
 # Paste unique cell values together, separate with ","
 func_paste2 <- function(x) paste(unique(x[!is.na(x)]),
                                  collapse = ",")

### 2019-06-11 notebooks/article_notebook2.Rmd
updated group listing

@@
 
-two_reads <- read.table("../data/no_of_reads.txt",
-                          sep = "\t",
-                          header = F,
-                          stringsAsFactors = F) %>%
-  dplyr::rename("ref" = V1,
-         "reads" = V2) %>%
-  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
-         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
-  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
-  group_by(ref) %>%
-  mutate(id = n()) %>%
-  {. ->> four_reads} %>%
-  filter(id == 2) %>%
-  spread(type, reads) %>%
-  dplyr::rename("R1-1" = R1,
-         "R2-1" = R2) %>%
-  mutate("R1-2" = NA,
-         "R2-2" = NA) %>%
-  select(-id) %>%
-  mutate_all(funs(as.character(.)))
-
-four_reads <- four_reads %>%
-  filter(id == 4) %>%
-  group_by(ref) %>%
-  mutate(id = 1:n()) %>%
-  spread(id, reads) %>%
-  summarise_all(funs(func_paste)) %>%
-  dplyr::rename("R1-1" = `1`,
-         "R2-1" = `2`,
-         "R1-2" = `3`,
-         "R2-2" = `4`) %>%
-  select(-type) %>%
-  mutate_all(funs(as.character(.)))
-
-no_of_reads <- as.data.frame(rbind(as.matrix(two_reads),
-                                   as.matrix(four_reads))) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-
 mic_data <- read.table("../data/mic_data.txt",
@@
 mic_data %>%
-  group_by(antres, CIP, NAL) %>%
+  left_join(isolate_data[,c("id","species")], by = "id") %>%
+  group_by(antres, CIP, NAL, species) %>%
   count() %>%

### 2019-06-12 notebooks/article_notebook2.Rmd
corrected for wrong excluded isolate

@@
   "2016-17-565-1-S",
-  "2016-02-415-2-S",
   "2016-02-428-2-S",
   "2016-02-486-2-S",
-  "2016-02-732-2-S"
+  "2016-02-732-2-S",
+  "2014-01-1741-1-S"
   )

### 2019-06-12 notebooks/Clustering analysis.Rmd
minor changes

@@
   "2016-17-565-1-S",
-  "2016-02-415-2-S",
   "2016-02-428-2-S",
@@
                            header = T,
-                           stringsAsFactors = F)
+                           stringsAsFactors = F) %>%
   filter(id %not_in% ex_samples)
@@
                          header = TRUE,
-                         stringsAsFactors = FALSE) %>%
-  filter(FILE != "2016-02-415_S208_pilon_spades.fasta")
+                         stringsAsFactors = FALSE)
 
@@
                                     stringsAsFactors = FALSE) %>%
-  filter(Genome != "2016-02-415_S208_pilon_spades.fasta")
+  mutate(sums = rowSums(.[,4:8]))
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 test <- detailed_cgmlst_stats %>%

### 2019-08-09 notebooks/Clustering analysis.Rmd

@@
 library(ggplot2)
-library(ggtree)
-library(cluster)
-library(Biostrings)
 library(ape)
@@
 library(distanceR)
+library(impoRt)
+library(readxl)
+library(treeio)
+library(kableExtra)
 ```
@@
 
-# cgMLST data
-cgmlst_stats <- read.table("../data/chewbbaca/complete_results/mdata_stats.tsv",
+```
+
+# Minimum spanning tree
+The EnteroBase cgMLST scheme [1] for _E. coli_ was used with chewBBACA [2] to generate the cgMLST data.
+A minimum spanning tree calculated from the cgMLST data with GrapeTree [3] using MSTreeV2.
+Number of allele differences are denoted as numbers on branches between nodes.
+The branches have been collapsed on ten differences or less.
+The number inside each node is the sequence type based on the EnteroBase 7-gene MLST scheme [4].
+```{r, fig.width=15, fig.height=12}
+include_graphics("../figures/GrapeTree_cgMLST.png")
+```
+
+# SNP distance statistics on clades of interest
+The table presents the summary statistics on the minimum SNP distance to the closest isolate per clade of interest.
+```{r}
+dist_data <- get_data("../data/Phylogeny/STs",
+                      "gubbins_snp_dists.tab",
+                      recursive = TRUE,
+                      df = FALSE)
+
+names(dist_data) <- c("ST10", "ST117", "ST162", "ST355", "ST58", "ST602", "ST744")
+
+dist_clean <- lapply(dist_data, function(x){
+  x %>%
+    dplyr::rename("isol1" = `snp-dists 0.6.3`) %>%
+    gather(isol2, value, -isol1) %>%
+    filter(isol1 != isol2) %>%
+    group_by(isol2) %>%
+    mutate(min_val = min(value))
+})
+
+summary_list <- lapply(dist_clean, function(x) round(summary(x$min_val),0)) %>%
+  bind_rows()
+
+summary_list$Type <- c("Min","1st.Q","Median","Mean","3rd.Q","Max")
+
+summary_list %>%
+  select(Type, everything()) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
+
+```
+
+# Maximum likelihood trees on clades of interest
+Each tree is midpoint rooted for visual purposes.
+The values on each node is the UF Bootstrap values calculated by IQTree.
+The tip point shape represents the year of isolation, and the color the animal species of origin.
+The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.
+```{r message=FALSE, warning=FALSE}
+# Import isolate information and tree metadata
+isol_info <- read.table("../data/id.txt",
                         sep = "\t",
@@
 
-detailed_cgmlst_stats <- read.table("../data/chewbbaca/results_statistics.tsv",
+location_data <- read_excel("../data/st_location_data.xlsx") %>%
+  select(id, Kommune, Fylke) %>%
+  mutate(fyl_id = group_indices(., Fylke),
+         kom_id = group_indices(., Kommune)) %>%
+  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
+  select(id, Location)
+
+tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                             sep = "\t",
-                                    header = TRUE,
-                                    stringsAsFactors = FALSE) %>%
-  mutate(sums = rowSums(.[,4:8]))
+                            header = TRUE) %>%
+  left_join(location_data, by = "id") %>%
+  mutate(Year = substr(id, 1, 4)) %>%
+  dplyr::rename("Species" = species)
 
-cgMLST_allele_data <- read.table("../data/chewbbaca/complete_results/cgMLST.tsv",
-                              sep = "\t",
-                              header = TRUE,
-                              stringsAsFactors = FALSE) %>%
-  mutate(FILE = sub("_pilon_spades.fasta", "", FILE)) %>%
-  dplyr::rename("ref" = FILE) %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref) %>%
-  filter(id %not_in% ex_samples)
-```
+# Import trees
+ST10_raw <- read.iqtree("../data/Phylogeny/STs/ST10/iqtree/iqtree.contree")
+ST355_raw <- read.iqtree("../data/Phylogeny/STs/ST355/iqtree/iqtree.contree")
+ST602_raw <- read.iqtree("../data/Phylogeny/STs/ST602/iqtree/iqtree.contree")
+ST744_raw <- read.iqtree("../data/Phylogeny/STs/ST744/iqtree/iqtree.contree")
 
-# Core genome MLST schema and quality control
-The _E. coli_ scheme from Enterobase [1] was used with chewBBACA [2] to get the cgMLST profile for each isolate.
-The number of missing loci in total (sum of LNF, NIPH, PLOT, ALM and ASM from chewBBACA) is plotted below for all isolates.
+raw_trees <- list(ST10_raw,
+                  ST355_raw,
+                  ST602_raw,
+                  ST744_raw)
 
-```{r cgMLST stats}
-boxplot(cgmlst_stats$number.of.missing.data,
-        ylab = "Number of missing loci")
-```
+names(raw_trees) <- c("ST10","ST355","ST602","ST744")
 
-Below follows the detailed results for all loci in the cgMLST scheme.
-EXC = excact matches, INF = novel allele, LNF = locus not found, NIPH = paralogous locus, PLOT = possible loci on tip of contig, ALM = allele 20 % larger than reference, ASM = allele 20 % smaller than reference.
-The numbers below each boxplot denote the mean.
+# Clean tree tip labels
+clean_trees <- lapply(
+  raw_trees,
+  function(x) fix_tip_labels(
+    x,
+    isol_info,
+    "_pilon_spades.fasta",
+    tree_type = "treedata"
+    )
+  )
 
-```{r, message=FALSE, warning=FALSE}
-test <- detailed_cgmlst_stats %>%
-  gather(metric, value, -Genome) %>%
-  group_by(metric) %>%
-  mutate(mean_val = round(mean(value), 1))
+# Specify color palette
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
 
-ggplot(test, aes(metric, value)) +
-  stat_boxplot(geom = "errorbar", width = 0.5) +
-  geom_boxplot() +
-  geom_text(data = subset(test,
-                          Genome == "2006-01-2184_S30_pilon_spades.fasta"),
-            aes(label = mean_val),
-            y = -60,
-            size = 3.8) +
-  scale_x_discrete(limits = c("EXC",
-                              "INF",
-                              "LNF",
-                              "NIPH",
-                              "PLOT",
-                              "ALM",
-                              "ASM")) +
-  labs(y = "Number of loci") +
-  theme(axis.title.x = element_blank(),
-        axis.ticks.x = element_blank())
-```
+# Generate trees
+annotate_tree(
+  clean_trees$ST10,
+  tree_metadata,
+  layout = "rectangular",
+  tree_type = "treedata",
+  midroot = TRUE,
+  line_width = 0.5,
+  label_variable = "Location",
+  color_variable = "Species",
+  shape_variable = "Year",
+  clade_label_node = 22,
+  clade_label = "10 SNP diff",
+  cladelabel_offset = 0.00055,
+  bootstrap_lab = TRUE,
+  tippoint_size = 4,
+  nodepoint_size = 2,
+  label_offset = 0.0001,
+  own_palette = palette
+) +
+  xlim(0, 0.005) +
+  ggtitle("ST10 Clade")
 
-# Minimum spanning tree
-A minimum spanning tree calculated from the cgMLST data with GrapeTree [3] using MSTreeV2.
-Number of allele differences are denoted as numbers on branches between nodes.
-The branches have been collapsed on ten differences or less.
-The number inside each node is the sequence type based on the Warwick 7-gene MLST scheme [4].
-```{r, fig.width=15, fig.height=12}
-include_graphics("../figures/GrapeTree_cgMLST.png")
-```
+annotate_tree(
+  clean_trees$ST355,
+  tree_metadata,
+  layout = "rectangular",
+  tree_type = "treedata",
+  midroot = TRUE,
+  line_width = 0.5,
+  label_variable = "Location",
+  color_variable = "Species",
+  shape_variable = "Year",
+  clade_label_node = 23,
+  clade_label = "3 SNP diff",
+  cladelabel_offset = 0.0013,
+  bootstrap_lab = TRUE,
+  tippoint_size = 4,
+  nodepoint_size = 2,
+  label_offset = 0.0003,
+  own_palette = palette
+) +
+  xlim(0, 0.0165) +
+  ggtitle("ST355 Clade")
 
-# Allele distance distribution
-Basic statistics are presented in the table.
-```{r, message=FALSE, warning=FALSE}
-# calculates distances from cgMLST data and returns
-# the lower triangular of the matrix, with (diag = FALSE)
-# or without (diag = TRUE) the diagonal
-lower_tri_dist_calc <- function(data, diag = TRUE, data_frame = FALSE) {
-  dist <- as.matrix(daisy(data, metric = "gower"))
-  if (diag == TRUE) {
-    dist[upper.tri(dist, diag = TRUE)] <- NA
-  } else {
-    dist[upper.tri(dist, diag = FALSE)] <- NA
-  }
-  dist_vec <- as.vector(as.matrix(dist))
-  dist_compl <- dist_vec[!is.na(dist_vec)]
+annotate_tree(
+  clean_trees$ST602,
+  tree_metadata,
+  layout = "rectangular",
+  tree_type = "treedata",
+  midroot = TRUE,
+  line_width = 0.5,
+  label_variable = "Location",
+  color_variable = "Species",
+  shape_variable = "Year",
+  clade_label_node = 10,
+  clade_label = "18 SNP diff",
+  cladelabel_offset = 0.001,
+  bootstrap_lab = TRUE,
+  tippoint_size = 4,
+  nodepoint_size = 2,
+  label_offset = 0.0002,
+  own_palette = palette
+) +
+  xlim(0, 0.009) +
+  ggtitle("ST602 Clade")
 
-  if (data_frame == TRUE) {
-    dist_compl <- as.data.frame(as.matrix(dist))
-  }
-  return(dist_compl)
-}
-
-broiler_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_broiler.tsv",
-                           sep = "\t",
-                           header = TRUE,
-                           colClasses = "factor",
-                           row.names = 1)
-
-pig_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_pig.tsv",
-                           sep = "\t",
-                           header = TRUE,
-                           colClasses = "factor",
-                           row.names = 1)
-
-fox_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_fox.tsv",
-                           sep = "\t",
-                           header = TRUE,
-                           colClasses = "factor",
-                           row.names = 1)
-
-bird_cgMLST <- read.table("../data/chewbbaca/complete_results/cgMLST_bird.tsv",
-                           sep = "\t",
-                           header = TRUE,
-                           colClasses = "factor",
-                           row.names = 1)
-
-broiler_dist <- lower_tri_dist_calc(broiler_cgMLST)
-pig_dist <- lower_tri_dist_calc(pig_cgMLST)
-fox_dist <- lower_tri_dist_calc(fox_cgMLST)
-bird_dist <- lower_tri_dist_calc(bird_cgMLST)
-
-# Distance plots
-par(mfrow = c(2,2))
-
-x1 <- hist(broiler_dist,
-     breaks = 20,
-     plot = FALSE)
-
-x1$density = x1$counts/sum(x1$counts)*100
-
-plot(x1,
-     freq = FALSE,
-     main = "Broiler distance frequency",
-     xlab = "Distance",
-     ylim = c(0, 50))
-
-x2 <- hist(pig_dist,
-     breaks = 20,
-     plot = FALSE)
-
-x2$density = x2$counts/sum(x2$counts)*100
-
-plot(x2,
-     freq = FALSE,
-     main = "Pig distance frequency",
-     xlab = "Distance",
-     ylim = c(0, 50))
-
-x3 <- hist(fox_dist,
-     breaks = 20,
-     plot = FALSE)
-
-x3$density = x3$counts/sum(x3$counts)*100
-
-plot(x3,
-     freq = FALSE,
-     main = "Red fox distance frequency",
-     xlab = "Distance",
-     ylim = c(0, 50))
-
-x4 <- hist(bird_dist,
-     breaks = 20,
-     plot = FALSE)
-
-x4$density = x4$counts/sum(x4$counts)*100
-
-plot(x4,
-     freq = FALSE,
-     main = "Wild bird distance frequency",
-     xlab = "Distance",
-     ylim = c(0, 50))
-
-# generate data frame from allele data
-alleles <- cgMLST_allele_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(species) %>%
-  mutate(n = 1:n(),
-         new_id = case_when(species == "Broiler" ~ paste("BR", n, sep = "_"),
-                            species == "Pig" ~ paste("PI", n, sep = "_"),
-                            species == "Red fox" ~ paste("RF", n, sep = "_"),
-                            species == "Wild bird" ~ paste("WB", n, sep = "_"))) %>%
-  ungroup() %>%
-  select(new_id, everything(), -c(species, id)) %>%
-  mutate_all(funs(as.factor)) %>%
-  column_to_rownames("new_id")
-
-# generate data frame for iteration analysis
-run_df <- lower_tri_dist_calc(alleles, data_frame = TRUE) %>%
-  rownames_to_column("id") %>%
-  gather(isolate1, value, -id) %>%
-  dplyr::rename("isolate2" = id) %>%
-  mutate(group1 = substr(isolate1, start = 1, stop = 2),
-         group2 = substr(isolate2, start = 1, stop = 2)) %>%
-  select(isolate1, group1, isolate2, group2, value) %>%
-  filter(is.na(value) == FALSE,
-         group1 == group2) %>%
-  arrange(group1, group2)
-
-run_df %>%
-  mutate(total_mean = round(mean(value), 3)) %>%
-  group_by(group1) %>%
-  mutate(mean = round(mean(value), 3),
-         var = round(var(value), 3)) %>%
-  select(group1, total_mean, mean, var) %>%
-  summarise_all(funs(func_paste))
-```
-
-# Distribution of iterated distances
-Distribution of percentages of distances < 0.1 or > 0.8.
-The distribution is based on 1000 iterations of the distance data.
-Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird.
-The mean value is denoted as a vertical black line.
-```{r, warning=FALSE, message=FALSE, dpi=300}
-# functions used in iteration analyses
-
-## Calculates the sum of observations lower than dist_value1 and
-## bigger than dist_value2, and calculates the percentage of
-## observed values for each group
-without <- function(df, run, dist_value1 = 0.1, dist_value2 = 0.8) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = value < dist_value1 | value > dist_value2) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Within", "Without")) %>%
-    spread(n, nn) %>%
-    mutate(Percent = round(Within / (Within + Without) * 100, 2),
-           iter = run)
-  return(df)
-}
-
-## Calculates the sum of observations between dist_value1 and dist_value2
-## and calculates the percentage of observed values for each group
-within <- function(df, run, dist_value1 = 0.1, dist_value2 = 0.8) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = value > dist_value1 & value < dist_value2) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Within", "Without")) %>%
-    spread(n, nn) %>%
-    mutate(Percent = round(Within / (Within + Without) * 100, 2),
-           iter = run)
-  return(df)
-}
-
-## Calculates the sum of observations below dist_value1
-## for each group
-below_calc <- function(df, run, dist_value1 = 0.1) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = value < dist_value1) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Below", "Above")) %>%
-    spread(n, nn) %>%
-    mutate(Below = ifelse(is.na(Below) == TRUE, 0, Below)) %>%
-    mutate(Percent_below = round(Below / (Above + Below) * 100, 2),
-           iter = run)
-  return(df)
-}
-
-## Calculates the sum of observations above dist_value1
-## for each group
-above_calc <- function(df, run, dist_value1 = 0.9) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(n = value > dist_value1) %>%
-    group_by(group1, group2, n) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(n = ifelse(n == TRUE, "Above", "Below")) %>%
-    spread(n, nn) %>%
-    mutate(Above = ifelse(is.na(Above) == TRUE, 0, Above)) %>%
-    mutate(Percent_above = round(Above / (Above + Below) * 100, 2),
-           iter = run)
-  return(df)
-}
-
-## Calculates the mean of observations for each group
-mean_calc <- function(df, run) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(metric = round(mean(value), 2)) %>%
-    select(group1, group2, metric) %>%
-    summarise_all(funs(func_paste)) %>%
-    mutate(iter = run,
-           metric = as.numeric(metric))
-}
-
-## Calculates the median of observations for each group
-median_calc <- function(df, run) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(metric = round(median(value), 2)) %>%
-    select(group1, group2, metric) %>%
-    summarise_all(funs(func_paste)) %>%
-    mutate(iter = run,
-           metric = as.numeric(metric))
-}
-
-## Calculates the quantiles of the observations for each group
-quantile_calc <- function(df, run) {
-  df <- df %>%
-    group_by(group1, group2) %>%
-    mutate(quant_0 = quantile(value)[1],
-           quant_25 = quantile(value)[2],
-           quant_50 = quantile(value)[3],
-           quant_75 = quantile(value)[4],
-           quant_100 = quantile(value)[5]) %>%
-    select(group1, group2, contains("quant")) %>%
-    summarise_all(funs(func_paste)) %>%
-    mutate_at(vars(contains("quant")),
-              funs(as.numeric)) %>%
-    mutate(iter = run)
-}
-
-# Iteration function
-iterate_samples <- function(df, run, method = "mean", dist_value1, dist_value2) {
-
-  # randomize the value column, regardless of groups
-  df <- df %>%
-    mutate(value = sample(value))
-
-  # run specified function
-  if (method == "mean") {
-    df <- mean_calc(df, run)
-    }
-
-  if (method == "below") {
-   df <- below_calc(df, run, dist_value1)
-    }
-
-  if (method == "above") {
-   df <- above_calc(df, run, dist_value1)
-    }
-
-  if (method == "within") {
-    df <- within(df, run, dist_value1, dist_value2)
-    }
-
-  if (method == "without") {
-    df <- without(df, run, dist_value1, dist_value2)
-    }
-
-  if (method == "median") {
-    df <- median_calc(df, run)
-    }
-
-  if (method == "quantile") {
-    df <- quantile_calc(df, run)
-  }
-
-    return(df)
-}
-
-# run iterations on selected data frame
-run_iterations <- function(df, runs = 1000, seed = 10, method = "mean", dist_value1, dist_value2) {
-  # set seed
-  set.seed(seed)
-
-  # create output list
-  output <- list()
-
-  # create expected values
-  if (method == "mean") {
-    orig <- mean_calc(df, 0)
-    }
-
-  if (method == "below") {
-   orig <- below_calc(df, 0, dist_value1)
-    }
-
-  if (method == "above") {
-   orig <- above_calc(df, 0, dist_value1)
-    }
-
-  if (method == "within") {
-    orig <- within(df, 0, dist_value1, dist_value2)
-    }
-
-  if (method == "without") {
-    orig <- without(df, 0, dist_value1, dist_value2)
-    }
-
-  if (method == "median") {
-    orig <- median_calc(df, 0)
-    }
-
-  if (method == "quantile") {
-    orig <- quantile_calc(df, 0)
-  }
-
-  output <- c(output, list(orig))
-
-  # run iterations
-  for (i in 1:runs) {
-    result <- iterate_samples(df,
-                              i,
-                              method = method,
-                              dist_value1 = dist_value1,
-                              dist_value2 = dist_value2)
-    output <- c(output, list(result))
-    }
-  return(output)
-}
-
-# Run functions
-
-complete_data <- run_iterations(run_df,
-                                runs = 1000,
-                                method = "without",
-                                dist_value1 = 0.1,
-                                dist_value2 = 0.8) %>%
-  bind_rows() %>%
-  mutate(Percent = as.numeric(Percent))
-
-lower_data <- run_iterations(run_df,
-                             runs = 1000,
-                             method = "below",
-                             dist_value1 = 0.1) %>%
-  bind_rows() %>%
-  mutate(Percent_below = as.numeric(Percent_below))
-
-# Plots
-
-ggplot(complete_data, aes(Percent)) +
-  geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
-                fill = "grey80") +
-  stat_function(fun = dnorm,
-                args = with(complete_data, c(mean = mean(Percent), sd = sd(Percent))),
-                color = "red") +
-  geom_segment(aes(x = complete_data$Percent[1],
-                   xend = complete_data$Percent[1], y = 0.08, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#4575b4") +
-  geom_segment(aes(x = complete_data$Percent[2],
-                   xend = complete_data$Percent[2], y = 0.08, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#74add1") +
-  geom_segment(aes(x = complete_data$Percent[3],
-                   xend = complete_data$Percent[3], y = 0.08, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#f46d43") +
-  geom_segment(aes(x = complete_data$Percent[4],
-                   xend = complete_data$Percent[4], y = 0.08, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#fdae61") +
-  geom_text(label = "p < 0.001",
-            x = 83.30,
-            y = 0.16,
-            angle = 90) +
-  geom_vline(xintercept = mean(complete_data$Percent),
-             size = 1) +
-  labs(x = "Percent (%) distances < 0.1 or > 0.8",
-       y = "Density") +
-  ggtitle(label = "Distribution of percentages",
-          subtitle = "Number of iterations: 1000") +
-  theme(plot.title = element_text(hjust = 0))
-
-ggplot(lower_data, aes(Percent_below)) +
-  geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "black",
-                fill = "grey80") +
-  stat_function(fun = dnorm,
-                args = with(lower_data, c(mean = mean(Percent_below), sd = sd(Percent_below))),
-                color = "red") +
-  geom_segment(aes(x = lower_data$Percent_below[1],
-                   xend = lower_data$Percent_below[1], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#4575b4") +
-  geom_segment(aes(x = lower_data$Percent_below[2],
-                   xend = lower_data$Percent_below[2], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#74add1") +
-  geom_segment(aes(x = lower_data$Percent_below[3],
-                   xend = lower_data$Percent_below[3], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#f46d43") +
-  geom_segment(aes(x = lower_data$Percent_below[4],
-                   xend = lower_data$Percent_below[4], y = 0.15, yend = 0),
-               arrow = arrow(length = unit(0.3, "cm"),
-                             type = "closed"),
-               color = "#fdae61") +
-  geom_text(label = "p < 0.001",
-            x = 5.87,
-            y = 0.32,
-            angle = 90) +
-  geom_vline(xintercept = mean(lower_data$Percent_below),
-             size = 1) +
-  labs(x = "Percent (%) distances < 0.1",
-       y = "Density") +
-  ggtitle(label = "Distribution of Percentages",
-          subtitle = "Number of iterations: 1000") +
-  theme(plot.title = element_text(hjust = 0))
+annotate_tree(
+  clean_trees$ST744,
+  tree_metadata,
+  layout = "rectangular",
+  tree_type = "treedata",
+  midroot = TRUE,
+  line_width = 0.5,
+  label_variable = "Location",
+  color_variable = "Species",
+  shape_variable = "Year",
+  clade_label_node = 12,
+  clade_label = "50 SNP diff",
+  cladelabel_offset = 0.00024,
+  bootstrap_lab = TRUE,
+  tippoint_size = 4,
+  nodepoint_size = 2,
+  label_offset = 0.00007,
+  own_palette = palette
+) +
+  xlim(0, 0.0030) +
+  ggtitle("ST744 Clade")
 

### 2019-08-12 notebooks/Clustering analysis.Rmd
adjusted ST trees

@@
 library(treeio)
+library(adephylo)
+library(RColorBrewer)
 library(kableExtra)
@@
 
+mlst_data <- read.table("../data/mlst_results.txt",
+                        sep = "\t",
+                        header = TRUE,
+                        stringsAsFactors = F)
+
 ```
@@
 include_graphics("../figures/GrapeTree_cgMLST.png")
+```
+
+# Patristic distances
+Patristic distances were calculated from the core gene alignment SNP tree.
+```{r fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
+iqtree <- read.tree("../data/IQTree/core_genes_roary.treefile")
+iqtree_clean <- fix_tip_labels(iqtree, id_data, "_pilon_spades.fasta", tree_type = "phylo")
+
+pat_dist <- as.matrix(distTips(iqtree_clean, method = "patristic"))
+
+pat_dist_bin <- as.data.frame(ifelse(pat_dist < 0.04, "1", "0")) %>%
+  rownames_to_column("isol1")
+
+row_val <- as.data.frame(pat_dist_bin$isol1) %>%
+  rename("id" = `pat_dist_bin$isol1`) %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  group_by(ST) %>%
+  mutate(n = 1:n(),
+         new_id = paste(ST, n, sep = "_"))
+
+col_val <- as.data.frame(names(pat_dist_bin)) %>%
+  rename("id" = `names(pat_dist_bin)`) %>%
+  left_join(mlst_data[,c("id","ST")], by = "id") %>%
+  group_by(ST) %>%
+  mutate(n = 1:n(),
+         new_id = paste(ST, n, sep = "_")) %>%
+  ungroup() %>%
+  mutate(new_id = ifelse(new_id == "NA_1", "isol1", new_id),
+         ST = ifelse(is.na(ST) == TRUE, "isol1", ST))
+
+pat_dist_clean <- pat_dist_bin
+names(pat_dist_clean) <- col_val$new_id
+pat_dist_clean$isol1 <- row_val$new_id
+
+plot_data <- pat_dist_clean %>%
+  gather(isol2, value, -isol1) %>%
+  filter(isol1 != isol2)
+
+ggplot(plot_data, aes(reorder(isol1, -as.numeric(value)), reorder(isol2, -as.numeric(value)), fill = value)) +
+  geom_tile() +
+  scale_fill_manual(values = c("1" = "#80b1d3",
+                               "0" = "grey95"),
+                    labels = c("Pat. dist > 0.04",
+                               "Pat. dist < 0.04")) +
+  labs(fill = NULL) +
+  theme(axis.text = element_text(size = 3),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
+        panel.background = element_blank(),
+        axis.ticks = element_blank(),
+        axis.title = element_blank()) +
+  coord_fixed()
 ```
@@
 The values on each node is the UF Bootstrap values calculated by IQTree.
-The tip point shape represents the year of isolation, and the color the animal species of origin.
+The tip point shape represent the animal species of origin, and the color the year of isolation.
 The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.
@@
 # Import isolate information and tree metadata
-isol_info <- read.table("../data/id.txt",
-                        sep = "\t",
-                        header = TRUE,
-                        stringsAsFactors = FALSE)
-
 location_data <- read_excel("../data/st_location_data.xlsx") %>%
@@
     x,
-    isol_info,
+    id_data,
     "_pilon_spades.fasta",
@@
 
-# Specify color palette
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
+# Specify color and shape palettes
+palette_shape <- c(
+  "Broiler" = 15,
+  "Pig" = 18,
+  "Red fox" = 16,
+  "Wild bird" = 17
+)
+
+palette_color <- brewer.pal(11, "Set3")
+
+year_val <- c(
+  "2006",
+  "2007",
+  "2008",
+  "2009",
+  "2010",
+  "2011",
+  "2012",
+  "2013",
+  "2014",
+  "2015",
+  "2016")
+
+names(palette_color) <- year_val
 
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 22,
+  color_variable = "Year",
+  shape_variable = "Species",
+  clade_label_node = 19,
   clade_label = "10 SNP diff",
@@
   bootstrap_lab = TRUE,
+  bootlab_size = 2,
   tippoint_size = 4,
-  nodepoint_size = 2,
   label_offset = 0.0001,
-  own_palette = palette
+  shape_palette = palette_shape,
+  color_palette = palette_color
 ) +
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 23,
-  clade_label = "3 SNP diff",
+  color_variable = "Year",
+  shape_variable = "Species",
+  clade_label_node = 26,
+  clade_label = "6 SNP diff",
   cladelabel_offset = 0.0013,
   bootstrap_lab = TRUE,
+  bootlab_size = 2,
   tippoint_size = 4,
-  nodepoint_size = 2,
   label_offset = 0.0003,
-  own_palette = palette
+  shape_palette = palette_shape,
+  color_palette = palette_color
 ) +
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
+  color_variable = "Year",
+  shape_variable = "Species",
   clade_label_node = 10,
@@
   bootstrap_lab = TRUE,
+  bootlab_size = 2,
   tippoint_size = 4,
-  nodepoint_size = 2,
   label_offset = 0.0002,
-  own_palette = palette
+  shape_palette = palette_shape,
+  color_palette = palette_color
 ) +
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
+  color_variable = "Year",
+  shape_variable = "Species",
   clade_label_node = 12,
@@
   bootstrap_lab = TRUE,
+  bootlab_size = 2,
   tippoint_size = 4,
-  nodepoint_size = 2,
   label_offset = 0.00007,
-  own_palette = palette
+  shape_palette = palette_shape,
+  color_palette = palette_color
 ) +

### 2019-08-12 notebooks/article_notebook2.Rmd
added package ref

@@
 | viridis       | 0.5.1   |   [20]    |
+| adephylo      | 1.1-11  |   [21]    |
 
@@
 [20] Garnier S. viridis: Default Color Maps from “matplotlib” [Internet]. 2018. Available from: https://cran.r-project.org/package=viridis
+
+[21] Jombart T., Dray S. (2008) adephylo: exploratory analyses for the phylogenetic comparative method.

### 2019-08-14 figure_script.R
updated with new figures

@@ -136,21 +136,124 @@ library(treeio)
 library(ggtree)
 library(phytools)
 library(readxl)
+library(RColorBrewer)
+library(ggplot2)
+
+## Total tree
+isol_info <- read.table("data/id.txt",
+                        sep = "\t",
+                        header = TRUE,
+                        stringsAsFactors = FALSE)
+
+mlst_info <- read.table("data/mlst_results.txt",
+                        sep = "\t",
+                        header = TRUE,
+                        stringsAsFactors = FALSE)
+
+isolate_data <- read.table("data/isolate_data.txt",
+                           sep = "\t",
+                           header = TRUE,
+                           stringsAsFactors = FALSE)
+
+metadata <- isolate_data[,c("id","species")] %>%
+  left_join(mlst_info[,c("id","ST")], by = "id") %>%
+  rename("Species" = species)
+
+heatmap_data <- read.table("data/chewbbaca/complete_results/tree_heatmap_data.txt",
+                           sep = "\t",
+                           header = TRUE,
+                           stringsAsFactors = FALSE)
+
+iqtree <- read.iqtree("data/Phylogeny/Roary/iqtree_roary.contree")
+iqtree_clean <- fix_tip_labels(iqtree, isol_info, "_pilon_spades.fasta", tree_type = "treedata")
+
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+
+
+palette2 <- c("> 95" = "#6a51a3",
+                  "80 - 94" = "#9e9ac8",
+                  "50 - 79" = "#cbc9e2",
+                  "< 50" = "#f2f0f7",
+                  "1-A" = "#fc8d59",
+                  "1-I" = "#80b1d3",
+                  "0" = "grey95")
+
+iqtree_clean@phylo <- midpoint(iqtree_clean@phylo, labels = "support")
+
+iqtree_clean@data$Bootstrap <- factor(
+  case_when(
+    iqtree_clean@data$UFboot >= 95 ~ "> 95",
+    iqtree_clean@data$UFboot < 95 &
+      iqtree_clean@data$UFboot >= 80 ~ "80 - 94",
[…]

### 2019-08-14 notebooks/Clustering analysis.Rmd

--- 2019-08-09 notebooks/Clustering analysis.Rmd
@@
 library(ggplot2)
+library(ggtree)
 library(ape)
@@
 library(readxl)
+library(readr)
+library(iteRate)
 library(treeio)
+library(adephylo)
+library(phangorn)
+library(RColorBrewer)
 library(kableExtra)
@@
 
+# Patristic distances
+Patristic distances were calculated from the core gene alignment SNP tree.
+```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
+location_data <- read_excel("../data/st_location_data.xlsx") %>%
+  select(id, Kommune, Fylke) %>%
+  mutate(fyl_id = group_indices(., Fylke),
+         kom_id = group_indices(., Kommune)) %>%
+  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
+  select(id, Location)
+
+tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
+                            sep = "\t",
+                            header = TRUE) %>%
+  left_join(location_data, by = "id") %>%
+  mutate(Year = substr(id, 1, 4)) %>%
+  dplyr::rename("Species" = species)
+
+iqtree <- read.iqtree("../data/Phylogeny/Roary/iqtree_roary.contree")
+iqtree_clean <- fix_tip_labels(iqtree, id_data, "_pilon_spades.fasta", tree_type = "treedata")
+
+pat_dist <- as.matrix(distTips(iqtree_clean@phylo, method = "patristic"))
+
+iqtree_clean@phylo <- midpoint(iqtree_clean@phylo, labels = "support")
+
+pat_dist_id <- as.data.frame(pat_dist) %>%
+  rownames_to_column("isol1") %>%
+  gather(isol2, value, -isol1) %>%
+  filter(isol1 != isol2) %>%
+  left_join(tree_metadata, by = c("isol1" = "id")) %>%
+  rename("ST1" = ST,
+         "Species1" = Species,
+         "Location1" = Location,
+         "Year1" = Year) %>%
+  left_join(tree_metadata, by = c("isol2" = "id")) %>%
+  rename("ST2" = ST,
+         "Species2" = Species,
+         "Location2" = Location,
+         "Year2" = Year) %>%
+  filter(value < 0.005) %>%
+  mutate(selection = case_when(Species1 == Species2 & Location1 != Location2 ~ "ss",
+                               Species1 != Species2 & is.na(Location1) == FALSE ~ "as")) %>%
+  filter(selection %in% c("ss", "as"))
+
+tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
+                            sep = "\t",
+                            header = TRUE) %>%
+  left_join(location_data, by = "id") %>%
+  mutate(Year = substr(id, 1, 4)) %>%
+  dplyr::rename("Species" = species) %>%
+  mutate(ST_new = ifelse(ST %in% pat_dist_id$ST1, ST, "Other"))
+
+pat_dist_bin <- as.data.frame(ifelse(pat_dist < 0.003, "< 0.003", "> 0.003"))
+
+col_palette <- c("< 0.003" = "#bebada",
+                 "> 0.003" = "grey95")
+
+palette <- c("Broiler" = "#4575b4",
+             "Pig" = "#74add1",
+             "Red fox" = "#f46d43",
+             "Wild bird" = "#fdae61")
+
+p <- ggtree(iqtree_clean) %<+% tree_metadata +
+  geom_tiplab(aes(label = ST),
+              size = 1,
+              align = TRUE) +
+  scale_color_manual(values = palette) +
+  geom_tippoint(aes(color = Species))
+
+gheatmap(p,
+         pat_dist_bin,
+         color = NULL,
+         offset = 0.01,
+         colnames = FALSE) +
+  scale_fill_manual(values = col_palette) + 
+  annotate("label", x = 0.27, y = 270, label = "Clade A") +
+  annotate("label", x = 0.29, y = 251, label = "Clade B") +
+  annotate("label", x = 0.39, y = 167, label = "Clade C") +
+  annotate("label", x = 0.38, y = 138, label = "Clade D") +
+  annotate("label", x = 0.32, y = 57, label = "Clade E") +
+  annotate("label", x = 0.31, y = 31, label = "Clade F")
+
+```
+
 # SNP distance statistics on clades of interest
@@
 
-names(dist_data) <- c("ST10", "ST117", "ST162", "ST355", "ST58", "ST602", "ST744")
+names(dist_data) <- c("Clade D", "Clade F", "Clade A", "Clade E", "Clade B", "Clade C")
 
@@
 summary_list %>%
-  select(Type, everything()) %>%
+  select(Type,
+         `Clade A`,
+         `Clade B`,
+         `Clade C`,
+         `Clade D`,
+         `Clade E`,
+         `Clade F`) %>%
   kable() %>%
@@
   select(id, Location)
-
-tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
-                            sep = "\t",
-                            header = TRUE) %>%
-  left_join(location_data, by = "id") %>%
-  mutate(Year = substr(id, 1, 4)) %>%
-  dplyr::rename("Species" = species)
 
@@
 
-names(raw_trees) <- c("ST10","ST355","ST602","ST744")
+names(raw_trees) <- c("Clade D","Clade E","Clade B","Clade C")
 
@@
 
-# Specify color palette
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
+clean_trees <- lapply(clean_trees,
+                      function(x) {
+                        x@data$Bootstrap <- factor(
+                          case_when(
+                            x@data$UFboot >= 95 ~ "> 95",
+                            x@data$UFboot < 95 &
+                              x@data$UFboot >= 80 ~ "80 - 94",
+                            x@data$UFboot < 80 &
+                              x@data$UFboot >= 50 ~ "50 - 79",
+                            x@data$UFboot < 50 ~ "< 50"
+                          ),
+                          ordered = TRUE,
+                          levels = c("> 95",
+                                     "80 - 94",
+                                     "50 - 79",
+                                     "< 50")
+                        )
+                        return(x)
+                      })
+
+# Specify color and shape palettes
+palette_shape <- c(
+  "Broiler" = 21,
+  "Pig" = 22,
+  "Red fox" = 23,
+  "Wild bird" = 24
+)
+
+palette_color <- brewer.pal(11, "Set3")
+
+year_val <- c(
+  "2006",
+  "2007",
+  "2008",
+  "2009",
+  "2010",
+  "2011",
+  "2012",
+  "2013",
+  "2014",
+  "2015",
+  "2016")
+
+names(palette_color) <- year_val
+
+boot_palette <- c("> 95" = "#6a51a3",
+              "80 - 94" = "#9e9ac8",
+              "50 - 79" = "#cbc9e2",
+              "< 50" = "#f2f0f7")
 
@@
 annotate_tree(
-  clean_trees$ST10,
+  clean_trees$`Clade B`,
   tree_metadata,
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 22,
-  clade_label = "10 SNP diff",
-  cladelabel_offset = 0.00055,
-  bootstrap_lab = TRUE,
+  color_variable = "Year",
+  shape_variable = "Species",
+  bootstrap_var = "Bootstrap",
+  clade_label_node = 10,
+  clade_label = "18 SNP diff",
+  cladelabel_offset = 0.001,
+  bootstrap_lab = FALSE,
   tippoint_size = 4,
-  nodepoint_size = 2,
-  label_offset = 0.0001,
-  own_palette = palette
+  label_offset = 0.0002,
+  shape_palette = palette_shape,
+  color_palette = palette_color,
+  node_palette = boot_palette
 ) +
-  xlim(0, 0.005) +
-  ggtitle("ST10 Clade")
+  xlim(0, 0.009) +
+  ggtitle("Clade B, ST602")
 
 annotate_tree(
-  clean_trees$ST355,
+  clean_trees$`Clade C`,
   tree_metadata,
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 23,
-  clade_label = "3 SNP diff",
-  cladelabel_offset = 0.0013,
-  bootstrap_lab = TRUE,
+  color_variable = "Year",
+  shape_variable = "Species",
+  bootstrap_var = "Bootstrap",
+  clade_label_node = 12,
+  clade_label = "50 SNP diff",
+  cladelabel_offset = 0.00024,
+  bootstrap_lab = FALSE,
   tippoint_size = 4,
-  nodepoint_size = 2,
-  label_offset = 0.0003,
-  own_palette = palette
+  label_offset = 0.00007,
+  shape_palette = palette_shape,
+  color_palette = palette_color,
+  node_palette = boot_palette
 ) +
-  xlim(0, 0.0165) +
-  ggtitle("ST355 Clade")
+  xlim(0, 0.0030) +
+  ggtitle("Clade C, ST744")
 
 annotate_tree(
-  clean_trees$ST602,
+  clean_trees$`Clade D`,
   tree_metadata,
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 10,
-  clade_label = "18 SNP diff",
-  cladelabel_offset = 0.001,
-  bootstrap_lab = TRUE,
+  color_variable = "Year",
+  shape_variable = "Species",
+  bootstrap_var = "Bootstrap",
+  clade_label_node = 19,
+  clade_label = "10 SNP diff",
+  cladelabel_offset = 0.00055,
+  bootstrap_lab = FALSE,
+  bootlab_size = 2,
   tippoint_size = 4,
-  nodepoint_size = 2,
-  label_offset = 0.0002,
-  own_palette = palette
+  label_offset = 0.0001,
+  shape_palette = palette_shape,
+  color_palette = palette_color,
+  node_palette = boot_palette
 ) +
-  xlim(0, 0.009) +
-  ggtitle("ST602 Clade")
+  xlim(0, 0.005) +
+  ggtitle("Clade D, ST10")
 
 annotate_tree(
-  clean_trees$ST744,
+  clean_trees$`Clade E`,
   tree_metadata,
@@
   label_variable = "Location",
-  color_variable = "Species",
-  shape_variable = "Year",
-  clade_label_node = 12,
-  clade_label = "50 SNP diff",
-  cladelabel_offset = 0.00024,
-  bootstrap_lab = TRUE,
+  color_variable = "Year",
+  shape_variable = "Species",
+  bootstrap_var = "Bootstrap",
+  clade_label_node = 26,
+  clade_label = "6 SNP diff",
+  cladelabel_offset = 0.0013,
+  bootstrap_lab = FALSE,
   tippoint_size = 4,
-  nodepoint_size = 2,
-  label_offset = 0.00007,
-  own_palette = palette
+  label_offset = 0.0003,
+  shape_palette = palette_shape,
+  color_palette = palette_color,
+  node_palette = boot_palette
 ) +
-  xlim(0, 0.0030) +
-  ggtitle("ST744 Clade")
+  xlim(0, 0.0165) +
+  ggtitle("Clade E, ST355")
+
+```
+
+# P-value calculations on SNP distances within animal species
+The table presents the non-exact p-values based on 1000 iterations of minimum SNP distances to closest isolate within animal species.
+The figure shows the distribution of the expected values with the observed values for each animal species denoted as an arrow.
+Red fox results is excluded from the plot as the observed distance values were too high to visualize.
+```{r}
+# Import distance data
+total_dist <- read_delim("../data/snp_dists/final_Roary_gubbins_SNP_dists.tab",
+                         delim = "\t") %>%
+  rename("isol1" = `snp-dists 0.6.3`)
+
+# Import isolate data
+id_data <- read.table("../data/id.txt",
+                      sep = "\t",
+                      header = TRUE,
+                      stringsAsFactors = F)
+
+isolate_data <- read.table("../data/isolate_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+# Set row/column values to unique species ID
+row_id <- as.data.frame(total_dist$isol1) %>%
+  rename("ref" = `total_dist$isol1`) %>%
+  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id) %>%
+  left_join(isolate_data[,c("id", "species")], by = "id") %>%
+  group_by(species) %>%
+  mutate(n = 1:n(),
+         isol1 = paste(species, n, sep = "_")) %>%
+  ungroup() %>%
+  select(isol1)
+
+col_id <- as.data.frame(names(total_dist)) %>%
+  rename("ref" = `names(total_dist)`) %>%
+  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id) %>%
+  left_join(isolate_data[,c("id", "species")], by = "id") %>%
+  group_by(species) %>%
+  mutate(n = 1:n(),
+         isol1 = paste(species, n, sep = "_")) %>%
+  ungroup() %>%
+  select(isol1) %>%
+  mutate(isol1 = ifelse(isol1 == "NA_1", "isol1", isol1))
+
+total_dist$isol1 <- row_id$isol1
+names(total_dist) <- col_id$isol1
+
+# Create data frame for iteration analysis
+run_df <- total_dist %>%
+  gather(isol2, value, -isol1) %>%
+  filter(isol1 != isol2) %>%
+  mutate(species1 = sub("_.+", "", isol1),
+         species2 = sub("_.+", "", isol2)) %>%
+  filter(species1 == species2) %>%
+  group_by(isol2) %>%
+  mutate(min_val = min(value)) %>%
+  ungroup() %>%
+  select(isol1, isol2, min_val, species1, species2) %>%
+  rename("value" = min_val,
+         "group1" = species1,
+         "group2" = species2)
+
+# Run iterations
+df <- iterate(run_df, runs = 1000, method = "median")
+
+# Filter out the observed values
+segment_data <- filter(df, iter == 0,
+                       group1 %in% c("Broiler", "Pig", "Wild bird"))
+
+# Calculate non-exact p-values
+df2 <- df %>%
+  group_by(group1) %>%
+  mutate(test = case_when(group1 == "Broiler" & metric <= 12 ~ 1,
+                          group1 == "Wild bird" & metric <= 53 ~ 1,
+                          group1 == "Pig" & metric <= 212 ~ 1,
+                          group1 == "Red fox" & metric <= 1388.5 ~ 1,
+                          TRUE ~ 0)) %>%
+  split(f = .$group1)
+
+lapply(df2, function(x) round(sum(x$test)/length(x$test), 3)) %>%
+  bind_rows() %>%
+  mutate(Iterations = 1000) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
+
+# Plot results, exclude red fox (too high values for the plot to be visible)
+plot_df <- df %>%
+  filter(group1 %in% c("Broiler", "Pig", "Wild bird"))
+
+ggplot(plot_df, aes(metric)) +
+  geom_histogram(
+    aes(y = ..density..),
+    binwidth = 0.2,
+    color = "black",
+    fill = "grey80"
+  ) +
+  stat_function(fun = dnorm,
+                args = with(df, c(
+                  mean = mean(metric), sd = sd(metric)
+                )),
+                color = "red") +
+  geom_segment(
+    segment_data,
+    mapping = aes(
+      x = metric,
+      xend = metric,
+      y = 0.5,
+      yend = 0,
+      color = group1
+    ),
+    arrow = arrow(length = unit(0.3, "cm"),
+                  type = "closed")
+  ) +
+  geom_text(segment_data,
+            mapping = aes(label = group1,
+                          x = metric),
+            y = 1,
+            angle = 90) +
+  guides(color = FALSE) +
+  labs(x = "Min. number of SNPs to closest isolate",
+       y = "Density") +
+  ggtitle(label = "Distribution of expected values",
+          subtitle = paste("Number of iterations: ", max(df$iter))) +
+  theme(plot.title = element_text(hjust = 0)) +
+  scale_color_manual(values = palette)
 

### 2019-08-16 figure_script.R
corrected bootstrap values

@@ -369,8 +369,6 @@ p2 <- annotate_tree(
   xlim(0, 0.047) +
   ggtitle("Clade A, ST162")
 
-
-
 ggsave("figures/CladeA_ST162.png",
        p2,
        device = "png",

### 2019-08-16 notebooks/Clustering analysis.Rmd

@@
   scale_fill_manual(values = col_palette) + 
-  annotate("label", x = 0.27, y = 270, label = "Clade A") +
-  annotate("label", x = 0.29, y = 251, label = "Clade B") +
-  annotate("label", x = 0.39, y = 167, label = "Clade C") +
-  annotate("label", x = 0.38, y = 138, label = "Clade D") +
-  annotate("label", x = 0.32, y = 57, label = "Clade E") +
-  annotate("label", x = 0.31, y = 31, label = "Clade F")
+  annotate("label", x = 0.26, y = 274, label = "Clade A") +
+  annotate("label", x = 0.27, y = 262, label = "Clade B") +
+  annotate("label", x = 0.40, y = 167, label = "Clade C") +
+  annotate("label", x = 0.39, y = 138, label = "Clade D") +
+  annotate("label", x = 0.33, y = 57, label = "Clade E") +
+  annotate("label", x = 0.32, y = 31, label = "Clade F")
 
@@
 
-names(dist_data) <- c("Clade D", "Clade F", "Clade A", "Clade E", "Clade B", "Clade C")
+names(dist_data) <- c("Clade D", "Clade F", "Clade A", "Clade B", "Clade E", "Clade C")
 
@@
 ST10_raw <- read.iqtree("../data/Phylogeny/STs/ST10/iqtree/iqtree.contree")
+ST162_A_raw <- read.iqtree("../data/Phylogeny/STs/ST162_A/iqtree/iqtree.contree")
 ST355_raw <- read.iqtree("../data/Phylogeny/STs/ST355/iqtree/iqtree.contree")
-ST602_raw <- read.iqtree("../data/Phylogeny/STs/ST602/iqtree/iqtree.contree")
 ST744_raw <- read.iqtree("../data/Phylogeny/STs/ST744/iqtree/iqtree.contree")
@@
 raw_trees <- list(ST10_raw,
+                  ST162_A_raw,
                   ST355_raw,
-                  ST602_raw,
                   ST744_raw)
 
-names(raw_trees) <- c("Clade D","Clade E","Clade B","Clade C")
+names(raw_trees) <- c("Clade D","Clade A","Clade E","Clade C")
 
@@
 annotate_tree(
-  clean_trees$`Clade B`,
+  clean_trees$`Clade A`,
   tree_metadata,
@@
   bootstrap_var = "Bootstrap",
-  clade_label_node = 10,
+  clade_label_node = 27,
   clade_label = "18 SNP diff",
@@
   xlim(0, 0.009) +
-  ggtitle("Clade B, ST602")
+  ggtitle("Clade A, ST162")
 

### 2019-08-20 figure_script.R
changed figures

@@ -136,6 +136,7 @@ library(treeio)
 library(ggtree)
 library(phytools)
 library(readxl)
+library(phangorn)
 library(RColorBrewer)
 library(ggplot2)
 
@@ -248,12 +249,12 @@ isol_info <- read.table("data/id.txt",
                         stringsAsFactors = FALSE)
 
 ST117_raw <- read.iqtree("data/Phylogeny/STs/ST117/iqtree/iqtree.contree")
-ST162_raw <- read.iqtree("data/Phylogeny/STs/ST162/iqtree/iqtree.contree")
+ST162_B_raw <- read.iqtree("data/Phylogeny/STs/ST162_B/iqtree/iqtree.contree")
 
 raw_trees <- list(ST117_raw,
-                  ST162_raw)
+                  ST162_B_raw)
 
-names(raw_trees) <- c("Clade F","Clade A")
+names(raw_trees) <- c("Clade F","Clade B")
 
 clean_trees <- lapply(
   raw_trees,
@@ -329,9 +330,9 @@ p1 <- annotate_tree(
   label_variable = "Location",
   color_variable = "Year",
   shape_variable = "Species",
-  clade_label_node = 23,
-  clade_label = "37 SNP diff",
-  cladelabel_offset = 0.00025,
+  clade_label_node = 21,
+  clade_label = "3 SNP diff",
+  cladelabel_offset = 0.0003,
   bootstrap_lab = FALSE,
   bootstrap_var = "Bootstrap",
   nodepoint_size = 1.5,
@@ -341,11 +342,11 @@ p1 <- annotate_tree(
   node_palette = boot_palette,
   shape_palette = palette_shape
 ) +
-  xlim(0, 0.0025) +
+  xlim(0, 0.0028) +
   ggtitle("Clade F, ST117")
 
 p2 <- annotate_tree(
-  clean_trees_new$`Clade A`,
+  clean_trees_new$`Clade B`,
   tree_metadata,
   layout = "rectangular",
   tree_type = "treedata",
@@ -354,23 +355,23 @@ p2 <- annotate_tree(
   label_variable = "Location",
   color_variable = "Year",
   shape_variable = "Species",
-  clade_label_node = 29,
+  clade_label_node = 14,
   clade_label = "12 SNP diff",
-  cladelabel_offset = 0.0055,
+  cladelabel_offset = 0.0012,
   bootstrap_lab = FALSE,
   bootstrap_var = "Bootstrap",
[…]

### 2019-08-20 notebooks/Clustering analysis.Rmd
removed cgmlst results and added evolutionary models to the trees

@@
 library(tibble)
-library(knitr)
 library(vegan)
@@
 
-```
-
-# Minimum spanning tree
-The EnteroBase cgMLST scheme [1] for _E. coli_ was used with chewBBACA [2] to generate the cgMLST data.
-A minimum spanning tree calculated from the cgMLST data with GrapeTree [3] using MSTreeV2.
-Number of allele differences are denoted as numbers on branches between nodes.
-The branches have been collapsed on ten differences or less.
-The number inside each node is the sequence type based on the EnteroBase 7-gene MLST scheme [4].
-```{r, fig.width=15, fig.height=12}
-include_graphics("../figures/GrapeTree_cgMLST.png")
 ```
@@
 The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.
+
+Evolutionary models:
+
+- Clade A: K2P+ASC+R2
+
+- Clade C: K3P+ASC
+
+- Clade D: K2P+ASC+R2
+
+- Clade E: K3P+ASC
+
 ```{r message=FALSE, warning=FALSE}
@@
 ```
-
-# References
-[1] Alikhan N-F et al.
-A genomic overview of the population structure of Salmonella.
-Casadesús J, editor.
-PLOS Genet [Internet].
-2018 Apr 5;14(4):e1007261.
-Available from: https://dx.plos.org/10.1371/journal.pgen.1007261
-
-[2] Machado MP et al. chewBBACA: A complete suite for gene-by-gene schema creation and strain identification. Microb Genomics. 2018;1–7.
-
-[3] Z Zhou et al. (2018) “GrapeTree: Visualization of core genomic relationships among 100,000 bacterial pathogens”, Genome Res. doi: https://doi.org/10.1101/gr.232397.117
-
-[4] Wirth T et al. (2006) "Sex and virulence in Escherichia coli : an evolutionary perspective", Molecular Microbiology (2006) 60 (5), 1136–1151, doi: 10.1111/j.1365-2958.2006.05172.x

### 2019-08-20 notebooks/Clustering analysis.Rmd

@@
                           case_when(
-                            x@data$UFboot >= 95 ~ "> 95",
+                            x@data$UFboot >= 95 ~ ">= 95",
                             x@data$UFboot < 95 &
@@
                           ordered = TRUE,
-                          levels = c("> 95",
+                          levels = c(">= 95",
                                      "80 - 94",
@@
 
-boot_palette <- c("> 95" = "#6a51a3",
+boot_palette <- c(">= 95" = "#6a51a3",
               "80 - 94" = "#9e9ac8",

### 2019-08-21 figure_script.R
removed midrooting of small trees

@@ -304,60 +304,64 @@ year_val <- c(
 
 names(palette_color) <- year_val
 
-boot_palette <- c("> 95" = "#6a51a3",
+boot_palette <- c(">= 95" = "#6a51a3",
               "80 - 94" = "#9e9ac8",
               "50 - 79" = "#cbc9e2",
               "< 50" = "#f2f0f7")
 
 clean_trees_new <- lapply(clean_trees, function(x){
-  x@data$Bootstrap <- factor(case_when(x@data$UFboot >= 95 ~ "> 95",
+  x@data$Bootstrap <- factor(case_when(x@data$UFboot >= 95 ~ ">= 95",
                                 x@data$UFboot <95 & x@data$UFboot >= 80 ~ "80 - 94",
                                 x@data$UFboot < 80 & x@data$UFboot >= 50 ~ "50 - 79",
                                 x@data$UFboot < 50 ~ "< 50"),
                              ordered = TRUE,
-                             levels = c("> 95", "80 - 94", "50 - 79", "< 50"))
+                             levels = c(">= 95", "80 - 94", "50 - 79", "< 50"))
   
   return(x)
 })
+
 
 p1 <- annotate_tree(
   clean_trees_new$`Clade F`,
   tree_metadata,
   layout = "rectangular",
   tree_type = "treedata",
-  midroot = TRUE,
   line_width = 0.5,
   label_variable = "Location",
   color_variable = "Year",
   shape_variable = "Species",
-  clade_label_node = 21,
+  clade_label_node = 26,
   clade_label = "3 SNP diff",
-  cladelabel_offset = 0.0003,
+  cladelabel_offset = 0.0005,
   bootstrap_lab = FALSE,
   bootstrap_var = "Bootstrap",
   nodepoint_size = 1.5,
   tippoint_size = 4,
-  label_offset = 0.00005,
+  label_offset = 0.00008,
   color_palette = palette_color,
   node_palette = boot_palette,
   shape_palette = palette_shape
 ) +
-  xlim(0, 0.0028) +
-  ggtitle("Clade F, ST117")
+  xlim(0, 0.0045) +
+  ggtitle("Clade F, ST117") +
+  scale_color_manual(values = boot_palette,
+                     breaks = c(">= 95",
+                                "80 - 94",
+                                "50 - 79",
+                                "< 50"))
 
 p2 <- annotate_tree(
   clean_trees_new$`Clade B`,
   tree_metadata,
[…]

### 2019-08-21 notebooks/Clustering analysis.Rmd

@@
   tree_type = "treedata",
-  midroot = TRUE,
   line_width = 0.5,
@@
   bootstrap_var = "Bootstrap",
-  clade_label_node = 27,
+  clade_label_node = 26,
   clade_label = "18 SNP diff",
-  cladelabel_offset = 0.001,
+  cladelabel_offset = 0.0014,
   bootstrap_lab = FALSE,
   tippoint_size = 4,
-  label_offset = 0.0002,
+  label_offset = 0.00018,
   shape_palette = palette_shape,
@@
 ) +
-  xlim(0, 0.009) +
-  ggtitle("Clade A, ST162")
+  xlim(0, 0.0125) +
+  ggtitle("Clade A, ST162") +
+  scale_color_manual(values = boot_palette,
+                     breaks = c(">= 95",
+                                "80 - 94",
+                                "50 - 79",
+                                "< 50"))
 
@@
   bootstrap_var = "Bootstrap",
-  clade_label_node = 12,
+  clade_label_node = 14,
   clade_label = "50 SNP diff",
-  cladelabel_offset = 0.00024,
+  cladelabel_offset = 0.00028,
   bootstrap_lab = FALSE,
@@
 ) +
-  xlim(0, 0.0030) +
-  ggtitle("Clade C, ST744")
+  xlim(0, 0.0032) +
+  ggtitle("Clade C, ST744") +
+  scale_color_manual(values = boot_palette,
+                     breaks = c(">= 95",
+                                "80 - 94",
+                                "50 - 79",
+                                "< 50"))
 
@@
   bootstrap_var = "Bootstrap",
-  clade_label_node = 19,
-  clade_label = "10 SNP diff",
+  clade_label_node = 24,
+  clade_label = "40 SNP diff",
   cladelabel_offset = 0.00055,
@@
 ) +
-  xlim(0, 0.005) +
-  ggtitle("Clade D, ST10")
+  xlim(0, 0.006) +
+  ggtitle("Clade D, ST10") +
+  scale_color_manual(values = boot_palette,
+                     breaks = c(">= 95",
+                                "80 - 94",
+                                "50 - 79",
+                                "< 50"))
 
@@
   clade_label = "6 SNP diff",
-  cladelabel_offset = 0.0013,
+  cladelabel_offset = 0.002,
   bootstrap_lab = FALSE,
@@
 ) +
-  xlim(0, 0.0165) +
-  ggtitle("Clade E, ST355")
+  xlim(0, 0.03) +
+  ggtitle("Clade E, ST355") +
+  scale_color_manual(values = boot_palette,
+                       breaks = c(">= 95",
+                                  "80 - 94",
+                                  "50 - 79",
+                                  "< 50"))
 

### 2019-08-23 notebooks/Clustering analysis.Rmd
added shared genome % table to selected clades chapter

@@
 
-# Maximum likelihood trees on clades of interest
-Each tree is midpoint rooted for visual purposes.
+# Maximum likelihood trees on selected clades
+The table presents the evolutionary model predicted by IQTree, and the percent shared genome calculated by ParSNP for each selected clade
+
+| Clade | Evolutionary model | Shared genome |
+|-------|--------------------|---------------|
+| A     | K2P+ASC+R2         | 88.4%         |
+| C     | K3P+ASC            | 88.4%         |
+| D     | K2P+ASC+R2         | 87.2%         |
+| E     | K3P+ASC            | 90.2%         |
+
 The values on each node is the UF Bootstrap values calculated by IQTree.
@@
 The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.
-
-Evolutionary models:
-
-- Clade A: K2P+ASC+R2
-
-- Clade C: K3P+ASC
-
-- Clade D: K2P+ASC+R2
-
-- Clade E: K3P+ASC
-
 ```{r message=FALSE, warning=FALSE}

### 2019-08-29 notebooks/article_notebook2.Rmd
added table on prokka refs
minor changes in text

@@
 This table presents the groups created in the isolate selection.
-The "Additional Resistances" column contains how many additional antimicrobials the isolates in that group is resistant to, other than CIP and/or NAL.
+The "Additional Resistances" column contains how many additional antimicrobials the isolates in that group is resistant to, in addition to CIP and/or NAL.
 ```{r}
@@
 ```
+
+# Prokka reference genomes
+The genomes listed below were used as references for the Prokka annotation, and downloaded with ncbi-genome-download (https://github.com/kblin/ncbi-genome-download) with the following commands:
+
+--refseq-category reference
+
+--assembly-level complete
+
+--genus "Escherichia coli" bacteria
+
+| Accession number           |  Information                 |
+|----------------------------|------------------------------|
+| GCF_000005845.2_ASM584v2   | _E. coli_ K12                |
+| GCF_000008865.2_ASM886v2   | _E. coli_ O157:H7 Sakai      |
+| GCF_000026345.1_ASM2634v1  | _E. coli_ IAI39              |
+| GCF_000183345.1_ASM18334v1 | _E. coli_ O83:H1 NRG 857C    |
+| GCF_000299455.1_ASM29945v1 | _E. coli_ O104:H4 2011C-3493 |
 

### 2019-08-29 notebooks/Clustering analysis.Rmd

@@
 # Patristic distances
-Patristic distances were calculated from the core gene alignment SNP tree.
+Patristic distances were calculated from the core gene alignment SNP tree. The labelled clades are the groups further investigated.
 ```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
@@
 | E     | K3P+ASC            | 90.2%         |
+
+\
+\
+\
+
+**Subclade trees**
 
@@
 # P-value calculations on SNP distances within animal species
-The table presents the non-exact p-values based on 1000 iterations of minimum SNP distances to closest isolate within animal species.
+The table presents the non-exact p-values based on 1000 iterations of minimum SNP distances to the closest isolate within each animal species.
 The figure shows the distribution of the expected values with the observed values for each animal species denoted as an arrow.

### 2019-10-22 figure_script.R
major changes to figures

@@ -9,112 +9,112 @@ library(cluster)
 library(ape)
 
 ## res per species
-gene_names <- names(mut_report)
+palette <- c("Acquired" = "#FB8C6F",
+             "Intrinsic" = "#73607D")
 
-gene_names <- gene_names[-c(1,2)]
+acquired_report <- read.table("data/ariba/resfinder_quin_report.txt",
+                              sep = "\t",
+                              header = TRUE,
+                              stringsAsFactors = FALSE) %>%
+  select(id, contains("qnr"), qepA4)
 
-palette <- c("Broiler" = "#4575b4",
-             "Pig" = "#74add1",
-             "Red fox" = "#f46d43",
-             "Wild bird" = "#fdae61")
+acquired_report <- acquired_report[,colSums(acquired_report != 0) > 0]
 
-ggsave(
- "C:/Users/vi1511/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/res_per_species.tiff",
-  mut_report %>%
-   filter(id %not_in% ex_samples) %>%
-   left_join(isolate_data, by = "id") %>%
-   gather(key, value, gene_names) %>%
-   group_by(key, value, species) %>%
-   dplyr::count() %>%
-   ungroup() %>%
-   mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-   spread(value, n, fill = 0) %>%
-   rowwise() %>%
-   mutate(
-     Total = Present + Absent,
-     Percent = round(Present / Total * 100, 1),
-     lwr = round(get_binCI(Present, Total)[1], 1),
-     upr = round(get_binCI(Present, Total)[2], 1)
-   ) %>%
-   mutate(species = factor(species,
-                           levels = c("Broiler","Pig","Wild bird","Red fox")),
-          key = paste0(toupper(substr(key, 1, 1)),
-                       substr(key, 2, nchar(key))),
-          group = if_else(key %in% c("GyrA","GyrB","ParC","ParE"), 1, 2)) %>%
-   ggplot(aes(key, Percent, fill = species)) +
-   geom_col(color = "black",
-            position = position_dodge(1)) +
-   scale_fill_manual(values = palette) +
-   scale_x_discrete(limits = c("GyrA","GyrB","ParC","ParE",
-                               "MarA","MarR","SoxR","RpoB",
-                               "RobA")) +
-   labs(y = "Percent (%) of isolates") +
-   guides(fill = FALSE) +
[…]

### 2019-10-22 scripts/correlation_tests.R
removed notebook, added correlation script

@@ -1,0 +1,236 @@
+# Correlation test script
+
+library(dplyr)
+library(tidyr)
+library(readr)
+library(tibble)
+library(ggplot2)
+
+ex_samples <- c(
+  "2016-17-565-1-S",
+  "2016-02-428-2-S",
+  "2016-02-486-2-S",
+  "2016-02-732-2-S",
+  "2014-01-1741-1-S"
+)
+
+id_data <- read.table("data/id.txt",
+                      sep = "\t",
+                      header = TRUE,
+                      stringsAsFactors = F)
+
+isolate_data <- read.table("data/isolate_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F) %>%
+  filter(!id %in% ex_samples)
+
+acquired_data <- read.table("data/VAMPIR/amr_ac/acquired_gene_report.tsv",
+                            sep = "\t",
+                            header = TRUE,
+                            stringsAsFactors = FALSE) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+
+merged_acquired_data <- read.table("merged_acquired_report.txt",
+                                   sep = ",",
+                                   header = TRUE,
+                                   stringsAsFactors = FALSE)
+
+intrinsic_genes <- read.table("data/ariba/mut_report.txt",
+                             sep = "\t",
+                             header = TRUE,
+                             stringsAsFactors = FALSE)
+
+
+
+all_mechanisms <- acquired_data %>%
+  left_join(intrinsic_genes, by = "id") %>%
+  select(-id)
+
+merged_mechanisms <- merged_acquired_data %>%
+  left_join(intrinsic_genes, by = "id") %>%
+  select(-id)
+
+results <- list()
+results_merged <- list()
+
+for (i in names(all_mechanisms)) {
+  results[[i]] <- apply(all_mechanisms, 2, cor.test, all_mechanisms[[i]], method="pearson", exact = FALSE)
+  
+}
+
+for (i in names(merged_mechanisms)) {
+  results_merged[[i]] <- apply(merged_mechanisms, 2, cor.test, merged_mechanisms[[i]], method="pearson", exact = FALSE)
+  
+}
+
+library(broom)
+
+test <- lapply(results, function(x) lapply(x, tidy))
+
+
+test2 <- lapply(test, function(x) bind_rows(x, .id = "ref2"))
+
+
[…]

### 2019-10-22 acquired_genes_script.R
moved files

@@ -1,533 +1,0 @@
-library(readr)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(cowplot)
-
-# Import data
-
-func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
-
-ex_samples <- c(
-  "2016-17-565-1-S",
-  "2016-02-428-2-S",
-  "2016-02-486-2-S",
-  "2016-02-732-2-S",
-  "2014-01-1741-1-S"
-)
-
-isolate_data <- read.table("data/isolate_data.txt",
-                           sep = "\t",
-                           header = T,
-                           stringsAsFactors = F) %>%
-  filter(!id %in% ex_samples)
-
-id_data <- read.table("data/id.txt",
-                      sep = "\t",
-                      header = TRUE,
-                      stringsAsFactors = F)
-
-mic_data <- read.table("data/mic_data.txt",
-                       sep = "\t",
-                       header = T,
-                       stringsAsFactors = F) %>%
-  filter(!id %in% ex_samples) %>%
-  select(-antres) %>%
-  mutate(SMX_R = if_else(SMX > 64, 1, 0),
-         TMP_R = if_else(TMP > 2, 1, 0),
-         CIP_R = if_else(CIP > 0.06, 1, 0),
-         TET_R = if_else(TET > 8, 1, 0),
-         NAL_R = if_else(NAL > 16, 1, 0),
-         CTX_R = if_else(CTX > 0.25, 1, 0),
-         CHL_R = if_else(CHL > 16, 1, 0),
-         AMP_R = if_else(AMP > 8, 1, 0),
-         GEN_R = if_else(GEN > 2, 1, 0)) %>%
-  select(id, contains("_R"))
-
-names(mic_data) <- sub("_R", "", names(mic_data))
-
-acquired_data <- read_delim("data/VAMPIR/amr_ac/acquired_gene_report.tsv",
-                            delim = "\t") %>%
-  left_join(id_data, by = "ref") %>%
-  select(id, everything(), -ref)
-
-
-
-## Create groups from acquired gene data
-
-aac3II_names <- grep("aac3", names(acquired_data), value = TRUE, ignore.case = TRUE)
-aadA_names <- grep("aadA", names(acquired_data), value = TRUE, ignore.case = TRUE)
-aph_names <- grep("aph", names(acquired_data), value = TRUE, ignore.case = TRUE)
-blaCTXM_names <- grep("ctx", names(acquired_data), value = TRUE, ignore.case = TRUE)
[…]

### 2019-10-22 figure_script.R

@@ -1,586 +1,0 @@
-## Script for generating figures for the article
-
-library(ggplot2)
-library(dplyr)
-library(ggtree)
-library(tibble)
-library(vegan)
-library(cluster)
-library(ape)
-
-## res per species
-palette <- c("Acquired" = "#FB8C6F",
-             "Intrinsic" = "#73607D")
-
-acquired_report <- read.table("data/ariba/resfinder_quin_report.txt",
-                              sep = "\t",
-                              header = TRUE,
-                              stringsAsFactors = FALSE) %>%
-  select(id, contains("qnr"), qepA4)
-
-acquired_report <- acquired_report[,colSums(acquired_report != 0) > 0]
-
-mut_report <- read.table("data/ariba/megares_quin_report.txt",
-                         sep = "\t",
-                         header = TRUE,
-                         stringsAsFactors = FALSE)
-
-intrinsic_genes <- names(mut_report)
-
-test <- mut_report %>%
-  left_join(acquired_report, by = "id") %>%
-  select(-nt_change) %>%
-  left_join(isolate_data[, c("id", "species")], by = "id") %>%
-  gather(key, value,-c(id, species)) %>%
-  group_by(key, value, species) %>%
-  dplyr::count() %>%
-  ungroup() %>%
-  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
-  spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present / Total * 100, 1)) %>%
-  mutate(
-    key2 = ifelse(key %in% intrinsic_genes, paste0(toupper(substr(key, 1, 1)), substr(key, 2, nchar(key))), key),
-    species = factor(species,
-                     levels = c("Broiler", "Pig", "Wild bird", "Red fox")),
-    group = if_else(
-      key %in% intrinsic_genes,
-      "Intrinsic",
-      "Acquired"
-    )
-  )
-
-res_plot <- ggplot(test, aes(key2, Percent, fill = group)) +
-  geom_col(color = "black",
-           size = 1,
-           position = position_dodge(1)) +
-  scale_fill_manual(values = palette) +
-  scale_x_discrete(
-    limits = c(
-      "GyrA",
-      "GyrB",
-      "ParC",
-      "ParE",
-      "MarA",
-      "MarR",
-      "SoxR",
-      "RpoB",
[…]

### 2019-10-22 scripts/figure_script.R
moved files

@@ -1,0 +1,586 @@
+## Script for generating figures for the article
+
+library(ggplot2)
+library(dplyr)
+library(ggtree)
+library(tibble)
+library(vegan)
+library(cluster)
+library(ape)
+
+## res per species
+palette <- c("Acquired" = "#FB8C6F",
+             "Intrinsic" = "#73607D")
+
+acquired_report <- read.table("data/ariba/resfinder_quin_report.txt",
+                              sep = "\t",
+                              header = TRUE,
+                              stringsAsFactors = FALSE) %>%
+  select(id, contains("qnr"), qepA4)
+
+acquired_report <- acquired_report[,colSums(acquired_report != 0) > 0]
+
+mut_report <- read.table("data/ariba/megares_quin_report.txt",
+                         sep = "\t",
+                         header = TRUE,
+                         stringsAsFactors = FALSE)
+
+intrinsic_genes <- names(mut_report)
+
+test <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  select(-nt_change) %>%
+  left_join(isolate_data[, c("id", "species")], by = "id") %>%
+  gather(key, value,-c(id, species)) %>%
+  group_by(key, value, species) %>%
+  dplyr::count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present / Total * 100, 1)) %>%
+  mutate(
+    key2 = ifelse(key %in% intrinsic_genes, paste0(toupper(substr(key, 1, 1)), substr(key, 2, nchar(key))), key),
+    species = factor(species,
+                     levels = c("Broiler", "Pig", "Wild bird", "Red fox")),
+    group = if_else(
+      key %in% intrinsic_genes,
+      "Intrinsic",
+      "Acquired"
+    )
+  )
+
+res_plot <- ggplot(test, aes(key2, Percent, fill = group)) +
+  geom_col(color = "black",
+           size = 1,
+           position = position_dodge(1)) +
+  scale_fill_manual(values = palette) +
+  scale_x_discrete(
+    limits = c(
+      "GyrA",
+      "GyrB",
+      "ParC",
+      "ParE",
+      "MarA",
+      "MarR",
+      "SoxR",
+      "RpoB",
[…]

### 2019-10-22 notebooks/supplementary.Rmd
added supplementary notebook

---
title: "Supplementary material"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    pdf_document:
      toc: yes
      toc_depth: '2'
---

```{r message = FALSE, warning=FALSE, echo=FALSE}
# Libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(ape)
library(tibble)
library(vegan)
library(distanceR)
library(impoRt)
library(readxl)
library(readr)
library(iteRate)
library(treeio)
library(adephylo)
library(phangorn)
library(RColorBrewer)
library(kableExtra)
```

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r Import data, include = FALSE}
# Vector of excluded samples
ex_samples <- c(
  "2016-17-565-1-S",
  "2016-02-428-2-S",
  "2016-02-486-2-S",
  "2016-02-732-2-S",
  "2014-01-1741-1-S"
  )

# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

"%not_in%" <- Negate("%in%")

id_data <- read.table("../data/id.txt",
                      sep = "\t",
                      header = TRUE,
                      stringsAsFactors = F)

isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F) %>%
  filter(id %not_in% ex_samples)

mlst_data <- read.table("../data/mlst_results.txt",
                        sep = "\t",
                        header = TRUE,
                        stringsAsFactors = F)

acquired_report <- read.table("../data/ariba/resfinder_quin_report.txt",
                              sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE)

mut_report <- read.table("../data/ariba/megares_quin_report.txt",
                              sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE)

```

\newpage
# Section 1: Supplementary tables
## Table S1: Group listing
This table presents the groups created in the isolate selection. The "Additional Resistances" column contains how many additional antimicrobials the isolates in that group is resistant to, in addition to CIP and/or NAL.
\
```{r,echo=FALSE}
mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(antres, CIP, NAL, species) %>%
  count() %>%
  dplyr::rename("Additional Resistances" = antres) %>%
  ungroup() %>%
  mutate(Group = 1:n()) %>%
  dplyr::select(Group, everything()) %>%
  kable(longtable = TRUE, booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                latex_options = c("hold_position", "repeat_header"))
```

\newpage
## Table S2: Prokka reference genomes
The genomes listed below were used as references for the Prokka annotation, and downloaded with ncbi-genome-download (https://github.com/kblin/ncbi-genome-download) with the following commands:

--refseq-category reference

--assembly-level complete

--genus "Escherichia coli" bacteria
\

| Accession number           |  Information                 |
|----------------------------|------------------------------|
| GCF_000005845.2_ASM584v2   | _E. coli_ K12                |
| GCF_000008865.2_ASM886v2   | _E. coli_ O157:H7 Sakai      |
| GCF_000026345.1_ASM2634v1  | _E. coli_ IAI39              |
| GCF_000183345.1_ASM18334v1 | _E. coli_ O83:H1 NRG 857C    |
| GCF_000299455.1_ASM29945v1 | _E. coli_ O104:H4 2011C-3493 |

\newpage
## Table S3: R and R-Packages
R version 3.6.0 [1] was used in the present study. The following packages were used:
\

| Package       | Version | Reference |
|---------------|---------|-----------|
| dplyr         | 0.8.0.1 |   [2]     |
| tidyr         | 0.8.3   |   [3]     |
| ggplot2       | 3.1.1   |   [4]     |
| purrr         | 0.3.2   |   [5]     |
| tibble        | 2.1.1   |   [6]     |
| Biostrings    | 2.50.2  |   [7]     |
| vegan         | 2.5.4   |   [8]     |
| stringr       | 1.4.0   |   [9]     |
| cluster       | 2.0.8   |   [10]    |
| ape           | 5.3     |   [11]    |
| ggtree        | 1.14.6  |   [12]    |
| RColorBrewer  | 1.1-2   |   [13]    |
| cowplot       | 0.9.4   |   [14]    |
| knitr         | 1.22    |   [15]    |
| kableExtra    | 1.1.0   |   [16]    |
| ggsci         | 2.9     |   [17]    |
| fastqcr       | 0.1.2   |   [18]    |
| scales        | 1.0.0   |   [19]    |
| viridis       | 0.5.1   |   [20]    |
| adephylo      | 1.1-11  |   [21]    |

\newpage
## Table S4: Identified amino acid substitutions in the QRDR of GyrA, ParC and ParE.
The Reference nt column denote the nucleotide codon in the reference used in the MegaRes database, while the Contig nt denote the identified codon in the isolates.
\

| Gene | AA Substitution | Reference nt |	Contig nt	| n	  | Total | Percent |
|------|-----------------|--------------|-----------|-----|-------|---------|
| GyrA | S83L	           | TCG	        | TTG       |	204 |	280	  | 72.9    |
|	     | S83A	           | TCG	        | GCG       |	9	  | 280	  | 3.2     |
|	     | D87Y	           | GAC	        | TAC       |	10	| 280	  | 3.6     |
|	     | D87N	           | GAC	        | AAC       |	36	| 280	  | 12.9    | 
|	     | D87H	           | GAC	        | CAC       |	1	  | 280	  | 0.4     |
|	     | D87G	           | GAC	        | GGC       |	2	  | 280	  | 0.7     |
| ParC | S80R	           | AGC	        | AGA       |	2	  | 280	  | 0.7     |
|	     | S80I	           | AGC	        | ATC       |	37	| 280	  | 13.2    |
|	     | S80I	           | AGC	        | ATT       |	2	  | 280	  | 0.7     |
|	     | S57T	           | AGC	        | ACC       |	2	  | 280	  | 0.7     |
|	     | E84V	           | GAA	        | GTA       |	1	  | 280	  | 0.4     |
|	     | A56T	           | GCC	        | ACC       |	8	  | 280	  | 2.9     |
| ParE | S458A	         | TCG	        | GCG       |	10	| 280	  | 3.6     |
|	     | L488M	         | CTC	        | ATG       |	1	  | 280	  | 0.4     |
|	     | L416F	         | CTT	        | TTT       |	1	  | 280	  | 0.4     |
|	     | H516Y	         | CAC	        | TAC       |	1	  | 280	  | 0.4     |
|	     | D475E	         | GAC	        | GAA       |	15	| 280	  | 5.4     |
|	     | D463N	         | GAT	        | AAT       |	1	  | 280	  | 0.4     |
|	     | A512T	         | GCG	        | ACG       |	1	  | 280	  | 0.4     |

\newpage
## Table S5: QRDR substitutions in GyrA, ParC and ParE per animal species.
The percentage is based on the number of isolates per species with the mutation in question (n) in relation to the total number of isolates per animal species (Total).
\

| Gene | AA substitution | Species   | n  | Total | Percent |
|------|-----------------|-----------|----|-------|---------|
| GyrA | S83L	           | Broiler   | 83	| 88    |	94.3    |
|	     | S83L	           | Pig	     | 50	| 75    |	66.7    |
|	     | S83L	           | Red fox   | 37	| 52    |	71.2    |
|	     | S83L	           | Wild bird | 34	| 65    |	52.3    |
|	     | S83A	           | Pig	     | 4  |	75    |	5.3     |
|	     | S83A	           | Red fox	 | 2  |	52    |	3.8     |
|	     | S83A	           | Wild bird | 3  |	65    |	4.6     |
|	     | D87Y	           | Broiler	 | 2  |	88    |	2.3     |
|	     | D87Y	           | Pig	     | 2  |	75    |	2.7     |
|	     | D87Y	           | Red fox	 | 1  |	52    |	1.9     |
|	     | D87Y	           | Wild bird | 5  |	65    |	7.7     |
|	     | D87N	           | Broiler	 | 9  |	88    |	10.2    |
|	     | D87N	           | Pig	     | 5  |	75    |	6.7     |
|	     | D87N	           | Red fox	 | 6  |	52    |	11.5    |
|	     | D87N	           | Wild bird | 16 |	65    |	24.6    |
|	     | D87H	           | Broiler	 | 1  |	88    |	1.1     |
|	     | D87G	           | Red fox	 | 2  |	52    |	3.8     |
| ParC | S80R	           | Pig	     | 2  |	75    |	2.7     |
|	     | S80I	           | Broiler	 | 8  |	88    |	9.1     |
|	     | S80I	           | Pig	     | 5  |	75    |	6.7     |
|	     | S80I	           | Red fox   | 10 |	52    |	19.2    |
|	     | S80I	           | Wild bird | 16	| 65    |	24.6    |
|	     | S57T	           | Pig	     | 2  |	75    |	2.7     |
|	     | E84V	           | Red fox	 | 1  |	52    |	1.9     |
|	     | A56T	           | Pig	     | 4  |	75    |	5.3     |
|	     | A56T	           | Red fox	 | 1  |	52    |	1.9     |
|	     | A56T	           | Wild bird | 3  |	65    |	4.6     |
| ParE | S458A	         | Pig	     | 3  |	75    |	4.0     |
|	     | S458A	         | Red fox	 | 1  |	52    |	1.9     |
|	     | S458A	         | Wild bird | 6  |	65    |	9.2     |
|	     | L488M	         | Broiler	 | 1  |	88    |	1.1     |
|	     | L416F	         | Pig	     | 1  |	75    |	1.3     |
|	     | H516Y	         | Wild bird | 1  |	65    |	1.5     |
|	     | D475E	         | Broiler	 | 13 |	88    |	14.8    |
|	     | D475E	         | Red fox	 | 2	| 52    |	3.8     |
|	     | D463N	         | Pig	     | 1	| 75    |	1.3     |
|	     | A512T	         | Broiler	 | 1	| 88    |	1.1     |

\newpage
## Table S6: The most abundant (n > 2) substitutions in MarA, MarR, RpoB and SoxR.
The percentage is based on how many of the total isolates (Total) with substitutions in the respective gene (n) had the described substitution.
\

|Gene  | AA Substitution | n   | Total | Percent |
|------|-----------------|-----|-------|---------|
| MarA | S127N	         | 26  | 34	   | 76.5    |
|	     | L78F	           | 6	 | 34	   | 17.6    |
| MarR | G103S	         | 210 | 211   | 99.5    |
|	     | Y137H	         | 210 | 211   | 99.5    |
|	     | K62R            | 6	 | 211   | 2.8     |
|	     | S3N	           | 6	 | 211   | 2.8     |
|	     | D118N	         | 4	 | 211   | 1.9     |
| RpoB | E320D	         | 37	 | 48	   | 77.1    |
|	     | V261A	         | 4	 | 48	   | 8.3     |
| SoxR | G74R	           | 63	 | 71	   | 88.7    |
|	     | T38S	           | 36	 | 71	   | 50.7    |
|	     | I40V	           | 4	 | 71	   | 5.6     |
|	     | A111T	         | 3	 | 71	   | 4.2     |
|	     | N45D	           | 3	 | 71	   | 4.2     |

\newpage
## Table S7: Mechanisms in isolates without substitutions in QRDR
This table presents the mechanisms identified in isolates with no amino acid substitutions in either of GyrA, GyrB, ParC and/or ParE.
\
```{r,echo=FALSE}
specific_ac_genes <- acquired_report %>%
  gather(PMQR, value, -id) %>%
  filter(value == 1) %>%
  select(-value)

mut_report %>%
  # all gyrB were 0
  filter(gyrA == 0,
         parC == 0,
         parE == 0) %>%
  left_join(specific_ac_genes, by = "id") %>%
  select(-c(nt_change, id)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  dplyr::select(-c(gyrA,gyrB,parC,parE)) %>%
  kable(booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

\newpage
## Table S8: Other plasmid mediated genes
The table presents the plasmid mediated genes related to other resistance phenotypes among the isolates.
\
```{r, echo = FALSE, message=FALSE}
read_delim("../data/VAMPIR/amr_ac/acquired_gene_report.tsv",
                             delim = "\t") %>%
  gather(Gene, value, -ref) %>% 
  group_by(Gene, value) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 2)) %>%
  filter(!Gene %in% c("mdfA","qepA4")) %>%
  filter(grepl("qnr", Gene) == FALSE) %>%
  mutate(Included = if_else(grepl("mph|lnu|msr|erm",
                                  Gene,
                                  ignore.case = T) == TRUE, "No", "Yes")) %>%
  arrange(-Percent) %>%
  kable(booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(1, italic = T)

```

\newpage
## Table S9: Occurrence of antimicrobial resistance among the selected quinolone resistant isolates.
The table presents the percent (%) occurrence of AMR among the selected isolates (n=280) identified through MIC values in the Norwegian monitoring program from 2006 to 2017. The cut off values used were defined by EUCAST.
\

```{r, echo=FALSE, message=FALSE}
t1 <- read_delim("../data/phenotypic_resistance.txt",
           delim = "\t")

names(t1)[2] <- paste0(names(t1)[2],"\\textsuperscript{a}")
names(t1)[3] <- paste0(names(t1)[3],"\\textsuperscript{b}")
names(t1)[4] <- paste0(names(t1)[4],"\\textsuperscript{c}")
names(t1)[5] <- paste0(names(t1)[5],"\\textsuperscript{d}")
names(t1)[6] <- paste0(names(t1)[6],"\\textsuperscript{e}")

t1 %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling(full_width = TRUE, position = "center") %>%
  footnote(alphabet = c("87 isolates",
                        "75 isolates",
                        "52 isolates",
                        "66 isolates",
                        "280 isolates"))
```

\newpage
## Table S10: SNP distance statistics on isolates from the same species.
The table presents the summary statistics on the minimum SNP distance to the closest isolate from same animal species.
\

| Statistic | Broiler | Pig   | Red fox | Wild bird |
|-----------|---------|-------|---------|-----------|
| Min	      | 0	      | 0	    | 0	      | 0         |
| 1st.Q     | 2	      | 2	    | 9	      | 1         | 
| Median    | 12	    | 212	  | 1388    | 53        | 
| Mean	    | 3193	  | 4310  | 7210    | 5323      |
| 3rd.Q	    | 52	    | 3471  | 12945	  | 7922      |
| Max	      | 75534	  | 47826 | 46955	  | 33629     |

\newpage
# Section 2: Supplementary figures

## Figure S1: Patristic distances
Patristic distances were calculated from the core gene alignment SNP tree. The labelled clades are the groups further investigated.
\
\
\
```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE,echo=FALSE}
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species)

iqtree <- read.iqtree("../data/Phylogeny/Roary/iqtree_roary.contree")
iqtree_clean <- fix_tip_labels(iqtree, id_data, "_pilon_spades.fasta", tree_type = "treedata")

pat_dist <- as.matrix(distTips(iqtree_clean@phylo, method = "patristic"))

iqtree_clean@phylo <- midpoint(iqtree_clean@phylo, labels = "support")

pat_dist_id <- as.data.frame(pat_dist) %>%
  rownames_to_column("isol1") %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  left_join(tree_metadata, by = c("isol1" = "id")) %>%
  rename("ST1" = ST,
         "Species1" = Species,
         "Location1" = Location,
         "Year1" = Year) %>%
  left_join(tree_metadata, by = c("isol2" = "id")) %>%
  rename("ST2" = ST,
         "Species2" = Species,
         "Location2" = Location,
         "Year2" = Year) %>%
  filter(value < 0.005) %>%
  mutate(selection = case_when(Species1 == Species2 & Location1 != Location2 ~ "ss",
                               Species1 != Species2 & is.na(Location1) == FALSE ~ "as")) %>%
  filter(selection %in% c("ss", "as"))

tree_metadata <- read.table("../data/chewbbaca/complete_results/tree_metadata.txt",
                            sep = "\t",
                            header = TRUE) %>%
  left_join(location_data, by = "id") %>%
  mutate(Year = substr(id, 1, 4)) %>%
  dplyr::rename("Species" = species) %>%
  mutate(ST_new = ifelse(ST %in% pat_dist_id$ST1, ST, "Other"))

pat_dist_bin <- as.data.frame(ifelse(pat_dist < 0.003, "< 0.003", "> 0.003"))

col_palette <- c("< 0.003" = "#bebada",
                 "> 0.003" = "grey95")

palette <- c("Broiler" = "#4575b4",
             "Pig" = "#74add1",
             "Red fox" = "#f46d43",
             "Wild bird" = "#fdae61")

p <- ggtree(iqtree_clean) %<+% tree_metadata +
  geom_tiplab(aes(label = ST),
              size = 1,
              align = TRUE) +
  scale_color_manual(values = palette) +
  geom_tippoint(aes(color = Species))

gheatmap(p,
         pat_dist_bin,
         color = NULL,
         offset = 0.01,
         colnames = FALSE) +
  scale_fill_manual(values = col_palette) + 
  annotate("label", x = 0.26, y = 274, label = "Clade A") +
  annotate("label", x = 0.27, y = 262, label = "Clade B") +
  annotate("label", x = 0.40, y = 167, label = "Clade C") +
  annotate("label", x = 0.39, y = 138, label = "Clade D") +
  annotate("label", x = 0.33, y = 57, label = "Clade E") +
  annotate("label", x = 0.32, y = 31, label = "Clade F")

```

\newpage
## Figure S2: Distance distribution plot
The figure shows the distribution of the expected values with the observed values for each animal species denoted as an arrow. Red fox results is excluded from the plot as the observed distance values were too high to visualize.
\
\
\
```{r echo=FALSE}
# Import distance data
total_dist <- read_delim("../data/snp_dists/final_Roary_gubbins_SNP_dists.tab",
                         delim = "\t") %>%
  rename("isol1" = `snp-dists 0.6.3`)

# Set row/column values to unique species ID
row_id <- as.data.frame(total_dist$isol1) %>%
  rename("ref" = `total_dist$isol1`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1)

col_id <- as.data.frame(names(total_dist)) %>%
  rename("ref" = `names(total_dist)`) %>%
  mutate(ref = sub("_pilon_spades.fasta", "", ref)) %>%
  left_join(id_data, by = "ref") %>%
  select(id) %>%
  left_join(isolate_data[,c("id", "species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         isol1 = paste(species, n, sep = "_")) %>%
  ungroup() %>%
  select(isol1) %>%
  mutate(isol1 = ifelse(isol1 == "NA_1", "isol1", isol1))

total_dist$isol1 <- row_id$isol1
names(total_dist) <- col_id$isol1

# Create data frame for iteration analysis
run_df <- total_dist %>%
  gather(isol2, value, -isol1) %>%
  filter(isol1 != isol2) %>%
  mutate(species1 = sub("_.+", "", isol1),
         species2 = sub("_.+", "", isol2)) %>%
  filter(species1 == species2) %>%
  group_by(isol2) %>%
  mutate(min_val = min(value)) %>%
  ungroup() %>%
  select(isol1, isol2, min_val, species1, species2) %>%
  rename("value" = min_val,
         "group1" = species1,
         "group2" = species2)

# Run iterations
df <- iterate(run_df, runs = 1000, method = "median")

# Filter out the observed values
segment_data <- filter(df, iter == 0,
                       group1 %in% c("Broiler", "Pig", "Wild bird"))

# Calculate non-exact p-values
df2 <- df %>%
  group_by(group1) %>%
  mutate(test = case_when(group1 == "Broiler" & metric <= 12 ~ 1,
                          group1 == "Wild bird" & metric <= 53 ~ 1,
                          group1 == "Pig" & metric <= 212 ~ 1,
                          group1 == "Red fox" & metric <= 1388.5 ~ 1,
                          TRUE ~ 0)) %>%
  split(f = .$group1)

# Plot results, exclude red fox (too high values for the plot to be visible)
plot_df <- df %>%
  filter(group1 %in% c("Broiler", "Pig", "Wild bird"))

ggplot(plot_df, aes(metric)) +
  geom_histogram(
    aes(y = ..density..),
    binwidth = 0.2,
    color = "black",
    fill = "grey80"
  ) +
  stat_function(fun = dnorm,
                args = with(df, c(
                  mean = mean(metric), sd = sd(metric)
                )),
                color = "red") +
  geom_segment(
    segment_data,
    mapping = aes(
      x = metric,
      xend = metric,
      y = 0.5,
      yend = 0,
      color = group1
    ),
    arrow = arrow(length = unit(0.3, "cm"),
                  type = "closed")
  ) +
  geom_text(segment_data,
            mapping = aes(label = group1,
                          x = metric),
            y = 1,
            angle = 90) +
  guides(color = FALSE) +
  labs(x = "Min. number of SNPs to closest isolate",
       y = "Density") +
  ggtitle(label = "Distribution of expected values",
          subtitle = paste("Number of iterations: ", max(df$iter))) +
  theme(plot.title = element_text(hjust = 0)) +
  scale_color_manual(values = palette)

```

\newpage

## Figure S3: Stressplot
Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types. The stressplot visualizes the goodness of fit.
\
\
\
```{r echo=FALSE}
res_mechanisms <- read.table("../data/MDS_data.txt",
                       sep = "\t",
                       header = TRUE,
                       stringsAsFactors = FALSE) %>%
  column_to_rownames("id")

bin_dist <- dist(res_mechanisms,
                 method = "binary")

nmds_df <- metaMDS(bin_dist,
                   trymax = 200,
                   trace = FALSE) # silence the output

stressplot(nmds_df)

```

\newpage
## Figure S4: Correlation between chromosomal genes and plasmid mediated genes
The figure presents the Pearson correlation values for amino acid substitutions in GyrA, ParC, ParE, MarA, MarR, RpoB, and SoxR against all  plasmid mediated genes identified. Grey = negative correlation, green = positive correlation. Black lines denote 95% confidence intervals. Red stars denote significant correlations. The plasmid mediated genes are grouped based on gene family: _dfrA_ = _dfrA1_, _dfrA5_, _dfrA8_, _dfrA12_, _dfrA14_, and _dfrA17_. _tet_ = _tetA_, _tetB_, and _tetD_. _sul_ = _sul1 - 3_. _aph_ = _aph3Ia_, _aph3Ib_, and _aph6Id_. _aadA_ = _aadA1_, _aadA2_, _aadA5_, _aadA12_, _aadA13_, and _aadA22_. _AAC(3')-II_ = _AAC(3')-IIa_ and _AAC(3')-IId_. _bla_~TEM~ = _bla_~TEM-1-A~ – _bla_~TEM-1-C~. _bla_~SHV~ = _bla_~SHV-2~ and _bla_~SHV-12~. _bla_~CTX-M~ = _bla_~CTX-M-1~, _bla_~CTX-M-15~, _bla_~CTX-M-32~, and _bla_~CTX-M-55~. _qnr_ = _qnrA1_, _qnrB19_, _qnrS1_, _qnrS2_, and _qnrS4_.

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
knitr::include_graphics("../figures/corr_chrom_vs_plasmid.png")
```

\newpage

## Figure S5: Maximum likelihood tree, Clade A
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K2P+ASC+R2. Shared genome: 88.4%.
\
\
\
```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeA_ST162A.png")
```

\newpage
## Figure S6: Maximum likelihood tree, Clade C
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K3P+ASC. Shared genome: 88.4%.
\
\
\
```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeC_ST744.png")
```

\newpage
## Figure S7: Maximum likelihood tree, Clade D
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K2P+ASC+R2. Shared genome: 87.2%.
\
\
\
```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeD_ST10.png")
```

\newpage
## Figure S8: Maximum likelihood tree, Clade E
The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. Evolutionary model: K3P+ASC. Shared genome: 90.2%.
\
\
\
```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
knitr::include_graphics("../figures/complete_st_trees/CladeE_ST355.png")
```

\newpage
# Section 3: Metadata and reference sequences

The whole of the supplementary material can also be accessed on github (https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/).

## 3.1: Isolate information and metadata
Metadata for all 280 included isolates can be downloaded here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/total_data.xlsx

## 3.2: FastQC analysis results
Raw read quality control results can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/fastqc_analysis.html

## 3.3: Mash screen results
Contaminant screening results can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/mash_analysis.html

## 3.4: QRDR determination of GyrB, ParC and ParE
The K12 reference sequence on the whole protein is listed as the first entry in each file. The rest of the fasta entries are taken from Jiménez Gómez et al. 2004.

GyrB:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/gyrB_QRDR_ref.fasta

ParC:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/parC_QRDR_ref.fasta

ParE:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/parE_QRDR_ref.fasta

## 3.5: ARIBA flag selection
Data on ARIBA flag selection can be downloaded here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/ariba_flag_selection.xlsx

## 3.6: MEGARes and ResFinder reference sequences

MEGARes:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/total_megares_references.fa

ResFinder:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/reference_genes/total_resfinder_references.fa

## 3.7: QUAST results
Assembly metrics can be accessed here:

https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/notebooks/assembly_metrics.html

\newpage
# References
[1] RCoreTeam. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/. 2018.

[2] Wickham H, François R, Henry L, Müller K. dplyr: A Grammar of Data Manipulation [Internet]. 2019. Available from: https://cran.r-project.org/package=dplyr

[3] Wickham H, Henry L. tidyr: Easily Tidy Data with “spread()” and “gather()” Functions [Internet]. 2019. Available from: https://cran.r-project.org/package=tidyr

[4] Wickham H, Chang W, Henry L, Pedersen TL, Takahashi K, Wilke C, et al. ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics [Internet]. 2019. Available from: https://cran.r-project.org/package=ggplot2

[5] Henry L, Wickham H. purrr: Functional Programming Tools [Internet]. 2019. Available from: https://cran.r-project.org/package=purrr

[6] Müller K, Wickham H. tibble: Simple Data Frames [Internet]. 2019. Available from: https://cran.r-project.org/package=tibble

[7] Pagès H, Aboyoun P, Gentleman R, DebRoy S. Biostrings: Efficient manipulation of biological strings. 2019.

[8] Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, et al. vegan: Community Ecology Package [Internet]. 2019. Available from: https://cran.r-project.org/package=vegan

[9] Wickham H. stringr: Simple, Consistent Wrappers for Common String Operations [Internet]. 2019. Available from: https://cran.r-project.org/package=stringr

[10] Maechler M, Rousseeuw P, Struyf A, Hubert M. cluster: “Finding Groups in Data”: Cluster Analysis Extended Rousseeuw et al. [Internet]. 2019. Available from: https://cran.r-project.org/package=cluster

[11] Paradis E, Blomberg S, Bolker B, Brown J, Claude J, Cuong HS, et al. ape: Analyses of Phylogenetics and Evolution [Internet]. 2019. Available from: https://cran.r-project.org/package=ape

[12] Yu G, Smith DK, Zhu H, Guan Y, Lam TT-Y. Ggtree: a package for visualization and annotation of phylogenetic trees with their covariates and other associated data. McInerny G, editor. Methods Ecol Evol [Internet]. 2017 Jan;8(1):28–36. Available from: http://doi.wiley.com/10.1111/2041-210X.12628

[13] Neuwirth E. RColorBrewer: ColorBrewer Palettes [Internet]. 2014. Available from: https://cran.r-project.org/package=RColorBrewer

[14] Wilke CO. cowplot: Streamlined Plot Theme and Plot Annotations for “ggplot2” [Internet]. 2019. Available from: https://cran.r-project.org/package=cowplot

[15] Xie Y. knitr: A General-Purpose Package for Dynamic Report Generation in R [Internet]. 2019. Available from: https://cran.r-project.org/package=knitr

[16] Zhu H. kableExtra: Construct Complex Table with “kable” and Pipe Syntax [Internet]. 2019. Available from: https://cran.r-project.org/package=kableExtra

[17] Xiao N. ggsci: Scientific Journal and Sci-Fi Themed Color Palettes for “ggplot2” [Internet]. 2018. Available from: https://cran.r-project.org/package=ggsci

[18] Kassambara A. fastqcr: Quality Control of Sequencing Data [Internet]. 2019. Available from: https://cran.r-project.org/package=fastqcr

[19] Wickham H. scales: Scale Functions for Visualization [Internet]. 2018. Available from: https://cran.r-project.org/package=scales

[20] Garnier S. viridis: Default Color Maps from “matplotlib” [Internet]. 2018. Available from: https://cran.r-project.org/package=viridis

[21] Jombart T., Dray S. (2008) adephylo: exploratory analyses for the phylogenetic comparative method.

### 2019-11-17 notebooks/supplementary.Rmd
last adjustments

@@
 
-mic_data <- read.table("../data/mic_data.txt",
+mic_data <- read.table("../data/new_mic_data.txt",
                            sep = "\t",
@@
   filter(id %not_in% ex_samples)
+
+mic_data2 <- read.table("../data/new_mic_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
 
@@
 ## Table S1: Group listing
-This table presents the groups created in the isolate selection.
-The "Additional Resistances" column contains how many additional antimicrobials the isolates in that group is resistant to, in addition to CIP and/or NAL.
+The table presents the different phenotypic groups used in the isolate selection, and the number of isolates from each animal species in each group.
+The number of resistances within each group is listed in the column "No.
+of resistances".
 \
 ```{r,echo=FALSE}
-mic_data %>%
-  left_join(isolate_data[,c("id","species")], by = "id") %>%
-  group_by(antres, CIP, NAL, species) %>%
-  count() %>%
-  dplyr::rename("Additional Resistances" = antres) %>%
+
+isolate_data2 <- read.table("../data/isolate_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+mic_data2 %>%
+  left_join(isolate_data2[,c("id","species")], by = "id") %>%
+  select(antres, CIP, NAL, species) %>%
+  group_by(antres, CIP, NAL) %>%
+  mutate(n = n(),
+         Group = group_indices()) %>%
+  group_by(species, Group) %>%
+  mutate(n_species = n(),
+         new_id = paste0(species, " (", n_species, ")")) %>%
+  group_by(Group) %>%
+  summarise_all(list(func_paste)) %>%
+  select(Group, antres, CIP, NAL, new_id, n) %>%
+  dplyr::rename("No. of resistances" = antres,
+                "Total isolates" = n,
+                "Isolates per species" = new_id) %>%
   ungroup() %>%
-  mutate(Group = 1:n()) %>%
-  dplyr::select(Group, everything()) %>%
   kable(longtable = TRUE, booktabs = TRUE) %>%
@@
 \newpage
-## Table S2: Prokka reference genomes
+## Table S2: Occurrence of antimicrobial resistance among the selected quinolone resistant isolates.
+The table presents the percent (%) occurrence of antimicrobial resistance among the selected isolates (_n_ = 280) identified through minimum inhibitory concentration values in the Norwegian monitoring programme from 2006 to 2017.
+The epidemiological cut off values used were defined by EUCAST.
+Azithromycin was excluded as no epidemiological cutoff value is currently available.
+\
+
+```{r, echo=FALSE, message=FALSE}
+t1 <- read_delim("../data/phenotypic_resistance.txt",
+           delim = "\t")
+
+names(t1)[2] <- paste0(names(t1)[2],"\\textsuperscript{a}")
+names(t1)[3] <- paste0(names(t1)[3],"\\textsuperscript{b}")
+names(t1)[4] <- paste0(names(t1)[4],"\\textsuperscript{c}")
+names(t1)[5] <- paste0(names(t1)[5],"\\textsuperscript{d}")
+names(t1)[6] <- paste0(names(t1)[6],"\\textsuperscript{e}")
+
+t1 %>%
+  kable(booktabs = TRUE, escape = FALSE) %>%
+  kable_styling(full_width = TRUE, position = "center") %>%
+  footnote(alphabet = c("87 isolates",
+                        "75 isolates",
+                        "52 isolates",
+                        "66 isolates",
+                        "280 isolates"))
+```
+
+\newpage
+## Table S3: Prokka reference genomes
 The genomes listed below were used as references for the Prokka annotation, and downloaded with ncbi-genome-download (https://github.com/kblin/ncbi-genome-download) with the following commands:
@@
 \newpage
-## Table S3: R and R-Packages
-R version 3.6.0 [1] was used in the present study. The following packages were used:
-\
-
-| Package       | Version | Reference |
-|---------------|---------|-----------|
-| dplyr         | 0.8.0.1 |   [2]     |
-| tidyr         | 0.8.3   |   [3]     |
-| ggplot2       | 3.1.1   |   [4]     |
-| purrr         | 0.3.2   |   [5]     |
-| tibble        | 2.1.1   |   [6]     |
-| Biostrings    | 2.50.2  |   [7]     |
-| vegan         | 2.5.4   |   [8]     |
-| stringr       | 1.4.0   |   [9]     |
-| cluster       | 2.0.8   |   [10]    |
-| ape           | 5.3     |   [11]    |
-| ggtree        | 1.14.6  |   [12]    |
-| RColorBrewer  | 1.1-2   |   [13]    |
-| cowplot       | 0.9.4   |   [14]    |
-| knitr         | 1.22    |   [15]    |
-| kableExtra    | 1.1.0   |   [16]    |
-| ggsci         | 2.9     |   [17]    |
-| fastqcr       | 0.1.2   |   [18]    |
-| scales        | 1.0.0   |   [19]    |
-| viridis       | 0.5.1   |   [20]    |
-| adephylo      | 1.1-11  |   [21]    |
-
-\newpage
 ## Table S4: Identified amino acid substitutions in the QRDR of GyrA, ParC and ParE.
-The Reference nt column denote the nucleotide codon in the reference used in the MegaRes database, while the Contig nt denote the identified codon in the isolates.
+The table presents the different amino acid substitutions identified in GyrA, ParC and ParE, their respective nucleotide changes, and percent occurrence in total for each amino acid substitution.
+The Reference nt column denote the nucleotide codon in the reference used from the MegaRes database, while the Contig nt denote the identified codon in the isolates.
 \
@@
 ## Table S5: QRDR substitutions in GyrA, ParC and ParE per animal species.
+The table presents each identified amino acid substitution in GyrA, ParC and ParE and their occurrence in isolates from each animal species.
 The percentage is based on the number of isolates per species with the mutation in question (n) in relation to the total number of isolates per animal species (Total).
@@
 |------|-----------------|-----------|----|-------|---------|
-| GyrA | S83L	           | Broiler   | 83	| 88    |	94.3    |
+| GyrA | S83L	           | Broiler   | 83	| 87    |	95.4    |
 |	     | S83L	           | Pig	     | 50	| 75    |	66.7    |
 |	     | S83L	           | Red fox   | 37	| 52    |	71.2    |
-|	     | S83L	           | Wild bird | 34	| 65    |	52.3    |
+|	     | S83L	           | Wild bird | 34	| 66    |	51.5    |
 |	     | S83A	           | Pig	     | 4  |	75    |	5.3     |
 |	     | S83A	           | Red fox	 | 2  |	52    |	3.8     |
-|	     | S83A	           | Wild bird | 3  |	65    |	4.6     |
-|	     | D87Y	           | Broiler	 | 2  |	88    |	2.3     |
+|	     | S83A	           | Wild bird | 3  |	66    |	4.5     |
+|	     | D87Y	           | Broiler	 | 2  |	87    |	2.3     |
 |	     | D87Y	           | Pig	     | 2  |	75    |	2.7     |
 |	     | D87Y	           | Red fox	 | 1  |	52    |	1.9     |
-|	     | D87Y	           | Wild bird | 5  |	65    |	7.7     |
-|	     | D87N	           | Broiler	 | 9  |	88    |	10.2    |
+|	     | D87Y	           | Wild bird | 5  |	66    |	7.6     |
+|	     | D87N	           | Broiler	 | 9  |	87    |	10.3    |
 |	     | D87N	           | Pig	     | 5  |	75    |	6.7     |
 |	     | D87N	           | Red fox	 | 6  |	52    |	11.5    |
-|	     | D87N	           | Wild bird | 16 |	65    |	24.6    |
-|	     | D87H	           | Broiler	 | 1  |	88    |	1.1     |
+|	     | D87N	           | Wild bird | 16 |	66    |	24.2    |
+|	     | D87H	           | Broiler	 | 1  |	87    |	1.1     |
 |	     | D87G	           | Red fox	 | 2  |	52    |	3.8     |
 | ParC | S80R	           | Pig	     | 2  |	75    |	2.7     |
-|	     | S80I	           | Broiler	 | 8  |	88    |	9.1     |
+|	     | S80I	           | Broiler	 | 8  |	87    |	9.2     |
 |	     | S80I	           | Pig	     | 5  |	75    |	6.7     |
 |	     | S80I	           | Red fox   | 10 |	52    |	19.2    |
-|	     | S80I	           | Wild bird | 16	| 65    |	24.6    |
+|	     | S80I	           | Wild bird | 16	| 66    |	24.2    |
 |	     | S57T	           | Pig	     | 2  |	75    |	2.7     |
@@
 |	     | A56T	           | Red fox	 | 1  |	52    |	1.9     |
-|	     | A56T	           | Wild bird | 3  |	65    |	4.6     |
+|	     | A56T	           | Wild bird | 3  |	66    |	4.5     |
 | ParE | S458A	         | Pig	     | 3  |	75    |	4.0     |
 |	     | S458A	         | Red fox	 | 1  |	52    |	1.9     |
-|	     | S458A	         | Wild bird | 6  |	65    |	9.2     |
-|	     | L488M	         | Broiler	 | 1  |	88    |	1.1     |
+|	     | S458A	         | Wild bird | 6  |	66    |	9.1     |
+|	     | L488M	         | Broiler	 | 1  |	87    |	1.1     |
 |	     | L416F	         | Pig	     | 1  |	75    |	1.3     |
-|	     | H516Y	         | Wild bird | 1  |	65    |	1.5     |
-|	     | D475E	         | Broiler	 | 13 |	88    |	14.8    |
+|	     | H516Y	         | Wild bird | 1  |	66    |	1.5     |
+|	     | D475E	         | Broiler	 | 13 |	87    |	14.9    |
 |	     | D475E	         | Red fox	 | 2	| 52    |	3.8     |
 |	     | D463N	         | Pig	     | 1	| 75    |	1.3     |
-|	     | A512T	         | Broiler	 | 1	| 88    |	1.1     |
+|	     | A512T	         | Broiler	 | 1	| 87    |	1.1     |
 
@@
 ## Table S6: The most abundant (n > 2) substitutions in MarA, MarR, RpoB and SoxR.
+The table presents the most abundant identified amino acid substitutions in MarA, MarR, RpoB and SoxR.
 The percentage is based on how many of the total isolates (Total) with substitutions in the respective gene (n) had the described substitution.
@@
                                   ignore.case = T) == TRUE, "No", "Yes")) %>%
+    mutate(gene2 = case_when(
+      grepl("catA|cmlA|floR|aadA|aph|dfrA|sul|tet|lnuF|msrE|mph|ermB|floR", Gene) == TRUE ~ paste0("\\textit{", Gene, "}"),
+                           Gene == "blaCMY2" ~ "\\textit{bla}\\textsubscript{CMY-2}",
+                           Gene == "blaTEM1A" ~ "\\textit{bla}\\textsubscript{TEM-1A}",
+                           Gene == "blaTEM1B" ~ "\\textit{bla}\\textsubscript{TEM-1B}",
+                           Gene == "blaTEM1C" ~ "\\textit{bla}\\textsubscript{TEM-1C}",
+                           Gene == "blaSHV2" ~ "\\textit{bla}\\textsubscript{SHV-2}",
+                           Gene == "blaSHV12" ~ "\\textit{bla}\\textsubscript{SHV-12}",
+                           Gene == "blaCTXM1" ~ "\\textit{bla}\\textsubscript{CTX-M-1}",
+                           Gene == "blaCTXM15" ~ "\\textit{bla}\\textsubscript{CTX-M-15}",
+                           Gene == "blaCTXM32" ~ "\\textit{bla}\\textsubscript{CTX-M-32}",
+                           Gene == "blaCTXM55" ~ "\\textit{bla}\\textsubscript{CTX-M-55}",
+                           Gene == "aac3IIa" ~ "\\textit{aac(3')-IIa}",
+                           Gene == "aac3IId" ~ "\\textit{aac(3')-IId}")) %>%
   arrange(-Percent) %>%
-  kable(booktabs = TRUE) %>%
-  kable_styling(bootstrap_options = c("striped", "hover")) %>%
-  column_spec(1, italic = T)
+  select(gene2, everything(), -Gene) %>%
+  rename("Gene" = gene2) %>%
+  kable(booktabs = TRUE, escape = FALSE) %>%
+  kable_styling()
 
@@
 \newpage
-## Table S9: Occurrence of antimicrobial resistance among the selected quinolone resistant isolates.
-The table presents the percent (%) occurrence of AMR among the selected isolates (n=280) identified through MIC values in the Norwegian monitoring program from 2006 to 2017.
-The cut off values used were defined by EUCAST.
-\
 
-```{r, echo=FALSE, message=FALSE}
-t1 <- read_delim("../data/phenotypic_resistance.txt",
-           delim = "\t")
+## Table S9: Co-resistance
+The table presents the number of isolates with each combination of plasmid mediated resistance genes, excluding combinations represented by only a single isolate.
 
-names(t1)[2] <- paste0(names(t1)[2],"\\textsuperscript{a}")
-names(t1)[3] <- paste0(names(t1)[3],"\\textsuperscript{b}")
-names(t1)[4] <- paste0(names(t1)[4],"\\textsuperscript{c}")
-names(t1)[5] <- paste0(names(t1)[5],"\\textsuperscript{d}")
-names(t1)[6] <- paste0(names(t1)[6],"\\textsuperscript{e}")
+```{r, message=FALSE}
 
-t1 %>%
+merged_plasmid_genes <- read_delim("../merged_acquired_report.txt",
+                                   delim = ",") %>%
+  select(-c("mph","msrE","ermB","lnuF"))
+
+mic_data2 <- read_delim("../data/new_mic_data.txt",
+                       delim = "\t") %>%
+  filter(!id %in% ex_samples) %>%
+  select(-antres) %>%
+  mutate(SMX_R = if_else(SMX > 64, 1, 0),
+         TMP_R = if_else(TMP > 2, 1, 0),
+         CIP_R = if_else(CIP > 0.06, 1, 0),
+         TET_R = if_else(TET > 8, 1, 0),
+         NAL_R = if_else(NAL > 16, 1, 0),
+         CTX_R = if_else(CTX > 0.25, 1, 0),
+         CHL_R = if_else(CHL > 16, 1, 0),
+         AMP_R = if_else(AMP > 8, 1, 0),
+         GEN_R = if_else(GEN > 2, 1, 0)) %>%
+  select(id, contains("_R"))
+
+AM_data <- mic_data2 %>%
+  mutate(quin_R = if_else(CIP_R == 1 | NAL_R == 1, 1, 0)) %>%
+  select(-c(CIP_R, NAL_R)) %>%
+  mutate(antres = rowSums(.[,2:9])) %>%
+  mutate(MDR = if_else(antres >= 3, 1, 0)) %>%
+  left_join(isolate_data[,c("id","species")]) %>%
+  left_join(mic_data, by = "id") %>%
+  left_join(merged_plasmid_genes)
+
+AM_data[,c(23:39)] %>%
+  group_by_all() %>%
+  mutate(group = group_indices(),
+         n = length(group)) %>%
+  ungroup() %>%
+  mutate(PMQR = if_else(qnr == 1 | qepA == 1, 1, 0)) %>%
+  gather(gene, value, -c(n, PMQR, group)) %>%
+  filter(value == 1) %>%
+  mutate(gene2 = case_when(gene %in% c("catA","cmlA","floR",
+                                       "qepA","aadA","aph",
+                                       "dfrA","qnr","sul",
+                                       "tet") ~ paste0("\\textit{", gene, "}"),
+                           gene == "blaCMY" ~ "\\textit{bla}\\textsubscript{CMY}",
+                           gene == "blaTEM" ~ "\\textit{bla}\\textsubscript{TEM}",
+                           gene == "blaSHV" ~ "\\textit{bla}\\textsubscript{SHV}",
+                           gene == "blaCTXM" ~ "\\textit{bla}\\textsubscript{CTX-M}",
+                           gene == "aac3II" ~ "\\textit{aac(3')-II}")) %>%
+  group_by(PMQR, group) %>%
+  summarise_all(list(func_paste)) %>%
+  ungroup() %>%
+  mutate(PMQR = if_else(PMQR == 1, "PMQR positive", "PMQR negative"),
+         n = as.numeric(n)) %>%
+  filter(n > 1) %>%
+  select(PMQR, gene2, n) %>%
+  mutate(gene_n = sapply(strsplit(gene2,","),FUN=function(x){length(x[x!="Null"])})) %>%
+  arrange(PMQR, gene_n) %>%
+  select(-c(PMQR, gene_n)) %>%
+  rename("Combination of genes" = gene2,
+         "Number of isolates" = n) %>%
   kable(booktabs = TRUE, escape = FALSE) %>%
-  kable_styling(full_width = TRUE, position = "center") %>%
-  footnote(alphabet = c("87 isolates",
-                        "75 isolates",
-                        "52 isolates",
-                        "66 isolates",
-                        "280 isolates"))
+  kable_styling() %>%
+  pack_rows("PMQR Negative", 1,33, latex_gap_space = "1em") %>%
+  pack_rows("PMQR Positive", 34,43, latex_gap_space = "1em")
+
 ```
@@
 
-## Figure S1: Patristic distances
-Patristic distances were calculated from the core gene alignment SNP tree. The labelled clades are the groups further investigated.
-\
-\
-\
 ```{r fig.width=12, fig.height=12, warning=FALSE, message=FALSE,echo=FALSE}
@@
 
+## Figure S1: Clade selection based on patristic distances
+The figure presents the overall phylogenetic tree (left) and the calculated patristic distances (right) as less than (<) or higher than (>) the specified cutoff value (0.003).
+The denoted clades were further investigated with phylogenetic analyses.
+Patristic distances were calculated from the core gene SNP tree.
+
 \newpage
-## Figure S2: Distance distribution plot
-The figure shows the distribution of the expected values with the observed values for each animal species denoted as an arrow.
-Red fox results is excluded from the plot as the observed distance values were too high to visualize.
-\
-\
-\
+
+```{r echo=FALSE}
+res_mechanisms <- read.table("../data/MDS_data.txt",
+                       sep = "\t",
+                       header = TRUE,
+                       stringsAsFactors = FALSE) %>%
+  column_to_rownames("id")
+
+bin_dist <- dist(res_mechanisms,
+                 method = "binary")
+
+nmds_df <- metaMDS(bin_dist,
+                   trymax = 200,
+                   trace = FALSE) # silence the output
+
+stressplot(nmds_df)
+
+```
+
+## Figure S2: Stressplot
+Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types.
+The stressplot visualizes the goodness of fit.
+
+\newpage
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
+knitr::include_graphics("../figures/corr_chrom_vs_plasmid.png")
+```
+
+## Figure S3: Correlations between chromosomal genes and plasmid mediated genes
+The figure presents the Pearson correlation values for amino acid substitutions in GyrA, ParC, ParE, MarA, MarR, RpoB, and SoxR against all  plasmid mediated genes identified.
+Grey = negative correlation, green = positive correlation.
+Black lines denote 95% confidence intervals.
+Red stars denote significant correlations.
+The plasmid mediated genes are grouped based on gene family: _dfrA_ = _dfrA1_, _dfrA5_, _dfrA8_, _dfrA12_, _dfrA14_, and _dfrA17_.
+_tet_ = _tetA_, _tetB_, and _tetD_.
+_sul_ = _sul1 - 3_.
+_aph_ = _aph3Ia_, _aph3Ib_, and _aph6Id_.
+_aadA_ = _aadA1_, _aadA2_, _aadA5_, _aadA12_, _aadA13_, and _aadA22_.
+_AAC(3')-II_ = _AAC(3')-IIa_ and _AAC(3')-IId_.
+_bla_~TEM~ = _bla_~TEM-1-A~ – _bla_~TEM-1-C~.
+_bla_~SHV~ = _bla_~SHV-2~ and _bla_~SHV-12~.
+_bla_~CTX-M~ = _bla_~CTX-M-1~, _bla_~CTX-M-15~, _bla_~CTX-M-32~, and _bla_~CTX-M-55~.
+_qnr_ = _qnrA1_, _qnrB19_, _qnrS1_, _qnrS2_, and _qnrS4_.
+
+\newpage
+
 ```{r echo=FALSE}
@@
   rename("isol1" = `snp-dists 0.6.3`)
-
 # Set row/column values to unique species ID
@@
   select(isol1)
-
 col_id <- as.data.frame(names(total_dist)) %>%
@@
   mutate(isol1 = ifelse(isol1 == "NA_1", "isol1", isol1))
-
 total_dist$isol1 <- row_id$isol1
 names(total_dist) <- col_id$isol1
-
 # Create data frame for iteration analysis
@@
          "group2" = species2)
-
 # Run iterations
 df <- iterate(run_df, runs = 1000, method = "median") 
-
 # Filter out the observed values
@@
                        group1 %in% c("Broiler", "Pig", "Wild bird"))
-
 # Calculate non-exact p-values
@@
   split(f = .$group1)
-
 # Plot results, exclude red fox (too high values for the plot to be visible)
@@
   guides(color = FALSE) +
-  labs(x = "Min. number of SNPs to closest isolate",
+  labs(x = "SNP distance",
        y = "Density") +
@@
   scale_color_manual(values = palette)
+```
 
-```
+## Figure S4: Iteration analysis results
+The figure presents the observed median minimum SNP distance values for each animal species (coloured arrows) in relation to the expected values from the iteration analysis.
+Red fox results is excluded from the plot as the observed distance values were too high to visualize.
+The figure shows that isolates from broilers are more closely aggregated in the phylogenetic tree than what is randomly expected.
 
@@
 
-## Figure S3: Stressplot
-Non-metric multidimensional scaling (NMDS) was used to see if the distribution of quinolone resistance mechanisms from some sequence types were more homogenous than in other sequence types.
-The stressplot visualizes the goodness of fit.
-\
-\
-\
-```{r echo=FALSE}
-res_mechanisms <- read.table("../data/MDS_data.txt",
-                       sep = "\t",
-                       header = TRUE,
-                       stringsAsFactors = FALSE) %>%
-  column_to_rownames("id")
-
-bin_dist <- dist(res_mechanisms,
-                 method = "binary")
-
-nmds_df <- metaMDS(bin_dist,
-                   trymax = 200,
-                   trace = FALSE) # silence the output
-
-stressplot(nmds_df)
-
+```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
+knitr::include_graphics("../figures/complete_st_trees/CladeA_ST162A.png")
 ```
-
-\newpage
-## Figure S4: Correlation between chromosomal genes and plasmid mediated genes
-The figure presents the Pearson correlation values for amino acid substitutions in GyrA, ParC, ParE, MarA, MarR, RpoB, and SoxR against all  plasmid mediated genes identified.
-Grey = negative correlation, green = positive correlation.
-Black lines denote 95% confidence intervals.
-Red stars denote significant correlations.
-The plasmid mediated genes are grouped based on gene family: _dfrA_ = _dfrA1_, _dfrA5_, _dfrA8_, _dfrA12_, _dfrA14_, and _dfrA17_.
-_tet_ = _tetA_, _tetB_, and _tetD_.
-_sul_ = _sul1 - 3_.
-_aph_ = _aph3Ia_, _aph3Ib_, and _aph6Id_.
-_aadA_ = _aadA1_, _aadA2_, _aadA5_, _aadA12_, _aadA13_, and _aadA22_.
-_AAC(3')-II_ = _AAC(3')-IIa_ and _AAC(3')-IId_.
-_bla_~TEM~ = _bla_~TEM-1-A~ – _bla_~TEM-1-C~.
-_bla_~SHV~ = _bla_~SHV-2~ and _bla_~SHV-12~.
-_bla_~CTX-M~ = _bla_~CTX-M-1~, _bla_~CTX-M-15~, _bla_~CTX-M-32~, and _bla_~CTX-M-55~.
-_qnr_ = _qnrA1_, _qnrB19_, _qnrS1_, _qnrS2_, and _qnrS4_.
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
-knitr::include_graphics("../figures/corr_chrom_vs_plasmid.png")
-```
-
-\newpage
 
@@
 Shared genome: 88.4%.
-\
-\
-\
+
+\newpage
+
 ```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
-knitr::include_graphics("../figures/complete_st_trees/CladeA_ST162A.png")
+knitr::include_graphics("../figures/complete_st_trees/CladeC_ST744.png")
 ```
 
-\newpage
 ## Figure S6: Maximum likelihood tree, Clade C
@@
 Shared genome: 88.4%.
-\
-\
-\
+
+\newpage
+
 ```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
-knitr::include_graphics("../figures/complete_st_trees/CladeC_ST744.png")
+knitr::include_graphics("../figures/complete_st_trees/CladeD_ST10.png")
 ```
 
-\newpage
 ## Figure S7: Maximum likelihood tree, Clade D
@@
 Shared genome: 87.2%.
-\
-\
-\
+
+\newpage
+
 ```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
-knitr::include_graphics("../figures/complete_st_trees/CladeD_ST10.png")
+knitr::include_graphics("../figures/complete_st_trees/CladeE_ST355.png")
 ```
 
-\newpage
 ## Figure S8: Maximum likelihood tree, Clade E
@@
 Shared genome: 90.2%.
-\
-\
-\
-```{r message=FALSE, warning=FALSE, echo=FALSE, out.width="90%"}
-knitr::include_graphics("../figures/complete_st_trees/CladeE_ST355.png")
-```
 
@@
 # Section 3: Metadata and reference sequences
-
-The whole of the supplementary material can also be accessed on github (https://norwegianveterinaryinstitute.github.io/qm_wgs_4species/).
 
@@
 ## 3.4: QRDR determination of GyrB, ParC and ParE
-The K12 reference sequence on the whole protein is listed as the first entry in each file. The rest of the fasta entries are taken from Jiménez Gómez et al. 2004.
+The K12 reference sequence on the whole protein is listed as the first entry in each file. The rest of the fasta entries are taken from Jiménez Gómez et al. 2004 [1].
 
@@
 # References
-[1] RCoreTeam. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/. 2018.
 
-[2] Wickham H, François R, Henry L, Müller K. dplyr: A Grammar of Data Manipulation [Internet]. 2019. Available from: https://cran.r-project.org/package=dplyr
-
-[3] Wickham H, Henry L. tidyr: Easily Tidy Data with “spread()” and “gather()” Functions [Internet]. 2019. Available from: https://cran.r-project.org/package=tidyr
-
-[4] Wickham H, Chang W, Henry L, Pedersen TL, Takahashi K, Wilke C, et al.
-ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics [Internet].
-2019.
-Available from: https://cran.r-project.org/package=ggplot2
-
-[5] Henry L, Wickham H. purrr: Functional Programming Tools [Internet]. 2019. Available from: https://cran.r-project.org/package=purrr
-
-[6] Müller K, Wickham H. tibble: Simple Data Frames [Internet]. 2019. Available from: https://cran.r-project.org/package=tibble
-
-[7] Pagès H, Aboyoun P, Gentleman R, DebRoy S. Biostrings: Efficient manipulation of biological strings. 2019.
-
-[8] Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, et al. vegan: Community Ecology Package [Internet]. 2019. Available from: https://cran.r-project.org/package=vegan
-
-[9] Wickham H. stringr: Simple, Consistent Wrappers for Common String Operations [Internet]. 2019. Available from: https://cran.r-project.org/package=stringr
-
-[10] Maechler M, Rousseeuw P, Struyf A, Hubert M.
-cluster: “Finding Groups in Data”: Cluster Analysis Extended Rousseeuw et al.
-[Internet].
-2019.
-Available from: https://cran.r-project.org/package=cluster
-
-[11] Paradis E, Blomberg S, Bolker B, Brown J, Claude J, Cuong HS, et al. ape: Analyses of Phylogenetics and Evolution [Internet]. 2019. Available from: https://cran.r-project.org/package=ape
-
-[12] Yu G, Smith DK, Zhu H, Guan Y, Lam TT-Y.
-Ggtree: a package for visualization and annotation of phylogenetic trees with their covariates and other associated data.
-McInerny G, editor.
-Methods Ecol Evol [Internet].
-2017 Jan;8(1):28–36.
-Available from: http://doi.wiley.com/10.1111/2041-210X.12628
-
-[13] Neuwirth E. RColorBrewer: ColorBrewer Palettes [Internet]. 2014. Available from: https://cran.r-project.org/package=RColorBrewer
-
-[14] Wilke CO. cowplot: Streamlined Plot Theme and Plot Annotations for “ggplot2” [Internet]. 2019. Available from: https://cran.r-project.org/package=cowplot
-
-[15] Xie Y. knitr: A General-Purpose Package for Dynamic Report Generation in R [Internet]. 2019. Available from: https://cran.r-project.org/package=knitr
-
-[16] Zhu H. kableExtra: Construct Complex Table with “kable” and Pipe Syntax [Internet]. 2019. Available from: https://cran.r-project.org/package=kableExtra
-
-[17] Xiao N. ggsci: Scientific Journal and Sci-Fi Themed Color Palettes for “ggplot2” [Internet]. 2018. Available from: https://cran.r-project.org/package=ggsci
-
-[18] Kassambara A. fastqcr: Quality Control of Sequencing Data [Internet]. 2019. Available from: https://cran.r-project.org/package=fastqcr
-
-[19] Wickham H. scales: Scale Functions for Visualization [Internet]. 2018. Available from: https://cran.r-project.org/package=scales
-
-[20] Garnier S. viridis: Default Color Maps from “matplotlib” [Internet]. 2018. Available from: https://cran.r-project.org/package=viridis
-
-[21] Jombart T., Dray S. (2008) adephylo: exploratory analyses for the phylogenetic comparative method.
+[1] P.A. JG, J.E. G de los R, A. RM, P.  de PR, R. GA. 2004. Molecular basis of quinolone resistance in Escherichia coli from wild birds. Canadian Journal of Veterinary Research 68:229–231
